### 2020-02-19 documentation/systemModel/systemModel.tex
adding system model document

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx, neuralnetwork,tikz}
\usepackage{float}
\usetikzlibrary{positioning}

\bibliographystyle{alpha}
\graphicspath{ {./figures/} }
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

\tikzstyle{state}=[shape=circle,draw=blue!30,fill=blue!10]
\tikzstyle{observation}=[shape=rectangle,draw=orange!30,fill=orange!10]
\tikzstyle{lightedge}=[<-, dashed]
\tikzstyle{mainstate}=[state, thick]
\tikzstyle{mainedge}=[<-, thick]
\tikzstyle{block} = [draw,rectangle,thick,minimum height=2em,minimum width=2em]
\tikzstyle{sum} = [draw,circle,inner sep=0mm,minimum size=2mm]
\tikzstyle{connector} = [->,thick]
\tikzstyle{line} = [thick]
\tikzstyle{branch} = [circle,inner sep=0pt,minimum size=1mm,fill=black,draw=black]
\tikzstyle{guide} = []
\tikzstyle{snakeline} = [connector, decorate, decoration={pre length=0.2cm,
                         post length=0.2cm, snake, amplitude=.4mm,
                         segment length=2mm},thick, magenta, ->]

\title{Free Probability Analysis of Secrecy Capacity in Relay Networks}
\author{Peter Hartig}

\begin{document}
\maketitle

\bibliography{bibliography}
\end{document}

### 2020-03-10 documentation/systemModel/systemModel.tex
adding notes for background research and updating gitignore

@@
 
-\title{Free Probability Analysis of Secrecy Capacity in Relay Networks}
+\title{Free Probability Analysis of Capacity in Relay Networks}
 \author{Peter Hartig}
@@
 \maketitle
+\chapter{Introduction}
+The goal of this work is to develop a framework to analyze the capacity of communication networks that incorporate intelligent reflective surfaces. 
+\par
+Questions
+\begin{enumerate}
+\item Why does this not fit into the current tools we have for analyzing MIMO channels?
+\item Are there similar results for amplify and forward relay networks? If so can we use the same tools?
+\item For deterministic channels, do we know the capacity for this type of irs or amplify and forward relaying?
+Check for this in Teletar
+\item How is the IRS different than an amplify and forward relay? Add details 
+\item How would this be different from Lucinda's work? 
+They assume decode and forward?
+\item Why do we need free probability and why aren't results to normal RMT sufficient.
+\end{enumerate}
+\chapter{System Model}
+\section{Lucinda's System Model}
+\begin{itemize}
+\item Decode and forward relaying so at each relay, the original signal is perfectly detected and then retransmitted after a precoding. 
+\item Point to point with all relays having the same number of antennas and receiver usering having at least as many antennas as the transmitting user.
+\item Assumes a single hop in the system. TODO check if it is common for IRS systems to have multiple hops
+\item Allows for multiple relays to be in the single hop. L total relays
+\item Assumes Source knowns no CSI
+\item Assumes relays have perfect CSI of all channels. TOD Discuss if this is this realistic for such large systems and if not, if we can add uncertainty to model easily.
+\item Not power allocation to the different antennas is performed?
+\item 
+\end{itemize}
+
+\section{General notes}
+\begin{itemize}
+\item Why is the assumption of asymmetric important and what exactly does asymmetric mean?
+\item Can any of this be used in a MIMO broadcast setting where no processing can be done at the receiver?
+\item What CSI will be known at the IRS?
+\item Need to look at optimal matched filtering for precoding. Will not be possible for IRS and also might not even assumed that this CSI would be known.
+\item Need to look at how to detect at the final end point for amplify and forward methods and IRS methods in general.
+\item What kind of noise will be present at the detector if we can decode and forward
+\item TODO derive standard ergodic capacity of MIMO channel. How the standard form with the determinant decomposes into the eigenvalue product. Also look at LogDet in problem 5 of Tutorial 1 of RMT
+\item For Marcenco Pasteur, why is it reasonable to assumed variance is one over NR
+%\item 
+\end{itemize}
 

### 2020-03-10 documentation/systemModel/systemModel.tex
working on system model

@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt,a4paper]{report}
 \usepackage[utf8]{inputenc}
@@
 \begin{enumerate}
-\item Why does this not fit into the current tools we have for analyzing MIMO channels?
+%\item Why does this not fit into the current tools we have for analyzing MIMO channels? Aspects of this that we currently lack. 
 \item Are there similar results for amplify and forward relay networks? If so can we use the same tools?
@@
 \item Why do we need free probability and why aren't results to normal RMT sufficient.
+\item Should we even consider noise being added at IRS? Based on previous work, no noise should be added at the IRS.
+\item Explain why it is reasonable to consider a single coefficient of channel and no multipath.
+\item Can phases be design such that a specific receiver detects a single transmitted symbol?
+\item Is it reasonable to assume that an IRS will know sufficient CSI?
+\item Need to show if other user interference can be cancelled using IRS?
 \end{enumerate}
@@
 \item Why is the assumption of asymmetric important and what exactly does asymmetric mean?
+Means that the attenuation coefficient may not be the same in the link from relay to detector.
+This makes finding the distribution hard for some reason detailed in lucinda's paper.
 \item Can any of this be used in a MIMO broadcast setting where no processing can be done at the receiver?
@@
 \item Need to look at how to detect at the final end point for amplify and forward methods and IRS methods in general.
-\item What kind of noise will be present at the detector if we can decode and forward
+\item What kind of noise will be present at the detector if we cant decode and forward. 
 \item TODO derive standard ergodic capacity of MIMO channel. How the standard form with the determinant decomposes into the eigenvalue product. Also look at LogDet in problem 5 of Tutorial 1 of RMT
 \item For Marcenco Pasteur, why is it reasonable to assumed variance is one over NR
+\item IRS wouldnt have any impact in the case of a guassian channel unless it knows the incoming phases. The phase of a gaussian is uniform so adding a shift won't change this. 
+\item Generally will not be able to optimize the secrecy capacity of the phase matrix because finding the CSI to eavedroppers would be hard.
+This would also impact the resulting estimate for the SNR based upon any noise injected into the system.
+
 %\item 
@@
 
+\section{My System Model}
+\begin{enumerate}
+\item IRS: A surface made of individual elements each of which reflect the single signal received by the surface with a unique phase shift.
+\item Most general: Individual users transmit simultaneously to the first set of IRS's.
+These relays hop through an arbitrary number of IRS's.
+The final set of IRS's reflect the signal out to a set of independent users.
+Possible?
+(For half duplex need to make sure number of relay sets makes sense)
+\item How do we describe the signal coming off of an IRS? Large diagonal matrix with each reflecting element represented as one elements (one phase shift) along the diagonal. 
+\item
+Each element of the IRS can apply a phase shift to the incoming signal.
+\item 
+A set of $T$ transmitters is assumed to transmit $T$ statistically independent symbols in time period T1.
+This signal is then transmitted through $N$ sets of IRS's, each with $R_{n,i}$, $i \in [1...
+\; R^n_L]$ reflecting surfaces.
+The signal received at element $k$ of IRS $(n,i)$ is
+\begin{equation}
+s^{n}_{i,k} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{h}^{n-1}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
+\end{equation}
+with $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between the $j$th IRS in the $(n-1)$st set of IRS's and $k$th element of the $i$th IRS in the $n$th set of IRS's.
+
+\item
+The vector of received signals for all elements at IRS $(n,i)$ is therefore
+\begin{equation}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{H}^{n-1}_{j,i})^T\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
+\end{equation}
+
+with $\mathbf{s}^{n}_{i} \in \mathbb{C}^{n_L}$
+
+\end{enumerate}
+
 \bibliography{bibliography}

### 2020-03-10 documentation/systemModel/systemModel.tex
TODOs for tomorrow added

@@
 \item Are there similar results for amplify and forward relay networks? If so can we use the same tools?
-\item For deterministic channels, do we know the capacity for this type of irs or amplify and forward relaying?
-Check for this in Teletar
-\item How is the IRS different than an amplify and forward relay? Add details 
-\item How would this be different from Lucinda's work? 
-They assume decode and forward?
-\item Why do we need free probability and why aren't results to normal RMT sufficient.
-\item Should we even consider noise being added at IRS? Based on previous work, no noise should be added at the IRS.
-\item Explain why it is reasonable to consider a single coefficient of channel and no multipath.
+\item For deterministic channels, do we know the capacity for this type of irs or amplify and forward relaying? TODO
+\item Why do we need free probability and why aren't results to normal RMT sufficient. TODO
+\item Explain why it is reasonable to consider a single coefficient of channel and no multipath. TODO
 \item Can phases be design such that a specific receiver detects a single transmitted symbol?
 \item Is it reasonable to assume that an IRS will know sufficient CSI?
-\item Need to show if other user interference can be cancelled using IRS?
+\item Need to show if other user interference can be cancelled using IRS? yes for the case of complex 
+\item Is it reasonable to think of complex as a two channel real MIMO system?
 \end{enumerate}
@@
 \end{itemize}
-
-\section{General notes}
+My TODO
+\begin{enumerate}
+	\item
+		Really have nice formalization of the final channel whose AED is desired.
+	\item
+		Then start building up with SISO one hop first then relay and finally up to the point where some form of FPT is
+		required.
+	\item
+		Look at algo from Lucinda Paper
+	\item
+		Look at the other article Saba sent
+	\item
+		Start looking at how to incorporate ideas from the pilot decontamination into this
+\end{enumerate}
 \begin{itemize}
@@
 \item Can any of this be used in a MIMO broadcast setting where no processing can be done at the receiver?
-\item What CSI will be known at the IRS?
-\item Need to look at optimal matched filtering for precoding. Will not be possible for IRS and also might not even assumed that this CSI would be known.
+\item Need to look at optimal matched filtering for precoding. Will not be possible for IRS and also might not even assumed that this
+	CSI would be known. Indeed it won't be able to do this precoding, so what do we lose?
 \item Need to look at how to detect at the final end point for amplify and forward methods and IRS methods in general.
-\item What kind of noise will be present at the detector if we cant decode and forward. 
 \item TODO derive standard ergodic capacity of MIMO channel. How the standard form with the determinant decomposes into the eigenvalue product. Also look at LogDet in problem 5 of Tutorial 1 of RMT

### 2020-03-12 documentation/Meeting Notes/3_12_20.tex
added meeting notes from today

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx, neuralnetwork,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex

\bibliographystyle{alpha}

\author{Peter Hartig}

\begin{document}

\chapter{To Review}
\section{System Model For IRS Work}
\begin{enumerate}
\item 
	We assume non-frequency selective channels using techniques like OFDM. (Is this always going to be reasonable? As in the IRS/MIMO capacity optimization paper)

\item 
	We assume a transmitter without channel state information and thus power is distributed equally over all antennas such that
	\begin{equation}
	\mathbf{Q_{\mathbf{x}}} = \frac{P^{\text{Total}}}{N_T}
	\end{equation}
	in which $P^{\text{Total}}$ is the power available at the transmitter, and $N_T$ is the number of antennas at the transmitter.

\item Intelligent Reflective Surfaces (IRS) are placed between transmitter and surface made of individual elements each of which reflects the single signal received at the element with a unique phase shift.
\item Individual users transmit simultaneously to the first set of IRS's. This transmit signal hops through an arbitrary number of IRS's. The final set of IRS's reflect the signal to a set of independent users. Possible? (For half duplex need to make sure number of relay sets makes sense)

\item
Each element of the IRS can apply a phase shift to the incoming signal.
\item 
A set of $T$ transmitters is assumed to transmit $T$ statistically independent symbols in time period T1. This signal is then transmitted through $N$ sets of IRS's, each with $R_{n,i}$, $i \in [1... \; R^n_L]$ reflecting surfaces. The signal received at element $k$ of IRS $(n,i)$ is 
\begin{equation}
s^{n}_{i,k} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{h}^{n-1}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
\end{equation}
with $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between the $j$th IRS in the $(n-1)$st set of IRS's and $k$th element of the $i$th IRS in the $n$th set of IRS's.

\item
The vector of received signals for all elements at IRS $(n,i)$ is therefore
\begin{equation}
\mathbf{s}^{n}_{i} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{H}^{n-1}_{j,i})^T\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
\end{equation}

\item For a specific receiving user with potentially multiple receiver antennas, the above results in a recursive form
\begin{equation}
\mathbf{s}^{n}_{i} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{H}^{n-1}_{j,i})^T\boldsymbol{\Theta}^{n-1}_{j} \left[
\sum_{k = 1}^{R^{n-1}_L} (\mathbf{H}^{n-2}_{k,i})^T\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
\end{equation}
with

\begin{equation}
\mathbf{s}^{1}_{i} = (\mathbf{H}^{0}_{i})^T\boldsymbol{I}\mathbf{x}
\end{equation}
 in which $\boldsymbol{I}$ is the identify matrix and $\mathbf{x}$ is the original vector of transmitted symbols.
 Factoring out $\mathbf{x}$, we obtain an equivalent channel through which $\mathbf{x}$ passes.

\item
	Note that this model assumes sufficient spacing between sets of IRSs such that we consider the signal reflected in set $(n-1)$ only at set $n$ and not at any set $k$, for any $n<k<(n-1)$.
\item
	Each IRS is assumed to have channel knowledge of up and downlink???

\end{enumerate}

\section{Notes from Lucinda's System Model}
\begin{itemize}
\item Decode and forward relaying so at each relay, the original signal is perfectly detected and then re-transmitted after a precoding. 
\item Point to point with all relays having the same number of antennas and receiver having at least as many antennas as the transmitting user.
\item Assumes a single hop in the system. TODO check if it is common for IRS systems to have multiple hops
\item Allows for multiple relays to be in the single hop. L total relays
\item Assumes Source knows no CSI so naive power allocation is used
\item Assumes relays have perfect CSI of all channels. TODO Discuss if this is this realistic for such large systems and if not, if we can add uncertainty to model easily.
\end{itemize}

%\chapter{To Discuss}

\chapter{Questions}
\begin{enumerate}
\item Is it reasonable to think of complex as a two channel real MIMO system?
\end{enumerate}

My TODO
\begin{enumerate}
	\item
		Read Voiculescu paper and understand why these results cannot be applied to this situation based off of Tulino review.

	\item 
		Understand how Lucinda's paper get around this problem (also see the other paper she cites) and then see how this might be done differently as was done in the pilot decontamination paper.

	\item Begin translating the above system model into this framework to find AED.
	\item Are there similar results for amplify and forward relay networks? If so can we use the same tools?
	\item Adapt system model for tapped channel

\end{enumerate}

\bibliography{bibliography}
\end{document}

### 2020-03-12 documentation/systemModel/systemModel.tex
added meeting notes from today

@@
 \usetikzlibrary{positioning}
+\input defs.tex
 
@@
 \item Why do we need free probability and why aren't results to normal RMT sufficient. TODO
-\item Explain why it is reasonable to consider a single coefficient of channel and no multipath. TODO
-\item Can phases be design such that a specific receiver detects a single transmitted symbol?
+\item Can phases be designed such that a specific receiver detects a single transmitted symbol?
 \item Is it reasonable to assume that an IRS will know sufficient CSI?
-\item Need to show if other user interference can be cancelled using IRS? yes for the case of complex 
+\item Need to show if other user interference can be canceled using IRS? yes for the case of complex 
 \item Is it reasonable to think of complex as a two channel real MIMO system?
@@
 \chapter{System Model}
-\section{Lucinda's System Model}
+\section{Notes from Lucinda's System Model}
 \begin{itemize}
-\item Decode and forward relaying so at each relay, the original signal is perfectly detected and then retransmitted after a precoding. 
-\item Point to point with all relays having the same number of antennas and receiver usering having at least as many antennas as the transmitting user.
+\item Decode and forward relaying so at each relay, the original signal is perfectly detected and then re-transmitted after a precoding. 
+\item Point to point with all relays having the same number of antennas and receiver having at least as many antennas as the transmitting user.
 \item Assumes a single hop in the system. TODO check if it is common for IRS systems to have multiple hops
 \item Allows for multiple relays to be in the single hop. L total relays
-\item Assumes Source knowns no CSI
+\item Assumes Source knows no CSI
 \item Assumes relays have perfect CSI of all channels. TOD Discuss if this is this realistic for such large systems and if not, if we can add uncertainty to model easily.
@@
 \begin{enumerate}
-	\item
-		Really have nice formalization of the final channel whose AED is desired.
 	\item
@@
 
-\section{My System Model}
+\section{System Model For IRS Work}
 \begin{enumerate}
-\item IRS: A surface made of individual elements each of which reflect the single signal received by the surface with a unique phase shift.
-\item Most general: Individual users transmit simultaneously to the first set of IRS's.
-These relays hop through an arbitrary number of IRS's.
-The final set of IRS's reflect the signal out to a set of independent users.
+\item 
+	We assume non-frequency selective channels using techniques like OFDM. (Is this always going to be reasonable?)
+
+\item 
+	We assume a transmitter without channel state information and thus power is distributed equally over all antennas such that
+	\begin{equation}
+	\mathbf{Q_{\mathbf{x}}} = \frac{P^{\text{Total}}}{N_T}
+	\end{equation}
+	in which $P^{\text{Total}}$ is the power available at the transmitter, and $N_T$ is the number of antennas at the transmitter.
+
+\item Intelligent Reflective Surfaces (IRS) are placed between transmitter and surface made of individual elements each of which reflects the single signal received at the element with a unique phase shift.
+\item Individual users transmit simultaneously to the first set of IRS's.
+This transmit signal hops through an arbitrary number of IRS's.
+The final set of IRS's reflect the signal to a set of independent users.
 Possible?
 (For half duplex need to make sure number of relay sets makes sense)
-\item How do we describe the signal coming off of an IRS? Large diagonal matrix with each reflecting element represented as one elements (one phase shift) along the diagonal. 
+
 \item
@@
 
-with $\mathbf{s}^{n}_{i} \in \mathbb{C}^{n_L}$
+\item For a specific receiving user with potentially multiple receiver antennas, the above results in a recursive form
+\begin{equation}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{H}^{n-1}_{j,i})^T\boldsymbol{\Theta}^{n-1}_{j} \left[
+\sum_{k = 1}^{R^{n-1}_L} (\mathbf{H}^{n-2}_{k,i})^T\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
+\end{equation}
+with
+
+\begin{equation}
+\mathbf{s}^{1}_{i} = (\mathbf{H}^{0}_{i})^T\boldsymbol{I}\mathbf{x}
+\end{equation}
+ in which $\boldsymbol{I}$ is the identify matrix and $\mathbf{x}$ is the original vector of transmitted symbols.
+ Factoring out $\mathbf{x}$, we obtain an equivalent channel through which $\mathbf{x}$ passes.
+
+\item
+	Note that this model assumes sufficient spacing between sets of IRSs such that we consider the signal reflected in set $(n-1)$ only at set $n$ and not at any set $k$, for any $n<k<(n-1)$.
+\end{enumerate}
+\section{Background For IRS Work}
+Now that we known the characteristics of the channel for which we need the capacity, build of the tools to find this.
+\begin{enumerate}
+\item First show the capacity for a deterministic point-to-point communication channel (with eigenvalues included).
+\item Next show how this extends to the non-deterministic channel and how we end up needing the pdf of the eigenvalues for the final channel.
+\item For a non-deterministic channel in which the transmitter has no CSI knowledge and Gaussian distributed $\mathbf{x}$ the capacity ($\mathcal{C}$) of a point-to-point channel $\mathbf{H}$ is 
+\begin{equation}
+\Capacity{(H)} = \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P^{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
+\end{equation}
+%Given the distribution for the eigenvalues of $\mathbf{H}\mathbf{H}^H$ this can be rewritten as 
+%\begin{equation}
+%\Capacity{(H)} = N \left[\Log\left(|\mathbf{I}_{N_R}+\frac{P^{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
+%
+%\end{equation}
+
+\item Look at if there is an AED definition of capacity for amplify and forward relays.
+If so, consider the IRS as a special case of this?
+This should just be equivalent to point-to-point for the deterministic case as long as we can factor out the original transmitted symbol vector.
+This addition of noise at the relay may cause issues but if the channel is gaussian this might be okay as along as the final noise variance is adjusted.
 

### 2020-03-24 documentation/systemModel/systemModel.tex
further work to prepare to send

--- 2020-03-10 documentation/systemModel/systemModel.tex
@@
 \usepackage{amssymb}
+\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
+
 \usepackage{graphicx, neuralnetwork,tikz}
@@
 \maketitle
+\tableofcontents
 \chapter{Introduction}
-The goal of this work is to develop a framework to analyze the capacity of communication networks that incorporate intelligent reflective surfaces. 
+The goal of this work is to consider the IRS channel capacity and secrecy capacity using free probability theory.
+Specifically, this system is compared to two similar problems which have been shown to admit analysis using free probability.
 \par
-Questions
+
+\chapter{Problem Background}
+
+\section{Information Theory Tools}
+
 \begin{enumerate}
-%\item Why does this not fit into the current tools we have for analyzing MIMO channels? Aspects of this that we currently lack. 
-\item Are there similar results for amplify and forward relay networks? If so can we use the same tools?
-\item For deterministic channels, do we know the capacity for this type of irs or amplify and forward relaying?
-Check for this in Teletar
-\item How is the IRS different than an amplify and forward relay? Add details 
-\item How would this be different from Lucinda's work? 
-They assume decode and forward?
-\item Why do we need free probability and why aren't results to normal RMT sufficient.
-\item Should we even consider noise being added at IRS? Based on previous work, no noise should be added at the IRS.
-\item Explain why it is reasonable to consider a single coefficient of channel and no multipath.
-\item Can phases be design such that a specific receiver detects a single transmitted symbol?
-\item Is it reasonable to assume that an IRS will know sufficient CSI?
-\item Need to show if other user interference can be cancelled using IRS?
+\item For a random MIMO communication channel and CSI known at the transmitter and Gaussian distributed transmit signal $\mathbf{x}$, the capacity ($\mathcal{C}$) of a point-to-point channel $\mathbf{H}$ is given by
+\begin{equation}\label{capacity}
+\Capacity{(H)} = \Expect\left[\underset{tr(\mathbf{Q}_{\mathbf{xx}}) < P_{Total}}{\mathrm{max}} \;\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{\mathbf{xx}}\mathbf{H}^H|\right)\right].
+\end{equation}
+*Question: Without an analytic solution to the waterfilling problem used to choose $\mathbf{Q}_{\mathbf{xx}}$, how can we take an expectation?
+-> need to do a Monte Carlo simulation (in MIMO book) so probably will not be able to include the beamformer in any free probability solution.
+
+\item
+In the case in which no CSI is known at the transmitter and the power at the transmitter, $P_{\text{Total}}$, is distributed equally across transmit antennas Equation \ref{capacity} becomes
+\begin{equation}
+\Capacity{(H)} = \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
+\end{equation} 
+Assuming asymptotically large antenna arrays, and given the distribution for the eigenvalues of $\mathbf{H}\mathbf{H}^H$ which we denote by $p_{\lambda}(\lambda)$, Equation \ref{capacity} can be rewritten as
+\begin{equation}
+\Capacity{(H)} = N_r \Expect\left[\Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda\right)\right].
+\end{equation}
+ in which the expectation is taken over $p_{\lambda}(\lambda)$.
+
+\item The secrecy capacity of the general wire-tapped communication channel with perfect knowledge of the channels to both the legitimate receiver, $\mathbf{H} \in \mathbb{C}_{N_r \times N_t}$, and eavesdropper, $\hat{\mathbf{H}}\in \mathbb{C}_{N_e \times N_t}$, is given by
+
+\begin{gather}\label{secrecy_capacity}
+\Capacity{(H)} = \Expect\left[\underset{tr{\mathbf{Q}_{xx}} < P_{Total}}{\mathrm{max}} \;\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{xx}\mathbf{H}^H|\right) -
+\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\hat{\mathbf{H}}\mathbf{Q}_{xx}\hat{\mathbf{H}}^H|\right)
+\right] \right].
+\end{gather}
+Again assuming $P_{\text{Total}}$ is distributed equally across transmit antennas such that $\mathbf{Q}_{xx} = \frac{P_{\text{Total}}}{N_t}\mathbf{I}_{N_t}$, and that both the legitimate user and eavesdropper have sufficient antennas such that $N_r > N_t$ and $N_e >N_t$,  this can be written as
+
+\begin{gather}\label{secrecy_capacity}
+\Capacity{(H)} = \Expect\left[N_r \Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda \right) -
+N_e \Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\hat{\lambda} \right) \right]
+\end{gather}
+in which $\lambda$ and $\hat{\lambda}$ denote the eigenvalues of $\mathbf{H}\mathbf{H}^H$ and $\hat{\mathbf{H}}\hat{\mathbf{H}}^H$ respectively.
+Generally, finding capacity will require knowledge of the joint asymptotic eigenvalue distribution $p_{\lambda}(\lambda,\hat{\lambda})$.
+
 \end{enumerate}
-\chapter{System Model}
-\section{Lucinda's System Model}
+\section{Free Probability/Random Matrix Theory Tools}
+
+Here the current tools available from free probability analysis are reviewed in order to clarify the types of polynomials (composite channels) which can be analyzed.
+\begin{enumerate}
+\item 
+	Transformations of the AED
+	Detail how to recover the true pdf given the Steiltjes, R/S transforms
 	\begin{itemize}
-\item Decode and forward relaying so at each relay, the original signal is perfectly detected and then retransmitted after a precoding. 
-\item Point to point with all relays having the same number of antennas and receiver usering having at least as many antennas as the transmitting user.
-\item Assumes a single hop in the system. TODO check if it is common for IRS systems to have multiple hops
-\item Allows for multiple relays to be in the single hop. L total relays
-\item Assumes Source knowns no CSI
-\item Assumes relays have perfect CSI of all channels. TOD Discuss if this is this realistic for such large systems and if not, if we can add uncertainty to model easily.
-\item Not power allocation to the different antennas is performed?
+	\item
+		For general random matrices, we can still find the eigenvalue distributions using the problem of moments
+	\end{itemize}
+
+\item 
+	Relevant matrices with known AEDs.
+	\begin{itemize}
+	\item General zero-mean iid $n\timesn$ with variance 1/$N$ -> Full Circle
+	\item Sum of the matrix above with its adjoint. The result is self-adjoint so the eigenvalues are all real.
 	\item 
 	\end{itemize}
-
-\section{General notes}
+\item
+	Free Families
 	\begin{itemize}
-\item Why is the assumption of asymmetric important and what exactly does asymmetric mean?
-Means that the attenuation coefficient may not be the same in the link from relay to detector.
-This makes finding the distribution hard for some reason detailed in lucinda's paper.
-\item Can any of this be used in a MIMO broadcast setting where no processing can be done at the receiver?
-\item What CSI will be known at the IRS?
-\item Need to look at optimal matched filtering for precoding. Will not be possible for IRS and also might not even assumed that this CSI would be known.
-\item Need to look at how to detect at the final end point for amplify and forward methods and IRS methods in general.
-\item What kind of noise will be present at the detector if we cant decode and forward. 
-\item TODO derive standard ergodic capacity of MIMO channel. How the standard form with the determinant decomposes into the eigenvalue product. Also look at LogDet in problem 5 of Tutorial 1 of RMT
-\item For Marcenco Pasteur, why is it reasonable to assumed variance is one over NR
-\item IRS wouldnt have any impact in the case of a guassian channel unless it knows the incoming phases. The phase of a gaussian is uniform so adding a shift won't change this. 
-\item Generally will not be able to optimize the secrecy capacity of the phase matrix because finding the CSI to eavedroppers would be hard.
-This would also impact the resulting estimate for the SNR based upon any noise injected into the system.
-
-%\item 
+	\item blank
 	\end{itemize}
 
-\section{My System Model}
-\begin{enumerate}
-\item IRS: A surface made of individual elements each of which reflect the single signal received by the surface with a unique phase shift.
-\item Most general: Individual users transmit simultaneously to the first set of IRS's.
-These relays hop through an arbitrary number of IRS's.
-The final set of IRS's reflect the signal out to a set of independent users.
-Possible?
-(For half duplex need to make sure number of relay sets makes sense)
-\item How do we describe the signal coming off of an IRS? Large diagonal matrix with each reflecting element represented as one elements (one phase shift) along the diagonal. 
 \item
-Each element of the IRS can apply a phase shift to the incoming signal.
+	Matrix Composition: Detail requirements on the variables
+	\begin{itemize}
 	\item
-A set of $T$ transmitters is assumed to transmit $T$ statistically independent symbols in time period T1.
-This signal is then transmitted through $N$ sets of IRS's, each with $R_{n,i}$, $i \in [1...
-\; R^n_L]$ reflecting surfaces.
-The signal received at element $k$ of IRS $(n,i)$ is
-\begin{equation}
-s^{n}_{i,k} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{h}^{n-1}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
-\end{equation}
-with $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between the $j$th IRS in the $(n-1)$st set of IRS's and $k$th element of the $i$th IRS in the $n$th set of IRS's.
+	Additive free convolution using the R-Transform for two free variables with known R-Transforms.
+	\item
+	Multiplicitate free convolution using the S-Transform for two free variables with known S-Transforms.
 
-\item
-The vector of received signals for all elements at IRS $(n,i)$ is therefore
-\begin{equation}
-\mathbf{s}^{n}_{i} = \sum_{j = 1}^{R^{n-1}_L} (\mathbf{H}^{n-1}_{j,i})^T\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
-\end{equation}
-
-with $\mathbf{s}^{n}_{i} \in \mathbb{C}^{n_L}$
+	\end{itemize}
 
@@
 
+\chapter{Comparison to Previous Work}
+First, the two systems to which the IRS channel is compared are first given and relevant details are highlighted. 
+\section{Notes from Basic Relay Work}
+\subsection{System Model}
+
+\begin{itemize}
+\item Assuming decode and forwarding at relays and a matched filter precoding, the resulting received signal is given by 
+\begin{equation}
+\mathbf{y} = \mathbf{H}_{Total} \mathbf{s} +  \mathbf{n} = \sum_{r=1}^{R_L} \mathbf{H}_{r}\mathbf{H}_{r}^{\dagger}\mathbf{s} +  \mathbf{n}
+\end{equation}
+for which the composite channel is given by 
+\begin{equation}
+\mathbf{H}_{Total} = \mathbf{H}_{1}\mathbf{H}_{1}^{H} +\cdots +\mathbf{H}_{R_L}\mathbf{H}_{R_L}^{H}.
+\end{equation}
+Assuming the covariance matrix $\mathbf{Q}_s = \mathbf{I}_{N_R}$, the ergodic capacity is given by 
+\begin{equation}\label{ergodic_capacity}
+E\left[\text{log}_2(\mathbf{I}_{N_R} + \frac{P}{\sigma_n}\mathbf{H}_{Total}\mathbf{H}_{Total}^H)\right].
+\end{equation}
+We consider $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ as a polynomial in the self-adjoint variables $\mathbf{H}_{i}\mathbf{H}_{i}^{H}$.
+\end{itemize}
+\subsection{Free Probability Analysis}
+
+\section{Notes Pilot Decontamination Work}
+\subsection{System Model}
+Here we consider the signal 
+\begin{equation}
+\mY = \mH\mX + \mH_I \mX_I + \mW
+\end{equation}
+with covariance matrix 
+\begin{gather*}
+\mQ_{YY} = (\mH\mX + \mH_I \mX_I + \mW)(\mH\mX + \mH_I \mX_I + \mW)^H = 
+\\
+\mH\mX\mX^H\mH^H + \mH\mX\mX_I^H\mH_I^H + \mH\mX\mW^H +\\ \mH_I \mX_I \mX^H\mH^H + \mH_I \mX_I\mX_I^H\mH_I^H + \mH_I \mX_I\mW^H + \mW\mX^H\mH^H + \mW\mX_I^H\mH_I^H + \mW\mH_I^H.
+\end{gather*}
+Clearly, the individual terms of this polynomial are not self-adjoint.
+\subsection{Free Probability Analysis}
+Need to figure out how the actual steiltjes transform of this polynomial is found.
+- There is a fixed point equation for finding this transform.
+
+- Ultimately we want to have a form of the Steiltjes transform for the entire channel AED using a fixed point equation.
+\chapter{IRS Work}
+
+\section{System Model}\label{system_model}
+
+\begin{enumerate}
+\item 
+	We assume non-frequency selective channels using techniques like OFDM. (Is this always going to be reasonable?)
+
+%\item 
+%	We assume a transmitter without channel state information and thus power is distributed equally over all antennas such that
+%	\begin{equation}
+%	\mathbf{Q_{\mathbf{x}}} = \frac{P^{\text{Total}}}{N_T}
+%	\end{equation}
+%	in which $P^{\text{Total}}$ is the power available at the transmitter, and $N_T$ is the number of antennas at the transmitter. 
+\item 
+	We assume a transmitter with channel state information and beamforming capability such that the transmitted signal is given by 
+	\begin{equation}
+	\mathbf{x} = \mathbf{F}\mathbf{s}^0
+	\end{equation}
+	in  which $\mathbf{U}$ is a beamforming matrix, $\mathbf{s}^0$ is the normalized vector of transmitted symbols such that $\mathbf{Q_{\mathbf{s}^0,\mathbf{s}^0}} = \mathbf{I}_{N_R}$.
+
+\item Intelligent Reflective Surfaces (IRS) are placed between transmitter and receiver.
+The IRS is made of individual elements each of which can reflect the received signal with a unique, chosen phase shift.
+
+%\item Individual users transmit simultaneously to the first set of IRS's.
+This transmit signal hops through an arbitrary number of IRS's.
+The final set of IRS's reflect the signal to a set of independent users.
+Possible?
+(For half duplex need to make sure number of relay sets makes sense)
+
+\item 
+The signal $\mathbf{x}$ is transmitted through $N$ sets of IRS's.
+The set $n, \forall n = [1 \cdots N]$, has $L^n$ reflecting surfaces.
+Each reflecting surface in set $n$ is indexed by $R^n_{i}, \forall i = [1\cdots L^n]$ and has $K_{i}^{n}$ elements.
+\item 
+ The  signal received at element $k$ of IRS $R_i^n$ is 
+\begin{equation}
+s^{n}_{i,k} = \sum_{j = 1}^{L^{n-1}} (\mathbf{h}^{n-1}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
+\end{equation}
+with $\boldsymbol{\Theta}^{n-1}_{j}$ as the phases shifts applied by $R_j^{n-1}$ and $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between $R_j^{n-1}$ and element $k$ of $R_i^n$.
+
+\item
+The \emph{vector} of received signals for all elements at $R^n_{i}$, is given by
+\begin{equation}\label{received_vector}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
+\end{equation}
+in which $\mathbf{H}^{n-1}_{j,i} = [\mathbf{h}^{n-1}_{j,i,1} \cdots \mathbf{h}^{n-1}_{j,i,K_{j}^{n-1}}]^T$.
+
+\item For a specific receiving user with potentially multiple receiver antennas, the received signal is given in a recursive form by
+\begin{equation}\label{general_received}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
+\sum_{k = 1}^{L^{n-2}} \mathbf{H}^{n-2}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
+\end{equation}
+
+\item
+	Note that this model assumes sufficient spacing between set $n$ and set $n+2$ such that we consider the signal reflected in set $n$ only at set $n+1$ and not for any $n<k$.
+
+\item 
+	Just as in the model from Relay by Luci, we write the composite channel as
+
+\end{enumerate}
+To better analyze types of components within the polynomial characterizing the resulting, composite channel as in ..., we now consider two specific cases of this general system model.
+
+\subsection{Case in which $N=1$}
+First we consider the case in which there is only one set (one hop) of IRSs between the transmitter and receiver. 
+Equation \ref{general_received} now becomes 
+\begin{equation}\label{N=1_received}
+\mathbf{s}^{1} = \sum_{j = 1}^{L^{1}} \mathbf{H}^{1}_{j}\boldsymbol{\Theta}^{1}_{j}
+\mathbf{H}^{0}_{j} \mathbf{U}\mathbf{s}
+\end{equation}
+in which $\mathbf{s}$ is the original vector of transmitted symbols. In this case, the composite channel is given by 
+\begin{equation}
+\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{H}^{1}_{j}\boldsymbol{\Theta}^{1}_{j}
+\mathbf{H}^{0}_{j}\mathbf{U}.
+\end{equation}
+The individual terms of the polynomial in \eqref{ergodic_capacity} are not self-adjoint unless $L^{1} = 1$ and the precoding matrix is chosen such that 
+\begin{equation}
+\mathbf{U} = [\mathbf{H}^{1}_{1}\boldsymbol{\Theta}^{1}_{1}\mathbf{H}^{0}_{1}]^H.
+\end{equation}
+
+\subsection{Case in which $L_n=1$}
+We now consider the case with only a single IRS in each hop between the transmitter and receiver. 
+Equation \ref{general_received} now becomes 
+\begin{equation}\label{L_n=1_received}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
+\sum_{k = 1}^{L^{n-2}} \mathbf{H}^{n-2}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
+\end{equation}
+In this case, the composite channel is given by 
+\begin{equation}
+\mathbf{H}_{Total} = \mathbf{H}^{N}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{N-1}_{1}\boldsymbol{\Theta}^{N-1}_{1} \cdots  
+\mathbf{H}^{0}_{j}\mathbf{U}.
+\end{equation}
+Similar to the case \ref{N=1_received}, the variable of the composite channel polynomial are only self-adjoint if
+
+** But can we actually do something with a term with so many different variables?
+
+\section{System Model With Secrecy Capacity}
+The model from Section \ref{system_model} can be easily extended to the case in which an eavesdropper may be located at any point along the transmission path.
+We denote the received signal at the eavesdropper by
+
+\begin{equation}\label{general_received}
+\hat{\mathbf{s}}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \hat{\mathbf{H}}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \mathbf{s}^{n-1}_{j}.
+\end{equation}
+Note that the term $\mathbf{s}^{n-1}_{j}$ is unchanged from \ref{received_vector} because up until the final hop, the path for the eavesdropper and legitimate receiver are the same.
+
+* We should only need to known the difference between the two composite channels to find the secrecy capacity. Even if we don't know the individual ones?
+\section{Free Probability Analysis}
+
+\chapter{My TODO}
+
 \bibliography{bibliography}

### 2020-03-26 documentation/systemModel/systemModel.tex
working through specific cases of the system

@@
 \usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-
-\usepackage{graphicx, neuralnetwork,tikz}
+\usepackage{graphicx,tikz}
 \usepackage{float}
@@
 \input defs.tex
-
-\bibliographystyle{alpha}
+\bibliographystyle{ieeetr}
 \graphicspath{ {./figures/} }
-\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black]
-\tikzstyle{arrow} = [thick,->,>=stealth]
-
-\tikzstyle{state}=[shape=circle,draw=blue!30,fill=blue!10]
-\tikzstyle{observation}=[shape=rectangle,draw=orange!30,fill=orange!10]
-\tikzstyle{lightedge}=[<-, dashed]
-\tikzstyle{mainstate}=[state, thick]
-\tikzstyle{mainedge}=[<-, thick]
-\tikzstyle{block} = [draw,rectangle,thick,minimum height=2em,minimum width=2em]
-\tikzstyle{sum} = [draw,circle,inner sep=0mm,minimum size=2mm]
-\tikzstyle{connector} = [->,thick]
-\tikzstyle{line} = [thick]
-\tikzstyle{branch} = [circle,inner sep=0pt,minimum size=1mm,fill=black,draw=black]
-\tikzstyle{guide} = []
-\tikzstyle{snakeline} = [connector, decorate, decoration={pre length=0.2cm,
-                         post length=0.2cm, snake, amplitude=.4mm,
-                         segment length=2mm},thick, magenta, ->]
 
@@
 \chapter{Introduction}
-The goal of this work is to consider the IRS channel capacity and secrecy capacity using free probability theory.
+The goal of this work is to consider the intelligent reflective surface (IRS) channel capacity and secrecy capacity using free probability theory.
 Specifically, this system is compared to two similar problems which have been shown to admit analysis using free probability.
@@
 \begin{equation}\label{capacity}
-\Capacity{(H)} = \Expect\left[\underset{tr(\mathbf{Q}_{\mathbf{xx}}) < P_{Total}}{\mathrm{max}} \;\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{\mathbf{xx}}\mathbf{H}^H|\right)\right].
+\Capacity{(\mathbf{H})} = \Expect\left[\underset{tr(\mathbf{Q}_{\mathbf{xx}}) < P_{Total}}{\mathrm{max}} \;\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{\mathbf{xx}}\mathbf{H}^H|\right)\right].
 \end{equation}
 *Question: Without an analytic solution to the waterfilling problem used to choose $\mathbf{Q}_{\mathbf{xx}}$, how can we take an expectation?
--> need to do a Monte Carlo simulation (in MIMO book) so probably will not be able to include the beamformer in any free probability solution.
+Answer: need to do a Monte Carlo simulation (in MIMO book) so probably will not be able to include the beamformer in any free probability solution.
 
@@
 \begin{equation}
-\Capacity{(H)} = \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
+\Capacity{(\mathbf{H})} = \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
 \end{equation} 
@@
 \begin{equation}
-\Capacity{(H)} = N_r \Expect\left[\Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda\right)\right].
+\Capacity{(\mathbf{H})} = N_r \Expect\left[\Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda\right)\right].
 \end{equation}
@@
 \begin{gather}\label{secrecy_capacity}
-\Capacity{(H)} = \Expect\left[\underset{tr{\mathbf{Q}_{xx}} < P_{Total}}{\mathrm{max}} \;\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{xx}\mathbf{H}^H|\right) -
+\Capacity{(\mathbf{H})} = \Expect\left[\underset{tr{\mathbf{Q}_{xx}} < P_{Total}}{\mathrm{max}} \;\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{xx}\mathbf{H}^H|\right) -
 \Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\hat{\mathbf{H}}\mathbf{Q}_{xx}\hat{\mathbf{H}}^H|\right)
@@
 \begin{gather}\label{secrecy_capacity}
-\Capacity{(H)} = \Expect\left[N_r \Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda \right) -
+\Capacity{(\mathbf{H})} = \Expect\left[N_r \Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda \right) -
 N_e \Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\hat{\lambda} \right) \right]
@@
 \end{enumerate}
-\section{Free Probability/Random Matrix Theory Tools}
+\section{Random Matrix Theory and Free Probability Tools}
 
+\subsubsection{Matrix AEDs}
+For some matrices, we know the AED directly.
+For more general classes of matrices we find the AED using the known moments of the matrix to find the Stieltjes transform and then inverting the Stieltjes transform to find the true distribution.
+Generally, given the distributions of two matrices, the distribution for compositions of these matrices is difficult to find.
+
+\subsubsection{Free Probability}
+In case of asymptotically large matrices, we can consider the matrices and free variables and use the theories of free probability.
+The distribution resulting from the composition of free variables can be found using the R-transform and S-transform to linearize addition and multiplication respectively.
+
+\subsubsection{Other Tools from Free Probability}
 Here the current tools available from free probability analysis are reviewed in order to clarify the types of polynomials (composite channels) which can be analyzed.
-\begin{enumerate}
-\item 
-	Transformations of the AED
-	Detail how to recover the true pdf given the Steiltjes, R/S transforms
-	\begin{itemize}
-	\item
-		For general random matrices, we can still find the eigenvalue distributions using the problem of moments
-	\end{itemize}
 
-\item 
-	Relevant matrices with known AEDs.
-	\begin{itemize}
-	\item General zero-mean iid $n\timesn$ with variance 1/$N$ -> Full Circle
-	\item Sum of the matrix above with its adjoint. The result is self-adjoint so the eigenvalues are all real.
-	\item 
-	\end{itemize}
-\item
-	Free Families
-	\begin{itemize}
-	\item blank
-	\end{itemize}
-
-\item
-	Matrix Composition: Detail requirements on the variables
-	\begin{itemize}
-	\item
-	Additive free convolution using the R-Transform for two free variables with known R-Transforms.
-	\item
-	Multiplicitate free convolution using the S-Transform for two free variables with known S-Transforms.
-
-	\end{itemize}
-
-\end{enumerate}
+\section{Itelligent Reflective Surfaces}
+In the following work, the MIMO channel with IRSs is considered.
 
@@
 First, the two systems to which the IRS channel is compared are first given and relevant details are highlighted. 
-\section{Notes from Basic Relay Work}
+\section{Notes from Decode and Forward Relay Work in \cite{hadley2019capacity}}
 \subsection{System Model}
@@
 \begin{itemize}
-\item Assuming decode and forwarding at relays and a matched filter precoding, the resulting received signal is given by 
+\item Assuming decode and forwarding at relays and a matched filter precoding at the transmitter, the resulting received signal is given by 
 \begin{equation}
@@
 \subsection{Free Probability Analysis}
+The use of additive free convolution in
 
@@
 \begin{equation}
-s^{n}_{i,k} = \sum_{j = 1}^{L^{n-1}} (\mathbf{h}^{n-1}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
+s^{n}_{i,k} = \sum_{j = 1}^{L^{n-1}} (\mathbf{h}^{n}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
 \end{equation}
@@
 \begin{equation}\label{received_vector}
-\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n}_{j,i}\boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}.
 \end{equation}
-in which $\mathbf{H}^{n-1}_{j,i} = [\mathbf{h}^{n-1}_{j,i,1} \cdots \mathbf{h}^{n-1}_{j,i,K_{j}^{n-1}}]^T$.
+in which $\mathbf{H}^{n}_{j,i} = [\mathbf{h}^{n}_{j,i,1} \cdots \mathbf{h}^{n}_{j,i,K_{j}^{n}}]^T$.
 
-\item For a specific receiving user with potentially multiple receiver antennas, the received signal is given in a recursive form by
+\item For IRS $i$ in set $n$, $R^n_{i}$, the received signal vector is given in a recursive form as
 \begin{equation}\label{general_received}
-\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
-\sum_{k = 1}^{L^{n-2}} \mathbf{H}^{n-2}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
+\sum_{k = 1}^{L^{n-2}} \mathbf{H}^{n-1}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
 \end{equation}
-
+The received signal vector for the first set of IRSs, $R^n_{i}$ $\forall i = [1 \cdots L^{1}]$ is given by 
+\begin{equation}
+\mathbf{s}^{1}_{i} = \mathbf{H}^{1}_{i}\mathbf{F}\mathbf{s}^0
+\end{equation}
+\item 
+	Finally, using $\mathbf{G}_b$ and $\mathbf{G}_e$ as the channel matrices between the last set of IRSs and a legitimate users and an eavesdropper respectively, the received signals for the legitimate user is given by
+	\begin{equation}\label{legit_received}
+\mathbf{r}_{b} = \sum_{j = 1}^{L^{N}} \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\end{equation}
+	and the received signal for the eavesdropper is given by 
+	\begin{equation}\label{tap_received}
+\mathbf{r}_{e} = \sum_{j = 1}^{L^{N}} \mathbf{G}_{e}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\end{equation}
 \item
@@
 \item 
-	Just as in the model from Relay by Luci, we write the composite channel as
+	Note that by allowing for $L^{n} >1$ (more than one IRS in each step of the path) highlights the fact that the channels to each IRS may have a unique attenuation coefficient.
+If this were not the case, this system model could be equivalently modeled by a \emph{single} large IRS in each step of the path.
 
 \end{enumerate}
-To better analyze types of components within the polynomial characterizing the resulting, composite channel as in ..., we now consider two specific cases of this general system model.
+To better analyze types of components within the polynomial characterizing the resulting, composite channel we consider only the legitimate receiver channel for specific cases of this general system model.
+Before these examples, we first consider the optimization of the phase-shifting array at each IRS.
+
+\subsection{Optimization at IRS}
+Assuming full CSI at each IRS (detail what this means), one choice for optimization of the phase-shift array $\boldsymbol{\Theta}$ is given by
+
+    \begin{equation}
+    \begin{array}{ll}
+    \underset{\boldsymbol{\Theta}}{\text{maximize }}   & \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}\mathbf{H}_{Total}\mathbf{H}_{Total}^H|\right)\right]
+    \\
+    \mbox{subject to } & tr\left(
+    \left(\boldsymbol{\Theta}\boldsymbol{\Theta}^H - \mathbf{I}  \right)
+    \left(\boldsymbol{\Theta}\boldsymbol{\Theta}^H - \mathbf{I}  \right)^H
+    \right) = 0
+    \end{array}
+    \label{general_irs_opt}
+    \end{equation}
+
+in which $\boldsymbol{\Theta}$ represents the choice of all $\boldsymbol{\Theta}^n_i$ and $\mathbf{H}_{Total}$ is the composite channel from transmitter to receiver.
+The constraint enforces that $\boldsymbol{\Theta}^n_i$ is diagonal and that each element on the diagonal has unit modulus.
+Noting that terms in the matrix  $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ will generally have terms that are both convex and concave in the diagonal elements of $\boldsymbol{\Theta}^n_i$, this optimization problem is generally non-convex.
+A similar version of this problem is presented and analyzed in \cite{wu2019intelligent}.
+\par
+We now considering this problem in the case by case basis in the following examples.
+
+\subsection{Case in which $L_n=1$ and $N=1$}
+We now consider the case with only a single IRS in each hop between the transmitter and receiver. 
+Equation \ref{legit_received} now becomes 
+	\begin{equation}
+\mathbf{r}_{b} =  \mathbf{G}_{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}\mathbf{F}\mathbf{s}^0
+\end{equation}
+Assuming no CSI at the transmitter, the composite channel covariance matrix is given by 
+\begin{equation}
+\mathbf{Q}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{1*}\mathbf{H}^{1}]^H
+\end{equation}
+This is similar to the example given in \cite[Section 4.10]{muller2013applications}.
+As we are interested in the distribution of the non-zero eigenvalues of this matrix, we can instead consider the distribution of
+\begin{equation}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*}\mathbf{G}_{b}^H]\mathbf{H}^{1}\mathbf{H}^{1H}
+\end{equation}
+
+Noting that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes
+
+\begin{equation}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\mathbf{G}_{b}^H]\mathbf{H}^{1}\mathbf{H}^{1H}
+\end{equation}
+
+In order to use the same methods for finding the AED as \cite[Section 4.10]{muller2013applications}, it must first be shown that $\left(\left( \mathbf{G}_{b},\mathbf{G}_{b}^\dagger \right)
+,\left( \mathbf{H}^{1},\mathbf{H}^{1^\dagger} \right)
+\right)$
+form a free family. 
+\par
+Referring to Section \ref{irs_opt}, we find that given the assumed complete knowledge of the CSI at base stations,
+
+\subsection{Case in which $L_n=1$}
+We now generalize to a system with $N$ IRS hops but still a \emph{single} IRS in each hop between the transmitter and receiver. 
+Equation \ref{general_received} now becomes 
+\begin{equation}\label{}
+\mathbf{r}_{b} =  \mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{s}^{N}
+\end{equation}
+in which $\mathbf{s}^{N}$ is defined in Equation \ref{general_received}.
+
+In this case, the composite channel is given by 
+\begin{equation}
+\mathbf{H}_{Total} = \mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
+\end{equation}
+with covariance matrix 
+\begin{equation}
+\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
+\end{equation}
+Using the rotational property of the trace as in the previous case
+We can equivalently look for the AED of the matrix 
+\begin{equation}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
+\end{equation}
+Iterating this procedure and noting that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, this is exactly the case from \cite[Section 4.10]{muller2013applications}.
 
 \subsection{Case in which $N=1$}
-First we consider the case in which there is only one set (one hop) of IRSs between the transmitter and receiver. 
-Equation \ref{general_received} now becomes 
-\begin{equation}\label{N=1_received}
-\mathbf{s}^{1} = \sum_{j = 1}^{L^{1}} \mathbf{H}^{1}_{j}\boldsymbol{\Theta}^{1}_{j}
-\mathbf{H}^{0}_{j} \mathbf{U}\mathbf{s}
+We now consider the case in which there is one IRS hop between the transmitter and receiver with arbitrarily many IRSs. 
+Equation \ref{legit_received} now becomes 
+\begin{equation}
+\mathbf{r}_{b} = \sum_{j = 1}^{L^{1}} \mathbf{G}_{b}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}\mathbf{s}^0
 \end{equation}
-in which $\mathbf{s}$ is the original vector of transmitted symbols. In this case, the composite channel is given by 
+with composite channel
 \begin{equation}
-\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{H}^{1}_{j}\boldsymbol{\Theta}^{1}_{j}
-\mathbf{H}^{0}_{j}\mathbf{U}.
+\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{H}^{1}_{j}
 \end{equation}
-The individual terms of the polynomial in \eqref{ergodic_capacity} are not self-adjoint unless $L^{1} = 1$ and the precoding matrix is chosen such that 
+and covariance matrix
+
 \begin{equation}
-\mathbf{U} = [\mathbf{H}^{1}_{1}\boldsymbol{\Theta}^{1}_{1}\mathbf{H}^{0}_{1}]^H.
+\mathbf{Q}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1} + \cdots + \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}]
+[\mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1} + \cdots + \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}]^H
 \end{equation}
-
-\subsection{Case in which $L_n=1$}
-We now consider the case with only a single IRS in each hop between the transmitter and receiver. 
-Equation \ref{general_received} now becomes 
-\begin{equation}\label{L_n=1_received}
-\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \mathbf{H}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
-\sum_{k = 1}^{L^{n-2}} \mathbf{H}^{n-2}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
-\end{equation}
-In this case, the composite channel is given by 
-\begin{equation}
-\mathbf{H}_{Total} = \mathbf{H}^{N}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{N-1}_{1}\boldsymbol{\Theta}^{N-1}_{1} \cdots  
-\mathbf{H}^{0}_{j}\mathbf{U}.
-\end{equation}
-Similar to the case \ref{N=1_received}, the variable of the composite channel polynomial are only self-adjoint if
-
-** But can we actually do something with a term with so many different variables?
-
 \section{System Model With Secrecy Capacity}
@@
 
-\begin{equation}\label{general_received}
+\begin{equation}
 \hat{\mathbf{s}}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \hat{\mathbf{H}}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \mathbf{s}^{n-1}_{j}.
@@
 
-\chapter{My TODO}
+\chapter{TODO}
+\section{My TODO}
+\begin{itemize}
+\item 
+	Show why the channels are free. 
+\item
+	Use method from mueller paper to solve further.
+\end{itemize}
+\section{Questions}
+\begin{itemize}
+\item 
+	Just having the IRS present should give us a boost, do we need to actually consider selecting the phases?
+
+\item 
+	Confirm that the one example on how to pull out the hermetian matrices is a result of the trace property.
+
+\end{itemize}
 

### 2020-03-29 documentation/systemModel/systemModel.tex
a few improvements and further work on FPT analysis

@@
 \chapter{Introduction}
-The goal of this work is to consider the intelligent reflective surface (IRS) channel capacity and secrecy capacity using free probability theory.
-Specifically, this system is compared to two similar problems which have been shown to admit analysis using free probability.
+The goal of this work is to consider the intelligent reflective surface (IRS) channel capacity and secrecy capacity using the tools of free probability.
+First, two similar problems which have been shown to admit analysis using free probability are considered.
+These examples are then compared to the IRS channel.
 \par
@@
 \begin{enumerate}
-\item For a random MIMO communication channel and CSI known at the transmitter and Gaussian distributed transmit signal $\mathbf{x}$, the capacity ($\mathcal{C}$) of a point-to-point channel $\mathbf{H}$ is given by
+\item For a random MIMO communication channel with Gaussian distributed transmit signal $\mathbf{x}$, and CSI known at the transmitter,  the capacity ($\mathcal{C}$) of a point-to-point channel characterized by the flat-fading channel matrix $\mathbf{H}$ is given by
 \begin{equation}\label{capacity}
@@
 \end{equation}
-*Question: Without an analytic solution to the waterfilling problem used to choose $\mathbf{Q}_{\mathbf{xx}}$, how can we take an expectation?
-Answer: need to do a Monte Carlo simulation (in MIMO book) so probably will not be able to include the beamformer in any free probability solution.
+Solving for the optimial $\mathbf{Q}_{\mathbf{xx}}$ generally lacks an analytic solution and requires solving the  "waterfilling" problem to choose $\mathbf{Q}_{\mathbf{xx}}$.
+As a result, taking an expectation of this general expressiong requires Monte Carlo simulation (as seen in MIMO course book).
 
 \item
-In the case in which no CSI is known at the transmitter and the power at the transmitter, $P_{\text{Total}}$, is distributed equally across transmit antennas Equation \ref{capacity} becomes
-\begin{equation}
+In the case in which CSI is \textbf{not} known at the transmitter, the power at the transmitter, $P_{\text{Total}}$, is distributed equally across transmit antennas and Equation \ref{capacity} becomes
+\begin{equation}\label{capacity_no_csi}
 \Capacity{(\mathbf{H})} = \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}\mathbf{H}\mathbf{H}^H|\right)\right].
 \end{equation} 
-Assuming asymptotically large antenna arrays, and given the distribution for the eigenvalues of $\mathbf{H}\mathbf{H}^H$ which we denote by $p_{\lambda}(\lambda)$, Equation \ref{capacity} can be rewritten as
-\begin{equation}
+Assuming asymptotically large antenna arrays, and given the distribution for the eigenvalues of $\mathbf{H}\mathbf{H}^H$ which we denote by $p_{\lambda}(\lambda)$, Equation \ref{capacity_no_csi} can be rewritten as
+\begin{equation}\label{capacity_AED}
 \Capacity{(\mathbf{H})} = N_r \Expect\left[\Log\left(1+\frac{P^{\text{Total}}}{N_T \sigma_n}\lambda\right)\right].
 \end{equation}
- in which the expectation is taken over $p_{\lambda}(\lambda)$.
+ in which the expectation is taken over $p_{\lambda}(\lambda)$ known as the asymptotic eigenvalue distribution (AED).
 
-\item The secrecy capacity of the general wire-tapped communication channel with perfect knowledge of the channels to both the legitimate receiver, $\mathbf{H} \in \mathbb{C}_{N_r \times N_t}$, and eavesdropper, $\hat{\mathbf{H}}\in \mathbb{C}_{N_e \times N_t}$, is given by
+\item The secrecy capacity of the general wire-tapped communication channel with perfect knowledge of the channels to both the legitimate receiver, $\mathbf{H} \in \mathbb{C}_{N_r \times N_t}$, and an eavesdropper, $\hat{\mathbf{H}}\in \mathbb{C}_{N_e \times N_t}$, is given by
 
 \begin{gather}\label{secrecy_capacity}
-\Capacity{(\mathbf{H})} = \Expect\left[\underset{tr{\mathbf{Q}_{xx}} < P_{Total}}{\mathrm{max}} \;\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{xx}\mathbf{H}^H|\right) -
+\Capacity{(\mathbf{H})} = \Expect\left[\underset{tr({\mathbf{Q}_{xx}}) < P_{Total}}{\mathrm{max}} \;\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\mathbf{H}\mathbf{Q}_{xx}\mathbf{H}^H|\right) -
 \Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}\hat{\mathbf{H}}\mathbf{Q}_{xx}\hat{\mathbf{H}}^H|\right)
@@
 in which $\lambda$ and $\hat{\lambda}$ denote the eigenvalues of $\mathbf{H}\mathbf{H}^H$ and $\hat{\mathbf{H}}\hat{\mathbf{H}}^H$ respectively.
-Generally, finding capacity will require knowledge of the joint asymptotic eigenvalue distribution $p_{\lambda}(\lambda,\hat{\lambda})$.
+Generally, finding capacity will require knowledge of the \emph{joint} AED $p_{\lambda}(\lambda,\hat{\lambda})$.
 
@@
 \subsubsection{Free Probability}
-In case of asymptotically large matrices, we can consider the matrices and free variables and use the theories of free probability.
-The distribution resulting from the composition of free variables can be found using the R-transform and S-transform to linearize addition and multiplication respectively.
+By considering asymptotically large matrices as free random variables, the distribution resulting from the composition of free random variables can be found using the R-transform and S-transform to linearize addition and multiplication respectively.
 
-\subsubsection{Other Tools from Free Probability}
-Here the current tools available from free probability analysis are reviewed in order to clarify the types of polynomials (composite channels) which can be analyzed.
-
-\section{Itelligent Reflective Surfaces}
-In the following work, the MIMO channel with IRSs is considered.
-
-\chapter{Comparison to Previous Work}
-First, the two systems to which the IRS channel is compared are first given and relevant details are highlighted. 
-\section{Notes from Decode and Forward Relay Work in \cite{hadley2019capacity}}
+\chapter{Previous Work}
+First, two systems related to the IRS channel capacity problem are described and relevant details are highlighted. 
+\section{Notes from Decode and Forward Relay Work in \cite{hadley2019capacity}}\label{relay}
 \subsection{System Model}
@@
 \begin{itemize}
-\item Assuming decode and forwarding at relays and a matched filter precoding at the transmitter, the resulting received signal is given by 
+\item A decode and forward protocol is used at $R_L$ different relays between a transmitter and receiver.
+A matched filter precoding is applied at each transmitting relay such that the signal at the receiver is given by
 \begin{equation}
@@
 \subsection{Free Probability Analysis}
-The use of additive free convolution in
+It is claimed in the final paragraph of the "theory" section of \cite{hadley2019capacity} that the R-transform and S-transform cannot be used to analyze more complex polynomials.
+This seems to contradict the methods used in the decontamination paper (TO DISCUSS).
 
@@
 \subsection{System Model}
-Here we consider the signal 
+Disregarding the motivation and context of this problem, we want to find the AED of the signal 
 \begin{equation}
@@
 Clearly, the individual terms of this polynomial are not self-adjoint.
-\subsection{Free Probability Analysis}
-Need to figure out how the actual steiltjes transform of this polynomial is found.
-- There is a fixed point equation for finding this transform.
+\subsection{Free Probability Analysis}\label{decont_fpt}
+In order to analyzed the matrix polynomial above, the Stieltjes of the matrix
 
-- Ultimately we want to have a form of the Steiltjes transform for the entire channel AED using a fixed point equation.
+\begin{equation}
+\mathbf{D} = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
+\end{equation}
+with $\alpha_k \in \mathbb{R}$ is considered.
+Given the S-transform of $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$.
+Given the Stieltjes transform of each term, we then find the R-transform of each term $\mathbf{B}_k \mathbf{C}_k$ and exploiting the additive property of the R-transform, we find the R-transform of $\mathbf{D}$.
+Inverting the R-transform to get the Stieltjes transform of $\mathbf{D}$ which we can evaluate to find the AED.
+\par
+Note that the resulting Stieltjes transform is given as a fixed point equation which is then evaluated for each value of of interest to find the corresponding value of the AED.
+
+\begin{itemize}
+\item The IRS model needs to be checked for compatibility with \cite{nica2006lectures}.
+\item Need to solve for a similar fixed point equation.
+\item Evaluate the capacity by using the above distribution to perform an integration.
+\end{itemize} 
 \chapter{IRS Work}
@@
 \item 
-	We assume non-frequency selective channels using techniques like OFDM. (Is this always going to be reasonable?)
+	We assume non-frequency selective channels using techniques like OFDM. (Is this always going to be reasonable? TO DISCUSS)
 
@@
 	\end{equation}
-	in  which $\mathbf{U}$ is a beamforming matrix, $\mathbf{s}^0$ is the normalized vector of transmitted symbols such that $\mathbf{Q_{\mathbf{s}^0,\mathbf{s}^0}} = \mathbf{I}_{N_R}$.
+	in  which $\mathbf{F}$ is a beamforming matrix and $\mathbf{s}^0$ is the normalized vector of transmitted symbols such that $\mathbf{Q_{\mathbf{s}^0,\mathbf{s}^0}} = \mathbf{I}_{N_t}$.
 
@@
 The set $n, \forall n = [1 \cdots N]$, has $L^n$ reflecting surfaces.
-Each reflecting surface in set $n$ is indexed by $R^n_{i}, \forall i = [1\cdots L^n]$ and has $K_{i}^{n}$ elements.
+Each reflecting surface in set $n$ is indexed by $R^n_{i}, \forall i = [1\cdots L^n]$ and has $K_{i}^{n}$ reflecting elements.
+\item
+The set of $L_n$ individual IRS in set $n$ may be spaced differently and thus the attenuation coefficient for each IRS is given by $\alpha_i^n$.
 \item 
@@
 \begin{equation}
-s^{n}_{i,k} = \sum_{j = 1}^{L^{n-1}} (\mathbf{h}^{n}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
+s^{n}_{i,k} = \sum_{j = 1}^{L^{n-1}} \alpha_i^n (\mathbf{h}^{n}_{j,i,k})^T \boldsymbol{\Theta}^{n-1}_{j}\mathbf{s}^{n-1}_{j}
 \end{equation}
-with $\boldsymbol{\Theta}^{n-1}_{j}$ as the phases shifts applied by $R_j^{n-1}$ and $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between $R_j^{n-1}$ and element $k$ of $R_i^n$.
+with $\boldsymbol{\Theta}^{n-1}_{j}$ as the phase shifts applied by IRS $R_j^{n-1}$ and $\mathbf{h}^{n-1}_{j,i,k}$ as the vector of channel coefficients between $R_j^{n-1}$ and element $k$ of $R_i^n$.
 
+\item
+	To reduce notation in the following, the attenuation component,$\alpha_i^n$ is included directly in the channel such that $ \mathbf{h}^{n}_{j,i,k} = \alpha_i^n \mathbf{h}^{n}_{j,i,k}$
 \item
@@
 
-\item For IRS $i$ in set $n$, $R^n_{i}$, the received signal vector is given in a recursive form as
+\item For IRS $R^n_{i}$ the received signal vector is given in a recursive form as
 \begin{equation}\label{general_received}
@@
 \end{equation}
+
+\item 
 The received signal vector for the first set of IRSs, $R^n_{i}$ $\forall i = [1 \cdots L^{1}]$ is given by 
@@
 \item 
-	Finally, using $\mathbf{G}_b$ and $\mathbf{G}_e$ as the channel matrices between the last set of IRSs and a legitimate users and an eavesdropper respectively, the received signals for the legitimate user is given by
+	Finally, using $\mathbf{G}^b$ and $\mathbf{G}^e$ as the channel matrices between the last set of IRSs and a legitimate users and an eavesdropper respectively, the received signals for the legitimate user is given by
 	\begin{equation}\label{legit_received}
-\mathbf{r}_{b} = \sum_{j = 1}^{L^{N}} \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\mathbf{r}_{b} = \sum_{j = 1}^{L^{N}}\mathbf{G}^{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
 \end{equation}
@@
 	\begin{equation}\label{tap_received}
-\mathbf{r}_{e} = \sum_{j = 1}^{L^{N}} \mathbf{G}_{e}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\mathbf{r}_{e} = \sum_{j = 1}^{L^{N}} \mathbf{G}^{e}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
 \end{equation}
@@
 
-\item 
-	Note that by allowing for $L^{n} >1$ (more than one IRS in each step of the path) highlights the fact that the channels to each IRS may have a unique attenuation coefficient.
-If this were not the case, this system model could be equivalently modeled by a \emph{single} large IRS in each step of the path.
-
 \end{enumerate}
-To better analyze types of components within the polynomial characterizing the resulting, composite channel we consider only the legitimate receiver channel for specific cases of this general system model.
+To better analyze types of components within the polynomial characterizing the resulting composite channel, we consider only the legitimate receiver channel for specific cases of this general system model.
 Before these examples, we first consider the optimization of the phase-shifting array at each IRS.
@@
 \subsection{Optimization at IRS}
-Assuming full CSI at each IRS (detail what this means), one choice for optimization of the phase-shift array $\boldsymbol{\Theta}$ is given by
+We assume the transmitter uses equal power distribution for the reasons shown in \ref{capacity}.
+With full CSI at each IRS (detail what this means), one choice for optimization of the phase-shift array $\boldsymbol{\Theta}$ is given by
 
@@
 The constraint enforces that $\boldsymbol{\Theta}^n_i$ is diagonal and that each element on the diagonal has unit modulus.
-Noting that terms in the matrix  $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ will generally have terms that are both convex and concave in the diagonal elements of $\boldsymbol{\Theta}^n_i$, this optimization problem is generally non-convex.
+Noting that elements the matrix  $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ will generally have terms that are both convex and concave w.r.t.
+the diagonal elements of $\boldsymbol{\Theta}^n_i$, this optimization problem is generally non-convex.
 A similar version of this problem is presented and analyzed in \cite{wu2019intelligent}.
@@
 	\begin{equation}
-\mathbf{r}_{b} =  \mathbf{G}_{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}\mathbf{F}\mathbf{s}^0
+\mathbf{r}_{b} =  \mathbf{G}^{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}\mathbf{F}\mathbf{s}^0
 \end{equation}
@@
 \begin{equation}
-\mathbf{Q}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{1*}\mathbf{H}^{1}]^H
+\mathbf{Q}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{1*}\mathbf{H}^{1}]^H
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*}\mathbf{G}_{b}^H]\mathbf{H}^{1}\mathbf{H}^{1H}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{1} \boldsymbol{\Theta}^{1*}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\mathbf{G}_{b}^H]\mathbf{H}^{1}\mathbf{H}^{1H}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
 \end{equation}
 
-In order to use the same methods for finding the AED as \cite[Section 4.10]{muller2013applications}, it must first be shown that $\left(\left( \mathbf{G}_{b},\mathbf{G}_{b}^\dagger \right)
-,\left( \mathbf{H}^{1},\mathbf{H}^{1^\dagger} \right)
+In order to use the same methods for finding the AED as \cite[Section 4.10]{muller2013applications}, it must first be shown that $\left(\left( \mathbf{G}^{b}\mathbf{G}^{b^\dagger }\right)
+,\left( \mathbf{H}^{1}\mathbf{H}^{1^\dagger} \right)
 \right)$
@@
 \par
-Referring to Section \ref{irs_opt}, we find that given the assumed complete knowledge of the CSI at base stations,
+TODO need to investigate how the final AED is found in this work.
 
 \subsection{Case in which $L_n=1$}
-We now generalize to a system with $N$ IRS hops but still a \emph{single} IRS in each hop between the transmitter and receiver. 
+We now increase the generalization to a system with $N$ IRS hops but still a \emph{single} IRS in each hop between the transmitter and receiver. 
 Equation \ref{general_received} now becomes 
 \begin{equation}\label{}
-\mathbf{r}_{b} =  \mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{s}^{N}
+\mathbf{r}_{b} =  \mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{s}^{N}
 \end{equation}
@@
 \begin{equation}
-\mathbf{H}_{Total} = \mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
+\mathbf{H}_{Total} = \mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
 \end{equation}
@@
 \begin{equation}
-\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
+\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}_{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
 \end{equation}
@@
 \subsection{Case in which $N=1$}
-We now consider the case in which there is one IRS hop between the transmitter and receiver with arbitrarily many IRSs. 
+We now consider the case in which there is one IRS hop between the transmitter and receiver (i.e $N=1$) but with an  arbitrarily number of surfaces ($L_n$). 
 Equation \ref{legit_received} now becomes 
 \begin{equation}
-\mathbf{r}_{b} = \sum_{j = 1}^{L^{1}} \mathbf{G}_{b}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}\mathbf{s}^0
+\mathbf{r}_{b} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{b}_{j}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}\mathbf{s}^0
 \end{equation}
@@
 \begin{equation}
-\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{H}^{1}_{j}
+\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{b}_{j}\boldsymbol{\Theta}^{N}_{j}\mathbf{H}^{1}_{j}\mathbf{F}
 \end{equation}
@@
 \begin{equation}
-\mathbf{Q}_{bb} = [\mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1} + \cdots + \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}]
-[\mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1} + \cdots + \mathbf{G}_{b}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}]^H
+\mathbf{Q}_{bb} = [\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
+\cdots + \mathbf{G}^{b}_{N}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]
+[\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
+\cdots + \mathbf{G}^{b}_{N}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]^H
 \end{equation}
+
+This sum looks similar to the polynomial considered in Section \ref{decont_fpt}, however each term of the polynomial contains more than two variables.
+If we can iteratively apply the S-transform rule then perhaps we can deal with this in the same way (TO DISCUSS).
+\par
+Unlike the model with amplify and forward relays discussed in section \ref{relay} and \cite{hadley2019capacity}, the individual variables in each term of the 
+polynomial, $\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F}$ cannot be self-adjoint
+ assuming the channel matrices $\mathbf{H}^{1}_{j} \neq \mathbf{H}^{1}_{i} $ $\forall j \neq i$. 
+ This prevents using the methods proposed in \cite{belinschi2017analytic} for finding the AED of this polynomial.
+
 \section{System Model With Secrecy Capacity}
@@
 
-\chapter{TODO}
-\section{My TODO}
+\chapter{Next Steps}
+\section{Independent TODO items}
 \begin{itemize}
 \item 
-	Show why the channels are free. 
-\item
-	Use method from mueller paper to solve further.
+	Show why the channels are free as assumed in decontamination work.
 \end{itemize}
@@
 
+\item 
+	Examples of secrecy codes apart from the one presented in the original wire-tap paper
+
+\item 
+	How to actually evaluate a pdf given a fixed point equation for the stieltjes transform and an expression for the
+	AED in terms of the Steitjes transform.
+
+\item 
+	Can we iterate the process used in the decontamination work (i.e) use the S-Transform on terms with $>2$
+	 matrices (free variables).
 \end{itemize}

### 2020-03-31 documentation/systemModel/systemModel.tex
Corrections based on cancelling non-self-adjoint terms

@@
 in which $\lambda$ and $\hat{\lambda}$ denote the eigenvalues of $\mathbf{H}\mathbf{H}^H$ and $\hat{\mathbf{H}}\hat{\mathbf{H}}^H$ respectively.
-Generally, finding capacity will require knowledge of the \emph{joint} AED $p_{\lambda}(\lambda,\hat{\lambda})$.
+Generally, finding the capacity will require knowledge of the \emph{joint} AED $p_{\lambda}(\lambda,\hat{\lambda})$.
 
@@
 \end{gather*}
-Clearly, the individual terms of this polynomial are not self-adjoint.
+Some any of individual terms of this polynomial are not self-adjoint.
+(TO DISCUSS) If, however, we are under the assumption that the channel matrices are i.i.d and zero-mean then these mixed terms should cancel and we should be left with only the self-adjoint components such that
+
+\begin{gather*}
+\mQ_{YY} =
+\mH\mX\mX^H\mH^H + \mH_I \mX_I\mX_I^H\mH_I^H + \mW\mW^H.
+\end{gather*}
 \subsection{Free Probability Analysis}\label{decont_fpt}
-In order to analyzed the matrix polynomial above, the Stieltjes of the matrix
 
@@
 with $\alpha_k \in \mathbb{R}$ is considered.
-Given the S-transform of $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$.
+Given the S-transform of the *self-adjoint matrices $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$.
 Given the Stieltjes transform of each term, we then find the R-transform of each term $\mathbf{B}_k \mathbf{C}_k$ and exploiting the additive property of the R-transform, we find the R-transform of $\mathbf{D}$.
@@
 
-\begin{itemize}
-\item The IRS model needs to be checked for compatibility with \cite{nica2006lectures}.
-\item Need to solve for a similar fixed point equation.
-\item Evaluate the capacity by using the above distribution to perform an integration.
-\end{itemize} 
 \chapter{IRS Work}
@@
 	\begin{equation}\label{legit_received}
-\mathbf{r}_{b} = \sum_{j = 1}^{L^{N}}\mathbf{G}^{b}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\mathbf{r}_{B} = \sum_{j = 1}^{L^{N}}\mathbf{G}^{B}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
 \end{equation}
@@
 	\begin{equation}\label{tap_received}
-\mathbf{r}_{e} = \sum_{j = 1}^{L^{N}} \mathbf{G}^{e}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\mathbf{r}_{E} = \sum_{j = 1}^{L^{N}} \mathbf{G}^{E}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
 \end{equation}
@@
 
-\subsection{Case in which $L_n=1$ and $N=1$}
+\subsection{Case: $L_n=1$ and $N=1$}
 We now consider the case with only a single IRS in each hop between the transmitter and receiver. 
@@
 	\begin{equation}
-\mathbf{r}_{b} =  \mathbf{G}^{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}\mathbf{F}\mathbf{s}^0
+\mathbf{r}_{B} =  \mathbf{G}^{B}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}\mathbf{F}\mathbf{s}^0
 \end{equation}
@@
 \begin{equation}
-\mathbf{Q}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{1*}\mathbf{H}^{1}]^H
+\mathbf{Q}_{bb} = [\mathbf{G}^{B}\boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{1*}\mathbf{H}^{1}]^H
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{1} \boldsymbol{\Theta}^{1*}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{B}\boldsymbol{\Theta}^{1} \boldsymbol{\Theta}^{1*}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{B}\mathbf{G}^{bH}]\mathbf{H}^{1}\mathbf{H}^{1H}
 \end{equation}
 
-In order to use the same methods for finding the AED as \cite[Section 4.10]{muller2013applications}, it must first be shown that $\left(\left( \mathbf{G}^{b}\mathbf{G}^{b^\dagger }\right)
+In order to use the same methods for finding the AED as \cite[Section 4.10]{muller2013applications}, it must first be shown that $\left(\left( \mathbf{G}^{B}\mathbf{G}^{b^\dagger }\right)
 ,\left( \mathbf{H}^{1}\mathbf{H}^{1^\dagger} \right)
@@
 
-\subsection{Case in which $L_n=1$}
+\subsection{Case:$L_n=1$}
 We now increase the generalization to a system with $N$ IRS hops but still a \emph{single} IRS in each hop between the transmitter and receiver. 
@@
 \begin{equation}\label{}
-\mathbf{r}_{b} =  \mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{s}^{N}
+\mathbf{r}_{B} =  \mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{s}^{N}
 \end{equation}
@@
 \begin{equation}
-\mathbf{H}_{Total} = \mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
+\mathbf{H}_{Total} = \mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
 \end{equation}
 with covariance matrix 
-\begin{equation}
-\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
+\begin{equation}\label{polynomial_terms}
+\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
 \end{equation}
@@
 \begin{equation}
-\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}^{b}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
+\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
 \end{equation}
@@
 
-\subsection{Case in which $N=1$}
+\subsection{Case: $N=1$}
 We now consider the case in which there is one IRS hop between the transmitter and receiver (i.e $N=1$) but with an  arbitrarily number of surfaces ($L_n$). 
@@
 \begin{equation}
-\mathbf{r}_{b} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{b}_{j}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}\mathbf{s}^0
+\mathbf{r}_{B} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{B}_{j}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}\mathbf{s}^0
 \end{equation}
@@
 \begin{equation}
-\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{b}_{j}\boldsymbol{\Theta}^{N}_{j}\mathbf{H}^{1}_{j}\mathbf{F}
+\mathbf{H}_{Total} = \sum_{j = 1}^{L^{1}} \mathbf{G}^{B}_{j}\boldsymbol{\Theta}^{1}_{j}\mathbf{H}^{1}_{j}\mathbf{F}
 \end{equation}
@@
 \begin{equation}
-\mathbf{Q}_{bb} = [\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
-\cdots + \mathbf{G}^{b}_{N}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]
-[\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
-\cdots + \mathbf{G}^{b}_{N}\boldsymbol{\Theta}^{N}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]^H
+\mathbf{Q}_{bb} = [\mathbf{G}^{B}_{1}\boldsymbol{\Theta}^{1}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
+\cdots + \mathbf{G}^{B}_{L^{1}}\boldsymbol{\Theta}^{1}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]
+[\mathbf{G}^{B}_{1}\boldsymbol{\Theta}^{1}_{1}\mathbf{H}^{1}_{1}\mathbf{F} + 
+\cdots + \mathbf{G}^{B}_{L^{1}}\boldsymbol{\Theta}^{1}_{L^{1}}\mathbf{H}^{1}_{L^{1}}\mathbf{F}]^H
 \end{equation}
 
-This sum looks similar to the polynomial considered in Section \ref{decont_fpt}, however each term of the polynomial contains more than two variables.
-If we can iteratively apply the S-transform rule then perhaps we can deal with this in the same way (TO DISCUSS).
+This sum looks similar to the polynomial considered in Section \ref{decont_fpt}.
+Similar to \ref{decont_fpt}, if we assume the elements of the channel matrices are zero-mean such that the non-self-adjoint terms go to zero we get the polynomial   (TO DISCUSS).
+
+\begin{equation}
+\mathbf{Q}_{bb} = \sum_{i=1}^{L^{1}}[\mathbf{G}^{B}_{i}\boldsymbol{\Theta}^{1}_{i}\mathbf{H}^{1}_{i}\mathbf{F}]
+[\mathbf{G}^{B}_{i}\boldsymbol{\Theta}^{1}_{i}\mathbf{H}^{1}_{i}\mathbf{F}]^H
+\end{equation}
+
+If this is indeed the case, we can simply use the method used in the previous examples to find the Stieltjes transform of the individual components and then use the R-transform to find the R-transform and eventually the Stieltjes and AED of the sum of components.
 \par
-Unlike the model with amplify and forward relays discussed in section \ref{relay} and \cite{hadley2019capacity}, the individual variables in each term of the 
-polynomial, $\mathbf{G}^{b}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F}$ cannot be self-adjoint
- assuming the channel matrices $\mathbf{H}^{1}_{j} \neq \mathbf{H}^{1}_{i} $ $\forall j \neq i$. 
- This prevents using the methods proposed in \cite{belinschi2017analytic} for finding the AED of this polynomial.
+Because individual variables in each term of the 
+polynomial, $\mathbf{G}^{B}_{1}\boldsymbol{\Theta}^{N}_{1}\mathbf{H}^{1}_{1}\mathbf{F}$ are self-adjoint, this may also admit the analysis methods used in the model with amplify and forward relays discussed in section \ref{relay} and \cite{hadley2019capacity} and based on the results from \cite{belinschi2017analytic}.
+
+\section{Case: General} 
+Recalling the equation for the received signal
+	\begin{equation*}
+\mathbf{r}_{B} = \sum_{j = 1}^{L^{N}}\mathbf{G}^{B}\boldsymbol{\Theta}^{N}_{j}\mathbf{s}^{N}_{j}
+\end{equation*}
+and the recursive expression 
+\begin{equation*}
+\mathbf{s}^{n}_{i} = \sum_{j = 1}^{L^{n-1}}  \mathbf{H}^{n}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \left[
+\sum_{k = 1}^{L^{n-2}}  \mathbf{H}^{n-1}_{k,i}\boldsymbol{\Theta}^{n-2}_{k}\mathbf{s}^{n-2}_{k} \right].
+\end{equation*}
+If we assume the individual channel matrices are made of i.i.d zero-mean elements, any non-self-adjoint term in the polynomial will cancel.
+Thus we can consider concatenated channel to be the sum of each possible path through the network of IRS possible.
+Thus for the covariance matrix of the system is a polynomial with a number of terms corresponding to the number of paths from the transmitter to receiver
+$\prod_{n=1}^{N} L_n$.
+Each term of this polynomial will be equivalent to the covariance matrix in Equation \eqref{polynomial_terms}.
+As we already known how to find the Stieltjes transform of each term, we can then find the sum of these terms using the R-transform.
+\par
+One example is shown in figure \ref{irs_figure}
+
+\begin{figure}
+
+\caption{IRS Network with ...}
+\label{irs_figure}
+\end{figure}
 
 \section{System Model With Secrecy Capacity}
-The model from Section \ref{system_model} can be easily extended to the case in which an eavesdropper may be located at any point along the transmission path.
-We denote the received signal at the eavesdropper by
+Recalling Equation \eqref{tap_received} for the received signal at the eavesdropper,  and Equation \eqref{secrecy_capacity} for the secrecy capacity of the resulting system, we see that in order to find the secrecy capacity, we need the joint AED of the channel covariance matrices of the intended receiver and eavesdropper, $p_{\lambda}(\lambda,\hat{\lambda})$.
+Clearly, given the channels common to the paths of both the legitimate and eavesdropping receivers will ensure that $p_{\lambda}(\lambda,\hat{\lambda}) \neq p_{\lambda}(\lambda)p_{\lambda}(\hat{\lambda})$
+\par
+Instead we consider the decomposition 
+\begin{equation}
+p_{\lambda}(\lambda,\hat{\lambda}) = p_{\lambda}(\lambda)p_{\lambda}(\hat{\lambda}|\lambda).
+\end{equation}
+Assuming that we choose a system such that we can find $ p_{\lambda}(\lambda)$ or  $p_{\lambda}(\hat{\lambda})$ we must consider how to find the conditional probability of an AED. (TO DISCUSS)
 
-\begin{equation}
-\hat{\mathbf{s}}^{n}_{i} = \sum_{j = 1}^{L^{n-1}} \hat{\mathbf{H}}^{n-1}_{j,i}\boldsymbol{\Theta}^{n-1}_{j} \mathbf{s}^{n-1}_{j}.
-\end{equation}
-Note that the term $\mathbf{s}^{n-1}_{j}$ is unchanged from \ref{received_vector} because up until the final hop, the path for the eavesdropper and legitimate receiver are the same.
-
-* We should only need to known the difference between the two composite channels to find the secrecy capacity. Even if we don't know the individual ones?
-\section{Free Probability Analysis}
+%\section{Free Probability Analysis}
 
@@
 \item 
-	Show why the channels are free as assumed in decontamination work.
+	Show why the individual channel matrices are free as assumed in decontamination work.
+\item
+	Work on example of how to find AED given a fixed point equation for the Stieltjes transform.
 \end{itemize}
@@
 \begin{itemize}
+\item
+	In Pilot Decontamination paper, the final reference refers to the case in which each variable is self-adjoint.
+	Are the non-similar terms canceling out?
 \item 

### 2020-03-31 documentation/systemModel/systemModel.tex
eod

@@
 \end{gather}
+or equivalently 
+\begin{gather}\label{secrecy_capacity}
+\Capacity{(\mathbf{H})} = \Expect\left[\frac{N_r}{N_T }\Log\left(\frac{1+\frac{P^{\text{Total}}}{\sigma_n}\lambda }{1+\frac{P^{\text{Total}}}{\sigma_n}\hat{\lambda} }\right) \right]
+\end{gather}
+
 in which $\lambda$ and $\hat{\lambda}$ denote the eigenvalues of $\mathbf{H}\mathbf{H}^H$ and $\hat{\mathbf{H}}\hat{\mathbf{H}}^H$ respectively. 

### 2020-04-06 documentation/systemModel/systemModel.tex
adding larger polynomial

@@
 
-Noting that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes
+Noting that for diagonal matrices, matrix multiplication is commutative and that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes
 
@@
 \end{equation}
-Iterating this procedure and noting that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, this is exactly the case from \cite[Section 4.10]{muller2013applications}.
+Noting again that matrix multiplication with a diagonal matrix is commutative and that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, we iterate the above procedure just as in \cite[Section 4.10]{muller2013applications}.
 

### 2020-04-06 simulation/src/network_simulation_tools.py

@@ -1,0 +1,47 @@
+import numpy as np
+import copy
+from testing.circle_laws import *
+class Network:
+    def __init__(self, num_elements, network_size):
+        """
+        Assume that the final size is always 1
+        :param num_elements:
+        :param network_size:
+        """
+        if network_size[-1] != 1:
+            network_size.apend([1])
+        previous_size = 1
+        self.network_size = network_size
+        self.network_channels = []
+        for elements in network_size:
+            step_surfaces = []
+            for previous_surface in range(previous_size):
+                step_surfaces.append(IRS(previous_surface,  num_elements))
+            self.network_channels.append(step_surfaces)
+            previous_size = elements
+        self.covariance = np.zeros((num_elements, num_elements))
+
+    def get_covariance(self, path=[0]):
+        for ind, matrix in enumerate(self.network_channels):
+            if len(path)  == len(self.network_channels):
+                self.channel_from_path(path)
+            if len(path) < len(self.network_channels):
+                path.append(ind)
+                self.get_covariance(path=copy.deepcopy(path))
+        return self.covariance()
+
+
+    def channel_from_path(self, path):
+        for ind, surface_ind in enumerate(path):
+            if ind == 0:
+                matrix = self.network_channels[ind][surface_ind]
+            else:
+                matrix = matrix@self.network_channels[ind][surface_ind]
+        self.covariance += matrix
+
+class IRS:
+    def __init__(self, num_previous_surfaces, num_elements):
+        self.phases = 1j*np.ones(num_elements)
+        self.channels = []
+        for i in range(num_previous_surfaces):
+            self.channels.append(c_rand(num_elements,num_elements))

### 2020-04-06 simulation/testing/eigen_vector_comparison.py

@@ -1,0 +1,43 @@
+import numpy as np
+from testing.circle_laws import *
+import testing.circle_laws as cl
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn
+
+size = 500
+rows = 500
+cols = 500
+bins = 300
+
+fig, ax = plt.subplots()
+plt.title("Eigen Vector Comparison: i.i.d, $\mathbb{N}(0,1/N)$, H = H_1 \Phi H_2")
+
+average = 1
+for i in range(average):
+    H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
+    G = c_rand(rows, cols)
+    F = c_rand(rows, cols)
+    GG = G @ np.conj(G.T)
+    HH = H @ np.conj(H.T)
+    FF = F @ np.conj(F.T)
+    GH = G @ np.conj(H.T)
+    HG = H @ np.conj(G.T)
+    GF = G @ np.conj(F.T)
+    FG = F @ np.conj(G.T)
+    total_irs = GH + HG + HH + GG
+    total = GF + FG + FF + GG
+    check = GG + FF
+    cross = GH + HG
+    main = GG + HH
+    e_val_total, e_vec_total = np.linalg.eig(total)
+    e_val_total_irs, e_vec_total_irs = np.linalg.eig(total_irs)
+    total_max = np.max(e_val_total)
+    irs_max = np.max(e_val_total_irs)
+    ax.plot(e_val_total, label=f'no_irs_phase', color='r')
+    ax.plot(e_val_total_irs, label=f'irs path', color='g')
+
+plt.legend(loc="upper right")
+ax.grid(True, which='both')
+seaborn.despine(ax=ax, offset=0) # the important part here
+plt.show()

### 2020-04-08 documentation/systemModel/systemModel.tex
Some changes to code but mostly working on writing work

--- 2020-03-31 documentation/systemModel/systemModel.tex
@@
 \subsection{Free Probability Analysis}\label{decont_fpt}
+Consider a channel describe by the matrix sum
+\begin{equation}
+\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
+\end{equation}
+with $\alpha_k \in \mathbb{R}$.
+Ultimately the goal is to find the AED of the covariance matrix $C_{DD^H}$ using the Stieltjes transform  $G_{C_{DD^H}}(s)$.
+In order to do this the following steps will be used.
 
-\begin{equation}
-\mathbf{D} = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
-\end{equation}
-with $\alpha_k \in \mathbb{R}$ is considered.
 Given the S-transform of the *self-adjoint matrices $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$.
@@
 
-Noting that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes
+Assuming cross terms cancel and that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes
 
@@
 \end{equation}
-Iterating this procedure and noting that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, this is exactly the case from \cite[Section 4.10]{muller2013applications}.
+Assuming cross terms cancel and that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, we iterate the above procedure just as in \cite[Section 4.10]{muller2013applications}.
 
@@
 
+\section{Solving } 
+\subsection{General Approach }
+
+Using the model from the decontamination work we say the total channel is described by 
+\begin{equation}
+\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
+\end{equation}
+We now consider how to ultimately arrive at an expression for the AED of $DD^H$
+\begin{itemize}
+
+\item 
+	We first find the Stieltjes transform of each term in the total channel $G_{D_kD_k^H}(s)$.
+\item
+	Now we use a key property of the Stieljes transform
+	\begin{equation}
+	G_{\lambda^2}(s) = \frac{G_{\lambda}(\sqrt{s})- G_{\lambda}(-\sqrt{s})}{2\sqrt{s}}
+	\end{equation}
+	to find the Stieltjes transform of the symmetrized singular value decomposition of $D_k$.
+\item
+	Taking the R-transform, $\tilde{R}_{D_k}(w)$ of the result, we then use the free additive convolution property in order to 
+	find the R-transform of the entire equivalent channel $\tilde{R}_{D}(w)$. 
+\item
+	Using the key property of the Stieltjes transform again we can find the the R-transform of the 
+	covariance matrix of the symmetrized channel, $\tilde{R}_{DD^H}(w)$ . We can then find $\tilde{G}_{DD^H}(s)$ and 
+	use the inversion to find the symmetrized AED.
+
+\item 
+	Noting that a covariance matrix is positive semi-definite, all eigenvalues are are non-negative. Therefore,
+	the symmetrized AED is equal to the true AED when evaluated for non-negative values.
+
+\end{itemize}
+
+\subsection{Solving for $G_{D_kD_k^H}(s)$ } 
+For the model presented we have 
+\begin{equation}
+D_k = G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1
+\end{equation}
+and 
+\begin{equation}
+C_N = D_kD_k^H = [G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1]
+[H_1 \boldsymbol{\Theta}_1 \cdots H_N \boldsymbol{\Theta}_N G_B]
+\end{equation}
+\begin{itemize}
+\item 
+	Because commuting matrix multiplication does not change the non-zero eigen value distribution, we consider 
+	the matrix 
+\begin{equation}
+\tilde{C}_N = [\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N  \cdots \boldsymbol{\Theta}_1 H_1 H_1^H]
+\end{equation}
+
+\item 
+	We use the following key property of the Stieltjes transform and S-Transform. 
+\begin{equation}
+S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N})
+\end{equation}
+
+\item 
+	Iterating this procedure, we can see that ultimately the phase matrices will all cancel out.
+
+\end{itemize}
+
 \begin{figure}
@@
 \begin{itemize}
-\item 
-	Show why the individual channel matrices are free as assumed in decontamination work.
-\item
-	Work on example of how to find AED given a fixed point equation for the Stieltjes transform.
+\item Key is to look at 
 \end{itemize}

### 2020-04-08 simulation/src/circle_laws.py

@@ -2,15 +2,25 @@ import numpy as np
 
 pos = lambda x: (x > 0)*x
 delta = lambda x: x == 0
+semi_cicle_aed = lambda x: (1/2*np.pi)*np.sqrt(4-np.power(x,2))
 st_half_circle = lambda s: (s / 2) * np.sqrt(1 - 4 / np.power(s, 2)) - s / 2
 st_quarter_circle = lambda s: 2*np.sqrt(4-np.power(s, 2))/np.pi * np.log((2+np.sqrt(4-np.power(s, 2)))/(-s))-s/2-2/np.pi
 estimated_pdf = lambda x: 1/np.pi * np.imag(x)
 true_pdf = lambda x: 1 / (2 * np.pi)*np.sqrt(4-np.power(x, 2))
-marcenko = lambda x, beta: pos(1-beta)*delta(x) + \
+marcenko = lambda x, beta: pos(1-1/beta)*delta(x) + \
                            np.sqrt(pos(x-np.power((1-np.sqrt(beta)), 2))*pos(np.power((1+np.sqrt(beta)), 2)-x)
                                    / (2*np.pi*beta*x))
 
-c_rand = lambda rows, cols:  np.random.randn(rows, cols)/np.sqrt(rows) + 1j*np.random.randn(rows, cols)/np.sqrt(rows)
+
+def deformed_qc_eigen(x, alpha):
+    interval = np.logical_and(np.power(1-np.sqrt(alpha), 2) < x, x < np.power(1+np.sqrt(alpha), 2))
+    output = interval*np.sqrt((x-np.power(1-np.sqrt(alpha), 2))*(np.power(1+np.sqrt(alpha), 2)-x))/(2*np.pi*alpha*x)
+    nan = np.isnan(output)
+    output[nan] = 0
+    # check = np.logical_not(interval)*(1-1/alpha)*(alpha > 1)
+    output += np.logical_not(interval)*(1-1/alpha)*(alpha > 1)
+    return output
+
 
 def eigen_vector_compare(e_values, e_vectors):
     """

### 2020-04-08 simulation/src/network_simulation_tools.py

@@ -1,6 +1,8 @@
 import copy
 from src.circle_laws import *
 import cvxpy as cp
+from src.lin_alg import *
+
 
 class Network:
     def __init__(self, num_elements, network_size):
@@ -21,16 +23,32 @@ class Network:
             self.network_channels.append(step_surfaces)
             previous_size = elements
         self.covariance = 1j*np.zeros((num_elements, num_elements))
+        self.channel = 1j*np.zeros((num_elements, num_elements))
 
-    def get_covariance(self, path=[0]):
+
+    def get_channel_covariance(self):
+        """
+        Numerical Error will usually give some imaginary components here
+        :return:
+        """
+        self.get_channel()
+        self.covariance = self.channel @ np.conj(self.channel.T)
+        return self.covariance
+
+    def get_transmit_covariance(self, transmit_powers):
+        self.transmit_powers = transmit_powers
+        self.get_channel()
+        self.transmit_covariance = self.channel @ self.transmit_powers @ np.conj(self.channel.T)
+        return self.transmit_covariance
+
+    def get_channel(self, path=[0]):
         for ind, matrix in enumerate(self.network_channels[path[-1]]):
             if len(path) == len(self.network_channels)-1:
                 self.channel_from_path(path)
             if len(path) < len(self.network_channels)-1:
                 path.append(ind)
-                self.get_covariance(path=copy.deepcopy(path))
-        return self.covariance@np.conj(self.covariance.T)
-
+                self.get_channel(path=copy.deepcopy(path))
+        return self.channel
 
     def channel_from_path(self, path):
         for ind, surface_ind in enumerate(path):
@@ -40,22 +58,38 @@ class Network:
             else:
                 irs = self.network_channels[ind][surface_ind]
                 matrix = irs.channels[path[ind]]@matrix
-        self.covariance += matrix
+        self.channel += matrix
 
 class IRS:
     def __init__(self, num_previous_surfaces, num_elements):
-        self.phases = 1j*-1j*np.ones(num_elements)  #Strange but seems to be the way to initialize to real here
+        self.phases = 1j*-1j*np.ones(num_elements)  # Strange but seems to be the way to initialize to real here
         self.channels = []
         for i in range(num_previous_surfaces):
             self.channels.append(c_rand(num_elements,num_elements))
 
 
 def water_filling(covariance_matrix, power_constraint, sigma_square=1e-2):
[…]

### 2020-04-08 simulation/testing/eigen_vector_comparison.py

@@ -9,7 +9,7 @@ cols = 500
 bins = 300
 
 fig, ax = plt.subplots()
-plt.title("Eigen Vector Comparison: i.i.d, $\mathbb{N}(0,1/N)$, H = H_1 \Phi H_2")
+plt.title("Eigen value Comparison: i.i.d, $\mathbb{N}(0,1/N), H = H_1 \Phi H_2$")
 
 average = 1
 for i in range(average):
@@ -37,5 +37,5 @@ for i in range(average):
 
 plt.legend(loc="upper right")
 ax.grid(True, which='both')
-seaborn.despine(ax=ax, offset=0) # the important part here
+seaborn.despine(ax=ax, offset=0)  # the important part here
 plt.show()

### 2020-04-08 simulation/testing/irs_comparison.py

@@ -1,4 +1,5 @@
 from src.circle_laws import *
+from src.lin_alg import *
 import numpy as np
 import matplotlib.pyplot as plt
 import seaborn
@@ -9,13 +10,13 @@ cols = 500
 bins = 300
 
 fig, ax = plt.subplots()
-plt.title("Eigen Vector Comparison: i.i.d, $\mathbb{N}(0,1/N)$, H = H_1 \Phi H_2")
+plt.title("With/Without IRS Comparison: i.i.d, $\mathbb{N}(0,1/N), H = H_1 \Phi H_2$")
 
 no_irs = []
 irs = []
 cross = []
 cross_irs = []
-average = 1
+average = 10
 for i in range(average):
     H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
     G = c_rand(rows, cols)
@@ -31,8 +32,8 @@ for i in range(average):
     total = GF + FG + FF + GG
     check = GG + FF
     cross_irs_sum = GH + HG
-    # cross_irs_sum = GG + HH
-    # cross_no_irs_sum = FF + GG
+    # cross_no_irs_sum = GG + HH
+    # cross_irs_sum = FF + GG
     cross_no_irs_sum = GF + FG
     main = GG + HH
 
@@ -54,14 +55,14 @@ irs_AED, bins_irs = np.histogram(np.asarray(irs), bins=bins)
 cross_AED, bins_cross_AED = np.histogram(np.asarray(cross), bins=bins)
 cross_irs_AED, bins_cross_irs_AED = np.histogram(np.asarray(cross_irs), bins=bins)
 
-ax.plot(bins_cross_AED[1::], cross_AED/rows, label='cross no IRS')
-ax.plot(bins_cross_irs_AED[1::], cross_irs_AED/rows, label='cross with IRS')
-ax.plot(bins_no_irs[1::], no_irs_AED/rows, label='no IRS')
-ax.plot(bins_irs[1::], irs_AED/rows, label='with IRS')
+ax.plot(bins_cross_AED[1::], cross_AED/rows, label='cross terms: no IRS')
+ax.plot(bins_cross_irs_AED[1::], cross_irs_AED/rows, label='cross terms: with IRS')
+ax.plot(bins_no_irs[1::], no_irs_AED/rows, label='total: no IRS')
+ax.plot(bins_irs[1::], irs_AED/rows, label='total: with IRS')
 
[…]

### 2020-04-08 simulation/testing/mixed_terms_check.py

@@ -1,21 +1,20 @@
 from src.circle_laws import *
 import numpy as np
 import matplotlib.pyplot as plt
+from src.lin_alg import *
 import seaborn
 
-size = 500
-rows = 500
-cols = 500
-bins = 300
+size = 100
+rows = 100
+cols = 100
+bins = int(size/2)
 
-fig, ax = plt.subplots()
-plt.title("Eigen Vector Comparison: i.i.d, $\mathbb{N}(0,1/N)$, H = H_1 \Phi H_2")
 
 no_irs = []
 irs = []
-main_none_list = []
-main_irs_list = []
-average = 1
+cross = []
+cross_irs = []
+average = 10
 for i in range(average):
     H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
     G = c_rand(rows, cols)
@@ -29,35 +28,46 @@ for i in range(average):
     FG = F @ np.conj(G.T)
     total_irs = GH + HG + HH + GG
     total = GF + FG + FF + GG
+    check = GG + FF
     cross_irs_sum = GH + HG
-    cross_no_irs_sum = GF + FG
-    main = GG + FF
-    main_irs = GG + HH
+    cross_no_irs_sum = GG + HH
+    # cross_irs_sum = FF + GG
+    # cross_no_irs_sum = GF + FG
+    main = GG + HH
+
     e_val_total, e_vec_total = np.linalg.eig(total)
     e_val_total_irs, e_vec_total_irs = np.linalg.eig(total_irs)
-    e_val_main, e_vec_total = np.linalg.eig(main)
-    e_val_main_irs, e_vec_total = np.linalg.eig(main_irs)
+    e_val_cross, e_vec_cross = np.linalg.eig(cross_no_irs_sum)
+    e_val_cross_irs, e_vec_cross_irs = np.linalg.eig(cross_irs_sum)
+
+    cross_irs_vec = eigen_vector_compare_2(e_val_cross, e_vec_cross)
+    cross_vec = eigen_vector_compare_2(e_val_cross_irs, e_vec_cross_irs)
 
     no_irs.append(np.real(e_val_total))
     irs.append(np.real(e_val_total_irs))
-    main_none_list.append(np.real(e_val_main))
-    main_irs_list.append(np.real(e_val_main_irs))
+    cross.append(np.real(e_val_cross))
[…]

### 2020-04-08 simulation/testing/test_circle.py

@@ -1,31 +1,32 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from src.circle_laws import *
 size = 500
-matrix = np.random.randn(size, size)
-half_circle = matrix + np.conj(matrix).T
+matrix = c_rand(size, size)
+half_circle = (matrix + np.conj(matrix).T)/np.sqrt(2)
 evd = np.linalg.eig(half_circle)
-eigen_value, bins = np.histogram(evd[0], bins=50)
-plt.figure()
-plt.title("Half Circle")
-plt.bar(bins[0:eigen_value.size], eigen_value)
-plt.show()
+eigen_value, bins = np.histogram(evd[0], bins=200)
+# plt.figure()
+# plt.title("Half Circle")
+# plt.bar(bins[0:eigen_value.size], eigen_value/size)
+# plt.show()
 
 # Analytic comparison to real Steiltjes
-H = np.random.randn(size, size/2)
+H = c_rand(size, size)
 X = matrix
 P = H@np.conj(H.T)@X
 evd = np.linalg.eig(P)
 eigen_value, bins = np.histogram(evd[0], bins=50)
 plt.figure()
 plt.title("Half Circle")
-plt.bar(bins[0:eigen_value.size], eigen_value)
+plt.plot(bins[0:eigen_value.size], eigen_value)
 plt.show()
-
-plt.figure()
-quarter_circle = matrix@np.conj(matrix.T)
-evd = np.linalg.eig(quarter_circle)
-eigen_value, bins = np.histogram(evd[0], bins=50)
-plt.figure()
-plt.title("Quarter Circle")
-plt.bar(bins[0:eigen_value.size], eigen_value)
-plt.show()
+#
+# plt.figure()
+# quarter_circle = matrix@np.conj(matrix.T)
+# evd = np.linalg.eig(quarter_circle)
+# eigen_value, bins = np.histogram(evd[0], bins=50)
+# plt.figure()
+# plt.title("Quarter Circle")
+# plt.bar(bins[0:eigen_value.size], eigen_value)
+# plt.show()

### 2020-04-09 documentation/systemModel/formalized_work.tex
recursion and secrecy capacity fixed

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Solution Outline}
\author{Peter Hartig}

\begin{document}
\maketitle
%\tableofcontents
%\section{Solution Outline} 
\subsection{Model}

Using the model from the pilot decontamination work we say the total channel is described by 
\begin{equation}
\mathbf{D} = \sum_{k=1}^{K}\mathbf{\mathbf{D}}_k 
\end{equation}
We now consider how to ultimately arrive at an expression for the AED of the covariance matrix $\mathbf{D}\mathbf{D}^H$.
\subsection{Approach 1} 
The expanded covariance matrix $\mathbf{D}\mathbf{D}^H$ contains both Hermetian (e.g. $\mathbf{D}_k \mathbf{D}_k ^H$) and non-hermetian terms ($\mathbf{D}_k \mathbf{D}_i^H, i \neq k$). Noting that the use of multiplicative free convolution and additive free convolution both require that the involved operators are \emph{hermetian} we could attempt to find the R-Transforms of the terms $\mathbf{D}_k \mathbf{D}_k ^H$ and $(\mathbf{D}_k \mathbf{D}_i^H+\mathbf{D}_i \mathbf{D}_k^H)$ which are both self-adjoint. 
For the terms $\mathbf{D}_k \mathbf{D}_k ^H$, we can find the S-Tranform using the methods shown in \ref{Polynomial_terms} and \cite{muller2002asymptotic}. From this S-Transform, we can move to the R-Transform in order to use additive free convolution. Finding the R-Transform of the terms $(\mathbf{D}_k \mathbf{D}_i^H+\mathbf{D}_i \mathbf{D}_k^H)$ is, however, not clear because the components of this sum are not are not self-adjoint.
Instead we consider a second approach.

\subsection{Approach 2} 
This approach is largely based on the ideas presented in \cite{muller2012channel} and which are applied in \cite[Eqns. 54-56]{muller2014blind}. The exampled \cite{muller2012channel} uses a very similar model to the IRS work.  The idea behind this approach is based on the property of the Stieltjes transform 
	\begin{equation}\label{ST_property}
	G_{\lambda^2}(s) = \frac{G_{\lambda}(\sqrt{s})- G_{\lambda}(-\sqrt{s})}{2\sqrt{s}}.
	\end{equation} (*TO DISCUSS Can we find an equivalent to this property in the S-transform domain?)
	Note that this property also requies that the matrix $D$ is bi-unitarily invariant.
	Rather than looking for the Stieltjes transform of the eigenvalues of $\mathbf{D} \mathbf{D}^H$ we will
	first find the Stieltjes transform of the singular values of $\mathbf{D}$. 
	Given the definition of the singular values $\sigma(\mathbf{D}) = \sqrt{\lambda(\mathbf{D} \mathbf{D}^H)}$ and 
	noting that 
	for a covariance matrix (positive semi-definite) all eigenvalues are non-negative, the eigen
	values of $\mathbf{D} \mathbf{D}^H$ are the squared singular values of  $\mathbf{D}$. Therefore, if 
	we can find $G_{\mathbf{D}}(s)$, we can use Property \ref{ST_property} to find $G_{\mathbf{D} \mathbf{D}^H}(s)$.
	 This is performed using the following steps.

\begin{enumerate}

\item 
	We first find the Stieltjes transform, $G_{\mathbf{D}_{k}\mathbf{D}_{k}^H}(s)$, for the AED of the 
	covariance matrix of individual terms, $\mathbf{D}_{k}$.
	This is discussed further in Section \ref{Polynomial_terms} and \cite[Section II]{muller2002asymptotic}.
\item
	Using Property \ref{ST_property} we find the Stieltjes transform of the symmetrized 
	singular value decomposition of $\mathbf{D}_{k}$, $\tilde{G}_{\mathbf{D}_{k}}(s)$. The symmetrized AED defined 
	in \cite[Eqn. 38]{muller2012channel} is given by 
	\begin{equation}
	\tilde{p}(x) = \frac{p(x)+P(-x)}{2}
	\end{equation}
\item
	Taking the R-transform, $\tilde{R}_{\mathbf{D}_{k}}(w)$ of $\tilde{G}_{\mathbf{D}_{k}}(s)$, we then use the free additive convolution property in order to 
	find the R-transform of the entire equivalent channel $\tilde{R}_{D}(w) = \sum_{k=1}^{K} \tilde{R}_{\mathbf{D}_{k}}(w) $ . 
	* Note that while additive free convolution holds for $\tilde{R}_{\mathbf{D}_{k}}(w)$, multiplicative free convolution does
	 not.
\item
	Using $\tilde{R}_{D}(w)$ we then find $\tilde{G}_{D}(s)$. We then use property \eqref{ST_property} in the reverse
	 order to get $\tilde{G}_{DD^H}(s)$  which allows us to evaluate the  symmetrized AED.

\item 
	Because all eigenvalues are are non-negative for the covariance matrix $\mathbf{D}\mathbf{D}^H$
	the symmetrized AED is equal to the true AED when evaluated at non-negative values and can be divided by 2
	when evaluated at 0.

\end{enumerate}

\subsection{Solving for $G_{\mathbf{D}_{k}\mathbf{D}_{k}^H}(s)$}\label{Polynomial_terms}
Using the model from the IRS work, the individual terms of the channel are given by
\begin{equation}
\mathbf{D}_{k} = G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1
\end{equation}
and 
\begin{equation}
C_N = \mathbf{D}_{k}\mathbf{D}_{k}^H = [G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1]
[H_1 \boldsymbol{\Theta}_1 \cdots H_N \boldsymbol{\Theta}_N G_B]
\end{equation}
\begin{itemize}
\item 
	Because commuting matrix multiplication does not change the non-zero eigenvalue distribution, we consider 
	the matrix 
\begin{equation}
\tilde{C}_N = [\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1 H_1^H]
\end{equation}

\item 
	We use the rotational property of the Stieltjes transform and S-Transform. 
\begin{equation}
S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N}).
\end{equation}
For the two terms of $\tilde{C}_N$, 
$ H_1 H_1^H$ and $[\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1]$
we can use multiplicative free convolution .

\item
	For the gaussian IID case, $ H_1 H_1^H$  has the marcenko-pastur distribution which has a well defined S-Transform.

\item 
	Iterating the rotation property with the second term 
	\begin{equation}
	[\boldsymbol{\Theta}_1^H H_2 \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H]
	[G_B \boldsymbol{\Theta}_N H_N  \cdots H_2^H \boldsymbol{\Theta}_1]
	\end{equation}

	\begin{equation}
	[ \boldsymbol{\Theta}_N H_N  \cdots H_2^H \boldsymbol{\Theta}_1]
	[\boldsymbol{\Theta}_1^H H_2 \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H G_B]
	\end{equation}

	 it is seen that ultimately, the phase matrices will cancel out and the result is a product of well known S-Transforms.

\item 
*TO DISCUSS If there is no equivalent to the Property \ref{ST_property} in the S-Transform domain, how do we 
find a closed form of the Stieltjes transform given an arbitrary S-transform?
\end{itemize}

\bibliography{bibliography}
\end{document}

### 2020-04-09 documentation/systemModel/systemModel.tex
recursion and secrecy capacity fixed

@@
 \begin{equation}
-\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
+\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k 
 \end{equation}
@@
 \item 
-	We first find the Stieltjes transform of each term in the total channel $G_{D_kD_k^H}(s)$.
+	We first find the Stieltjes transform, $G_{D_kD_k^H}(s)$, for the AED of each term, $D_k$.
+	This is discussed further in Section \ref{Polynomial_terms}.
 \item
-	Now we use a key property of the Stieljes transform
-	\begin{equation}
+	Now we use a property of the Stieljes transform
+	\begin{equation}\label{ST_property}
 	G_{\lambda^2}(s) = \frac{G_{\lambda}(\sqrt{s})- G_{\lambda}(-\sqrt{s})}{2\sqrt{s}}
 	\end{equation}
-	to find the Stieltjes transform of the symmetrized singular value decomposition of $D_k$.
+	to find the Stieltjes transform of the symmetrized singular value decomposition of
+	 $D_k$, $\tilde{G}_{D_k}(s)$. Note that is property is reasonable to use because
+	 we consider Gaussian  $D_k$ which are bi-unitarily invariant.
 \item
@@
 \item
-	Using the key property of the Stieltjes transform again we can find the the R-transform of the 
+	Using property \eqref{ST_property}, we can find the the R-transform of the 
 	covariance matrix of the symmetrized channel, $\tilde{R}_{DD^H}(w)$ . We can then find $\tilde{G}_{DD^H}(s)$ and 
@@
 
-\subsection{Solving for $G_{D_kD_k^H}(s)$ } 
+\subsection{Solving for $G_{D_kD_k^H}(s)$}\label{Polynomial_terms}
 For the model presented we have 
@@
 \item 
-	We use the following key property of the Stieltjes transform and S-Transform. 
+	We use the rotational property of the Stieltjes transform and S-Transform. 
 \begin{equation}

### 2020-04-09 simulation/src/network_simulation_tools.py

@@ -38,39 +38,37 @@ class Network:
         Numerical Error will usually give some imaginary components here
         :return:
         """
+        self.covariance_eve = self.channel_eve @ np.conj(self.channel_eve.T)
+        self.covariance = self.channel @ np.conj(self.channel.T)
         if eve:
-            self.covariance_eve = self.channel_eve @ np.conj(self.channel_eve.T)
             return self.covariance, self.covariance_eve
         else:
-            self.covariance = self.channel @ np.conj(self.channel.T)
             return self.covariance
 
 
     def get_transmit_covariance(self, transmit_powers, eve=False):
         self.paths_check = 0
         self.transmit_powers = transmit_powers
+        self.transmit_covariance = self.channel @ self.transmit_powers @ np.conj(self.channel.T)
+        self.transmit_covariance_eve = self.channel_eve @ self.transmit_powers @ np.conj(self.channel_eve.T)
         if not eve:
-            self.transmit_covariance = self.channel @ self.transmit_powers @ np.conj(self.channel.T)
             return self.transmit_covariance
         if eve:
-            self.transmit_covariance_eve = self.channel_eve @ self.transmit_powers @ np.conj(self.channel_eve.T)
             return self.transmit_covariance, self.transmit_covariance_eve
 
 
-    def get_channel(self, path=[0]):
-        self.channel = 1j*np.zeros(self.channel.shape)
-        self.channel_eve = 1j*np.zeros(self.channel.shape)
+    def get_channel(self, path=[], i=0):
+        if len(path) == len(self.network_channels):
+            self.channel_from_path(path)
+        else:
+            for ind in range(self.network_size[i]):
+                if len(path) < len(self.network_channels):
+                    self.get_channel(path=path+[ind], i=i+1)
 
-        for ind, matrix in enumerate(self.network_channels[path[-1]]):
-            if len(path) == len(self.network_channels):
-                self.channel_from_path(path)
-            if len(path) < len(self.network_channels):
-                path.append(ind)
-                self.get_channel(path=copy.deepcopy(path))
-        return self.channel
 
     def channel_from_path(self, path):
         matrix = None
+        previous = 0
         self.paths_check += 1
         for ind, surface_ind in enumerate(path):
             if ind == 0:
@@ -78,7 +76,8 @@ class Network:
                 matrix = irs.channels[path[ind]]
             else:
                 irs = self.network_channels[ind][surface_ind]
[…]

### 2020-04-09 simulation/testing/network.py

@@ -9,7 +9,9 @@ plt.title("Number of IRS (N=1): i.i.d, $\mathbb{N}(0,1/N)$, H = H_1 \Phi H_2")
 surface_size = 300
 bins = int(surface_size/2)
 list_irs = [1, 1]
+capacity = []
 secrecy_capacities = []
+capacity_water = []
 secrecy_capacities_water = []
 for num_irs in [2, 4]:
     list_irs += [1]
@@ -18,20 +20,26 @@ for num_irs in [2, 4]:
     cov_water = net.get_transmit_covariance(np.diag(optiized_powers))
     e_val_water, e_vec_water = np.linalg.eig(cov_water)
     AED_water, AED_bins_water = np.histogram(np.asarray(e_val_water), bins=bins)
-    secrecy_capacities_water.append(net.get_capacity())
+    secrecy_capacities_water.append(net.get_capacity(secrecy=True))
+    capacity_water.append(net.get_capacity())
 
     cov = net.get_transmit_covariance(np.eye(surface_size)/surface_size)
     e_val, e_vec = np.linalg.eig(cov)
     AED, AED_bins = np.histogram(np.asarray(e_val), bins=bins)
-    secrecy_capacities.append(net.get_capacity())
+    secrecy_capacities.append(net.get_capacity(secrecy=True))
+    capacity.append(net.get_capacity())
 
 
     ax.plot(AED_bins[1::], AED/surface_size, label=f'Number of IRS: {num_irs}')
     ax.plot(AED_bins_water[1::], AED_water/surface_size, label=f'Water Filling: {num_irs}')
 
 
-print(f"no water filling {secrecy_capacities} \n")
-print(f"water filling {secrecy_capacities_water} \n")
+print(f"no water filling {capacity} \n")
+print(f"water filling {capacity_water} \n")
+print(f"sec no water filling {secrecy_capacities} \n")
+print(f"sec water filling {secrecy_capacities_water} \n")
+
+
 
 
 plt.legend(loc="upper right")

### 2020-04-12 simulation/testing/MISO_phase_optimization.py
added initial presentation work and work on MISO beamforming optmization

@@ -1,0 +1,21 @@
+import numpy as np
+import cvxpy as cp
+from src.lin_alg import *
+
+H_1 = c_rand(100, 100)
+h_2 = c_rand(100, 1)
+h = c_rand(100, 1)
+f = np.ones((100, 1))
+
+Thetas = cp.Variable((100, 1), complex=True)
+# constraint = [cp.abs(Thetas) <= np.ones(Thetas.shape)]
+constraint = [cp.abs(phase) <= 1 for phase in Thetas]
+utility = [cp.real(np.conj(h.T)@np.conj(H_1.T)@cp.diag(Thetas).H@h_2 + np.conj(h_2.T)@cp.diag(Thetas)@H_1@h)]
+# utility = [cp.norm2((np.conj(h_2.T)@cp.diag(Thetas)@H_1 + np.conj(h.T))@f)]
+prob = cp.Problem(cp.Maximize(cp.sum(utility)), constraint)
+prob.solve(verbose=True)
+Thetas = Thetas.value
+
+
+
+

### 2020-04-16 documentation/systemModel/formalized_work.tex
just some small changes while investigating impact of cross terms

@@
 
+\item
+*TO DISCUSS If the phases cancel out, then they cannot influence the final AED or the expected value. Does this mean that controlling these phases is not beneficial in the asymptotic limit?
+Note these terms might affect the AED of the cross terms in the polynomial but we hypothesis that these cancel out anyways.
+
 \bibliography{bibliography}

### 2020-04-16 simulation/testing/phase_optimization/MIMO_waterfill_phase.py

@@ -6,15 +6,15 @@ import seaborn
 
 fig, ax = plt.subplots()
 plt.title("")
-surface_size = 100
+surface_size = 1
 bins = int(surface_size/2)
 original_network = [1]
-list_irs = [1, 2]
+list_irs = [2]
 capacity_water = []
 capacity_water_phase = []
 for num_irs in [4]:
-    list_irs = original_network + [num_irs]
-    # list_irs = list_irs + [1]
+    # list_irs = original_network + [num_irs]
+    list_irs = list_irs + [1]
     net = Network(surface_size, copy.copy(list_irs))
     c1 = net.channel
     optimized_powers_1 = water_filling(net.channel, 1)

### 2020-04-25 documentation/systemModel/systemModel2.tex
added work for second system model

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Summary of Work}
Below is a summary of my work up to now.
\section{System Model}
I consider a flat-fading MIMO broadcast channel with a linear precoding at the transmitter. 
Using a basic model with an IRS component and line of sight (LOS) component gives the received signal 
	 \begin{equation}
		\mathbf{r} = \mathbf{H}_{Total}\mathbf{F}\mathbf{x}.
	\end{equation}
	The total channel, is given by
	\begin{equation}
	\mathbf{H}_{Total} =  \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}
	\end{equation}
	with both the IRS component $ \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}$ and 
	the LOS component $\mathbf{G}$. To simplify notation I will assume 
	$\mathbf{H}_{1} \in \mathbb{C}_{N \times N}$ ...

\section{Channel Capacity}
First the capacity of the total channel $\mathbf{H}_{Total}$ is considered. 
Following the methods from (non-hermetian), we can find the AED of the channel 
covariance matrix, $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ using the asymptotic distribution of the "symmetrized singular values" of $\mathbf{H}_{Total}$.
One problem I am currently working on is that the fixed point equation  provided in (non-hermetian) assumes that the
 elements of $ \mathbf{H}_{1}$ and  $\mathbf{H}_{2}$ have variance $\frac{1}{RS}$. Typically we want these
  matrices to have elements with variance $\frac{1}{R}$.
  If I were to use the equation from (non-hermetian), the IRS component would have no impact on the capacity of system. In order to resolve this, I will need to re-derive this which requires using subordination methods.

\subsection{Canceling Phases}
Additive free convolution of the R-Transform holds for symmetrized singular value distributions. Therefore we can consider $R_{\mathbf{H}_{Total}} = R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}} + R_{\mathbf{G}}$. In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ and $R_{\mathbf{G}}$ we will use the property 
\begin{equation}
\tilde{G}_{B}(s) = sG_{BB^{\dagger}}(s^2).
\end{equation}
For the term $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ we begin by finding 
$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ 
Using the method shown in (vector channel paper), the rotational property of the S-Transform
\begin{equation}
S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N}).
\end{equation}
is used with 
\begin{equation}
\tilde{C}_N = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
\end{equation}.
Applying the rotational property again to the term 
\begin{equation}
\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
\end{equation}
we get 
\begin{equation}
\boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
\end{equation}
with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.
From this, we can conclude that the present of the phase-shift matrix $\boldsymbol{\Phi}$ does not influence the
AED of the channel covariance matrix and thus does not affect the channel capacity.

\section{Optimization}

\section{Simulation Results}

\subsection{Evaluating Impact of Cross-terms on AED}
Here we wanted to see if the final AED is influenced by the cross terms in the polynomial describing the channel
covariance matrix given by $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$

\begin{figure}[H]
	\includegraphics[width=\textwidth]{results/cross}
\caption{Eigenvalue histogram for N=500. Cross terms represent the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H)$ and the non-cross terms are the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H)$}
\label{irs_figure}
\end{figure}

\subsection{Impact of Phases on total Capacity}
It is hypothesized that in the asymptotic limit, the phases will not impact the capacity. In order to check this numerically I evaluated the capacity while holding the channel values constant for different values for the IRS phases. Here statistics are shown for 50 random realizations of the phases at the IRS. 
Phases, $e^{j\phi}$, are generated using a uniform distribution of $\phi \in \text{Uniform}[0,2\pi]$ and.
\begin{itemize}
\item 
Capacity variance:
 8.968116152773685e-06
\item 
Capacity average:
 1.9860953144228657
\item 
Capacity min:
 1.9783515328782062
\item 
Capacity max:
 1.993849732554225
\end{itemize}

\bibliography{bibliography}
\end{document}

### 2020-04-25 simulation/testing/channel_testing/direct_scatter.py

@@ -106,13 +106,13 @@ print(f"theoretic capacity: {theoretic_capacity}")
 #   Plot results
 fig, ax = plt.subplots()
 plt.title("IRS and Line of Sight: i.i.d, $\mathbb{N}(0,1/N), H = H_1 \Phi H_2$")
-ax.bar(bins_irs[1::], irs_AED/rows, label='total: AED')
+ax.bar(bins_irs[1::], irs_AED/rows, label='simulated AED')
 # ax.bar(bins_irs_svd[1::], irs_SVD/rows, label='total: ASD')
-ax.plot(x_values, normalize(theoretic_pdf), label='theoretical distribution', c='r')
+ax.plot(x_values, normalize(theoretic_pdf), label='normalized theoretical aed', c='r')
 
 plt.legend(loc="upper right")
 ax.grid(True, which='both')
 seaborn.despine(ax=ax, offset=0)
-# plt.show()
+plt.show()
 
 

### 2020-04-25 simulation/testing/phase_optimization/phase_check_simulation.py

@@ -1,0 +1,38 @@
+from src.circle_laws import *
+import numpy as np
+import matplotlib.pyplot as plt
+from src.lin_alg import *
+import seaborn
+
+size = 500
+rows = 500
+cols = 500
+bins = int(size/2)
+
+
+capacities = []
+cross_sum = 0
+main_sum = 0
+
+average = 50
+H_1 = c_rand(rows, cols)
+H_2 = c_rand(rows, cols)
+G = c_rand(rows, cols)
+F = c_rand(rows, cols)
+for i in range(average):
+    H = H_2 @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ H_1
+    # F = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
+    GG = G @ np.conj(G.T)
+    HH = H @ np.conj(H.T)
+    GH = G @ np.conj(H.T)
+    HG = H @ np.conj(G.T)
+    cross = GH + HG
+    main = HH + GG
+    total_irs = cross + main
+    capacities.append(capacity(total_irs, 1/rows))
+
+print(f"Capacity variance:\n {np.var(capacities)}")
+print(f"Capacity average:\n {np.average(capacities)}")
+print(f"Capacity min:\n {np.min(capacities)}")
+print(f"Capacity max:\n {np.max(capacities)}")
+print(f"Capacities:\n {capacities}")

### 2020-04-26 documentation/systemModel/systemModel2.tex
updates of current report

@@
 Using a basic model with an IRS component and line of sight (LOS) component gives the received signal 
-	 \begin{equation}
+	 \begin{equation}\label{system_model}
 		\mathbf{r} = \mathbf{H}_{Total}\mathbf{F}\mathbf{x}.
@@
 	the LOS component $\mathbf{G}$. To simplify notation I will assume 
-	$\mathbf{H}_{1} \in \mathbb{C}_{N \times N}$ ...
+	$\mathbf{H}_{1},\mathbf{H}_{2}, \mathbf{G} \in \mathbb{C}_{N \times N}$
 
 \section{Channel Capacity}
-First the capacity of the total channel $\mathbf{H}_{Total}$ is considered. 
+First the capacity of the channel $\mathbf{H}_{Total}$ is considered without precoding at the transmitter. 
 Following the methods from (non-hermetian), we can find the AED of the channel 
@@
 Additive free convolution of the R-Transform holds for symmetrized singular value distributions.
-Therefore we can consider $R_{\mathbf{H}_{Total}} = R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}} + R_{\mathbf{G}}$.
-In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ and $R_{\mathbf{G}}$ we will use the property
-\begin{equation}
+Therefore we can consider $R_{\mathbf{H}_{Total}}(w) = R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w) + R_{\mathbf{G}}(w)$.
+In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w)$ and $R_{\mathbf{G}}(w)$, we use the property
+\begin{equation}\label{svd_aed_property}
 \tilde{G}_{B}(s) = sG_{BB^{\dagger}}(s^2).
@@
 For the term $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ we begin by finding 
-$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ 
-Using the method shown in (vector channel paper), the rotational property of the S-Transform
+$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ in order to use property \eqref{svd_aed_property} to find $G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(s)$
+\par
+To find $G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ we define
 \begin{equation}
+C_N = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H
+\end{equation}
+and the rotated version 
+\begin{equation}
+\tilde{C}_N = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
+\end{equation} in order to use the rotational property of the S-Transform shown in (vector channel paper)
+\begin{equation}\label{rotation_property}
 S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N}).
 \end{equation}
-is used with 
-\begin{equation}
-\tilde{C}_N = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
-\end{equation}.
+in which $\chi_N$ is the aspect ratio of the channel matrices. Here we assume $\chi_N = 1$
 Applying the rotational property again to the term 
@@
 with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.
-From this, we can conclude that the present of the phase-shift matrix $\boldsymbol{\Phi}$ does not influence the
+From this, we can conclude that the phase-shift matrix $\boldsymbol{\Phi}$ does not influence the
 AED of the channel covariance matrix and thus does not affect the channel capacity.
@@
 \section{Optimization}
+Now we consider how to choose the transmit covariance matrix $\mathbf{Q}_x$ and $\boldsymbol{\Phi}$ in order to optimize an objective.
+First, We will optimize the precoder and IRS phases to maximize the channel capacity.
+		    \begin{equation}
+    \begin{array}{ll}
+    \underset{\boldsymbol{\Phi}, \mathbf{Q}_x}{\text{maximize }}   & \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{P_{\text{Total}}}{N_T \sigma_n}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]\mathbf{Q}_x[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]^H|\right)\right]
+    \\
+    \mbox{subject to } & \|\phi_i\|^2 = 1
+    \\
+    & tr(\mathbf{Q}_x) \leq P_{\text{Total}}
+%        \\
+%    & \tilde{\mathbf{f}}_i \mathbf{h}_j = 0, \; i\neq j, \; \forall i,j \in N
+    \end{array}
+    \label{general_irs_opt}
+    \end{equation}
+
+If the proposal above is true and the phase matrices do no affect the capacity in the asymptotic case, then it would seem that standard water filling could be performed in order to choose the covariance matrix $\mathbf{Q}_x$.
 
@@
 Here we wanted to see if the final AED is influenced by the cross terms in the polynomial describing the channel
-covariance matrix given by $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$
+covariance matrix given by $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$. The case in which elements of the channel matrices are zero-mean with variance  $\frac{1}{N}$ is shown in \ref{cross}
 
@@
 Cross terms represent the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H)$ and the non-cross terms are the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H)$}
-\label{irs_figure}
+\label{cross}
+\end{figure}
+
+If the elements of the channel matrix are, however, not zero-mean these cross terms will impact the channel AED. 
+This scenario is shown in \ref{cross_non_zero_mean}.
+
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{results/cross_non_zero_mean}
+\caption{Eigenvalue histogram for N=500 with non-zero mean elements.}
+\label{cross_non_zero_mean}
 \end{figure}
@@
 In order to check this numerically I evaluated the capacity while holding the channel values constant for different values for the IRS phases.
-Here statistics are shown for 50 random realizations of the phases at the IRS.
-Phases, $e^{j\phi}$, are generated using a uniform distribution of $\phi \in \text{Uniform}[0,2\pi]$ and.
+Here statistics are shown for 100 random realizations of the phases at the IRS.
+Phases, $e^{j\phi}$, are generated using a uniform distribution of $\phi \in \text{Uniform}[0,2\pi]$. This is clearly not rigorous but perhaps provides some insight. 
+$N_t = N_r = M = 500$
 \begin{itemize}
@@
 Capacity variance:
- 8.968116152773685e-06
+ 1.1512987368371963e-05
  \item
 Capacity average:
- 1.9860953144228657
+ 1.990962024050068
  \item
 Capacity min:
- 1.9783515328782062
+ 1.9818993465470727
  \item
 Capacity max:
- 1.993849732554225
+ 1.9986482518130326
+\end{itemize}
+
+If, however, the aspect ratio of the channel matrices are not 1, the results below reflect that the phases play a more important role in determining the capacity.
+This supports the terms of the in equation \eqref{rotation_property} highlighted in red below.
+
+\begin{equation}\label{rotation_property}
+S_{C_N}(z) = \frac{z+1}{z+ \textcolor{red}{\chi_N}} S_{\tilde{C}_N}(\frac{z}{\textcolor{red}{\chi_N}}).
+\end{equation}
+
+As indicated in the simulation with $N_t = N_r = 4$ and $M = 100$
+
+\begin{itemize}
+\item 
+Capacity variance:
+ 0.04863896876118487
+ \item
+Capacity average:
+ 1.695035095091801
+ \item
+Capacity min:
+ 1.2342437149649095
+ \item
+Capacity max:
+ 2.320914548447325
 \end{itemize}

### 2020-04-26 simulation/src/lin_alg.py

@@ -11,7 +11,9 @@ def aed_capacity(value, value_probability, step, snr, number_receivers):
     return number_receivers*np.sum(np.log(1 + snr*value)*probabilities)
 
 
-c_rand = lambda rows, cols:  (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
+c_rand = lambda rows, cols: (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
+c_rand_mean = lambda rows, cols: 1e-1 + 1j*1e-1 + (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
+
 c_rand_2 = lambda rows, cols:  (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(cols*rows)
 
 # c_rand = lambda rows, cols: -1j*1j*np.eye(rows)

### 2020-04-27 documentation/systemModel/systemModel2.tex
some work on initial presentation

--- 2020-04-25 documentation/systemModel/systemModel2.tex
@@
 \section{System Model}
-I consider a flat-fading MIMO broadcast channel with a linear precoding at the transmitter. 
-Using a basic model with an IRS component and line of sight (LOS) component gives the received signal 
-	 \begin{equation}
-		\mathbf{r} = \mathbf{H}_{Total}\mathbf{F}\mathbf{x}.
-	\end{equation}
+I consider a flat-fading MIMO channel with an IRS component and a line of sight (LOS) component. In order to analyze the capacity, I consider the channel without a transmitter or receiver filter.
+
+\section{Channel Capacity}
+The capacity of the channel from the system
+
+	 \begin{equation}\label{system_model}
+		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x}.
+	\end{equation} is considered. 
 	The total channel, is given by
 	\begin{equation}
-	\mathbf{H}_{Total} =  \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}
+	\mathbf{H}_{Total} =  \underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}_{\text{IRS component}} + \underbrace{\mathbf{G}}_{\text{LOS component}}.
 	\end{equation}
-	with both the IRS component $ \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}$ and 
-	the LOS component $\mathbf{G}$. To simplify notation I will assume 
-	$\mathbf{H}_{1} \in \mathbb{C}_{N \times N}$ ...
+	with 
+	$\mathbf{H}_{1}\in \mathbb{C}_{S \times T},\mathbf{H}_{2} \in \mathbb{C}_{R \times S}, \mathbf{G} \in \mathbb{C}_{R \times T}$ and $\boldsymbol{\Phi}$ as a diagonal matrix with elements $e^{j\phi_i}$.
 
-\section{Channel Capacity}
-First the capacity of the total channel $\mathbf{H}_{Total}$ is considered. 
-Following the methods from (non-hermetian), we can find the AED of the channel 
-covariance matrix, $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$ using the asymptotic distribution of the "symmetrized singular values" of $\mathbf{H}_{Total}$.
-One problem I am currently working on is that the fixed point equation  provided in (non-hermetian) assumes that the
- elements of $ \mathbf{H}_{1}$ and  $\mathbf{H}_{2}$ have variance $\frac{1}{RS}$. Typically we want these
-  matrices to have elements with variance $\frac{1}{R}$.
-  If I were to use the equation from (non-hermetian), the IRS component would have no impact on the capacity of system.
-In order to resolve this, I will need to re-derive this which requires using subordination methods.
+Following the approach from \cite{muller2012channel}, we can find the AED of the channel 
+covariance matrix, $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$, using the asymptotic distribution of the "symmetrized singular values" of $\mathbf{H}_{Total}$.\cite{muller2012channel} provides a fixed point equation for the Stieltjes transform of this system.
+One problem, however, is that the system in this formulation considers
+ elements of $ \mathbf{H}_{1}$ and  $\mathbf{H}_{2}$ to have variance $\frac{1}{RS}$. We want these
+  matrices to have elements with variance $\frac{1}{s}$ and $\frac{1}{R}$ respectively.
+  If I were to use the equation from \cite{muller2012channel}, the IRS component of the channel would have no impact on the capacity of system because all of the eigenvalues would be so small.
+In order to resolve this, I will need to re-derive the fixed point equation which requires using subordination methods.
+I am currently working this.
+ \par
+ Some numerical results investigating the eigenvalues of the terms from the covariance matrix polynomial are show in 
+ section \ref{cross_terms}.
 
 \subsection{Canceling Phases}
-Additive free convolution of the R-Transform holds for symmetrized singular value distributions.
-Therefore we can consider $R_{\mathbf{H}_{Total}} = R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}} + R_{\mathbf{G}}$.
-In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ and $R_{\mathbf{G}}$ we will use the property
+While deriving the fixed point equation for the AED, one step includes using additive free convolution to find the R-Transform for symmetrized singular value distributions using the property 
 \begin{equation}
+\tilde{R}_{\mathbf{H}_{Total}}(w) = \tilde{R}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w) + \tilde{R}_{\mathbf{G}}(w).
+\end{equation} In order to find the individual terms $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(w)$ and $R_{\mathbf{G}}(w)$, we use the property relating the symmetrized singular value distribution to the eigenvalue distribution.
+\begin{equation}\label{svd_aed_property}
 \tilde{G}_{B}(s) = sG_{BB^{\dagger}}(s^2).
 \end{equation}
-For the term $R_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}$ we begin by finding 
-$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ 
-Using the method shown in (vector channel paper), the rotational property of the S-Transform
+We begin by finding the stieltjes transform of the eigenvalue distributions of the individual component,
+$G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$, in order to use property \eqref{svd_aed_property} to find $\tilde{G}_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}(s)$ (the symmetrized singular valued distribution).
+\par
+To find $G_{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}(s)$ we define
 \begin{equation}
-S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N}).
+C_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H
 \end{equation}
-is used with 
+and a rotated version 
 \begin{equation}
-\tilde{C}_N = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
+\tilde{C}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H
 \end{equation}. 
+These allow use of the rotational property of the S-Transform from \cite{muller2002asymptotic}
+\begin{equation}\label{rotation_property}
+S_{C_2}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_2}(\frac{z}{\chi_2}).
+\end{equation}
+in which $\chi_N$ is the aspect ratio of the channel matrices. In the given model $\chi_2 = \frac{S}{R}$ and
+$\chi_2 = \frac{S}{R}$.
 Applying the rotational property again to the term 
 \begin{equation}
-\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
+C_{1} = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
+\tilde{C}_{1} \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
 \end{equation}
 with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.
-From this, we can conclude that the present of the phase-shift matrix $\boldsymbol{\Phi}$ does not influence the
-AED of the channel covariance matrix and thus does not affect the channel capacity.
+
+From this we can conclude that for $R,S,T \rightarrow \infty $ the only impact of the IRS comes from the aspect ratio
+$\chi_N$ but the actual phases of the IRS do not impact the final AED.
+The results shown in \cite{zhang2019capacity} do not reflect the result only because 
+the values $R = 4, S = 100 , T = 4$ are chosen so these asymptotic results do not apply. 
+These results can be seem in
 
 \section{Optimization}
+For the system
+\begin{equation}
+		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x},
+	\end{equation}
+	with  $R,S,T \rightarrow \infty $,
+we now consider
+how to choose the transmit covariance matrix $\mathbf{Q}_x$ and $\boldsymbol{\Phi}$ in order to optimize an objective.
+First, We will optimize the precoder and IRS phases to maximize the channel capacity.
+		    \begin{equation}
+    \begin{array}{ll}
+    \underset{\boldsymbol{\Phi}, \mathbf{Q}_x}{\text{maximize }}   & \Expect\left[\Log\left(|\mathbf{I}_{N_R}+\frac{1}{\sigma_n}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]\mathbf{Q}_x[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]^H|\right)\right]
+    \\
+    \mbox{subject to } & \|\phi_i\|^2 = 1
+    \\
+    & tr(\mathbf{Q}_x) \leq P_{\text{Total}}
+%        \\
+%    & \tilde{\mathbf{f}}_i \mathbf{h}_j = 0, \; i\neq j, \; \forall i,j \in N
+    \end{array}
+    \label{general_irs_opt}
+    \end{equation}
+
+%We assume that  $R,S,T \rightarrow \inf $. In cases when  $\chi_N = 1$ it is shown that the phases do 
+%not impact the capacity and thus standard water-filling can be used. If, however, $\chi_N \neq 1$, 
+%Then the problem is non-convex and many approaches are possible.
 
@@
 
-\subsection{Evaluating Impact of Cross-terms on AED}
-Here we wanted to see if the final AED is influenced by the cross terms in the polynomial describing the channel
-covariance matrix given by $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$
+\subsection{Evaluating Impact of Cross-terms on AED}\label{cross_terms}
+The polynomial describing the channel
+covariance matrix given by 
+\begin{equation}
+\mathbf{H}_{Total}\mathbf{H}_{Total}^H = 
+\underbrace{
+\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}_{\text{cross terms}}
++ 
+\underbrace{
+\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H}_{\text{non-cross terms}}.
+\end{equation} 
+Here we wanted to see if the final AED is influenced by the cross terms.
+The case in which elements of the channel matrices are zero-mean with variance  $\frac{1}{N}$ is shown in \ref{cross}.
+In this case, the cross terms appear to have symmetry such that they will not influence the final AED.
 
@@
 Cross terms represent the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H)$ and the non-cross terms are the eigenvalues of the matrix sum $(\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H)$}
-\label{irs_figure}
+\label{cross}
 \end{figure}
 
-\subsection{Impact of Phases on total Capacity}
-It is hypothesized that in the asymptotic limit, the phases will not impact the capacity.
-In order to check this numerically I evaluated the capacity while holding the channel values constant for different values for the IRS phases.
-Here statistics are shown for 50 random realizations of the phases at the IRS.
-Phases, $e^{j\phi}$, are generated using a uniform distribution of $\phi \in \text{Uniform}[0,2\pi]$ and.
+If the elements of the channel matrix are, however, not zero-mean the cross terms will impact the channel AED. 
+This scenario is shown in \ref{cross_non_zero_mean}.
+
+\begin{figure}[H]
+	\includegraphics[width=\textwidth]{results/cross_non_zero_mean}
+\caption{Eigenvalue histogram for N=500 with non-zero mean elements.}
+\label{cross_non_zero_mean}
+\end{figure}
+
+\subsection{Impact of Phases on total Capacity}\label{phase_impact}
+It is hypothesized that in the asymptotic limit, if $\chi_n = 1$ the phases will not impact the capacity.
+In order to check this numerically I evaluated the capacity while holding the channel values constant for different values of the IRS phases.
+Here statistics are shown for 100 random realizations of the phases at the IRS.
+Phases, $e^{j\phi}$, are generated randomly over a uniform distribution $\phi \in [0,2\pi]$. This is clearly not rigorous but perhaps provides some insight. 
+$T = R = S = 500$
 \begin{itemize}
@@
 Capacity variance:
- 8.968116152773685e-06
+ 1.1512987368371963e-05
  \item
 Capacity average:
- 1.9860953144228657
+ 1.990962024050068
  \item
 Capacity min:
- 1.9783515328782062
+ 1.9818993465470727
  \item
 Capacity max:
- 1.993849732554225
+ 1.9986482518130326
+\end{itemize}
+Similarly, the results for $T = 500 R = 500$ and $S = 1000$ are shown 
+\begin{itemize}
+\item
+Capacity variance:
+ 1.730056273847992e-05
+ \item 
+Capacity average:
+ 1.9937729377849243
+ \item
+Capacity min:
+ 1.9855835490537985
+ \item
+Capacity max:
+ 2.0025075539557737
+\end{itemize}
+
+%\textcolor{red}{\chi_N}}
+
+In the non-asymptotic case with $T = R = 4$ and $M = 100$, the results below indicate that the phases play a more important role in determining the capacity.
+
+\begin{itemize}
+\item 
+Capacity variance:
+ 0.04863896876118487
+ \item
+Capacity average:
+ 1.695035095091801
+ \item
+Capacity min:
+ 1.2342437149649095
+ \item
+Capacity max:
+ 2.320914548447325
 \end{itemize}

### 2020-04-29 documentation/systemModel/systemModel2.tex
Finished content for initial presentation

@@
 \begin{equation}
-\tilde{C}_{1} \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
+\tilde{C}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
 \end{equation}
@@
 $\chi_N$ but the actual phases of the IRS do not impact the final AED.
-The results shown in \cite{zhang2019capacity} do not reflect the result only because 
-the values $R = 4, S = 100 , T = 4$ are chosen so these asymptotic results do not apply. 
-These results can be seem in
-
+The results shown in \cite{zhang2019capacity} do not reflect that the phases play no role in capacity because 
+the matrix dimensions chosen, $R = 4, S = 100 , T = 4$ are not applicable to asymptotic analysis. 
+These results can be seen in 
 \section{Optimization}
@@
 \underbrace{
-\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}_{\text{cross terms}}
+\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H + \mathbf{G}\mathbf{G}^H}_{\text{cross terms}}
 + 
 \underbrace{
-\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}\mathbf{G}^H}_{\text{non-cross terms}}.
+\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H + \mathbf{G}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H}_{\text{non-cross terms}}.
 \end{equation} 
@@
 \end{itemize}
-Similarly, the results for $T = 500 R = 500$ and $S = 1000$ are shown 
+Similarly, the results for $T = 500$, $ R = 500$ and $S = 1000$ are shown 
 \begin{itemize}

### 2020-04-29 simulation/testing/phase_optimization/phase_check_simulation.py

@@ -4,9 +4,9 @@ import matplotlib.pyplot as plt
 from src.lin_alg import *
 import seaborn
 
-size = 1000
-n_t = 500
-n_r = 500
+size = 100
+n_t = 4
+n_r = 4
 rows = size
 cols = size
 bins = int(size/2)
@@ -16,7 +16,7 @@ capacities = []
 cross_sum = 0
 main_sum = 0
 
-average = 50
+average = 1000
 H_1 = c_rand(size, n_t)
 H_2 = c_rand(n_r, size)
 G = c_rand(n_r, n_t)

### 2020-04-29 simulation/src/lin_alg.py
theoretic pdf working and generating results

@@ -1,13 +1,13 @@
 import numpy as np
 
 # capacity = lambda matrix: np.log(np.linalg.det(np.eye(matrix.shape[0]) + matrix)) #   Returns complex residual
-capacity = lambda matrix, snr: np.sum(np.log(1 + snr*np.linalg.eigvalsh(matrix)))
+capacity = lambda matrix, snr: np.sum(np.log2(1 + snr*np.linalg.eigvalsh(matrix)))
 
 
 def aed_capacity(x_value, value_probability, step, snr, number_receivers):
     probabilities = value_probability*step
     check = np.sum(probabilities)
-    return number_receivers*np.sum(np.log(1 + snr*x_value)*probabilities)
+    return number_receivers*np.sum(np.log2(1 + snr*x_value)*probabilities)
 
 
 c_rand = lambda rows, cols: (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)

### 2020-04-29 simulation/testing/channel_testing/direct_scatter.py

@@ -5,27 +5,18 @@ from src.lin_alg import *
 import seaborn
 
 #   Initialize System Parameters
-size = 500
+size = 100
 rows = size
 cols = size
-sigma = 0   # Relative attenuation of IRS portion
+sigma = 1   # Relative attenuation of IRS portion
 rho = 1     # S/R
-phi = 1     # phi=1 is full rank line of sight.  L/T
+phi = 1    # phi=1 is full rank line of sight.  L/T
 #   Plot Theoretical Distribution and find corresponding capacity
 
 
 def arg_function(s_val, g_k, sigma=1, rho=1):
     x_s = np.sqrt(s_val)
-    # return x_s - x_s*g_k/(1-s_val*np.power(g_k, 2))
     return x_s - (pow(sigma, 2)*rho*x_s*g_k)/(rho-pow(sigma, 2)*s_val*np.power(g_k, 2))
-
-
-def final_stieltjes(s_values, g_k, func_params):
-    sigma = func_params[0]
-    rho = func_params[1]
-    x_arg = arg_function(s_values, g_k, sigma=sigma, rho=rho)
-    G_BB = (1/2)*np.sqrt(1-(4/np.power(x_arg, 2)))-1/2
-    return (1/np.sqrt(s_values))*x_arg*G_BB
 
 
 def final_stieltjes_deformed(s_values, g_k, func_params):
@@ -34,7 +25,7 @@ def final_stieltjes_deformed(s_values, g_k, func_params):
     phi = func_params[2]
     x_arg = arg_function(s_values, g_k, sigma=sigma, rho=rho)
     # Correct
-    G_BB = 1/2 + (1-phi)/(2*np.power(x_arg, 2)) +\
+    G_BB = 1/2 + (1-phi)/(2*np.power(x_arg, 2)) -\
            np.sqrt(np.power((1-phi), 2)/(4*np.power(x_arg, 4)) - (1+phi)/(2*np.power(x_arg, 2)) + 1/4)
     # Correct
     # G_BB = np.sqrt(np.power((1-phi), 2)/(4*np.power(x_arg, 4)) - (1+phi)/(2*np.power(x_arg, 2))
@@ -60,8 +51,8 @@ irs_svd = []
 capacities = []
 cross_sum = 0
 main_sum = 0
-los_rank_matrix = reduced_rank(rows, rho)
-average = 1
+los_rank_matrix = reduced_rank(rows, phi)
+average = 3
 for i in range(average):
[…]

### 2020-04-29 simulation/testing/channel_testing/parameter_vary.py

@@ -1,0 +1,110 @@
+from src.circle_laws import *
+import numpy as np
+import matplotlib.pyplot as plt
+from src.lin_alg import *
+import seaborn
+
+#   Initialize System Parameters
+size = 100
+rows = size
+cols = size
+sigma = 1   # Relative attenuation of IRS portion
+rho = 1     # S/R
+phi = 1    # phi=1 is full rank line of sight.  L/T
+#   Plot Theoretical Distribution and find corresponding capacity
+
+
+def arg_function(s_val, g_k, sigma=1, rho=1):
+    x_s = np.sqrt(s_val)
+    return x_s - (pow(sigma, 2)*rho*x_s*g_k)/(rho-pow(sigma, 2)*s_val*np.power(g_k, 2))
+
+
+def final_stieltjes_deformed(s_values, g_k, func_params):
+    sigma = func_params[0]
+    rho = func_params[1]
+    phi = func_params[2]
+    x_arg = arg_function(s_values, g_k, sigma=sigma, rho=rho)
+    # Correct
+    G_BB = 1/2 + (1-phi)/(2*np.power(x_arg, 2)) -\
+           np.sqrt(np.power((1-phi), 2)/(4*np.power(x_arg, 4)) - (1+phi)/(2*np.power(x_arg, 2)) + 1/4)
+    # Correct
+    # G_BB = np.sqrt(np.power((1-phi), 2)/(4*np.power(x_arg, 4)) - (1+phi)/(2*np.power(x_arg, 2))
+    #                + 1/4) - 1/2 - (1-phi)/(2*np.power(x_arg, 2))
+    return (1/np.sqrt(s_values))*x_arg*G_BB
+
+
+
+
+
+x_values, step = np.linspace(1e-2, 10, 100, retstep=True)
+y = 1e-9
+s_vector = x_values + 1j*y
+
+numeric_capacity_list = []
+theoretic_capacity_list = []
+param_list = [0, .25 , .5 , 1]
+
+
+for val in param_list:
+    phi = val
+    # sigma = val
+    parameters = [sigma, rho, phi]
+    theoretic_pdf = estimated_pdf(fixed_point(final_stieltjes_deformed, s_vector, 100, 1e-3, 3,
+                                              func_param=parameters, unique_half_plane=True))
+    # theoretic_pdf = estimated_pdf(marcenko_st(s_vector, phi))
+    theoretic_capacity = aed_capacity(x_values, theoretic_pdf, step, 1/cols, rows)
+
+
[…]

### 2020-05-01 simulation/src/lin_alg.py
Corrected normalization details on numerical resuls for presentation

@@ -11,7 +11,12 @@ def aed_capacity(x_value, value_probability, step, snr, number_receivers):
     return number_receivers*np.sum(np.log2(1 + snr*x_value)*probabilities)
 
 
-c_rand = lambda rows, cols: (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
+def c_rand(rows, cols, var=None):
+    if var is not None:
+        return (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))*np.sqrt(var)
+    else:
+        return (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
+
 c_rand_mean = lambda rows, cols: 1e-1 + 1j*1e-1 + (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows)
 
 c_rand_2 = lambda rows, cols:  (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(cols*rows)

### 2020-05-01 simulation/testing/phase_optimization/phase_check_simulation.py

@@ -5,8 +5,8 @@ from src.lin_alg import *
 import seaborn
 
 size = 100
-n_t = 4
-n_r = 4
+n_t = 100
+n_r = 100
 rows = size
 cols = size
 bins = int(size/2)
@@ -17,10 +17,12 @@ cross_sum = 0
 main_sum = 0
 
 average = 1000
-H_1 = c_rand(size, n_t)
-H_2 = c_rand(n_r, size)
+# H_1 = c_rand(size, n_t)
+H_1 = c_rand(size, n_t, var=1/np.sqrt(n_r*size))
+H_2 = c_rand(n_r, size, var=1/np.sqrt(n_r*size))
+# H_1 = c_rand(size, n_t)
+# H_2 = c_rand(n_r, size)
 G = c_rand(n_r, n_t)
-F = c_rand(rows, cols)
 for i in range(average):
     H = H_2 @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, size))) @ H_1
     # F = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)

### 2020-05-03 simulation/testing/channel_testing/parameter_vary.py
some revisions from feedback on inital presentation slides

@@ -46,8 +46,8 @@ param_list = [0, .25 , .5 , 1]
 
 
 for val in param_list:
-    phi = val
-    # sigma = val
+    # phi = val
+    sigma = val
     parameters = [sigma, rho, phi]
     theoretic_pdf = estimated_pdf(fixed_point(final_stieltjes_deformed, s_vector, 100, 1e-3, 3,
                                               func_param=parameters, unique_half_plane=True))
@@ -95,13 +95,13 @@ for val in param_list:
 
 #   Plot results
 fig, ax = plt.subplots()
-plt.title("Capacity Vs. Line of Sight Rank")
-# plt.title("Capacity Vs. IRS Path Attenuation")
-ax.plot(param_list, theoretic_capacity_list, label='theoretical capacity', c='g')
-ax.plot(param_list, numeric_capacity_list, label='numerical capacity (N=100)', c='r')
+# plt.title("Capacity Vs. Line of Sight Rank")
+plt.title("Capacity Vs. IRS Path Attenuation")
+ax.plot(param_list, theoretic_capacity_list, label='Asymptotic Expression', c='g')
+ax.plot(param_list, numeric_capacity_list, label='Numerical Capacity (N=100)', c='r')
 ax.set_ylabel('Capacity (Bits)')
-# ax.set_xlabel('Relative Attenuation Coefficient')
-ax.set_xlabel('Line of Sight Channel Rank')
+ax.set_xlabel('Relative Attenuation Coefficient')
+# ax.set_xlabel('Line of Sight Channel Rank')
 plt.legend(loc="upper left")
 ax.grid(True, which='both')
 seaborn.despine(ax=ax, offset=0)

### 2020-05-26 documentation/Meeting Notes/5_13_20.tex
working on convergence for gamma in terms of s-transform values

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Current Questions}
\begin{enumerate}
\item 
System Model for Correlation 
	 \begin{equation}\label{system_model}
		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x}.
	\end{equation}
	for 
%	\begin{equation}
%	\mathbf{H}_{Total} =  \underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{C}_{3}\mathbf{G}\mathbf{P}_{3}}_{\text{LOS} }
%	\end{equation}
	\begin{equation}
	\mathbf{H}_{Total} =  \underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{D}}_{\text{LOS} }
	\end{equation}
	with $\mathbf{P}_{2}\in \mathbb{C}_{R \times S}, \mathbf{P}_{1}\in \mathbb{C}_{S \times T}$ 
%	and $\mathTbf{P}_{3}\in \mathbb{C}_{R \times T} $ 
	as rectangular projection matrices such that 
	$\mathbf{H}_{1}\in \mathbb{C}_{S \times S}, \mathbf{H}_{2} \in \mathbb{C}_{R \times R}$ 
%	and $  \mathbf{G} \in \mathbb{C}_{R \times R}$
	 are square matrices.
	 A deterministic Line of Sight portion of the channel is represented by the matrix $\mathbf{D}$.
	Correlations between antennas/surfaces are represented using deterministic, \emph{hermetian?} matrices
	 $\mathbf{C}_{2} $ and $\mathbf{C}_{1} $. 
	 Finally $\boldsymbol{\Phi}$ as a diagonal matrix with elements $e^{j\phi_i}$ representing the IRS.

So far, I have considered two, different paths to obtain the AED of $\Expect{[\mathbf{H}_{Total} \mathbf{H}_{Total}^H}]$. 
\begin{enumerate}
\item 
The first option is to use the fixed point subordination equation
\begin{equation}
\tilde{G}_{\text{Total}}(s) = \tilde{G}_{\text{LOS}}(s - \tilde{R}_{\text{IRS}}( - \tilde{G}_{\text{Total}}(s))
\end{equation} (Note that for this method the IRS component and LOS component can be swapped).
Because this equation is in fixed point, both $\tilde{G}_{\text{LOS}}(s)$ and $\tilde{R}_{\text{IRS}}(w)$ need to be given in close form, otherwise this is an infinite recursion. 
\item 
The second option is to find $\tilde{R}_{\text{Total}}(w) = \tilde{R}_{\text{IRS}}(w) + \tilde{R}_{\text{LOS}}(w)$, then use
$R(w) = G^{-1}(-w) - \frac{1}{w}$ and $\tilde{G}_{\mathbf{H}_{Total}}(s) = sG_{\mathbf{H}_{Total}\mathbf{H}_{Total}^H}(s^2)$ to find the true Stieltjes transform, $G_{\mathbf{H}_{Total}\mathbf{H}_{Total}^H}(s)$. In general, however, given $R(w)$ the corresponding steiltjes transform, $G(w)$ will be given in the form of a fixed-point equation. This means we must find both $\tilde{R}_{\text{IRS}}(w)$ and $\tilde{R}_{\text{LOS}}(w)$ in closed form.
In general, however, the S-Tranform is needed in order to find $\tilde{R}_{\text{IRS}}(w)$
This gives the same problem as case 1 because t
\end{enumerate}

 \item 
Incorporate correlation it the channel using 
	\begin{equation}
	\mathbf{H}_{1} =  \prod_{i=1}^{M} \mathbf{H}_{1,i}
	\end{equation}
 with $\mathbf{H}_{1}\in \mathbb{C}_{S \times S} $ with elements of variance $\frac{1}{S}$.

 \item
 Is is reasonable to handing low rank LOS matrices using a correlation matrix with all $\frac{1}{N}$

 \item 
Can handle by attenuation in general by jusing using the scaling laws of S-Transform

 \item
 Incorporating a ricean component to the LOS channel. If we were just considering the LOS portion, this could be
 easily handled because whatever mean we add, this will always just result in a single large eigenvalue corresponding to the normalized all ones vector. From simulation it does not seem like adding this mean affects the cross terms in the IRS path case as long as the IRS is zero mean (any non-zero mean components should be considered LOS).
 So if the result is essentially the sum of the two aeds then we can just add in the large eigenvalue for the LOS after but mathematically this is not allowed. One other important question is if I can model the reduction in rank by simply adding another correlation matrix since this is the method through which the rank would be reduced. In fact I think that the correlation method is the right way to do it because it properly represents that fact that there not the diversity but there is still increased power from multiple channels here.

\item 
To begin, I want to consider whether or not the elements in 
\begin{equation}
\underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{D}}_{\text{LOS} }
\end{equation}
are free. In order to use the singular value value/eigenvalue relationship.
Slide 76 of the lecture slides seems to show this but I need verification.

\item
In order to find full aed, I will need to consider the stieltjes transform of a deterministic matrix, $\mathbf{D}\mathbf{D}^H$. 
First I want to check the that current fixed point/capacity calculations work for this setup.
Finding the stieltjes transform and then inverting this seems to work although there is a slight issue with normalization of the pdf.
\par
Next I want to consider the case in which I first the S-Transform and then invert to find the AED. Being able to do this is important since I will eventually need to do this in order to find the aed of a matrix such as
$mathbf{C}\mathbf{H}$ in which $mathbf{C}$ is deterministic but $\mathbf{H}$ is random. 
This includes the following steps
\begin{itemize}
\item 
\begin{equation}
G_{\mathbf{C}\mathbf{C}^H}(s) = \frac{-1}{s}(1+\gamma(\frac{1}{s}))
\end{equation}
\item
Solve for $\gamma(z)$ using fixed point.
\begin{equation}
\gamma^{-1}(\gamma(z)) = z = \frac{\gamma(z)}{1+\gamma(z)} S(\gamma(z))
\end{equation}
\item
For each value of $S(\gamma(z))$ in the above iteration, solve
for 
\begin{equation}
S(z) = \frac{1}{1+z} \gamma^{-1}(z)
\end{equation}
\item 
Solve for $\gamma^{-1}(z)$ using the fixed point equation.???
\begin{equation}
\gamma(\gamma^{-1}(z)) = z = -1 - \frac{G_{\mathbf{C}\mathbf{C}^H}(\frac{1}{\gamma^{-1}(z)})}{\gamma^{-1}(z)}
\end{equation}
\end{itemize}
\end{enumerate}
At the moment, however, this is not converging for me. 
\bibliography{bibliography}
\end{document}

### 2020-05-26 simulation/src/fixed_point.py

@@ -10,7 +10,7 @@ def decontamination(G_k, s, K):
 
 
 def fixed_point(fixed_point_function, s_vector, iterations,
-                tolerance=1e-3, attempt_tol=5, func_param=None, unique_half_plane=False):
+                tolerance=1e-1, attempt_tol=1, func_param=None, unique_half_plane=False):
     """
     NOTE that I do not do perform the function concurrently. This allows for better
     tracking of the fixed-point convergence.
@@ -47,16 +47,16 @@ def fixed_point(fixed_point_function, s_vector, iterations,
                     val = fixed_point_function(s, G_k)
                 if unique_half_plane and np.imag(val) < 0:
                     val = np.conj(val)
-                if np.isinf(val) or np.isinf(val):
+                if np.isinf(val) or np.isinf(val) or np.isclose(0, val):
                     print("nan/inf")
                 epsilon.append(np.abs(G_k - val))
                 G_k = val
                 check.append(val)
+                ret_vec[ind] = G_k
                 if epsilon[-1] <= tolerance:
-                    ret_vec[ind] = G_k
                     epsilons.append(epsilon[-1])
                     tol_met = True
-                    break
+                    # break
             check1 = np.asarray(check)
             attempt += 1
         if tol_met:

### 2020-05-26 simulation/testing/Correlation_tests/correlation_fixed_point.py

@@ -5,26 +5,34 @@
 # is used then the correlated channels seem to have increasing capacity as correlation increases
 
 import matplotlib.pyplot as plt
-from src.fixed_point import *
 from src.lin_alg import *
-from src.circle_laws import estimated_pdf
+from src.stieltjes_eqn import *
 import seaborn
 from scipy.linalg import toeplitz
+from src.circle_laws import *
 
 rows = 100
 cols = 100
 H_cor = c_rand(rows, cols)
 block = toeplitz(np.array((1, .5)))
 correlation_matrix = block_matrix(block, H_cor.shape[0], normalize=True)
+correlation_matrix = np.sqrt(correlation_matrix)
 e_val_correlation = np.linalg.eigvalsh(correlation_matrix)
-H_cor = correlation_matrix@H_cor
+# H_cor = correlation_matrix@H_cor
 HH = H_cor@hermetian(H_cor)
 e_val = np.linalg.eigvalsh(HH)
 AED, bins = np.histogram(np.asarray(e_val), bins=50)
-x_values = np.linspace(1e-6, 5, 100)
+x_values, step = np.linspace(.01, 10, 100, retstep=True)
 s_values = x_values + 1j*1e-6
 
-def gamma_c_fixed(s, gamma_s):
+def rayleigh_steiltjes(z, gamma_inv_z):
+    total = - (1/2)*np.sqrt((1-4*gamma_inv_z)) - 1/2
+    total /= (1+z)
+    if np.isnan(total) or np.isinf(total):
+        print("nan")
+    return total
+
+def gamma_c_fixed(z, gamma_inv_z):
     """
     Using an out of scope variable here because I don't want to pass it though another fixed point equation.
     :param s:
@@ -34,40 +42,44 @@ def gamma_c_fixed(s, gamma_s):
     num = e_val_correlation.size
     total = 0
     for eigen_val in e_val_correlation:
-        total += (1/num)*(1/(eigen_val-1/gamma_s))
-    total /= (-(1+s))
+        total -= (1/num)*(1/(eigen_val-1/gamma_inv_z))
+    total /= (1+z)
     if np.isnan(total) or np.isinf(total):
         print("nan")
     return total
 
-def S_ch(gamma_s):
-    check = fixed_point(gamma_c_fixed, gamma_s, 100, tolerance=1e-3)*(1/gamma_s)
[…]

### 2020-05-26 simulation/testing/channel_testing/direct_scatter.py

@@ -1,4 +1,5 @@
 from src.circle_laws import *
+from src.stieltjes_eqn import *
 import numpy as np
 import matplotlib.pyplot as plt
 from src.lin_alg import *
@@ -41,7 +42,7 @@ parameters = [sigma, rho, phi]
 theoretic_pdf = estimated_pdf(fixed_point(final_stieltjes_deformed, s_vector, 100, 1e-3, 3,
                                           func_param=parameters, unique_half_plane=True))
 # theoretic_pdf = estimated_pdf(marcenko_st(s_vector, phi))
-theoretic_capacity = aed_capacity(x_values, theoretic_pdf, step, 1/cols, rows)
+theoretic_capacity = aed_capacity(x_values, theoretic_pdf, 1/cols, rows, step=step)
 
 
 #   Now check histogram of simulations

### 2020-05-26 simulation/testing/channel_testing/single_irs.py

@@ -5,22 +5,23 @@ This test is to check the fixed point expression for the concatenated channel fr
 import matplotlib.pyplot as plt
 from src.circle_laws import *
 from src.fixed_point import *
+from src.stieltjes_eqn import *
 from src.lin_alg import *
 import seaborn
 
-size = 500
-rows = 500
-cols = 500
+size = 100
+rows = 100
+cols = 100
 
 def gamma(z, gamma_z):
-    return np.power(gamma_z/z, 1/3) - 1
+    return np.power(gamma_z/z, 1/4) - 1
 
 y = 1e-6
-x, step = np.linspace(0.01, 10, 1000, retstep=True)
+x, step = np.linspace(0.01, 10, 100, retstep=True)
 s_vector = x + 1j*y
 steiltjes_values = (1/s_vector)*(1+fixed_point(gamma, 1/s_vector, 100, 1e-2))
 aed = estimated_pdf(steiltjes_values)
-theoretic_capacity = aed_capacity(x, aed, step, 1/cols, rows)
+theoretic_capacity = aed_capacity(x, aed, 1/cols, rows, step=step)
 
 
 bins = int(size/2)
@@ -30,7 +31,7 @@ irs = []
 numeric_cacacity = []
 average = 10
 for i in range(average):
-    H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
+    H = c_rand(rows, cols) @ c_rand(rows, cols)  @ c_rand(rows, cols)
     HH = H @ np.conj(H.T)
     total_irs = HH
     e_val_total_irs, e_vec_total_irs = np.linalg.eig(total_irs)
@@ -45,8 +46,10 @@ print(f"numeric_capacity: {np.average(numeric_cacacity)}")
 irs_AED, bins_irs = np.histogram(np.asarray(irs), bins=bins)
 fig, ax = plt.subplots()
 plt.title("IRS Channels Components: i.i.d, $\mathbb{N}(0,1/N), H = H_1 \Phi H_2$")
-ax.plot(bins_irs[1::], irs_AED/(rows*average), label='total: with IRS')
-ax.plot(x, aed/np.linalg.norm(aed), label='Theoretic AED')
+ax.plot(bins_irs[:-1], irs_AED/(rows*average), label='total: with IRS')
[…]

### 2020-05-26 simulation/testing/theory_testing/subordination_test.py

@@ -6,32 +6,6 @@ from src.fixed_point import *
 from scipy.linalg import toeplitz
 
 
-def deterministic_stieltjes(s_vals):
-    num = e_val_correlation.size
-    if type(s_vals) is not np.ndarray:
-        total = 1j * 0
-    else:
-        total = 1j*np.zeros(s_vals.shape)
-    for e_val in e_val_correlation:
-        total += 1/((e_val-s_vals)*num)
-    return total
-
-def inv_gamma_fixed(z, inv_gamma_z):
-    return - deterministic_stieltjes(1/inv_gamma_z)/(z+1)
-
-def inverse_gamma(z_vector):
-    check = fixed_point(inv_gamma_fixed, z_vector, 100, unique_half_plane=True)
-    return check
-
-def gamma_fixed(z, gamma_z):
-    check = (gamma_z + 1)*z/s_transform_inverse_gamma(gamma_z, inverse_gamma)
-    # check = gamma_z*s_transform_inverse_gamma(z, inverse_gamma) - z
-    return check
-
-def gamma(z):
-    check = fixed_point(gamma_fixed, z, 100)
-    return check
-
 
 #   First Generate a deterministic matric
 cols = 100
@@ -44,6 +18,8 @@ e_val_correlation = np.linalg.eigvalsh(correlation_matrix)
 x_vals, step = np.linspace(1e-6, 3, 100, retstep=True)
 # x_vals = np.unique(e_val_correlation) #    Need to get rid of duplicates here
 s_vals = x_vals + 1j*1e-6
+pdf = normalize_pdf(estimated_pdf(aed_from_svd(s_vals, g_tilde))) #   Is Normalizing this reasonable?
+
 check = estimated_pdf(deterministic_stieltjes(s_vals))
 # pdf = estimated_pdf(deterministic_stieltjes(s_vals)) #   Is Normalizing this reasonable?
 pdf = normalize_pdf(estimated_pdf(deterministic_stieltjes(s_vals))) #   Is Normalizing this reasonable?

### 2020-05-29 documentation/Meeting Notes/05_29_20.tex
added exponential model for correlation

--- 2020-05-26 documentation/Meeting Notes/5_13_20.tex
@@
 
-\section{Current Questions}
-\begin{enumerate}
-\item 
-System Model for Correlation 
-	 \begin{equation}\label{system_model}
-		\mathbf{r} = \mathbf{H}_{Total}\mathbf{x}.
-	\end{equation}
-	for 
-%	\begin{equation}
-%	\mathbf{H}_{Total} =  \underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{C}_{3}\mathbf{G}\mathbf{P}_{3}}_{\text{LOS} }
-%	\end{equation}
-	\begin{equation}
-	\mathbf{H}_{Total} =  \underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{D}}_{\text{LOS} }
-	\end{equation}
-	with $\mathbf{P}_{2}\in \mathbb{C}_{R \times S}, \mathbf{P}_{1}\in \mathbb{C}_{S \times T}$ 
-%	and $\mathTbf{P}_{3}\in \mathbb{C}_{R \times T} $ 
-	as rectangular projection matrices such that 
-	$\mathbf{H}_{1}\in \mathbb{C}_{S \times S}, \mathbf{H}_{2} \in \mathbb{C}_{R \times R}$ 
-%	and $  \mathbf{G} \in \mathbb{C}_{R \times R}$
-	 are square matrices.
-	 A deterministic Line of Sight portion of the channel is represented by the matrix $\mathbf{D}$.
-	Correlations between antennas/surfaces are represented using deterministic, \emph{hermetian?} matrices
-	 $\mathbf{C}_{2} $ and $\mathbf{C}_{1} $. 
-	 Finally $\boldsymbol{\Phi}$ as a diagonal matrix with elements $e^{j\phi_i}$ representing the IRS.
-
-So far, I have considered two, different paths to obtain the AED of $\Expect{[\mathbf{H}_{Total} \mathbf{H}_{Total}^H}]$. 
-\begin{enumerate}
-\item 
-The first option is to use the fixed point subordination equation
-\begin{equation}
-\tilde{G}_{\text{Total}}(s) = \tilde{G}_{\text{LOS}}(s - \tilde{R}_{\text{IRS}}( - \tilde{G}_{\text{Total}}(s))
-\end{equation} (Note that for this method the IRS component and LOS component can be swapped).
-Because this equation is in fixed point, both $\tilde{G}_{\text{LOS}}(s)$ and $\tilde{R}_{\text{IRS}}(w)$ need to be given in close form, otherwise this is an infinite recursion. 
-\item 
-The second option is to find $\tilde{R}_{\text{Total}}(w) = \tilde{R}_{\text{IRS}}(w) + \tilde{R}_{\text{LOS}}(w)$, then use
-$R(w) = G^{-1}(-w) - \frac{1}{w}$ and $\tilde{G}_{\mathbf{H}_{Total}}(s) = sG_{\mathbf{H}_{Total}\mathbf{H}_{Total}^H}(s^2)$ to find the true Stieltjes transform, $G_{\mathbf{H}_{Total}\mathbf{H}_{Total}^H}(s)$.
-In general, however, given $R(w)$ the corresponding steiltjes transform, $G(w)$ will be given in the form of a fixed-point equation.
-This means we must find both $\tilde{R}_{\text{IRS}}(w)$ and $\tilde{R}_{\text{LOS}}(w)$ in closed form.
-In general, however, the S-Tranform is needed in order to find $\tilde{R}_{\text{IRS}}(w)$
-This gives the same problem as case 1 because t
-\end{enumerate}
-
- \item 
-Incorporate correlation it the channel using 
-	\begin{equation}
-	\mathbf{H}_{1} =  \prod_{i=1}^{M} \mathbf{H}_{1,i}
-	\end{equation}
- with $\mathbf{H}_{1}\in \mathbb{C}_{S \times S} $ with elements of variance $\frac{1}{S}$.
-
- \item
- Is is reasonable to handing low rank LOS matrices using a correlation matrix with all $\frac{1}{N}$
-
- \item 
-Can handle by attenuation in general by jusing using the scaling laws of S-Transform
-
- \item
- Incorporating a ricean component to the LOS channel. If we were just considering the LOS portion, this could be
- easily handled because whatever mean we add, this will always just result in a single large eigenvalue corresponding to the normalized all ones vector.
-From simulation it does not seem like adding this mean affects the cross terms in the IRS path case as long as the IRS is zero mean (any non-zero mean components should be considered LOS).
- So if the result is essentially the sum of the two aeds then we can just add in the large eigenvalue for the LOS after but mathematically this is not allowed.
-One other important question is if I can model the reduction in rank by simply adding another correlation matrix since this is the method through which the rank would be reduced.
-In fact I think that the correlation method is the right way to do it because it properly represents that fact that there not the diversity but there is still increased power from multiple channels here.
-
-\item 
-To begin, I want to consider whether or not the elements in 
-\begin{equation}
-\underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{D}}_{\text{LOS} }
-\end{equation}
-are free. In order to use the singular value value/eigenvalue relationship.
-Slide 76 of the lecture slides seems to show this but I need verification.
-
-\item
-In order to find full aed, I will need to consider the stieltjes transform of a deterministic matrix, $\mathbf{D}\mathbf{D}^H$. 
-First I want to check the that current fixed point/capacity calculations work for this setup.
-Finding the stieltjes transform and then inverting this seems to work although there is a slight issue with normalization of the pdf.
-\par
-Next I want to consider the case in which I first the S-Transform and then invert to find the AED.
-Being able to do this is important since I will eventually need to do this in order to find the aed of a matrix such as
-$mathbf{C}\mathbf{H}$ in which $mathbf{C}$ is deterministic but $\mathbf{H}$ is random. 
-This includes the following steps
-\begin{itemize}
-\item 
-\begin{equation}
-G_{\mathbf{C}\mathbf{C}^H}(s) = \frac{-1}{s}(1+\gamma(\frac{1}{s}))
-\end{equation}
-\item
-Solve for $\gamma(z)$ using fixed point.
-\begin{equation}
-\gamma^{-1}(\gamma(z)) = z = \frac{\gamma(z)}{1+\gamma(z)} S(\gamma(z))
-\end{equation}
-\item
-For each value of $S(\gamma(z))$ in the above iteration, solve
-for 
-\begin{equation}
-S(z) = \frac{1}{1+z} \gamma^{-1}(z)
-\end{equation}
-\item 
-Solve for $\gamma^{-1}(z)$ using the fixed point equation.???
-\begin{equation}
-\gamma(\gamma^{-1}(z)) = z = -1 - \frac{G_{\mathbf{C}\mathbf{C}^H}(\frac{1}{\gamma^{-1}(z)})}{\gamma^{-1}(z)}
-\end{equation}
-\end{itemize}
-\end{enumerate}
-At the moment, however, this is not converging for me. 
-\bibliography{bibliography}
+\section{}
+\
 \end{document}

### 2020-05-29 documentation/Meeting Notes/5_28_20.tex
added exponential model for correlation

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\subsection{Questions}
\begin{itemize}
\item 
For the channel model given by 
	\begin{equation}
	\mathbf{H}_{Total} =  \underbrace{\mathbf{C}_{2}\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{C}_{1}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{D}}_{\text{LOS} }
	\end{equation}
should the power added by the correlation matrix be normalized such that rows of matrix $\mathbf{C}_{i}$ have norm 1?
This ensure that the elements would still have the same variance despite being correlated. 
\item 
I notice that for the correlated channels (normalized) that if I perform water-filling the power allocation, the capacity is greater than for 
the iid channel. Is this correct? This result seems strange.

\item 
In order to test the method discussed during our last meeting, I am using the following method to find the AED of the channel given by 
	\begin{equation}
	\mathbf{H}_{Total} = \mathbf{C}\mathbf{H}
	\end{equation}
	with C as a normalized correlation matrix with "blocks of correlation" and H is an iid gaussian matrix. 
	The fsolve function in matlab does not converge unless I use a closed form of the s-transform in which case the expected aed is found. This may be a result of having a distribution with delta masses*. If $\mathbf{C}$ is substituted with another gaussian iid matrix $\mathbf{H}_2 $, and I use the quarter circle stieltjes this method also does not work. This makes me think that something is wrong with equation \eqref{S_Stieltjes} below.
\begin{enumerate}
\item 
	\begin{equation}
	p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{lim} \; Img(G(x+jy))
	\end{equation}

\item 
	\begin{equation}
	G(s)=  \frac{1}{s} (1+\gamma\left(\frac{1}{s}\right))
	\end{equation}

\item 
	\begin{equation}
	\gamma(z) =  \frac{z(1+\gamma(z))}{S(\gamma(z))}
	\end{equation}
\item 
	\begin{equation}
	S(z) = S_{\text{correlation}}(z) S_{\text{rayleigh}}(z)
	\end{equation}

	\item 
	\begin{equation}\label{S_Stieltjes}
	S_{\text{correlation}}(z)  = \frac{z+1}{z} \gamma^{-1}(z)
	\end{equation}

\item 
	\begin{equation}
	\gamma^{-1}(z) = \frac{-G(\frac{1}{\gamma^{-1}(z))}}{z+1})
	\end{equation}
	In this case I am using the known AED of the correlation matrix to find $G_{\text{correlation}}(s)$ using
	\begin{equation}
	G(s) = \int \frac{1}{x-s} dPx
	\end{equation}

\end{enumerate}
\end{itemize}

\end{document}

### 2020-05-29 simulation/MATLAB/gamma_s_func.m

@@ -1,7 +1,0 @@
-function x = gamma_s_func(eval_point, init_point)
-x = fsolve(@root2d,init_point);
-      function F = root2d(gamma_z)
-%             F = gamma_z - eval_point.*power(1+gamma_z,2);
-            F = gamma_z - eval_point.*(1+gamma_z)./S_func(gamma_z);
-      end
-end

### 2020-05-29 simulation/MATLAB/useful_functions/normr_la.m

@@ -1,0 +1,7 @@
+function ret_mat = normr_la(matrix)
+    ret_mat = matrix;
+    rows = size(matrix,2);
+    for i = 1 : rows
+    	ret_mat(i,:) = matrix(i,:)./norm(matrix(i,:));
+    end
+end

### 2020-05-29 simulation/src/lin_alg.py

@@ -39,15 +39,25 @@ def block_matrix(block, size, normalize=False):
         return normalize_matrix(mat, rows=True)
 
 
+def exponential_correlation(size, r):
+    ret_mat = np.ones((size, size))
+    for row in range(ret_mat.shape[0]):
+        for col in range(ret_mat.shape[1]):
+            ret_mat[row, col] *= pow(r, abs(col-row))
+    for row in range(ret_mat.shape[0]):
+        ret_mat[row, :] /= np.linalg.norm(ret_mat[row, :], 2)
+    return ret_mat
+
+
 def normalize_matrix(matrix, rows=False):
     ret_mat = matrix
     if not rows:
         for col in range(matrix.shape[1]):
-            matrix[:, col]/np.linalg.norm(matrix[:, col])
+            ret_mat[:, col]/np.linalg.norm(matrix[:, col], 2)
         return ret_mat
     else:
         for row in range(matrix.shape[0]):
-            matrix[row, :] /= np.linalg.norm(matrix[row, :])
+            ret_mat[row, :] /= np.linalg.norm(matrix[row, :], 2)
         return ret_mat
 
 

### 2020-05-29 simulation/src/network_simulation_tools.py

@@ -111,4 +111,5 @@ def water_filling(channel_matrix, power_constraint, sigma_square=1):
     prob = cp.Problem(cp.Maximize(cp.sum(utility)), constraint)
     prob.solve()
     covariance_matrix = hermetian(V_H)@np.diag(variables.value)@V_H
+    check = np.trace(covariance_matrix)
     return channel_matrix@covariance_matrix@hermetian(channel_matrix)

### 2020-06-01 documentation/Meeting Notes/05_29_20.tex
added work for update

@@
 
-\section{}
-\
+\section{Channel Model}
+In this update I detail the channel model I am using as well as the current issue I am having in finding the resulting channel AED. 
+\par
+This IRS channel considered includes correlation at the transmitter, receiver and IRS, as well as a LOS component with a non-zero mean.
+First, I describe the general channel.
+Then I will begin with a simple channel in order to investigate the calculation of the AED as each component (correlation, IRS, LOS) is introduced.
+
+\begin{enumerate}
+\item 
+The general channel is given by 
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}(\underbrace{\mathbf{H}_{2}\mathbf{P}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}\mathbf{H}_{1}\mathbf{P}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}
+	\end{equation}
+	with $\mathbf{R}_{R}$, $\mathbf{R}_{T}$ and $\mathbf{R}_{S}$ representing deterministic correlation matrices at the receiver, transmitter and IRS, $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ have i.i.d zero mean Gaussian elements, $\mathbf{G}$ has \emph{non-zero} mean i.i.d Gaussian elements and $\boldsymbol{\Phi}$ representing the IRS phases.
+	This simplify notation, at this stage I will consider all matrices to have dimension $N \times N$.
+
+\item 
+In order to evaluate the impact of correlation on the system capacity, the power of the correlation matrices , $\mathbf{R}_{R}$, $\mathbf{R}_{T}$ and $\mathbf{R}_{S}$ must be normalized to prevent adding power to the received signal.
+If $\mathbf{H}$ has elements with variance $\frac{1}{N}$ then the channel $\mathbf{G} = \mathbf{R}\mathbf{H}$ with correlation $\mathbf{R}$ must also have elements with variance $\frac{1}{N}$.
+Elements of $\mathbf{G}$ are given by  $g_{i,j} = \bar{\mathbf{r}}^T_i \mathbf{h}_j$ in which $\bar{\mathbf{r}}^T_i$ is the $i^{\text{th}}$ row of $\mathbf{R}$.
+ Normalizing these \emph{rows} such that $\| \bar{\mathbf{r}}_i \| ^2_2 = 1$ ensures that $E[\|g_{i,j}\|^2] = 
+ E[\text{trace}(\mathbf{h}_j^T \bar{\mathbf{r}}_i \bar{\mathbf{r}}^T_i \mathbf{h}_j)] 
+ =  \frac{1}{N}E[\text{trace}(\bar{\mathbf{r}}_i \bar{\mathbf{r}}^T_i)]= \frac{1}{N}$.
+ For the case in which the correlation is a pre-multiplication, $\mathbf{G} = \mathbf{H}\mathbf{R}$, the \emph{columns} of 
+ $\mathbf{R}$ should be normalized to such that  $\|\mathbf{r}_j\| ^2_2 = 1$
+
+\item 
+	Before considering the general channel,  I will go through the calculation of the AED for an 
+	asymptotically large  MIMO channel with correlation given by
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{G}
+	\end{equation}
+	in which $\mathbf{R}_{R}$ is some row-normalized, deterministic matrix describing the correlation and
+	elements $\mathbf{g}_{i,j}$ are zero mean iid.
+	The following steps and equations are used to find the AED of $\mathbf{C} = \mathbf{H}_{Total} \mathbf{H}_{Total}^H$.
+	\begin{enumerate}
+	\item 
+		Find AED from the stieltjes transform.
+		\begin{equation}
+		p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
+		\end{equation}
+	\item 
+		Find Stieltjes transform from $\gamma\left(\right)$
+		\begin{equation}
+		G(s)=  \frac{1}{s} (1+\gamma\left(\frac{1}{s}\right))
+		\end{equation}
+	\item 
+		To find an expression for $\gamma(z)$, evaluate the equation
+		\begin{equation}
+		\gamma^{-1}(z) =  \frac{z}{1+z} S(z)
+		\end{equation}
+		at $z= \gamma(z)$
+		to get the fixed point equation
+		\begin{equation}
+		\gamma(z) = \frac{z(1+ \gamma(z))}{S(\gamma(z))}
+		\end{equation}
+		From this point I can find the AED so long as I can evaluate $S(z)$ is in closed form. 
+		In general, however, this is not the case (e.g when $S(z)$ includes a deterministic correlation matrix). 
+	\item 
+		To find an expression for evaluating $S(z)$ with a correlated channel, use the property
+		\begin{equation}\label{rotation_property}
+		S_{\mathbf{C}}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{\mathbf{C}}}\left(\frac{z}{\chi_2}\right).
+		\end{equation}
+		with the rotated channel $\tilde{\mathbf{C}} = \mathbf{R}_{R}^H\mathbf{R}_{R}\mathbf{G}\mathbf{G}^H$
+		to get
+		\begin{equation}
+		S_{\tilde{\mathbf{C}}}\left( z \right) = S_ {\mathbf{R}_{Rx}^H\mathbf{R}_{Rx}}(z)
+		\underbrace{S_ {\mathbf{G}\mathbf{G}^H}(z)}_{\text{closed form}}.
+		\end{equation}
+	\item 
+		To find the S-Transform of the correlation matrix, $S_ {\mathbf{R}_{R}^H\mathbf{R}_{R}}(z)$,
+		use the property 
+		\begin{equation}
+		 S(z) = \gamma^{-1}(z)\frac{1+z}{z}.
+		\end{equation}
+	\item
+		Now find an expression for $\gamma^{-1}(z)$ using the known Steiltjes transform of the deterministic
+		 correlation matrix. 
+		 \par
+		 If I use the procedure shown in the following, the value of $S(z)$ is 
+		generally not correct when compared to known S-Transforms (e.g. when I try this procedure for the 
+		quarter circle and compare it to the known S-Transform, it is not correct).
+	\item
+		Evaluate the the equation 
+		\begin{equation}
+		G(s)=  -\frac{1}{s} (1+\gamma\left(\frac{1}{s}\right))
+		\end{equation}
+		at $s = \gamma^{-1}(z)$ 
+		to find the fixed point equation
+		 \begin{equation}
+		\gamma^{-1}(z) = \frac{-G(\frac{1}{\gamma^{-1}(z)})}{z+1}
+		\end{equation}
+		* When I try this in MATLAB, I get a the error that fsolve does not converge
+		(possibly because $\gamma^{-1}(z))$ could have discontinuities?).
+	\item 
+		For the case of deterministic matrices 
+		\begin{equation}
+		G(s) = \int \frac{1}{x-s} dPx
+		\end{equation}
+		becomes 
+		\begin{equation}
+		G(s) = \sum_{i=1}^{N} \frac{1}{N (\lambda_i - s)}.
+		\end{equation}	
+	\item
+		Until I can handle this case I cannot move to a more general channel model. As I haven't been able to solve
+		 the above issue, I went onto develop the model so that once the above issue is resolved the next step
+		 will be in place. 
+	\end{enumerate}
+
+\item
+The next step the correlated LOS component with non-zero mean is added. 
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{G}
+	\end{equation}
+	with 
+	\begin{equation}
+	\mathbf{G} = \underbrace{\mathbf{H}}_{\mathcal{NC}(0,\sigma)} + \underbrace{\alpha\mathbf{I}}_{\text{scaled identity}}
+	\end{equation}
+	\begin{enumerate}
+	\item
+	Using the property 
+		\begin{equation}\label{svd_aed_property}
+		G_{\lambda\lambda}(s) = \frac{1}{\sqrt{s}}\tilde{G}_{\lambda}(\sqrt{s})
+		\end{equation}
+	avoids having to work with non-hermetian and non-free terms in the covariance polynomial $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$. 
+	\item 
+		Next obtain the symmetrized singular value from the corresponding R-Transform using
+		\begin{equation}
+		\tilde{G}_{\lambda}(\sqrt{s}) = \frac{1}{\tilde{R}(-\tilde{G}_{\lambda}(\sqrt{s})) - s}
+		\end{equation}
+		and 
+			\begin{equation}
+			\tilde{R}(w) = \tilde{R}_{\mathbf{R}\mathbf{H}}(w) + \tilde{R}_{\alpha\mathbf{R}\mathbf{I}}(w)
+			\end{equation}
+	\item 
+		In order to find the 
+	\end{enumerate}	
+\end{enumerate}
 \end{document}

### 2020-06-01 simulation/src/fixed_point.py

@@ -37,10 +37,7 @@ def fixed_point(fixed_point_function, s_vector, iterations,
         # non_negative = True
         # Use different initialization points if not converging fast enough
         while not tol_met and attempt < attempt_tol:
-            # Fixed point is unique in the upper half complex plane so initialize in this region
             G_k = np.random.uniform(0, 1) + 1j*np.random.uniform(0, 1)
-            G_k = G_k / np.linalg.norm(G_k) # Check to see if this is needed for convergenece
-            # G_k = .1
             for step in range(iterations):
                 if func_param is not None:
                     val = fixed_point_function(s, G_k, func_param)
@@ -50,11 +47,12 @@ def fixed_point(fixed_point_function, s_vector, iterations,
                     val = np.conj(val)
                 if np.isinf(val) or np.isinf(val) or np.isclose(0, val):
                     print("nan/inf")
+                check = np.abs(G_k - val)
                 epsilon.append(np.abs(G_k - val))
                 G_k = val
-                check.append(val)
-                ret_vec[ind] = G_k
                 if epsilon[-1] <= tolerance:
+                    check.append(val)
+                    ret_vec[ind] = G_k
                     epsilons.append(epsilon[-1])
                     tol_met = True
                     # break

### 2020-06-01 simulation/src/stieltjes_eqn.py

@@ -10,6 +10,9 @@ def estimated_pdf(x):
     """
     return 1/np.pi * np.imag(x)
 
+# Move from an R-Transform function to the Stieltjes domain
+def steiltjes_from_r_transform(s_vector, R_Transform, itr=200):
+    return fixed_point(R_Transform, s_vector, itr, 1e-3)
 
 # Move from a gamma function to the Stieltjes domain
 def steiltjes_from_gamma(s_vector, gamma_function, itr=200):

### 2020-06-01 simulation/testing/Correlation_tests/correlation_fixed_point.py

@@ -62,11 +62,11 @@ def S_ch(z):
     return check
 
 def gammafixed(z, gamma_z):
-    check = S_ch(gamma_z)
-    if np.isinf(check) or np.isnan(check):
-        print("inf/nan")
+    # check = S_ch(gamma_z)
+    # if np.isinf(check) or np.isnan(check):
+    #     print("inf/nan")
     return np.power(gamma_z/z, 1 / 2) - 1
-    # return np.power(gamma_z / z, 1 / 4) - 1
+    # return np.power(gamma_z / z, 1 / 3) - 1
     # return (check*gamma_z)/z-1
     # return z*(1+gamma_z)/check
 

### 2020-06-01 simulation/testing/Correlation_tests/correlation_setup.py

@@ -59,14 +59,14 @@ for i in range(average):
     irs_trace.append(np.trace(HH))
 
 
-    # G_cap.append(capacity(GG, 1/size))
-    # irs_cap.append(capacity(HH, 1/size))
-    # cor_cap.append(capacity(HH_cor, 1/size))
+    G_cap.append(capacity(GG, 1/size))
+    irs_cap.append(capacity(HH, 1/size))
+    cor_cap.append(capacity(HH_cor, 1/size))
 
-
-    G_cap.append(capacity_water_filled(water_filling(G, 1)))
-    irs_cap.append(capacity_water_filled(water_filling(H, 1)))
-    cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
+    #
+    # G_cap.append(capacity_water_filled(water_filling(G, 1)))
+    # irs_cap.append(capacity_water_filled(water_filling(H, 1)))
+    # cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
 
 irs_AED, bins_irs = np.histogram(np.asarray(irs), bins=bins)
 AED_cor, bins_cor = np.histogram(np.asarray(irs_cor), bins=bins)

### 2020-06-02 documentation/Meeting Notes/05_29_20.tex
working on update

--- 2020-05-29 documentation/Meeting Notes/05_29_20.tex
@@
 
-\section{}
-\
+\section{Channel Model}\label{section:channel}
+In this update I detail the channel model I am using as well as the current issue I am having in finding the resulting channel AED. 
+\par
+The channel considered includes correlation at the transmitter, receiver and IRS, as well as a LOS component with a non-zero mean.
+First, I describe the general channel.
+Then I will begin with a simple channel in order to investigate the calculation of the AED as each component (correlation, IRS, LOS) is introduced.
+\par
+The general channel is given by 
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}.
+	\end{equation}
+	 $\mathbf{R}_{R}$, $\mathbf{R}_{T}$ and $\mathbf{R}_{S}$ represent deterministic correlation matrices at the receiver, transmitter and IRS.
+$\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ have i.i.d elements in $\mathcal{NC}(0,\frac{1}{N})$, $\mathbf{G}$ has i.i.d elements in $\mathcal{NC}(\mu,\frac{1}{N})$ and $\boldsymbol{\Phi}$ represent the phase shifts at the IRS elements.
+	This simplify notation, I will consider all matrices to have dimension $N \times N$.
+
+\subsection{Correlation Normalization}
+
+In order to evaluate the impact of correlation on the system capacity, the power of the correlation matrices, $\mathbf{R}_{R}$, $\mathbf{R}_{T}$ and $\mathbf{R}_{S},$ must be normalized to prevent adding power to the received signal.
+For $\mathbf{H}$ with $h_{i,j} \in \mathcal{NC}(0,\frac{1}{N})$ and received signal $\mathbf{y} = \mathbf{H} \mathbf{x}$ with $E[\mathbf{x}\mathbf{x}^H] = \mathbf{I}$, the expected received power is given by
+$E[\text{trace}\left( \mathbf{y} \mathbf{y}^H \right) ]= E[\text{trace}\left(\mathbf{H} \mathbf{H}^H \right)] = N$.
+For the correlated channel given by $\mathbf{G} = \mathbf{R}\mathbf{H}$, the received power is given by $E[\text{trace}\left(\mathbf{R} \mathbf{H} \mathbf{H}^H \mathbf{R}^H\right)] = E[\text{trace}\left(\mathbf{R}\mathbf{R}^H\right)] $.
+This constraint is equivalent to requiring that the columns of $\mathbf{R}$, $\mathbf{r}_i$ are normalized such that $\|\mathbf{r}_i\|^2_2 =1$.
+ For the case in which the correlation is a pre-multiplication, $\mathbf{G} = \mathbf{H}\mathbf{R}$, the rotation property of the trace gives the same constraint on the columns of 
+ $\mathbf{R}$.
+
+*Need to discuss water filling at some point still.
+
+\section{Correlation Normalization}
+In order to use tools from free probability to analyze the AED of this channel, I first need to confirm that the components of the above channel are free.
+For the channel shown in \ref{section:channel} there are two components  to consider.
+\begin{enumerate}
+\item 
+	First just consider the Concatenated matrices with correlation
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}\mathbf{H}_{1}\mathbf{R}_{T}
+	\end{equation}
+
+\item 
+	Adding matrices with the same type of correlation	
+\end{enumerate}
+
+\section{Evaluating Channel Components}
+
+\subsection{Basic Correlated Channel}\label{ssection:basic_correlation}
+	Before considering the general channel,  I will go through the calculation of the AED the basic MIMO 
+	channel with correlation given by
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{G}
+	\end{equation}
+	in which $\mathbf{R}_{R}$ is a deterministic matrix with normalized columns describing the correlation and
+	elements $\mathbf{g}_{i,j} \in \mathcal{NC}(0,\frac{1}{N})$ and are iid.
+
+\subsubsection{Evaluating Freeness}
+Rotating the covariance matrix $\mathbf{C} = \mathbf{H}_{Total} \mathbf{H}_{Total}^H$ to get $ \tilde{\mathbf{C}}=\mathbf{R}_{R}^H \mathbf{R}_{R} \mathbf{G} \mathbf{G}^H
+$ gives a matrix which has the same non-zero eigenvalues. 
+$ \mathbf{R}_{R}^H \mathbf{R}_{R}$ is free from any matrix with upper bounded norm and a limit distribution.
+The correlation matrix $\mathbf{G} \mathbf{G}^H$ satisfies these conditions so these two matrices are free. 
+(Pg. 77 of Random Matrix Theory Notes. TODO: find literature source)
+
+\subsubsection{AED Expression}
+
+	The following steps and equations are used to find the AED of $\mathbf{C} = \mathbf{H}_{Total} \mathbf{H}_{Total}^H$.
+	\begin{enumerate}
+	\item 
+		Find AED from the stieltjes transform.
+		\begin{equation}
+		p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
+		\end{equation}
+	\item 
+		Find Stieltjes transform from $\gamma\left(\cdot \right)$
+		\begin{equation}
+		G(s)=  \frac{1}{s} (1+\gamma\left(\frac{1}{s}\right))
+		\end{equation}
+	\item 
+		To find an expression for $\gamma(z)$, evaluate the equation
+		\begin{equation}
+		\gamma^{-1}(z) =  \frac{z}{1+z} S(z)
+		\end{equation}
+		at $z= \gamma(z)$
+		to get the fixed point equation
+		\begin{equation}
+		\gamma(z) = \frac{z(1+ \gamma(z))}{S(\gamma(z))}
+		\end{equation}
+		I can find the AED using steps 1-3 as long as I can evaluate $S(z)$ is in closed form. 
+		In general, however, this is not the case (e.g when $S(z)$ includes a deterministic correlation matrix). 
+		In the following steps I show how I am trying to find $S(z)$ when I do not have a closed form expression. 
+	\item 
+		For the case with all square matrices, 
+		\begin{equation}\label{rotation_property}
+		S_{\mathbf{C}}(z) =  S_{\tilde{\mathbf{C}}}\left(z\right)
+		\end{equation}
+		with the rotated channel $\tilde{\mathbf{C}} = \mathbf{R}_{R}^H\mathbf{R}_{R}\mathbf{G}\mathbf{G}^H$.
+		This gives
+		\begin{equation}
+		S_{\tilde{\mathbf{C}}}\left( z \right) = S_ {\mathbf{R}_{Rx}^H\mathbf{R}_{Rx}}(z)
+		\underbrace{S_ {\mathbf{G}\mathbf{G}^H}(z)}_{\text{close form}}.
+		\end{equation}
+	\item 
+		To find the S-Transform of the correlation matrix, $S_ {\mathbf{R}_{R}^H\mathbf{R}_{R}}(z)$,
+		use the property 
+		\begin{equation}
+		 S(z) = \gamma^{-1}(z)\frac{1+z}{z}.
+		\end{equation}
+	\item
+		Now find an expression for $\gamma^{-1}(z)$ using the known Steiltjes transform of the deterministic
+		 correlation matrix. 
+		 \par
+		 If I use the procedure shown in the following, the value of $S(z)$ is 
+		generally not correct when compared to a known S-Transforms (e.g. when I try this procedure for the 
+		quarter circle and compare it to the known S-Transform, it is not correct).
+	\item
+		Evaluate the the equation 
+		\begin{equation}
+		G(s)=  -\frac{1}{s} (1+\gamma\left(\frac{1}{s}\right))
+		\end{equation}
+		at $s = \gamma^{-1}(z)$ 
+		to find the fixed point equation
+		 \begin{equation}
+		\gamma^{-1}(z) = \frac{-G(\frac{1}{\gamma^{-1}(z)})}{z+1}
+		\end{equation}
+		* When I try this in MATLAB, I get a the error that fsolve does not converge.
+	\item 
+		For the case of deterministic matrices 
+		\begin{equation}
+		G(s) = \int \frac{1}{x-s} dPx
+		\end{equation}
+		becomes 
+		\begin{equation}
+		G(s) = \sum_{i=1}^{N} \frac{1}{N (\lambda_i - s)}.
+		\end{equation}	
+	\end{enumerate}
+
+\subsection{LOS Component}\label{ssection:los}
+
+The next step the correlated LOS component with non-zero mean is added. 
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{G}
+	\end{equation}
+	with 
+	\begin{equation}
+	\mathbf{G} = \underbrace{\mathbf{H}}_{\mathcal{NC}(0,\sigma)} + \underbrace{\alpha\mathbf{I}}_{\text{scaled identity}}
+	\end{equation}
+	\begin{enumerate}
+	\item
+	Using the property 
+		\begin{equation}\label{svd_aed_property}
+		G_{\lambda\lambda}(s) = \frac{1}{\sqrt{s}}\tilde{G}_{\lambda}(\sqrt{s})
+		\end{equation}
+	avoids having to work with non-hermetian and non-free terms in the covariance polynomial $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$. 
+	\item 
+		Next obtain the symmetrized singular value from the corresponding R-Transform using
+		\begin{equation}
+		\tilde{G}_{\lambda}(\sqrt{s}) = \frac{1}{\tilde{R}(-\tilde{G}_{\lambda}(\sqrt{s})) - s}
+		\end{equation}
+		and 
+			\begin{equation}
+			\tilde{R}(w) = \tilde{R}_{\mathbf{R}\mathbf{H}}(w) + \tilde{R}_{\alpha\mathbf{R}\mathbf{I}}(w)
+			\end{equation}
+	\item 
+		In order to find the 
+	\end{enumerate}
+
+\subsection{Correlated LOS and IRS components}\label{ssection:los_correlated}
+Now I consider the case in which the same correlation is applied to both the LOS and IRS components of the channel.
+
 \end{document}

### 2020-06-02 documentation/Paper/paper.tex
working on update

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\input defs.tex
\bibliographystyle{alpha}
\graphicspath{ {./figures/} }
%\hyphenpenalty=10000

\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
\author{Peter Hartig}

\begin{document}
\maketitle
\begin{abstract}
This work investigates the impact of passive reflective surfaces on the capacity of MIMO systems. In particular, the system is analyzed in the asymptotic size domain in order to employ the tools of free probability. A system is considered first in the case of fully uncorrelated channels. In this case it is shown that the passive control of such a reflective surface has no impact on the capacity of the system. The system is then generalized to allow for correlation within the channel. In this case is it shown that the selection of the phases at the IRS still has an impact on the system. Finally we compare the capacity of the correlated channel with optimized phases at the IRS to the uncorrelated channel to show that....?
\end{abstract}
%
%\newpage
\tableofcontents

\chapter{Introduction}

\section{Notation}
The following notation is used throughout the remainder. 
$E\{x\}$ is the expected value of a random variable $x$.
Vectors are denoted by bold font, lower-case letters ($\mathbf{x}$) and are assumed to be column vectors.
Matrices are denoted by bold font, upper-case letters ($\mathbf{X}$). The transpose and hermetian of $\mathbf{X}$ are denoted by $\mathbf{X}^T$ and $\mathbf{X}^H$ respectively.
$\mathbf{I}$ denotes the identity matrix.

\section{Outline}
The following sections are now previewed. In section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel. In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed. 
\chapter{System Model}\label{system_model}
\section{Outline}

\section{Proof of Freeness}

\chapter{Proofs etc. }\label{Results}

\chapter{Numerical Results}\label{numerical}

\chapter{Conclusion}

\bibliography{bibliography}
\end{document}

### 2020-06-02 simulation/src/lin_alg.py

@@ -25,7 +25,7 @@ def c_rand(rows, cols, var=None, mean=0):
         return (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows) + mean
 
 
-def block_matrix(block, size, normalize=False):
+def block_matrix(block, size, rows=True):
     """
     Create correlation matrices with repeated correlation blocks
     :param block:
@@ -33,27 +33,22 @@ def block_matrix(block, size, normalize=False):
     """
     dim = size//block.shape[1]
     mat = np.kron(np.eye(dim), block)
-    if not normalize:
-        return mat
-    else:
-        return normalize_matrix(mat, rows=True)
+    return normalize_matrix(mat, rows=rows)
 
 
-def exponential_correlation(size, r):
+def exponential_correlation(size, r, rows=True):
     ret_mat = np.ones((size, size))
     for row in range(ret_mat.shape[0]):
         for col in range(ret_mat.shape[1]):
             ret_mat[row, col] *= pow(r, abs(col-row))
-    for row in range(ret_mat.shape[0]):
-        ret_mat[row, :] /= np.linalg.norm(ret_mat[row, :], 2)
-    return ret_mat
+    return normalize_matrix(ret_mat, rows)
 
 
 def normalize_matrix(matrix, rows=False):
     ret_mat = matrix
     if not rows:
         for col in range(matrix.shape[1]):
-            ret_mat[:, col]/np.linalg.norm(matrix[:, col], 2)
+            ret_mat[:, col] /= np.linalg.norm(matrix[:, col], 2)
         return ret_mat
     else:
         for row in range(matrix.shape[0]):

### 2020-06-03 documentation/Meeting Notes/06_03_20_update.tex
Working on background for report

--- 2020-06-02 documentation/Meeting Notes/05_29_20.tex
@@
 \par
-The channel considered includes correlation at the transmitter, receiver and IRS, as well as a LOS component with a non-zero mean.
-First, I describe the general channel.
-Then I will begin with a simple channel in order to investigate the calculation of the AED as each component (correlation, IRS, LOS) is introduced.
-\par
-The general channel is given by 
+The channel considered includes correlation at the transmitter, receiver and IRS, as well as a LOS component with a non-zero mean given by 
 	\begin{equation}
@@
 
-*Need to discuss water filling at some point still.
-
-\section{Correlation Normalization}
-In order to use tools from free probability to analyze the AED of this channel, I first need to confirm that the components of the above channel are free.
-For the channel shown in \ref{section:channel} there are two components  to consider.
-\begin{enumerate}
-\item 
-	First just consider the Concatenated matrices with correlation
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}\mathbf{H}_{1}\mathbf{R}_{T}
-	\end{equation}
-
-\item 
-	Adding matrices with the same type of correlation	
-\end{enumerate}
+%*Need to discuss water filling at some point still.
 
 \section{Evaluating Channel Components}
+\subsection{Note from Previous work}\label{ssection:basic_correlation}
+In \cite[Section II Eqn.
+4]{muller2012channel} it is mentioned that correlation can be added "easily" using free probability once the AED of the uncorrelated channel is known.
+Can we discuss these steps towards easily incorporating this correlation?
+They are not apparent to me.
+In the following I attempt to find the AED in a similar case but it seems quite tedious, I imagine there is an easier way.
 
 \subsection{Basic Correlated Channel}\label{ssection:basic_correlation}
-	Before considering the general channel,  I will go through the calculation of the AED the basic MIMO 
+	Before considering the general channel,  I will go through the calculation of the AED of a basic MIMO 
 	channel with correlation given by
@@
 
-\subsubsection{Evaluating Freeness}
-Rotating the covariance matrix $\mathbf{C} = \mathbf{H}_{Total} \mathbf{H}_{Total}^H$ to get $ \tilde{\mathbf{C}}=\mathbf{R}_{R}^H \mathbf{R}_{R} \mathbf{G} \mathbf{G}^H
-$ gives a matrix which has the same non-zero eigenvalues. 
-$ \mathbf{R}_{R}^H \mathbf{R}_{R}$ is free from any matrix with upper bounded norm and a limit distribution.
-The correlation matrix $\mathbf{G} \mathbf{G}^H$ satisfies these conditions so these two matrices are free. 
-(Pg. 77 of Random Matrix Theory Notes. TODO: find literature source)
+%\subsubsection{Evaluating Freeness}
+%First, I show that free probability can be used. 
+%Rotating the covariance matrix $\mathbf{C} = \mathbf{H}_{Total} \mathbf{H}_{Total}^H$ to get $ \tilde{\mathbf{C}}=\mathbf{R}_{R}^H \mathbf{R}_{R} \mathbf{G} \mathbf{G}^H
+%$ gives a matrix which has the same non-zero eigenvalues. The hermetian matrix
+%$ \mathbf{R}_{R}^H \mathbf{R}_{R}$ is free from any matrix with upper bounded norm and a limit distribution.
+%The correlation matrix $\mathbf{G} \mathbf{G}^H$ satisfies these conditions so these two matrices are free. 
+%(Pg. 77 of Random Matrix Theory Notes. TODO: find literature source)
 
@@
 	\item 
-		Find AED from the stieltjes transform.
+		Expression for AED given the stieltjes transform.
 		\begin{equation}
@@
 	\item 
-		Find Stieltjes transform from $\gamma\left(\cdot \right)$
+		Expression for  Stieltjes transform given $\gamma\left(\cdot \right)$
 		\begin{equation}
@@
 	\item 
-		To find an expression for $\gamma(z)$, evaluate the equation
+		To find an expression for $\gamma(z)$, evaluate 
 		\begin{equation}
@@
 		\end{equation}
-		I can find the AED using steps 1-3 as long as I can evaluate $S(z)$ is in closed form. 
+		I can find the AED using steps 1-3 as long as I can evaluate $S(z)$ in closed form. 
 		In general, however, this is not the case (e.g when $S(z)$ includes a deterministic correlation matrix). 
@@
 		S_{\tilde{\mathbf{C}}}\left( z \right) = S_ {\mathbf{R}_{Rx}^H\mathbf{R}_{Rx}}(z)
-		\underbrace{S_ {\mathbf{G}\mathbf{G}^H}(z)}_{\text{close form}}.
+		\underbrace{S_ {\mathbf{G}\mathbf{G}^H}(z)}_{\text{closed form}}.
 		\end{equation}
@@
 	\item
-		Now find an expression for $\gamma^{-1}(z)$ using the known Steiltjes transform of the deterministic
+		Now I find an expression for $\gamma^{-1}(z)$ using the known Steiltjes transform of the deterministic
 		 correlation matrix. 
@@
 		generally not correct when compared to a known S-Transforms (e.g. when I try this procedure for the 
-		quarter circle and compare it to the known S-Transform, it is not correct).
+		quarter circle eigenvalues and compare it to the known S-Transform, it is not correct).
 	\item
@@
 		\end{equation}	
+		for known eigenvalues $\lambda_i$.
 	\end{enumerate}
 
-\subsection{LOS Component}\label{ssection:los}
-
-The next step the correlated LOS component with non-zero mean is added. 
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}\mathbf{G}
-	\end{equation}
-	with 
-	\begin{equation}
-	\mathbf{G} = \underbrace{\mathbf{H}}_{\mathcal{NC}(0,\sigma)} + \underbrace{\alpha\mathbf{I}}_{\text{scaled identity}}
-	\end{equation}
-	\begin{enumerate}
-	\item
-	Using the property 
-		\begin{equation}\label{svd_aed_property}
-		G_{\lambda\lambda}(s) = \frac{1}{\sqrt{s}}\tilde{G}_{\lambda}(\sqrt{s})
-		\end{equation}
-	avoids having to work with non-hermetian and non-free terms in the covariance polynomial $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$. 
-	\item 
-		Next obtain the symmetrized singular value from the corresponding R-Transform using
-		\begin{equation}
-		\tilde{G}_{\lambda}(\sqrt{s}) = \frac{1}{\tilde{R}(-\tilde{G}_{\lambda}(\sqrt{s})) - s}
-		\end{equation}
-		and 
-			\begin{equation}
-			\tilde{R}(w) = \tilde{R}_{\mathbf{R}\mathbf{H}}(w) + \tilde{R}_{\alpha\mathbf{R}\mathbf{I}}(w)
-			\end{equation}
-	\item 
-		In order to find the 
-	\end{enumerate}
-
-\subsection{Correlated LOS and IRS components}\label{ssection:los_correlated}
-Now I consider the case in which the same correlation is applied to both the LOS and IRS components of the channel.
+%\subsection{LOS Component}\label{ssection:los}
+%
+%The next component of channel to consider is the correlated LOS channel given by  
+%	\begin{equation}
+%	\mathbf{H}_{Total} = \mathbf{R}\mathbf{G}
+%	\end{equation}
+%	with $\mathbf{R}$ as a deterministic correlation matrix and 
+%	\begin{equation}
+%	\mathbf{G} = 
+%	\underbrace{\mathbf{H}}_{\mathcal{NC}(0,\sigma)} + \underbrace{\mathbf{D}}_{\text{Deterministic}}.
+%	\end{equation}
+%	
+%\subsubsection{Evaluating Freeness}
+%
+%	
+%\subsubsection{AED Expression}
+%
+%	\begin{enumerate}
+%	\item
+%	Using the property 
+%		\begin{equation}\label{svd_aed_property}
+%		G_{\lambda\lambda}(s) = \frac{1}{\sqrt{s}}\tilde{G}_{\lambda}(\sqrt{s})
+%		\end{equation}
+%	avoids having to work with non-hermetian and non-free terms in the covariance polynomial $\mathbf{H}_{Total}\mathbf{H}_{Total}^H$. 
+%	\item 
+%		Next obtain the Stieltjes transform of the symmetric singular value from the corresponding R-Transform using
+%		\begin{equation}
+%		\tilde{G}_{\lambda}(\sqrt{s}) = \frac{1}{\tilde{R}(-\tilde{G}_{\lambda}(\sqrt{s})) - s}
+%		\end{equation}
+%		and 
+%			\begin{equation}
+%			\tilde{R}(w) = \tilde{R}_{\mathbf{R}\mathbf{H}}(w) + \tilde{R}_{\mathbf{R}\mathbf{D}}(w)
+%			\end{equation}
+%	\item 
+%		I will omit the next steps until I have a solution to the basic correlated channel because the basic
+%		correlated channel needs to be solved to handle this case. 
+%	\end{enumerate}	
+\bibliography{bibliography}
 

### 2020-06-03 documentation/Paper/paper.tex
Working on background for report

@@
 $E\{x\}$ is the expected value of a random variable $x$.
-Vectors are denoted by bold font, lower-case letters ($\mathbf{x}$) and are assumed to be column vectors.
-Matrices are denoted by bold font, upper-case letters ($\mathbf{X}$). The transpose and hermetian of $\mathbf{X}$ are denoted by $\mathbf{X}^T$ and $\mathbf{X}^H$ respectively.
+Vectors are denoted by bold font, lower-case letters ($\vx$) and are assumed to be column vectors.
+Matrices are denoted by bold font, upper-case letters ($\mX$). The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
 $\mathbf{I}$ denotes the identity matrix.
@@
 \chapter{System Model}\label{system_model}
-\section{Outline}
+\section{The Correlated IRS Channel}
+
+\section{Mutual Information}
+A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
+If the channel information is known at the receiver, this is given by
+\begin{equation}
+I(\vy;\vx) = h(\vy) - h(\vy|\vx).
+\end{equation}
+If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} $h(\vy) = \Log(|\pi e \mathbf{Q}_{y}|)$ with $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$.
+Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
+
+For a given transmit covariance matrix, $\mathbf{Q_x}$, the mutual information is given by 
+\begin{equation}\label{mut_inf}
+\Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
+\end{equation}
+Using the model for $\mHt$ from the previous section, and maximizing with respect to the
+transmit covariance matrix and the phases at the IRS gives and expression for the capacity of the system
+
+\begin{equation}\label{capacity}
+\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
+\end{equation}
+Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
+
+If $\mHt$ is not known at the transmitter, the transmit power can be allocated
+equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
+\begin{equation}\label{no_csi_capacity}
+\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
+\end{equation}
+where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
+
+Note that because $\Log$ is a concave function, we can upper bound the mutual information with
+\begin{equation}\label{mut_inf}
+\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda_i])
+\end{equation}
+Now discuss capacity and consider equal power but optimize over phase
+We now want to consider the case in
 
 \section{Proof of Freeness}
+\begin{itemize}
 
-\chapter{Proofs etc. }\label{Results}
+\item 
+	Brief introduction to free probability.
+	Use the example of a degenerate matrix with rank 1 being multiplied by a random matrix. 
+	In this case the resulting pdf is not just the convolution of the original distributions. 
+	In order to introduce the tools that will be used in the following, an example is now given to
+	exhibit this case in which classic probability (commutative probability) no longer holds.
+	Consider the Gaussian i.i.d. matrix $\mH$ with $ h_{i,j} \in \CN$ and the diagonal matrix $\mP$ with 
+	$p_{i,i} = 1$ for $i = 1 \cdots K$ with $K \leq N$ and both matrices of dimension $N \times N$. 
+	We are interested the AED for $\mP\mH \mH^H \mP^H$ and immediately note  that the matrices
+	$ \mP^H \mP \mH \mH^H = \mP \mH \mH^H$ have the same non-zero eigenvalues (show why!). 
+	Given that both $ \mP$ and  $\mH \mH^H$ have converging AEDs as $N \rightarrow \infty$
+	classical probability would tell us that the AED for $\mP \mH \mH^H$ is the convolution of the two AEDs. 
+	This, however, is not the case non-commutative distributions. This can be seen without the tools of free probability by noting
+	that the resulting product has $\Rank = \min[\Rank(\mP), \Rank(\mH \mH^H)]$. In the case that $\Rank(\mP) = 1$
+	the resulting product must have rank 1 (i.e a single non-zero eigenvalue) with corresponding AED
+	$p_{\mP}(x) = \frac{1}{N-1}\delta(x)  + \frac{1}{N}\delta(x-1)$. Following classical probability, the convolution of 
+	a general AED with $p_{\mP}(x)$, however, can be continuous. 
+	\par 
+	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
 
-\chapter{Numerical Results}\label{numerical}
+\item 
+	Show that at least for uncorrelated case, the matrices in the sum are free.
+
+\end{itemize}
+
+\chapter{Results}\label{Results}
+\begin{itemize}
+\item 
+	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
+	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
+	cancels out of \eqref{capacity}.
+
+\item
+	Show that numerical agrees with theoretical for uncorrelated 
+\item
+	Now consider the case where the assumptions of free probability no longer apply 
+	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
+	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
+	correlated case, free probability assumptions no longer hold. 
+\end{itemize}
 

### 2020-06-03 simulation/src/stieltjes_eqn.py

@@ -45,6 +45,7 @@ def final_stieltjes_deformed(s_values, g_k, func_params):
     # Correct
     # G_BB = np.sqrt(np.power((1-phi), 2)/(4*np.power(x_arg, 4)) - (1+phi)/(2*np.power(x_arg, 2))
     #                + 1/4) - 1/2 - (1-phi)/(2*np.power(x_arg, 2))
+    G_BB_tilde = 0
     return (1/np.sqrt(s_values))*x_arg*G_BB
 
 

### 2020-06-03 simulation/testing/channel_testing/parameter_vary.py

@@ -21,9 +21,6 @@ rho = 1     # S/R
 phi = 1    # phi=1 is full rank line of sight.  L/T
 #   Plot Theoretical Distribution and find corresponding capacity
 
-
-
-
 x_values, step = np.linspace(1e-2, 10, 100, retstep=True)
 y = 1e-9
 s_vector = x_values + 1j*y

### 2020-06-04 documentation/Paper/paper.tex
working on free probability background of paper

@@
 \chapter{Introduction}
+Wireless communication has progressively moved to higher frequencies in recent standards in order to utilize large bands of unoccupied resource
+and to enable increased antenna density (CITE). One hurdle to higher frequencies is increased radio frequency energy (power) attenuation (CITE). 
+In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
+and supplementing the transmission power using relays (CITE).
+Two key costs to deploying relays are the radio frequency chains used for receiving and transmitter, and the power used for transmission.
+(Would be nice to cite some actual costs here).
+As the power attenuation problem has been increased by X-Fold moving from freq1 to freq2, new solutions have been considered.
+One solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
+In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
+\par
+The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
+associated with a relay.
+The absence of an rf-chain in the RIS offers additional benefits over relays.
+First, reflection does not introduce the noise associated with the rf-chain.
+Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum.
+Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay.
+The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
+For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is proportional to $N^2$ (?
+Does this hold if we remove the ability to choose the angle?).
+For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
+(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
+Example of how these surfaces are implemented (Cite).
+Discuss how it has be looked at from MISO and MIMO perspective.
+\par
+One area that has not be investigated, (at least up to the author's knowledge) is the impact of the RIS for 
+asymptotically large MIMO systems. In the following, the impact of the RIS on a general MIMO system is explored.
 
@@
 The following notation is used throughout the remainder. 
-$E\{x\}$ is the expected value of a random variable $x$.
-Vectors are denoted by bold font, lower-case letters ($\vx$) and are assumed to be column vectors.
-Matrices are denoted by bold font, upper-case letters ($\mX$). The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
-$\mathbf{I}$ denotes the identity matrix.
+$\Expect[x]$ is the expected value of a random variable $x$.
+Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
+The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
+$\mI_N$ denotes the identity matrix of dimension $N$.
 
@@
 \item 
-	Brief introduction to free probability.
-	Use the example of a degenerate matrix with rank 1 being multiplied by a random matrix. 
-	In this case the resulting pdf is not just the convolution of the original distributions. 
-	In order to introduce the tools that will be used in the following, an example is now given to
-	exhibit this case in which classic probability (commutative probability) no longer holds.
-	Consider the Gaussian i.i.d. matrix $\mH$ with $ h_{i,j} \in \CN$ and the diagonal matrix $\mP$ with 
-	$p_{i,i} = 1$ for $i = 1 \cdots K$ with $K \leq N$ and both matrices of dimension $N \times N$. 
-	We are interested the AED for $\mP\mH \mH^H \mP^H$ and immediately note  that the matrices
-	$ \mP^H \mP \mH \mH^H = \mP \mH \mH^H$ have the same non-zero eigenvalues (show why!). 
-	Given that both $ \mP$ and  $\mH \mH^H$ have converging AEDs as $N \rightarrow \infty$
-	classical probability would tell us that the AED for $\mP \mH \mH^H$ is the convolution of the two AEDs. 
-	This, however, is not the case non-commutative distributions. This can be seen without the tools of free probability by noting
-	that the resulting product has $\Rank = \min[\Rank(\mP), \Rank(\mH \mH^H)]$. In the case that $\Rank(\mP) = 1$
-	the resulting product must have rank 1 (i.e a single non-zero eigenvalue) with corresponding AED
-	$p_{\mP}(x) = \frac{1}{N-1}\delta(x)  + \frac{1}{N}\delta(x-1)$. Following classical probability, the convolution of 
-	a general AED with $p_{\mP}(x)$, however, can be continuous. 
+	A brief introduction to free probability.
+	In order to introduce the tools that will be used in the following, an example of non-commutative probability now given to
+	exhibit a case in which the tools of classic probability are not sufficient for finding a probability density function (pdf).
 	\par 
+	Free probability is the counterpart to classical probability for the case in which the random variables (in this case random matrices) do 
+	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
+	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables (It is important to note here
+	 that by changing this nature of the random variables from commutative to non-commutive (is this embedded in kolmogorov's axioms?) 
+	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context 
+	but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
 	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
@@
 	Show that at least for uncorrelated case, the matrices in the sum are free.
+
+\item 
+	Give proofs in an appendix?
 
@@
 
+\chapter{Appendix}
+Proofs
+\begin{enumerate}
+\item
+	Prove freeness of the relevant points. 
+\item 
+	rotations have same non-zero eigenvalues
+\end{enumerate}
+
 \bibliography{bibliography}

### 2020-06-04 simulation/testing/theory_testing/free_probability_examples.py

@@ -1,0 +1,38 @@
+"""
+This test is to check the fixed point expression for the concatenated channel from the IRS.
+"""
+
+import matplotlib.pyplot as plt
+from src.lin_alg import *
+import seaborn
+
+size = 100
+rows = 100
+cols = 100
+
+bins = int(size/2)
+main_terms = []
+no_irs = []
+irs = []
+numeric_cacacity = []
+average = 10
+for i in range(average):
+    H = c_bernoulli(rows, cols, .5)
+    HH = H @ np.conj(H.T)
+    total_irs = HH
+    e_val_total_irs, e_vec_total_irs = np.linalg.eig(total_irs)
+    irs.append(np.real(e_val_total_irs))
+
+irs_AED, bins_irs = np.histogram(np.asarray(irs), bins=bins)
+fig, ax = plt.subplots()
+plt.title("AED")
+ax.plot(bins_irs[:-1], irs_AED/(rows*average), label='total: with IRS')
+# ax.plot(x, normalize_pdf(aed), label='Theoretic AED')   #Only normalizing for visualization but not for capacity calc.
+# ax.plot(x, aed, label='Theoretic AED')   #Only normalizing for visualization but not for capacity calc.
+
+plt.legend(loc="upper right")
+ax.grid(True, which='both')
+seaborn.despine(ax=ax, offset=0)  # the important part here
+plt.show()
+
+

### 2020-06-08 documentation/Paper/paper.tex
adding equations for s-transform of teoplitz from new paper

@@
 \usepackage{graphicx}
+\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
 \usepackage{float}
@@
 Wireless communication has progressively moved to higher frequencies in recent standards in order to utilize large bands of unoccupied resource
-and to enable increased antenna density (CITE). One hurdle to higher frequencies is increased radio frequency energy (power) attenuation (CITE). 
+and to enable increased antenna density \cite{akyildiz2018combating}.
+A primary hurdle towards communicating over higher frequencies is the increased radio frequency energy (power) attenuation.
+For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
+has been shown to exceed y.
+Given that increasing current wireless communication capacity and reliability rely on moving to these high frequencies, mitigating or countering these power losses on the order of 100 dB is highly relevant.
 In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
@@
 $\mI_N$ denotes the identity matrix of dimension $N$.
+For square $\mX$ with dimension $N\times N$ the normalized trace is denoted by $\Tr(\mX)  = \underset{N \rightarrow \infty}{\text{lim}}
+\Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
 
@@
 In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
+
 \chapter{System Model}\label{system_model}
@@
 
-\section{Mutual Information}
+\section{Mutual Information}\label{sectiond:mut_info}
 A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel information is known at the receiver, this is given by
-\begin{equation}
+If the channel $\mHt$ is known at the receiver, the mutual information is given by
+\begin{equation}\label{mut_ent}
 I(\vy;\vx) = h(\vy) - h(\vy|\vx).
 \end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} $h(\vy) = \Log(|\pi e \mathbf{Q}_{y}|)$ with $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$.
-Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
+If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
+\begin{equation}\label{entropy}
+h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
+\end{equation}
+ for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
 
-For a given transmit covariance matrix, $\mathbf{Q_x}$, the mutual information is given by 
+For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect[\vn\vn^H] = \mI$, substituting equation \eqref{entropy} into equation
+\eqref{mut_ent} gives
 \begin{equation}\label{mut_inf}
@@
 \end{equation}
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
+Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
 
@@
 where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
+Expanding the expectation gives the expression
+\begin{equation}\label{no_csi_capacity_aed}
+\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
+\end{equation}
+in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
 
@@
 
+\section{Motivation for AED as a projection of Matrices}
+Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
+For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = Expect[\Tr{(mHt \mHt^H})] = $.
+One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
+the channel.
+For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
+\begin{equation}\label{sigular_eigen}
+\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
+\end{equation}
+This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
 \section{Proof of Freeness}
@@
 	A brief introduction to free probability.
-	In order to introduce the tools that will be used in the following, an example of non-commutative probability now given to
-	exhibit a case in which the tools of classic probability are not sufficient for finding a probability density function (pdf).
 	\par 
-	Free probability is the counterpart to classical probability for the case in which the random variables (in this case random matrices) do 
+	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
 	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables (It is important to note here
-	 that by changing this nature of the random variables from commutative to non-commutive (is this embedded in kolmogorov's axioms?) 
-	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context 
-	but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
+	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
+	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
+	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
+	for this investigation but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
+	 \begin{itemize}
+	 \item 
+	 	The cornerstone of free probability is the concept freeness of two random variables in the context of a projection. In our case, freeness between two 
+	 	matrices is given by 
+	 	We now introduce useful tools from free probability  that can be applied to free random variables.
+
+	 	\item 
+	 	Stieltjes Transform: 
+	 	Generally, the projection of matrix is not given in closed form. The projection can be uniquely represented using
+		\begin{equation}\label{stieltjes}
+		G(s) = \int{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
+		\end{equation}	
+		This representation will allow us to find the projection of the sum and project of matrices using addition and multiplication.
+		\item 
+		R-Transform:
+		Mention self-adjoint
+		Given the Stieltjes transform of a RV, the corresponding R-Transform given by 
+		X. 
+		The R-Transform of the sum of two, free RVs can be found using
+		Y
+		This process use known as free additive convolution.
+		\item
+		S-Transform:
+		Mention self-adjoint
+		Similarly, given the Stieltjes transform of a RV, the S-Transform given by 
+		X, X;
+		The S-Transform of the product of two, free RVs can be found using
+		Y
+		This process is known as free multiplicative convolution.
+	 	\item 
+	 	If there is a case in which finding the singular value distribution is easier than the eigenvalue distribution we can use equation  
+	 	 \eqref{sigular_eigen} to move between the two. Considering a symmetric function of the singular value distribution
+	 	 \begin{equation}\label{symmetric}
+	 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2}
+	 	 \end{equation}
+	 	 the definition of the Steiltjes transform gives the following steps, leading to a relationship between the Stieltjes transforms. 
+	 	 \begin{align*}
+	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}
+	 	 \\&  =  
+	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
+	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
+	 	 \\&  =  
+	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
+	  	 \\&  =  
+	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
+ 	 	  	 \\&  =  
+ 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
+ 	 	 \\&  =  
+			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
+	 	 \\&  = 
+	 	 	-s G_{\lambda}(s^2)
+	 	 \end{align*}
+	 	\item 
+	 		Another useful property that will be used is the relationship between the eigenvalues of a matrix and its rotation.
+	 		Specifically, consider the covariance matrix given by $\mH_2\mH_1 \mH1^H \mH2^H$.
+	 		By the rotational property of the trace, 
+	 		\begin{align*}
+	 		trace({\mH_2\mH_1 \mH1^H \mH2^H }) = \trace({\mH_1^H \mH_2^H \mH_2 \mH_1 }).
+	 		\end{align*}
+	 		... Not sure if I need to prove this here.  
+	 	\item 
+	  	build up to show that phases cancel via s-transform
+	 	Stieltjes of AED (IE for hermetian matrices)
+	 	Stieltjes of SVD (when they are components covariance matrix)
+	 	Getting Steiltes of SVD using R-Transform
+	 	Finding R-Transform of Individual components
+	 	Using Stieltjes of individuals to get R-transforms
+	 	Using aed stieltjes to get svd individual
+	 	Using S-transform to get stieltjes of individual aed.
+	\end{itemize}
+
 	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
@@
 \chapter{Results}\label{Results}
+\section{Canceling Phases}
+We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
+phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
+By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
+the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
+\par
+While the realized capacity of the asymptotic channel converge towards the asymptotic capacity, the rate of this convergence is particularly relevant in this application as this implies the number of rf-chains at the transmitter and receiver approach infinity.
+One notable observation is that this rate
+decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
 \begin{itemize}
 \item
+
 	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters

### 2020-06-08 simulation/MATLAB/fixed_point_test.m

@@ -4,32 +4,52 @@ steps = 100;
 bottom = .01;
 top = 10; 
 channel = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
-block_size = 2;
-block = toeplitz(ones(1,rows/block_size));
-correlation = kron(eye(block_size), block);
-correlation = normr_la(correlation);
+% block_size = 2;
+% block = toeplitz(ones(1,rows/block_size));
+% correlation = kron(eye(block_size), block);
+% correlation = normr_la(correlation);
 % correlation = correlation*correlation';
-% total_cov = (correlation*(channel*channel')*correlation');
-total_cov = (channel*channel');
+rho = .9;
+correlation = exponential_correlation(rows, rho);
+total_cov = (correlation*(channel*channel')*correlation');
+% total_cov = (channel*channel');
 det_eigen_values = eig(correlation);
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
-s_values = x_values + 1i*1e-4;
-stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, det_eigen_values));
+s_values = x_values + 1i*1e-6;
+% stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, det_eigen_values));
+stieltjes_values = (1./s_values).*(1+gamma_s(s_values));
 pdf = 1/pi .* imag(stieltjes_values);
 num_capacity = MIMO_capacity(total_cov, 1/cols)
 asymptotic_capacity  = aed_capacity(x_values, pdf, 1/cols, rows, step_size)
 plot(x_values, pdf)
 title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)])
 
-function output = gamma_s(input, eval)
+function output = gamma_s(input)
     out = zeros(size(input), 'like', input);
     for i = 1:length(input)
-        out(i) =  gamma_s_func(input(i), eval);
+%         out(i) =  gamma_s_func(input(i), eval);
+        out(i) =  new(input(i));
     end
     output = out;
 end
 
+function x = inverse_gamma_cor(eval_point)
+    rho = .9;
+    alpha = (1+rho^2)/(1-rho^2);
[…]

### 2020-06-08 simulation/testing/phase_optimization/phase_check_simulation.py

@@ -24,7 +24,7 @@ cross_sum = 0
 main_sum = 0
 
 average = 1000
-block = toeplitz(np.ones(int(rows / 1)))
+block = toeplitz(np.ones(int(rows / 4)))
 correlation_matrix = block_matrix(block, rows)
 # correlation_matrix = exponential_correlation(size, .5)
 # H_1 = c_rand(size, n_t)
@@ -42,7 +42,8 @@ for i in range(average):
     HG = H @ np.conj(G.T)
     cross = GH + HG
     main = HH + GG
-    total_irs = cross + main
+    # total_irs = cross + main
+    total_irs = HH
     # total_irs = GG
     capacities.append(capacity(total_irs, 1/n_t))
 

### 2020-06-11 documentation/Paper/paper.tex
working on background

@@
 \chapter{Introduction}
-Wireless communication has progressively moved to higher frequencies in recent standards in order to utilize large bands of unoccupied resource
-and to enable increased antenna density \cite{akyildiz2018combating}.
-A primary hurdle towards communicating over higher frequencies is the increased radio frequency energy (power) attenuation.
+\section{Background}\label{Background}
+The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards, projected to reach even the Terahertz range by the 6G standard (CITE).
+This expansion is motivated by large bands of currently unused frequency
+and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
+A primary hurdle towards efficient communication at these frequencies is the increased radio frequency (power) attenuation.
 For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
-has been shown to exceed y.
-Given that increasing current wireless communication capacity and reliability rely on moving to these high frequencies, mitigating or countering these power losses on the order of 100 dB is highly relevant.
+has been shown to exceed 100 dB.
+With increases in current wireless system capacity and reliability relying on the use of these high frequencies, mitigating or countering this power loss  is highly relevant.
+\par
 In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
 and supplementing the transmission power using relays (CITE).
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitter, and the power used for transmission.
+Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
 (Would be nice to cite some actual costs here).
-As the power attenuation problem has been increased by X-Fold moving from freq1 to freq2, new solutions have been considered.
-One solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
+Deploying this solution on a scale to counter
+the attenuation levels relevant to new standards would pose impractical costs.
+One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
 In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
@@
 First, reflection does not introduce the noise associated with the rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum.
+Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
 Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay.
 The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is proportional to $N^2$ (?
+For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
 Does this hold if we remove the ability to choose the angle?).
@@
 \par
-One area that has not be investigated, (at least up to the author's knowledge) is the impact of the RIS for 
-asymptotically large MIMO systems. In the following, the impact of the RIS on a general MIMO system is explored.
+Now discuss all of the previous work looking at the optimization of the phases. 
+General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
+Need to mention anyone who has considered correlation at end in order to transition. 
+\par
+Early investigation leading to the great interest in MIMO systems showed the key property of capacity in an $N \times N$ MIMO system scaling linearly in $N$ (teletar).
+While this scaling applies to many models for the channel (cite) one key property that can impact this scaling is correlation between channels of a MIMO system.
+In order to evaluate MIMO system performance in a real-world application, correlation between between the channels must be accounted for.
+In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?).
+(Is this a bigger problem for higher frequencies?).
+Correlations cause reduction in the capacity of the system and often make the calculation of the capacity more difficult.
+However, with the possibility for arrays with (x) co-located antenna and RIS with (y) individually controlled reflective elements, the impact of channel correlation on system performance must be well understood before deployment.
+\par
+Detail all of the work looking at correlated MIMO. 
+1. Looking at correlation for finite size systems.
+Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
+2. Extending to asymptotic system sizes. 
+One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. This method has been used 
+extensively since. 
+Now list of the different types of channels/methods that have been used here. 
+\cite{loyka2001channel}
+\par
+From the above, it should be clear that investigating the RIS  system for 
+asymptotically large MIMO systems with correlation, an topic not yet explicitly investigated to the author's knowledge, is relevant to the design of upcoming wireless systems.
 
@@
 \section{The Correlated IRS Channel}
+In the most general case we consider the channel given by
+
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+	\end{equation}
 
@@
 		\begin{equation}\label{stieltjes}
-		G(s) = \int{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
+		G(s) = \int_{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
 		\end{equation}	
@@
 \chapter{Results}\label{Results}
-\section{Canceling Phases}
+\section{Asymptotic Results}
 We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
@@
 \par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity, the rate of this convergence is particularly relevant in this application as this implies the number of rf-chains at the transmitter and receiver approach infinity.
+While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity.
 One notable observation is that this rate
@@
 \end{itemize}
+\subsection{Asymptotic Assumptions Under Correlation}
+First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.
+
+\section{Non-Asymptotic Results}
+In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
+RIS has no impact on the capacity of the system.
+Second, that for matrices of random elements, the  convergence of matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
+As seen in \ref{Background} current large systems have been shown for
+sizes (X). We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).
 
 \chapter{Conclusion}
-
 \chapter{Appendix}
@@
 	rotations have same non-zero eigenvalues
+\item 
+Proof of split using singular values.\\
+Consider the channel given by 
+\begin{equation}
+\mH_{Total} = \mH_{2} + \mH_{1}.
+\end{equation}
+with covariance matrix 
+\begin{equation}
+\mC =
+\mH_{2}\mathbf{H}_{2}^H + 
+\mH_{1}\mathbf{H}_{1}^H + 
+\mH_{2}\mathbf{H}_{1}^H +
+\mH_{1}\mathbf{H}_{2}^H .
+\end{equation}
+Because the components of this sum are not free (CITE) (and some are non-hermetian), additive free convolution cannot be used.
+We note that using the singular value decomposition $\mH_{Total} = \mU \boldsymbol{\Sigma} \mV^H$ we can
+see that the covariance matrix can be written 
+\begin{equation}
+\mC = \mU \boldsymbol{\Sigma}\boldsymbol{\Sigma}^H \mU^H
+= \mU \boldsymbol{\Lambda} \mU^H.
+\end{equation}
+With $\mC \succeq 0 \implies \lambda(\mC\mC)_i \geq 0$ we see that the definition of the singular values 
+$\sigma(\mC)_i = \sqrt{\lambda(\mC\mC)_i}$ leads to an expression relating the density function of the singular values of a channel matrix
+, $p_{\sigma}(x)$, and the eigenvalues of the channel covariance matrix, $ p_{\lambda}(x)$,  given by $p_{\sigma}(x^2) = p_{\lambda}(x)$.
+Defining a new, symmetric density function of the singular value distribution
+ 	 \begin{equation}\label{symmetric}
+ 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2},
+ 	 \end{equation}
+ 	 we now substitute $\tilde{p}_{\sigma}(x)$ into the definition of the Stieljes transform \eqref{stieltjes}
+ 	 to give the relationship
+ 	 	 	 \begin{align*}
+	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}dx
+	 	 \\&  =  
+	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
+	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
+	 	 \\&  =  
+	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
+	  	 \\&  =  
+	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
+ 	 	  	 \\&  =  
+ 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
+ 	 	 \\&  =  
+			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
+	 	 \\&  = 
+	 	 	-s G_{\lambda}(s^2)
+	 	 \end{align*}
+With the R-transform definition unchanged for the symmetric distribution, and the additional useful property (CITE)
+\begin{equation}
+\tilde{R}_{\mH_{Total}}(w) = \tilde{R}_{\mH_{2}}(w) + \tilde{R}_{\mH_{1}}
+\end{equation}
+We find that we only need the Stieltjes transform for the covariance matrix of each component of the sum which can in general be found using the tools
+of multiplicative free convolution.
+
+\item 
+Proof of canceling phases.\\
+Consider the channel given by 
+\begin{equation}
+\mathbf{H}_{Total} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}.
+\end{equation}
+This product of matrices suggests the use of multiplicative free convolution to solve for the AED using the S-Transform. 
+Following the same procedure show in \cite{muller2002asymptotic}, we begin with the covariance matrix
+\begin{equation}
+\mC_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H
+\end{equation}
+and note that this has the same non-zero eigenvalues as the matrix (Proof)
+\begin{equation}
+\tilde{\mC}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H.
+\end{equation}
+The relationship between the S-Transform of $\mC_2$ and $\tilde{\mC}_2$ is given by 
+\begin{equation}\label{rotation_property}
+S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
+\end{equation}
+With \begin{equation}
+\mC_1 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
+\end{equation}
+we see that $S_{\tilde{\mC}_2}(z) = S_{\mC_1}(z) S_{\mathbf{H}_{1}\mathbf{H}_{1}^H}(z)$.  Repeating the above procedure on $\mC_1$ we find
+\begin{equation}
+\tilde{\mC}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
+\end{equation}
+with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$ we are left with 
+$S_{\mC_1}(z) =  S_{\mathbf{H}_{2}^H\mathbf{H}_{2}}(z)$. As a result we see that 
+the phase matrix cancels. The same canceling applies to any unitary matrix $\mathbf{A}$ such that   $\mathbf{A}\mathbf{A}^\dagger =\mathbf{A}^\dagger \mathbf{A} = \mI$.
 \end{enumerate}
-
 \bibliography{bibliography}

### 2020-06-15 documentation/Paper/correlation_review.tex
finishing correlation background and moving onto RIS optimization background

\par
Investigation leading to the current interest in MIMO systems showed the key property that (with constant transmit power), capacity (and mutual information) in an $N \times N$ MIMO system with independent channels scales linearly in $N$ \cite{foschini1998limits}. However the assumption of independent channels between each pair of transmit and receive antennas is not generally reasonable. This prompted much investigation into the impact of correlation to this scaling. In the following, a number of studies on this are reviewed. The chosen works use for channel correlation models similar to the one we will develop for our system model. In particular, this models the correlation of a linear array for which correlation decreases exponentially with distance.
\par
1. Looking at correlation for finite size systems.
In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?). (Is this a bigger problem for higher frequencies?).
Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
\par
One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. 
This is motivated by the observation that the eigenvalues of a channel covariance matrix converge to a deterministic distribution for many random 
matrices as  $N \rightarrow \infty$. In the remainder, this will be referred to as the Asymptotic Eigenvalue Distribution (AED). 
By making this assumption the impact of $N$ is only a scaling for some deterministic value. 
\par
In \cite{loyka2001channel}, an upper bound on the capacity is found using Jensen's inequality. For this upper bound, increasing correlation between antennas can be interpreted as a decrease in the average receiver SNR. 
Similarly, \cite{skupch2005free} used multiplicative free convolution from free probability theory to analyze the exponential correlation model at either the transmitting or receiving antenna. Part of this result will be used in our analysis.

\cite{taricco2008asymptotic} analyzed the correlated ricean fading channel using the Replica Method. 
In general it is observed that the slope of the linear scaling of (mutual information) is reduced by the addition of correlation. 
One important point to note is that the above works do not consider the case of CSIT. When the water-filling capacity of the 
correlated MIMO channel is considered,  \cite{chuah2002capacity} notes an exception to this trend. For the case with low
SNR (what does this mean?) the water-filling capacity is actually increased with correlation.
\par
Before moving on to use any of these asymptomatic results, it is fitting to consider whether or not the assumption of  $N \rightarrow \infty$ is reasonable. Intuitively, we would expect that under correlation this convergence may occur only at very high $N$. This result was confirmed in  \cite{martin2004asymptotic} using the exponential correlation model at both the transmitter and receiver.

### 2020-06-15 documentation/Paper/paper.tex
finishing correlation background and moving onto RIS optimization background

@@
 General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
-Need to mention anyone who has considered correlation at end in order to transition. 
+
 \par
-Early investigation leading to the great interest in MIMO systems showed the key property of capacity in an $N \times N$ MIMO system scaling linearly in $N$ (teletar).
-While this scaling applies to many models for the channel (cite) one key property that can impact this scaling is correlation between channels of a MIMO system.
-In order to evaluate MIMO system performance in a real-world application, correlation between between the channels must be accounted for.
-In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?).
-(Is this a bigger problem for higher frequencies?).
-Correlations cause reduction in the capacity of the system and often make the calculation of the capacity more difficult.
-However, with the possibility for arrays with (x) co-located antenna and RIS with (y) individually controlled reflective elements, the impact of channel correlation on system performance must be well understood before deployment.
+\input{correlation_review}
 \par
-Detail all of the work looking at correlated MIMO. 
-1. Looking at correlation for finite size systems.
-Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
-2. Extending to asymptotic system sizes. 
-One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. This method has been used 
-extensively since. 
-Now list of the different types of channels/methods that have been used here. 
-\cite{loyka2001channel}
-\par
-From the above, it should be clear that investigating the RIS  system for 
-asymptotically large MIMO systems with correlation, an topic not yet explicitly investigated to the author's knowledge, is relevant to the design of upcoming wireless systems.
+Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
 
@@
 where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
+Notice that in this form, the linear scaling of capacity with N is clear.
 Expanding the expectation gives the expression

### 2020-06-15 simulation/MATLAB/fixed_point_test.m

@@ -3,16 +3,19 @@ cols = 100;
 steps = 100;
 bottom = .001;
 top = 10; 
-channel = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
-rho = .9;
-correlation = exponential_correlation(rows, rho);
-total_cov = (correlation*(channel*channel')*correlation');
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
 s_values = x_values + 1i*1e-6;
-% stieltjes_values = (1./s_values).*(1+gamma_s(s_values));
-stieltjes_values = marcenko_pastur(s_values,1);
+channel = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
+rho = .98;
+correlation = exponential_correlation(rows, rho);
+% check = correlation*correlation';
+total_cov = (correlation*(channel*channel')*correlation');
+% stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values));
+stieltjes_values = (1./s_values).*(1+gamma_s(s_values));
+% stieltjes_values = marcenko_pastur(s_values,1);
 pdf = 1/pi .* imag(stieltjes_values);
+% pdf = imag(stieltjes_values);
 num_capacity = MIMO_capacity(total_cov, 1/cols)
 asymptotic_capacity  = aed_capacity(x_values, pdf, 1/cols, rows, step_size)
 plot(x_values, pdf)
@@ -27,43 +30,31 @@ function output = gamma_s(input)
     output = out;
 end
 
-function x = inverse_gamma_cor(eval_point)
-    rho = .9;
-    alpha = (1+rho^2)/(1-rho^2);
-    val = (alpha*z^2 + z*sqrt(alpha^2*z^2 - (z^2 -1)))/((z^2-1)*(z+1))
-end
-
 function x = new(eval_point)
-init_point = rand() + .1j*rand();
-rho = .5;
+init_point = rand() + 1j*abs(rand());
+rho = .1;
 alpha = (1+rho^2)/(1-rho^2);
+beta =1;
 x = fsolve(@root1,init_point);
-    function F = root1(gamma_inv_s)
-        F = ((gamma_inv_s + 1)^2*(gamma_inv_s^2 - 1))/(eval_point^2) ...
-            - gamma_inv_s^2*(2*alpha*(gamma_inv_s + 1)/eval_point - 1);
+    function F = root1(z)
[…]

### 2020-06-24 documentation/Meeting Notes/06_24_20.tex
report for update and some comparison scripts

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Summary}
I found a variety of different methods for considering correlation in the literature (especially for the exponential model). I will detail these methods in the next section. First I want to note 2 particular points that I think may be relevant. 
\begin{enumerate}
\item
	Intuitively, adding correlation will slow the rate of convergence towards an asymptotic AED. 
	In \cite[Fig. 2]{martin2004asymptotic} the asymptotic capacity for a MIMO channel with the exponential correlation model is found (by  using methods for estimating the AED of a sample covariance matrix) and compared with numerical results for systems of increasing size. The convergence of the numerical result to the asymptotic result is shown to slow significantly under realistic correlation values ($\rho = 0.7$ with the exponential model). This is important because even though I can show that the phase matrix at the IRS ,$\phase$, will cancel for the asymptotic case, the choice of $\phase$ will still be important for small system sizes or systems with high correlation.
	It  might be useful to give an estimate for the system size needed in order for the asymptotic assumption to hold even when the channel is correlated with the exponential correlation model.

\item	
	A second interesting note is from \cite{chuah2002capacity} which investigates a \emph{unitarily invariant} MIMO channel (eg $\CNO$) with the exponential correlation model. 
	This work compares both the mutual information and capacity (with water-filling) of the system. It is shown that for a low SNR system, increasing the correlation ($\rho$)
	actually increases the capacity. In contrast, increasing correlation always reduces capacity for equal power allocation and for the water-filling case with high SNR.

\end{enumerate}
\section{Methods for Analyzing Correlation}
\subsection{Review of Empirical Results}
First I reviewed MIMO correlation measurement studies to get some idea of realistic correlation coefficient that I can use to parameterize the exponential model. 
\begin{enumerate}
\item
In test data collected for a point-to-point, $4 \times 4$ MIMO system  operating at 1.9 GHz (30 kHz bandwidth), \cite{martin2000multiple} showed correlation coefficient between channels could be as high as $0.5$. 
The measurements in \cite{lee1973effects} show similar results for channels at 836 MHz over varying antenna spacing. This work indicates that the above correlation coefficient of $0.5$ is indeed reasonable and that much higher correlation coefficients are observed even when antennas are distanced by multiple wavelengths.

\end{enumerate}
To Discuss: I could not find correlation measurement results for the exponential model. There seems to be a lot of measurement work looking at the capacity with correlation and some that measure the correlation coefficients but I did not find work in which measurements were used to parameterize any correlation models.

\subsection{Review of Asymptotic Results}
Next I reviewed (and tried to simulate) works for asymptotic mutual information and capacity for the correlated MIMO channel. I've outlined some of the different approaches I case across below.
\begin{enumerate}

\item
In \cite{loyka2001channel}, a deterministic and normalized channel with the exponential correlation model is investigated. An upper bound on the capacity is found using Jensen's inequality and an manipulated representation of the determinant. Using this upper bound, increasing correlation between antennas can be interpreted as a decrease in the average receiver SNR. 
\par
To Discuss: This would be an appropriate method to investigating just the impact of correlation but may not work for investigating other aspects of the system.

\item
In \cite{skupch2005free} multiplicative free convolution from free probability theory is used to analyze the exponential correlation model at the transmitting or receiving antenna. In their free probability framework, \cite{skupch2005free} provides a closed form expression for the S-transform of the matrix representing the exponential correlation model (for a linear array). This should be exactly what I need. However, when I use this S-Transform equation I get inconsistent results. In particular, when I attempt to solve for $G(s)$ (the stieltjes transform) from $\gamma(s)$ using the inverse S-Transform, the results are clearly incorrect and not consistent with those in the paper. 
\par
To Discuss: Difference between my approach and what is used in the paper.

\item 
As mentioned above  \cite{chuah2002capacity} also considered the MIMO channel with unitarily invariant elements and the exponential correlation model at both the transmitter
and receiver. The problem with this model is that it will not extend to the case of a LOS component since the channel elements would no longer be unitarily invariant.
This approach is based primarily on results from random matrix theory for random determinants but also uses some concepts of free probability.

\item 
In \cite{taricco2008asymptotic} and \cite{moustakas2003mimo} the MIMO channel with correlated ricean (LOS) fading elements is analyzed using the Replica Method to find the first and second moments of the capacity. I haven't tried to repeat the results from either of these methods yet because they seem quite complicated.

\end{enumerate}
To summarize, I can previous work for asymptotic results for similar system models but not my exact system model. 
In the next section I detail a couple of resulting questions.
\section{Questions to Discuss }
\begin{enumerate}
\item
Since we can show that the IRS phase contribution will cancel out in any asymptotic system, the resulting channel is just a correlated MIMO channel with a set of scatterers and a LOS component that may be low rank. The problem is, none of the above results use this system model. Should I continue trying to find results for the correlated channel that is relevant to this system model even though there is no further insight to the IRS system?
\item
If power allocation becomes more important as the system size increases (for the low SNR case), is it worthwhile to find results for the equal power allocation case? 
\item
I noticed that in the works looking at IRS phase optimization, it is always assumed that power allocation can also be performed. Is it worthwhile to look at the realizable rates using phase optimization without power allocation and compare this to the asymptotic, equal power allocation case in which the phases cancel?
\end{enumerate}
\bibliography{bibliography}

\end{document}

### 2020-06-24 documentation/Paper/abstract.tex
report for update and some comparison scripts

--- 2020-06-08 documentation/Paper/paper.tex
@@
-\documentclass[12pt,a4paper]{report}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{float}
-\input defs.tex
-\bibliographystyle{alpha}
-\graphicspath{ {./figures/} }
-%\hyphenpenalty=10000
-
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
-\author{Peter Hartig}
-
-\begin{document}
-\maketitle
-\begin{abstract}
 This work investigates the impact of passive reflective surfaces on the capacity of MIMO systems.
@@
 Finally we compare the capacity of the correlated channel with optimized phases at the IRS to the uncorrelated channel to show that....?
-\end{abstract}
-%
-%\newpage
-\tableofcontents
 
-\chapter{Introduction}
-Wireless communication has progressively moved to higher frequencies in recent standards in order to utilize large bands of unoccupied resource
-and to enable increased antenna density \cite{akyildiz2018combating}.
-A primary hurdle towards communicating over higher frequencies is the increased radio frequency energy (power) attenuation.
-For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
-has been shown to exceed y.
-Given that increasing current wireless communication capacity and reliability rely on moving to these high frequencies, mitigating or countering these power losses on the order of 100 dB is highly relevant.
-In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
-and supplementing the transmission power using relays (CITE).
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitter, and the power used for transmission.
-(Would be nice to cite some actual costs here).
-As the power attenuation problem has been increased by X-Fold moving from freq1 to freq2, new solutions have been considered.
-One solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-\par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-The absence of an rf-chain in the RIS offers additional benefits over relays.
-First, reflection does not introduce the noise associated with the rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum.
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?).
-For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
-(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
-Example of how these surfaces are implemented (Cite).
-Discuss how it has be looked at from MISO and MIMO perspective.
-\par
-One area that has not be investigated, (at least up to the author's knowledge) is the impact of the RIS for 
-asymptotically large MIMO systems. In the following, the impact of the RIS on a general MIMO system is explored.
-
-\section{Notation}
-The following notation is used throughout the remainder. 
-$\Expect[x]$ is the expected value of a random variable $x$.
-Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
-The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
-$\mI_N$ denotes the identity matrix of dimension $N$.
-For square $\mX$ with dimension $N\times N$ the normalized trace is denoted by $\Tr(\mX)  = \underset{N \rightarrow \infty}{\text{lim}}
-\Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
-
-\section{Outline}
-The following sections are now previewed.
-In section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
-
-\chapter{System Model}\label{system_model}
-\section{The Correlated IRS Channel}
-
-\section{Mutual Information}\label{sectiond:mut_info}
-A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel $\mHt$ is known at the receiver, the mutual information is given by
-\begin{equation}\label{mut_ent}
-I(\vy;\vx) = h(\vy) - h(\vy|\vx).
-\end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
-\begin{equation}\label{entropy}
-h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
-\end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
-
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect[\vn\vn^H] = \mI$, substituting equation \eqref{entropy} into equation
-\eqref{mut_ent} gives
-\begin{equation}\label{mut_inf}
-\Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to the
-transmit covariance matrix and the phases at the IRS gives and expression for the capacity of the system
-
-\begin{equation}\label{capacity}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
-
-If $\mHt$ is not known at the transmitter, the transmit power can be allocated
-equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
-\end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-Expanding the expectation gives the expression
-\begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
-\end{equation}
-in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
-
-Note that because $\Log$ is a concave function, we can upper bound the mutual information with
-\begin{equation}\label{mut_inf}
-\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda_i])
-\end{equation}
-Now discuss capacity and consider equal power but optimize over phase
-We now want to consider the case in
-
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{Proof of Freeness}
-\begin{itemize}
-
-\item 
-	A brief introduction to free probability.
-	\par 
-	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
-	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
-	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
-	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
-	for this investigation but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
-	 \begin{itemize}
-	 \item 
-	 	The cornerstone of free probability is the concept freeness of two random variables in the context of a projection. In our case, freeness between two 
-	 	matrices is given by 
-	 	We now introduce useful tools from free probability  that can be applied to free random variables.
-
-	 	\item 
-	 	Stieltjes Transform: 
-	 	Generally, the projection of matrix is not given in closed form. The projection can be uniquely represented using
-		\begin{equation}\label{stieltjes}
-		G(s) = \int{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
-		\end{equation}	
-		This representation will allow us to find the projection of the sum and project of matrices using addition and multiplication.
-		\item 
-		R-Transform:
-		Mention self-adjoint
-		Given the Stieltjes transform of a RV, the corresponding R-Transform given by 
-		X. 
-		The R-Transform of the sum of two, free RVs can be found using
-		Y
-		This process use known as free additive convolution.
-		\item
-		S-Transform:
-		Mention self-adjoint
-		Similarly, given the Stieltjes transform of a RV, the S-Transform given by 
-		X, X;
-		The S-Transform of the product of two, free RVs can be found using
-		Y
-		This process is known as free multiplicative convolution.
-	 	\item 
-	 	If there is a case in which finding the singular value distribution is easier than the eigenvalue distribution we can use equation  
-	 	 \eqref{sigular_eigen} to move between the two. Considering a symmetric function of the singular value distribution
-	 	 \begin{equation}\label{symmetric}
-	 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2}
-	 	 \end{equation}
-	 	 the definition of the Steiltjes transform gives the following steps, leading to a relationship between the Stieltjes transforms. 
-	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-	 	\item 
-	 		Another useful property that will be used is the relationship between the eigenvalues of a matrix and its rotation.
-	 		Specifically, consider the covariance matrix given by $\mH_2\mH_1 \mH1^H \mH2^H$.
-	 		By the rotational property of the trace, 
-	 		\begin{align*}
-	 		trace({\mH_2\mH_1 \mH1^H \mH2^H }) = \trace({\mH_1^H \mH_2^H \mH_2 \mH_1 }).
-	 		\end{align*}
-	 		... Not sure if I need to prove this here.  
-	 	\item 
-	  	build up to show that phases cancel via s-transform
-	 	Stieltjes of AED (IE for hermetian matrices)
-	 	Stieltjes of SVD (when they are components covariance matrix)
-	 	Getting Steiltes of SVD using R-Transform
-	 	Finding R-Transform of Individual components
-	 	Using Stieltjes of individuals to get R-transforms
-	 	Using aed stieltjes to get svd individual
-	 	Using S-transform to get stieltjes of individual aed.
-	\end{itemize}
-
-	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
-
-\item 
-	Show that at least for uncorrelated case, the matrices in the sum are free.
-
-\item 
-	Give proofs in an appendix?
-
-\end{itemize}
-
-\chapter{Results}\label{Results}
-\section{Canceling Phases}
-We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
-phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
-By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
-the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
-\par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity, the rate of this convergence is particularly relevant in this application as this implies the number of rf-chains at the transmitter and receiver approach infinity.
-One notable observation is that this rate
-decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
-\begin{itemize}
-\item
-
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
-\end{itemize}
-
-\chapter{Conclusion}
-
-\chapter{Appendix}
-Proofs
-\begin{enumerate}
-\item
-	Prove freeness of the relevant points. 
-\item 
-	rotations have same non-zero eigenvalues
-\end{enumerate}
-
-\bibliography{bibliography}
-\end{document}
+This work provides an upper bound on the capacity of the RIS assisted MIMO system.
+This upper bound is a useful tool with which to analyze the performance of the various sub-optimal optimization approaches used.
+Question: would the capacity of the finite , phase-optimized system with correlation ever exceed that of the asymptotic case?

### 2020-06-24 documentation/Paper/correlation_review.tex
report for update and some comparison scripts

@@
 This prompted much investigation into the impact of correlation to this scaling.
-In the following, a number of studies on this are reviewed.
-The chosen works use for channel correlation models similar to the one we will develop for our system model.
-In particular, this models the correlation of a linear array for which correlation decreases exponentially with distance.
+In the following, a number of these studies are reviewed.
+These chosen works use channel correlation models similar to the one we will develop for our system model.
+In particular, this models the correlation between antennas as decreasing  exponentially with the distance between antennas.
+Definitely need to show this is a valid model (Cite).
 \par

### 2020-06-24 documentation/Paper/paper.tex
report for update and some comparison scripts

@@
 \begin{abstract}
-This work investigates the impact of passive reflective surfaces on the capacity of MIMO systems.
-In particular, the system is analyzed in the asymptotic size domain in order to employ the tools of free probability.
-A system is considered first in the case of fully uncorrelated channels.
-In this case it is shown that the passive control of such a reflective surface has no impact on the capacity of the system.
-The system is then generalized to allow for correlation within the channel.
-In this case is it shown that the selection of the phases at the IRS still has an impact on the system.
-Finally we compare the capacity of the correlated channel with optimized phases at the IRS to the uncorrelated channel to show that....?
+\input{abstract}
 \end{abstract}
@@
 \tableofcontents
-
 \chapter{Introduction}
 \section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards, projected to reach even the Terahertz range by the 6G standard (CITE).
+The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards.
+Industry research for 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
 This expansion is motivated by large bands of currently unused frequency
 and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-A primary hurdle towards efficient communication at these frequencies is the increased radio frequency (power) attenuation.
-For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
-has been shown to exceed 100 dB.
-With increases in current wireless system capacity and reliability relying on the use of these high frequencies, mitigating or countering this power loss  is highly relevant.
+A primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
+For frequency bands relevant for upcoming wireless standards, energy attenuation due to free space path loss  molecular absorption loss
+has been shown to exceed 100 dB. Countering this loss is a key steps towards meeting the increased data rates and reliability required for future networks.
 \par
-In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
-and supplementing the transmission power using relays (CITE).
+In previous generations of wireless communications, the problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
+and supplementing the transmission power using relays \cite{dahlman20134g}.
 Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
@@
 Deploying this solution on a scale to counter
-the attenuation levels relevant to new standards would pose impractical costs.
+the attenuation levels relevant to new standards would pose impractical costs (why).
 One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
@@
 \par
-Now discuss all of the previous work looking at the optimization of the phases. 
-General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
-
+Need to show how many surfaces are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
+\par
+The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will requires systems 
+sizes on the order of x. Now motivate using asymptotic analysis.
 \par
@@
 Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
+
+\par
+Now discuss all of the previous work looking at the optimization of the phases. 
+General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
 

### 2020-06-24 simulation/testing/channel_testing/increasing_antennas.py

@@ -57,11 +57,8 @@ for val in param_list:
         irs_svd.append(s_val_total_irs)
         capacities.append(capacity(total_irs, 1/cols))
         water_capacity.append(capacity_water_filled(Q_x))
-
-        Q_H = water_filling(H, 1)
-        Q_G = water_filling(G, 1)
-        H_capacity.append(capacity_water_filled(H@Q_H@hermetian(H)))
-        G_capacity.append(capacity_water_filled(G@Q_G@hermetian(G)))
+        H_capacity.append(capacity_water_filled(water_filling(H, 1)))
+        G_capacity.append(capacity_water_filled(water_filling(G, 1)))
 
 # Compare Estimated to True Capacity
     numeric_capacity_list.append(np.average(capacities))

### 2020-06-29 simulation/MATLAB/fixed_point_test.m
corrected fixed point so results seem better but still needs some work

@@ -6,25 +6,35 @@ top = 10;
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
 s_values = x_values + 1i*1e-6;
-channel = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
-rho = .98;
-correlation = exponential_correlation(rows, rho);
-% check = correlation*correlation';
-total_cov = (correlation*(channel*channel')*correlation');
-stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values));
-% stieltjes_values = (1./s_values).*(1+gamma_s(s_values));
-% stieltjes_values = marcenko_pastur(s_values,1);
-pdf = 1/pi .* imag(stieltjes_values);
-% pdf = imag(stieltjes_values);
-num_capacity = MIMO_capacity(total_cov, 1/cols)
-asymptotic_capacity  = aed_capacity(x_values, pdf, 1/cols, rows, step_size)
-plot(x_values, pdf)
-title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)]);
+rhos = linspace(.5,.99,5);
+numeric_capacity = zeros(size(rhos));
+asymptotic_capacity = zeros(size(rhos));
+for i = 1:length(rhos)
+    stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, rhos(i)));
+    % stieltjes_values = (1./s_values).*(1+gamma_s(s_values));
+    % stieltjes_values = marcenko_pastur(s_values,1);
+    pdf = 1/pi .* imag(stieltjes_values);
+    % pdf = imag(stieltjes_values);
+    numeric_capacity(i) = average_numeric_capacity(rows, cols, rhos(i), 20);
+    asymptotic_capacity(i)  = aed_capacity(x_values, pdf, 1/cols, rows, step_size);
+%     plot(x_values, pdf)
+%     title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)]);
+end
 
-function output = gamma_s(input)
+plot(rhos, numeric_capacity, '-s')
+hold on
+plot(rhos, asymptotic_capacity, '-o')
+legend('numeric_capacity','asymptotic_capacity')
+title('Numeric Vs. Asyptotic Correlated Capacity');
+
+
+
+
+
+function output = gamma_s(input, rho)
     out = zeros(size(input), 'like', input);
     for i = 1:length(input)
[…]

### 2020-06-29 simulation/testing/channel_testing/mixed_terms_check.py

@@ -23,12 +23,14 @@ los_path = []
 cross_sum = 0
 main_sum = 0
 
-average = 5
+average = 50
 for i in range(average):
     # H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
     H = c_rand(rows, cols) @ np.diag(np.exp(1j * np.random.uniform(0, 2 * np.pi, rows))) @ c_rand(rows, cols)
     # G = c_rand(rows, cols, mean=.1)
-    G = c_rand(rows, cols)
+    G = np.ones((rows, cols))*.05   
+    # G = np.zeros((rows, cols))*-.01
+    # G = c_rand(rows, cols)
     GG = G @ np.conj(G.T)
     HH = H @ np.conj(H.T)
     GH = G @ np.conj(H.T)

### 2020-07-01 documentation/Presentations/Mid-Presentation/mid_term_presentation.tex
prep for OFDM check

\documentclass[10pt,tgadventor, onlymath]{beamer}

\usepackage{graphicx,amsmath,amssymb,tikz,psfrag,neuralnetwork, fontawesome, xcolor}

%\input defs.tex
\graphicspath{ {./figures/} }
\input defs.tex

%% formatting

\mode<presentation>
{
\usetheme{default}
\usetheme{Berlin}

\usecolortheme{seahorse}
}
\setbeamertemplate{navigation symbols}{}
\usecolortheme[rgb={0.03,0.28,0.59}]{structure}
\setbeamertemplate{itemize subitem}{--}
\setbeamertemplate{frametitle} {
	\begin{center}
	  {\large\bf \insertframetitle}
	\end{center}
}

\newcommand\footlineon{
  \setbeamertemplate{footline} {
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,leftskip=.8cm,rightskip=.6cm]{structure}
      \footnotesize \insertsection
      \hfill
      {\insertframenumber}
    \end{beamercolorbox}
    \vskip 0.45cm
  }
}
\footlineon

\AtBeginSection[] 
{ 
	\begin{frame}<beamer> 
		\frametitle{Outline} 
		\tableofcontents[currentsection,currentsubsection] 
	\end{frame} 
}

\tikzstyle{state}=[shape=circle,draw=blue!30,fill=blue!10]
\tikzstyle{observation}=[shape=rectangle,draw=orange!30,fill=orange!10]
\tikzstyle{lightedge}=[<-, dashed]
\tikzstyle{mainstate}=[state, thick]
\tikzstyle{mainedge}=[<-, thick]
\tikzstyle{block} = [draw,rectangle,thick,minimum height=2em,minimum width=2em]
\tikzstyle{sum} = [draw,circle,inner sep=0mm,minimum size=2mm]
\tikzstyle{connector} = [->,thick]
\tikzstyle{line} = [thick]
\tikzstyle{branch} = [circle,inner sep=0pt,minimum size=1mm,fill=black,draw=black]
\tikzstyle{guide} = []
\tikzstyle{snakeline} = [connector, decorate, decoration={pre length=0.2cm,
                         post length=0.2cm, snake, amplitude=.4mm,
                         segment length=2mm},thick, magenta, ->]

%% begin presentation

\title{\large \bfseries Asymptotic Capacity of Systems with Intelligent Reflective Surfaces}
\author{Peter Hartig \\ \and Supervisor: Saba Asaad}

\date{\today}

\begin{document}

\frame{
\thispagestyle{empty}
\titlepage
}

\section{Background}

\begin{frame}
\frametitle{Relaying}

	\begin{itemize}
		\item 			
			Base Stations Vs. Relay $\rightarrow$ reduced core network footprint/cost
%		\item 
%			Modes of operation provide cost/performance trade-off (decode and forward, amplify and forward ...)
		\item 
			Extending wireless coverage is relevant to high frequencies used in 5G +
	\end{itemize}

\end{frame}

\begin{frame}
\frametitle{The basic MIMO channel}
Much easier if each channel has a constant frequency response over the relevant band.
	\begin{itemize}
		\item

	\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Relay vs. Intelligent Reflective Surface}
\begin{columns}
\begin{column}{0.5\linewidth}
\centering 
	\underline{Relay: Amplify and Forward}
	\\
	\begin{equation*}
	\mathbf{y} = (\mathbf{H}_2\underbrace{\mathbf{F}}_{\text{Relay}}\mathbf{H}_1 + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{x}
	\end{equation*}

	\begin{itemize}
	\item 
		$\mathbf{F}$ is diagonal
	\item 
		Requires expensive RF hardware
	\item 
		Can often ignore line of sight
	\item 
		Typically half-duplex
	\end{itemize}
\end{column}
\begin{column}{0.5\linewidth}
\centering

	\underline{IRS}
	\\
	\begin{equation*}
	\mathbf{y} = (\mathbf{H}_2\underbrace{\boldsymbol{\Phi}}_{\text{IRS}}\mathbf{H}_1 + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{x}
	\end{equation*} 
	\begin{itemize}
	\item 
		$\boldsymbol{\Phi}$ is diagonal with $| \phi_i | =1$
	\item 
		No RF chain required (low cost)
	\item
		Shorter distances $\rightarrow$ keep LOS
	\item 
		Inherently full-duplex
	\end{itemize}
\end{column}
\end{columns}

\end{frame}

%\begin{frame}
%\frametitle{IRS: Modes of Operation}
%\begin{columns}
%\begin{column}{0.5\linewidth}
%	Mode 1
%	\begin{figure}
%%			\centering
%		\includegraphics[scale=1]{irs}
%	\end{figure}\end{column}
%\begin{column}{0.5\linewidth}
%	Mode 2 (The one we will use)
%\end{column}
%\end{columns}
%\end{frame}
\begin{frame}
\frametitle{System Model}
\centering
\begin{equation*}
	\mathbf{y} = (\mathbf{H}_2\boldsymbol{\Phi}\mathbf{H}_1 + \mathbf{G})\mathbf{x}
\end{equation*}
with 
$\mathbf{H}_{1}\in \mathbb{C}_{S \times T},\mathbf{H}_{2} \in \mathbb{C}_{R \times S}, \mathbf{G} \in \mathbb{C}_{R \times T}$ and $\boldsymbol{\Phi}$ as $S \times S$ a diagonal matrix with $| \phi_i | =1$.

	\begin{figure}
		\centering
		\includegraphics[scale=.3]{irs_correct}
	\end{figure}
\end{frame}

\section{Goals}
\begin{frame}
\frametitle{Project Goals}
\begin{enumerate}
\item
	Find capacity of asymptotic IRS systems
	\begin{itemize}
	\item 
		Optimize phase matrix
	\item
		Generalize results
	\end{itemize}

\item 
	Confirm findings with numerical results throughout

\item
	* Time permitting - Consider secrecy capacity of IRS system
\end{enumerate}
\end{frame}

\section{Current Status}

\begin{frame}
\frametitle{Channel Capacity}
\begin{equation*}
	\mathbf{y} = (\mathbf{H}_2\boldsymbol{\Phi}\mathbf{H}_1 + \mathbf{G})\mathbf{x}
\end{equation*}

With $\mathbf{x}\mathbf{x}^H = \mathbf{I}$ and fixed $\boldsymbol{\Phi}$, capacity is given by 
\begin{equation*}\label{capacity}
\Expect\left[\Log\left(|\mathbf{I}_{R}+\frac{P_{\text{Total}}}{T \sigma_n}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}][\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}]^H|\right)\right].
\end{equation*}
Using 
\begin{equation*}
\mathbf{H}_{\text{Total}} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1} + \mathbf{G}
\end{equation*}
and simplifying, \eqref{capacity} becomes
\begin{equation*}
N \int_0^{\infty}\Log\left(1+\frac{P_{\text{Total}}}{T \sigma_n} \lambda_{\mathbf{H}_{\text{Total}}\mathbf{H}_{\text{Total}}^H}\right) \textcolor{red}{p_{\lambda_{\mathbf{H}_{\text{Total}}\mathbf{H}_{\text{Total}}^H}}(\lambda)} d\lambda
\end{equation*}
with $N$ = rank($\mathbf{H}_{\text{Total}}$).
Need Asymptotic Eigenvalue Distribution (AED).
\end{frame}

\subsubsection{Previous Work + Free Probability}

\begin{frame}
\frametitle{Using Free Probability to find $p_{\lambda_{\mathbf{H}_{\text{Total}}\mathbf{H}_{\text{Total}}^H}}(\lambda)$}
Free probability provides a set of tools to find the AED of polynomials in free, self-adjoint operators (matrices).

\begin{itemize}
\item 
	For $\mathbf{A}$ and $\textbf{B}$ consider $\mathbf{C} = \mathbf{A} + \textbf{B}$ and $\mathbf{D} = \mathbf{A}\textbf{B}$
\item 
	Additive free convolution: $R_{\mathbf{C}}(w) = R_{\mathbf{A}}(w) + R_{\mathbf{B}}(w)$
\item 
	Multiplicative free convolution: $S_{\mathbf{D}}(z) = S_{\mathbf{A}}(z)S_{\mathbf{B}}(z)$
\end{itemize}

\begin{equation*}
\mathbf{H}_{\text{Total}}\mathbf{H}_{\text{Total}}^H 
= 
\underbrace{
\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H +
\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{G}^H}_{\color{red}{\text{non-hermetian}}}+
\underbrace{\mathbf{G}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H}_{\color{red}{\text{non-hermetian}}}+
\mathbf{G}\mathbf{G}^H
}_
{\color{red}{\text{not free}}}
\end{equation*}
%Need to show that these are not free!
\end{frame}

\begin{frame}
\frametitle{Free Probability}
The symmetrized density function.
\begin{equation*}
\tilde{p}(x) = \frac{p(x) + p(-x)}{2}
\end{equation*}
The symmetrized singular value, eigenvalue relationship.
\begin{equation*}\label{svd_aed_property}
\tilde{G}_{H}(s) = sG_{HH^{\dagger}}(s^2).
\end{equation*}
Additive free convolution still applies for singular values.
We only need the self-adjoint terms of the polynomial!
\end{frame}

\begin{frame}
\frametitle{Free Probability: Canceling Phase Terms}
Say 
\begin{equation*}
C_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}[\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}]^H
\end{equation*}
and a rotated version 
\begin{equation*}
\tilde{C}_2 = \underbrace{\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}}_{C_{1}}\mathbf{H}_{1}\mathbf{H}_{1}^H.
\end{equation*}
Use the property 
\begin{equation*}\label{rotation_property}
S_{C_2}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_2}(\frac{z}{\chi_2}).
\end{equation*}
\pause
Repeat for
\begin{equation*}
C_{1} = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
\end{equation*}
\begin{equation*}
\tilde{C}_{1} = \textcolor{red}{\boldsymbol{\Phi}\boldsymbol{\Phi}^H}\mathbf{H}_{2}^H\mathbf{H}_{2}
\end{equation*}
with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$.
\end{frame}

\begin{frame}
	\frametitle{Capacity of constant channels with random phase
	}
	\centering 
	$\mathbf{H}_{2} \in \mathbb{C}_{R \times T}$, 
	$\mathbf{H}_{1} \in \mathbb{C}_{S \times T}$, with $\mathcal{CN}(0,\frac{1}{\sqrt{RS}})$
	and $ \mathbf{G} \in \mathbb{C}_{R \times T}$ with $\mathcal{CN}(0,\frac{1}{R})$. 
	Statistics from $1000$ random realizations of $\boldsymbol{\Phi}$.
	\bigskip

\begin{columns}
\begin{column}{0.5\linewidth}
\centering
Asymptotic case \\ $T = 100$, $ R = 100$ and $S = 100$
\begin{itemize}
\item
Capacity variance:
 0.0038458296727919886
 \item
Capacity average:
 6.764651253343401
 \item
Capacity min:
 6.574540172102809
 \item
Capacity max:
 6.98960781887363
\end{itemize}
%\textcolor{red}{\chi_N}}
\end{column}
\begin{column}{0.5\linewidth}
\centering
Non-asymptotic case\\ $T = R = 4$ and $S = 100$
\begin{itemize}
\item
Capacity variance:
 0.4157628312024758
 \item
Capacity average:
 4.014673474337778
 \item
Capacity min:
 2.146914646653352
 \item
Capacity max:
 6.130275960052021
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Numerical Results: Asymptotic Capacity}

	\begin{equation*}
	\mathbf{y} = (\mathbf{H}_2\underbrace{\boldsymbol{\Phi}}_{\text{IRS}}\mathbf{H}_1 + \underbrace{\mathbf{G}}_{\text{LOS: Rank $\leq 1$}})\mathbf{x}
	\end{equation*} 
\begin{figure}
\includegraphics[width= .75\linewidth, height = 5.5cm]{los}
\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Numerical Results: Asymptotic Capacity}
		\begin{equation*}
	\mathbf{y} = (\underbrace{\sigma}_{\text{attenuation}}\mathbf{H}_2\underbrace{\boldsymbol{\Phi}}_{\text{IRS}}\mathbf{H}_1 + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{x}
	\end{equation*} 
\begin{figure}
\includegraphics[width= .75\linewidth, height = 5.5cm]{attenuation}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Current Status}
\begin{enumerate}
\item
	Find capacity of asymptotic IRS systems  \faCheck
	\begin{itemize}
	\item 
		Optimize phase matrix	\faQuestion
	\item
		Generalize results
	\end{itemize}

\item 
	Confirm findings with numerical results throughout

\item
	* Time permitting - Consider secrecy capacity of IRS system
\end{enumerate}
\end{frame}

\section{Conclusion}

\begin{frame}
  \centering \Large
  \emph{Thank You.}
  \\
	\bigskip
    \centering \Large
  \emph{Questions or Comments?}

\end{frame}

\end{document}

### 2020-07-02 documentation/Meeting Notes/07_01_20.tex
Added function and script for projection results. TODO noralize power with projection

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Numerical Results}
First I confirm that the methods I use to obtain the AED from a known, closed-form S-Transform works.
The numerical results  used from comparison are generated by averaging 100 realizations of the system for size $N=100 \times 100$.
\begin{itemize}
\item
	Channel 
	\begin{equation}
	\mH  \sim \CNO
	\end{equation}
	For equal power allocation 
	\\
	Numeric Capacity = $  1.4286$ 
	\\
	Asymptotic Capacity = $ 1.4129 $

\item
	Channel 
	\begin{equation}
	\mH \phase \mH  
	\end{equation}
	with $\mH \sim \CNO$.
	For equal power allocation 
	\\
	Numeric Capacity = $ 1.4227$ 
	\\
	Asymptotic Capacity = $ 1.4060 $

\item
	Channel 
	\begin{equation}
	\mC\mH  
	\end{equation}
	with $\mC$ as the exponential correlation matrix. 
	For equal power allocation the capacities are plotted below. Because the s-transform for the correlation has a quadratic form, I show the solutions for both the $+/-$ forms of the sign.

	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison}
	  \caption{Equal power capacity with correlation form 1.}
	\end{figure}
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison2}
	  \caption{Equal power capacity with correlation form 2.}
\end{figure}
\item
	These are the steps I use to move from the closed-form S-Transform of the channel to the AED.
	The S-Tranform of the exponential correlation model is derived in \cite{skupch2005free}
	\begin{enumerate}
	\item
		Beginning with the S-Transform, $S(z)$, I find $\gamma(s)$ using the fixed-point equation
		\begin{equation}
			S(\gamma(z))\gamma(z) - z - z\gamma(z) = 0 
		\end{equation}
	\item
		I find the Stieltjes transform using
		\begin{equation}
			G(s) = \frac{1}{s} (1+\gamma(\frac{1}{s}))
		\end{equation}			
	\item 
		I recover the AED using 
\begin{equation}
p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
\end{equation}
	\end{enumerate}

\item
	I checked the equation provided for the gamma transform of the the exponential correlation model and observed the following AEDs.
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/exponential_correlation_aed}
	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
	\end{figure}

\end{itemize}
\section{Questions}
\bibliography{bibliography}

\end{document}

### 2020-07-02 simulation/MATLAB/fixed_point_test.m

@@ -1,30 +1,30 @@
 rows = 100;
 cols = 100;
-steps = 100;
+steps = 400;
 bottom = 1e-3;
 top = 10; 
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
-% s_values = x_values + 1i*1e-6;
-s_values = x_values;
-% rhos = linspace(0,.9,3);
-rhos = linspace(0,1-1e-4,10);
+s_values = x_values + 1i*1e-3;
+% s_values = x_values;
+rhos = linspace(0, 0.8, 8);
+% rhos = linspace(0,1-1e-4,10);
 numeric_capacity = zeros(size(rhos));
 asymptotic_capacity = zeros(size(rhos));
-figure(1)
+% figure(1)
 for i = 1:length(rhos)
-%     stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, rhos(i)));
-    stieltjes_values = (1./s_values).*(1+gamma_s(s_values, rhos(i)));
-%     imag(stieltjes_values)
+    numeric_capacity(i) = average_numeric_capacity(rows, cols, rhos(i), 100);
+    stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, rhos(i)));
+%     stieltjes_values = (1./s_values).*(1+gamma_s(s_values, rhos(i)));
     pdf = 1/pi .* imag(stieltjes_values);
     pdf = abs(pdf);
-    numeric_capacity(i) = average_numeric_capacity(rows, cols, rhos(i), 100);
     asymptotic_capacity(i)  = aed_capacity(x_values, pdf, 1/cols, rows, step_size);
-    plot(x_values, pdf)
-    title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)]);
-    hold on
+%     plot(x_values,pdf);
+%     title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)]);
+%     hold on
 end
-
+numeric_capacity
+asymptotic_capacity
 figure(2)
 plot(rhos, numeric_capacity, '-s');
 hold on
@@ -40,60 +40,44 @@ clear all;
 function output = gamma_s(input, rho)
     out = zeros(size(input), 'like', input);
     for i = 1:length(input)
-%         out(i) =  gamma_s_func(input(i), rho);
-        out(i) =  new(input(i), rho);
+        out(i) =  gamma_s_func(input(i), rho);
+%         out(i) =  new(input(i), rho);
[…]

### 2020-07-02 simulation/MATLAB/useful_functions/projection_matrix.m

@@ -1,0 +1,5 @@
+function ret_mat = projection_matrix(size, beta)
+    vec = ones(1,size);
+    vec(round(size*(beta)):end) = 0;
+    ret_mat = diag(vec);
+end

### 2020-07-02 simulation/src/stieltjes_eqn.py

@@ -19,6 +19,7 @@ def steiltjes_from_gamma(s_vector, gamma_function, itr=200):
     return (1/s_vector)*(1+fixed_point(gamma_function, 1/s_vector, itr, 1e-2))
 
 
+
 # Move from inverse gamme to s-transform
 def s_transform_inverse_gamma(z, inverse_gamma):
     return ((1+z)/z)*inverse_gamma(z)

### 2020-07-02 simulation/testing/Correlation_tests/optimal_pdf_correlation.py

@@ -24,6 +24,7 @@ for cor_rho in correlations:
         powers.append(water_filling(H_cor, 1, vals=True))
     density, power_bin = np.histogram(np.asarray(powers), bins=bins)
     plt.plot(power_bin[:-1], density/average, '-o', label='power allocation')
+plt.legend()
 plt.show()
 
 

### 2020-07-06 documentation/Presentations/Mid-Presentation/mid_term_presentation.tex
adding for mid-term presentation

@@
 \section{Background}
-
 \begin{frame}
-\frametitle{Relaying}
-
-	\begin{itemize}
-		\item 			
-			Base Stations Vs. Relay $\rightarrow$ reduced core network footprint/cost
-%		\item 
-%			Modes of operation provide cost/performance trade-off (decode and forward, amplify and forward ...)
-		\item 
-			Extending wireless coverage is relevant to high frequencies used in 5G +
-	\end{itemize}
-
+\frametitle{Massive MIMO}
+				\centering
+		\begin{figure}
+		\includegraphics[scale=.18]{Pictures/mimo_antenna}
+	\end{figure}
 \end{frame}
@@
 \begin{frame}
-\frametitle{The basic MIMO channel}
-Much easier if each channel has a constant frequency response over the relevant band.
+\frametitle{Simplifying the MIMO Channel: OFDM}
+Assuming flat-fading channels between each transmit and receive antenna.
 	\begin{itemize}
 		\item 			
-
+			Interpret each symbol to be the complex coefficient of a complex sinusoid of specific frequency. 
+		\item
+			These sinusoids pass through a channel whose frequency response is flat.
+		\item
+			By considering both channel and transmit signal in the frequency domain, we have a system given by
+			\begin{equation}
+				\my = \mH \mx.
+			\end{equation}
+		\item 
+			Justify if there are any difficulties for OFDM when moving to the high frequencies required for massive MIMO.
 	\end{itemize}
-
 \end{frame}
@@
 
-%\begin{frame}
-%\frametitle{IRS: Modes of Operation}
-%\begin{columns}
-%\begin{column}{0.5\linewidth}
-%	Mode 1
-%	\begin{figure}
-%%			\centering
-%		\includegraphics[scale=1]{irs}
-%	\end{figure}\end{column}
-%\begin{column}{0.5\linewidth}
-%	Mode 2 (The one we will use)
-%\end{column}
-%\end{columns}
-%\end{frame}
+\begin{frame}
+\frametitle{Channel Correlation}
+	\begin{itemize}
+		\item 
+			FIR filters can represent correlation and their AED can be found. (Can just consider subset for FIR length?)
+		\item
+			The exponential correlation model is one such FIR filter with a single parameter. (Closed form expression for the S-Transform)
+		\item 
+			Measurement results for similar models show that $\rho$ can take on values up to $.6$ in $x$ GHz (a band relevant to 5G)
+	\end{itemize}
+\end{frame}
+
 \begin{frame}
@@
 \begin{equation*}
-	\mathbf{y} = (\mathbf{H}_2\boldsymbol{\Phi}\mathbf{H}_1 + \mathbf{G})\mathbf{x}
+	\mathbf{y} = (\mathbf{H}_2\boldsymbol{\Phi}\mathbf{H}+ \mathbf{G})\mathbf{x}
 \end{equation*}
@@
 
+\begin{frame}
+\frametitle{Channel Normalization}
+	\begin{itemize}
+		\item 
+			If transmit power is held constant as $N \rightarrow \infty$, the linear increase in capacity of the MIMO system depends on constant channel statistics.
+		\item 
+			In order to have a converging AED, we need to normalize the power of the channel.
+		\item
+			Show how can these results extend to the case without statistics changing? (is this even relevant?)
+	\end{itemize}
+\end{frame}
+
 \section{Goals}
@@
 \item
-	Find capacity of asymptotic IRS systems
+	Generalize capacity analysis for a channel model relevant to the IRS.
+\item 
+	Optimize phase matrix at the IRS (when relevant).
 	\begin{itemize}
 	\item
-		Optimize phase matrix
+		Investigate the cases in which phase optimization is worthwhile. 
+	\end{itemize}
 \item
-		Generalize results
-	\end{itemize}
-
+	* Time permitting - Consider secrecy capacity of IRS system (phase optimization may be relevant for some cases to increase secrecy capacity).
 \item 
-	Confirm findings with numerical results throughout
-
-\item
-	* Time permitting - Consider secrecy capacity of IRS system
+	Confirm findings with numerical results throughout.
 \end{enumerate}
 \end{frame}
-
 \section{Current Status}

### 2020-07-06 simulation/MATLAB/fixed_point_test.m

@@ -1,13 +1,14 @@
 rows = 100;
 cols = 100;
-steps = 400;
+steps = 500;
 bottom = 1e-3;
-top = 10; 
+% top = 200; 
+top = 100; 
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
 s_values = x_values + 1i*1e-3;
 % s_values = x_values;
-rhos = linspace(0, 0.8, 8);
+rhos = linspace(0, 0.9, 10);
 % rhos = linspace(0,1-1e-4,10);
 numeric_capacity = zeros(size(rhos));
 asymptotic_capacity = zeros(size(rhos));
@@ -17,7 +18,7 @@ for i = 1:length(rhos)
     stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, rhos(i)));
 %     stieltjes_values = (1./s_values).*(1+gamma_s(s_values, rhos(i)));
     pdf = 1/pi .* imag(stieltjes_values);
-    pdf = abs(pdf);
+%     pdf = abs(pdf);
     asymptotic_capacity(i)  = aed_capacity(x_values, pdf, 1/cols, rows, step_size);
 %     plot(x_values,pdf);
 %     title(['Capacity for this PDF is: ', num2str(asymptotic_capacity)]);
@@ -86,8 +87,6 @@ function ave = average_numeric_capacity(rows, cols, rho, average)
     for i = 1:length(vals)
         channel1 = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
         channel2 = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
-        projector1 = projection_matrix(rows, rows/2);
-        projector2 = projection_matrix(rows, rows/2);
         correlation = exponential_correlation(rows, rho);
 %         total_channel = channel1;
 %         total_channel = channel2*channel1;

### 2020-07-06 simulation/MATLAB/useful_functions/projection_matrix.m

@@ -1,5 +1,5 @@
 function ret_mat = projection_matrix(size, beta)
     vec = ones(1,size);
-    vec(round(size*(beta)):end) = 0;
+    vec(max(1,round(size*(beta))):end) = 0;
     ret_mat = diag(vec);
 end

### 2020-07-06 simulation/MATLAB/useful_functions/rayleigh_channel.m

@@ -1,3 +1,4 @@
  function matrix = rayleigh_channel(rows, cols, var)
 	matrix = normrnd(0, var, rows, cols) + 1i*normrnd(0, var, rows, cols);
+%     matrix = normrnd(0, var, rows, cols);
  end

### 2020-07-06 simulation/testing/Correlation_tests/correlation_convergence.py

@@ -25,7 +25,7 @@ ave_plt = fig.add_subplot(2, 1, 1)
 var_plt = fig.add_subplot(2, 1, 2)
 average = 100
 sizes = [5, 10, 20, 50, 100, 200]
-correlations = [.5, .8, .9, .98]
+correlations = [0, .5, .8, .9, .98]
 for cor_rho in correlations:
     capacity_variance = []
     capacity_average = []

### 2020-07-06 simulation/testing/Correlation_tests/correlation_setup.py

@@ -31,6 +31,7 @@ correlated_capacity_averages = np.zeros(len(correlation_block_sizes))
 average = 5
 for i in range(average):
     G = c_rand(rows, cols)
+    G2 = c_rand(rows, cols)
     GG = G @ np.conj(G.T)
     e_val_G = np.linalg.eigvalsh(GG)
     no_irs.append(e_val_G)
@@ -49,8 +50,9 @@ for i in range(average):
     # block = toeplitz(np.ones(int(rows / 2)))
     # correlation_matrix = block_matrix(block, rows, rows=True)
     t_correlation_matrix = exponential_correlation(size, .98)
+    s_correlation_matrix = exponential_correlation(size, .98)
     r_correlation_matrix = exponential_correlation(size, .98)
-    H_cor = r_correlation_matrix@G@t_correlation_matrix
+    H_cor = r_correlation_matrix@G2@s_correlation_matrix@G@t_correlation_matrix
     HH_cor = H_cor@hermetian(H_cor)
     e_val_cor = np.linalg.eigvalsh(HH_cor)
     irs_cor.append(np.real(e_val_cor))
@@ -60,9 +62,9 @@ for i in range(average):
     irs_trace.append(np.trace(HH))
 
 
-    # G_cap.append(capacity(GG, 1/size))
-    # irs_cap.append(capacity(HH, 1/size))
-    # cor_cap.append(capacity(HH_cor, 1/size))
+    G_cap.append(capacity(GG, 1/size))
+    irs_cap.append(capacity(HH, 1/size))
+    cor_cap.append(capacity(HH_cor, 1/size))
 
 
     # G_cap.append(capacity_water_filled(water_filling(G, 1)))

### 2020-07-06 simulation/testing/channel_testing/increasing_antennas.py

@@ -11,13 +11,11 @@ sigma = 1   # Relative attenuation of IRS portion
 rho = 1     # S/R
 phi = 1    # phi=1 is full rank line of sight.  L/T
 #   Plot Theoretical Distribution and find corresponding capacity
-
-
 numeric_capacity_list = []
 water_fill_capacity = []
 G_fill_capacity = []
 H_fill_capacity = []
-param_list = [1, 10, 50, 100, 200, 300]
+param_list = [5, 10, 50, 100]
 
 
 for val in param_list:
@@ -34,29 +32,18 @@ for val in param_list:
     los_rank_matrix = reduced_rank(rows, phi)
     average = 1
     for i in range(average):
-        # H = sigma*c_rand(rows, cols) @random_phase(rows)@ c_rand(rows, cols)
         H = sigma*c_rand(rows, cols) @ c_rand(rows, cols)
         G = c_rand(rows, cols)@los_rank_matrix
-        # check = np.var(H)
-        # check1 = np.var(G)
-        F = c_rand(rows, cols)
         GG = G @ np.conj(G.T)
         HH = H @ np.conj(H.T)
-        FF = F @ np.conj(F.T)
         GH = G @ np.conj(H.T)
         HG = H @ np.conj(G.T)
-        GF = G @ np.conj(F.T)
-        FG = F @ np.conj(G.T)
-        total_irs = GH + HG + GG + HH
-        # total_irs = GG
-        total_channel = H + G
-        Q_x = water_filling(total_channel, 1)
-        e_val_total_irs = np.linalg.eigvalsh(total_irs)
-        s_val_total_irs = np.linalg.svd(total_channel)[1]
-        irs.append(e_val_total_irs)
-        irs_svd.append(s_val_total_irs)
-        capacities.append(capacity(total_irs, 1/cols))
-        water_capacity.append(capacity_water_filled(Q_x))
+        # total_cov = GH + HG + GG + HH
+        # total_channel = H + G
+        total_cov = GG
+        total_channel = G
+        capacities.append(capacity(total_cov, 1/cols))
+        water_capacity.append(capacity_water_filled(water_filling(total_channel, 1)))
         H_capacity.append(capacity_water_filled(water_filling(H, 1)))
         G_capacity.append(capacity_water_filled(water_filling(G, 1)))
 
@@ -72,8 +59,8 @@ plt.title("Capacity Vs. Num Antennas")
[…]

### 2020-07-06 simulation/testing/sandbox.py

@@ -7,29 +7,10 @@ from scipy.linalg import toeplitz
 from scipy.optimize import fixed_point, root
 from scipy.optimize import fsolve
 
-
-
-one = np.eye(4)*2
-two = np.eye(4)*1
-# two[:,0] = 4
-check1 = np.linalg.eig(one)
-check2 = np.linalg.eig(two)
-print("done")
-
-
-
-
-def f_solve_check(gamma_z):
-    return 1/7*(pow(gamma_z, 3)+2)
-
-func = lambda x : np.power(x, 3) - 7*x +2
-x_values, step = np.linspace(.01, 10, 100, retstep=True)
-check = root(func, .1)
-val = check.x
-plt.plot(x_values, func(x_values))
-plt.scatter(val, func(val))
-plt.show()
-
-
-func = lambda x: (1+1j)*np.power(x, 3) - (2j+7)*x +2
-check = root(func)
+capacities = []
+average = 10000
+for i in range(average):
+    check = pow(np.abs(np.random.randn(1)), 2) + 1
+    # capacities.append(check)
+    capacities.append(np.log2(check))
+print(np.average(capacities))

### 2020-07-10 documentation/Meeting Notes/07_07_20.tex
Primary addition is finding multiple roots to the gamma polynomial. Also includes some checks on OFDM implementation

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Asymptotic Vs. Numerical Results}
First I confirm that the methods I use to obtain the AED given a closed-form S-Transform work for basic cases.
The numerical results  used for comparison are generated by averaging 100 realizations of the system for size $N=100 \times 100$.
As seen in the figures below, the solution becomes incorrect if there is too much correlation in the channel. 
\begin{itemize}
\item
	Channel 
	\begin{equation}
	\mH  \sim \CNO
	\end{equation}
	For equal power allocation 
	\\
	Numeric Capacity = $  1.4286$ 
	\\
	Asymptotic Capacity = $ 1.4129 $

\item
	Channel 
	\begin{equation}
	\mH_2 \phase \mH_1  
	\end{equation}
	with $\mH \sim \CNO$.
	For equal power allocation 
	\\
	Numeric Capacity = $ 1.4227$ 
	\\
	Asymptotic Capacity = $ 1.4060 $

	* This difference between numeric and asymptotic  can be reduced by increasing the AED sampling density.

\item
	The single Rayleigh channel with receiver correlation given by 
	\begin{equation}
	\mC\mH  
	\end{equation}
	with $\mC$ as the exponential correlation matrix  with elements $c_{i,j} = \rho^{i-j}$. 
	For equal power allocation the capacities are plotted below. Because the s-transform for the correlation has a quadratic form, I show the solutions for both the $+/-$ forms of the quadratic expression.

	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison}
	  \caption{Equal power capacity with correlation form 1.}
	\end{figure}
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison2}
	  \caption{Equal power capacity with correlation form 2.}
\end{figure}

\item
	The IRS channel without LOS and correlation only at receiver
	\begin{equation}
	\mC\mH_2 \phase\mH_1
	\end{equation}
	with $\mC$ as the exponential correlation matrix. 
	For equal power allocation the capacities are plotted below. 
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/equation1_Rx_only}
	  \caption{Equal power capacity for IRS channel without LOS and correlation only at receiver.}
	\end{figure}
\item
	The IRS channel without LOS and correlation at receiver and IRS elements
	\begin{equation}
	\mC_2 \mH_2\phase\mC_1 \mH_1
	\end{equation}
	with $\mC_i$ as an exponential correlation matrix. 
	For equal power allocation the capacities are plotted below. 
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/Rx_IRS_Correlation}
	  \caption{Equal power capacity for IRS channel without LOS and correlation at receiver and IRS.}
	\end{figure}

\item
	The IRS channel without LOS and correlation at receiver, IRS elements and transmitter
	\begin{equation}
	\mC_3 \mH_2\mC_2 \phase\mC_1 \mH_1
	\end{equation}
	with $\mC_i$ as the exponential correlation matrix. 
	For equal power allocation the capacities are plotted below. 
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/Rx_IRS_Tx_Correlation}
	  \caption{Equal power capacity for IRS channel without LOS and correlation at receiver, IRS and transmitter.}
	\end{figure}

\item
	These are the steps I use to move from the closed-form S-Transform of the channel to the AED.
	The S-Tranform of the exponential correlation model is derived in \cite{skupch2005free}
	\begin{enumerate}
	\item
		Beginning with the S-Transform, $S(z)$, I find $\gamma(s)$ using the fixed-point equation
		\begin{equation}
			S(\gamma(z))\gamma(z) - z - z\gamma(z) = 0 
		\end{equation}
	\item
		I find the Stieltjes transform using
		\begin{equation}
			G(s) = \frac{1}{s} (1+\gamma(\frac{1}{s}))
		\end{equation}			
	\item 
		I recover the AED using 
\begin{equation}
p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
\end{equation}
	\end{enumerate}

\item
	I checked the equation provided for the gamma transform of just the the exponential correlation model and observed the following AEDs.
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/exponential_correlation_aed}
	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
	\end{figure}

\end{itemize}
\section{Capacity Convergence}
As there is not a method for evaluating the rate of convergence of finite systems towards the asymptotic behavior I investigate this numerically in the following figures by looking at the variance of the capacity of numerical simulations for the IRS channel given by 
	\begin{equation}
	\mC_3 \mH_2\mC_2 \phase\mC_1 \mH_1
	\end{equation}
Figures \ref{equal_power} and \ref{water_filling} show the average capacity and capacity variance over different, random, values of  $\phase$ with all channels held constant for equal power allocation and optimal power allocation respectively. Note that \emph{neither} uses any optimization of the phases at the IRS. 
These results reflect the intuition that smaller finite system sizes with significant correlation are not well represented by the asymptotic result that shows the phase matrix at the IRS canceling. It might be interesting to see if an optimization of the IRS phases reduces this variance in the high correlation case.

		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_equal_power_capacity}
	  \caption{Capacity average and variance for system size $100$ equal power allocation over $N=1000$ channel realizations}
	  \label{equal_power}
	\end{figure}

		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_water_filling_capacity}
	  \caption{Capacity average and variance for system size $100$ equal power allocation over $N=100$ channel realizations}
	  	  \label{water_filling}
	\end{figure}

\section{Questions}
\begin{enumerate}
\item
	Because we can show that the phases cancel for the asymptotic analysis, it seems that any further look at this is just extending capacity results without 
	IRS to a more complicated channel model. Can I focus the remainder of the work on just looking at the when the asymptotic assumption is valid?
\item
	To start considering the "zero-tapping" that was suggested in our last meeting.I began by using the S-Transform of the "projector" matrix (essentially resulting in the deformed quarter circle law for $\mH \in \complex^{N \times K}$ with $N \leq K$). There are two points, I was hoping to discuss on this.
	\begin{enumerate}
	\item
		First, one problem is that this is also changing the power of the received signal to $\frac{N}{K}$ 
		 because the deformed quarter circle assumes elements with variance $\frac{1}{K}$ rather than $\frac{1}{N}$.
\item
		One other strange thing I noticed is that when I perform numerical simulations with the deformed quarter circle the distributions I have, as shown in Figure \ref{AED} 
		appear to not integrate to 1 but nevertheless  the resulting capacity they agree with numerical results as shown in Figure \ref{projection}.
				\begin{figure}[H]
	\includegraphics[width= 12cm,height = 8cm]{figures/projection}
	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
	  \label{projection}
	\end{figure}
		\begin{figure}[H]
	\includegraphics[width= 12cm,height = 8cm]{figures/AED}
	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
	  	  \label{AED}
	\end{figure}

	\end{enumerate}
\item
I was also preparing some slides for the mid-term presentation and a question on the background.
\begin{itemize}
\item
	Will it be okay to just assume flat-fading using OFDM rather than including a model with multiple delays? Will OFDM still be possible at very high frequencies?
\end{itemize}
\end{enumerate}
\bibliography{bibliography}

\end{document}

### 2020-07-10 documentation/Presentations/Mid-Presentation/mid_term_presentation.tex
Primary addition is finding multiple roots to the gamma polynomial. Also includes some checks on OFDM implementation

@@
 \section{Background}
+\subsection{MIMO Channel}
+
 \begin{frame}
@@
 	\end{figure}
+	\cite{argos2020}
 \end{frame}
@@
 		\item 			
-			Interpret each symbol to be the complex coefficient of a complex sinusoid of specific frequency. 
+			Interpret each symbol to be the complex coefficient in frequency of a specific complex sinusoid (a sub-carrier). 
 		\item
@@
 		\item
-			By considering both channel and transmit signal in the frequency domain, we have a system given by
+			By considering both the channel and transmit signals in the frequency domain for each "sub-carrier" in OFDM, we have a system given by
 			\begin{equation}
-				\my = \mH \mx.
+				\vy = \mH \vx.
 			\end{equation}
 		\item 
-			Justify if there are any difficulties for OFDM when moving to the high frequencies required for massive MIMO.
+			Forms of OFDM are defined for wireless protocols in 5G and 802.11
 	\end{itemize}
@@
 
+\begin{frame}
+\frametitle{Channel Correlation}
+	\begin{itemize}
+		\item 
+			From OFDM perspective this implies correlation between the same sub-carrier at different antennas.
+		\item 
+			FIR filters can represent some relevant types of correlation.
+			Toeplitz can be modeled asymptotically as circulant matrices. AED of circulant matrices is DFT of 1st row
+		\item
+			The exponential correlation model is one such FIR filter with elements $c_{i,j} = \rho^{i-j} $ with $ 0 \leq \rho \leq 1$.
+		\item 
+			Measurement results for similar models show that $\rho$ can take on values up to $.6$ in $x$ GHz (a band relevant to 5G)
+	\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Channel Normalization}
+\begin{itemize}
+\item
+	We want to analyze the channel behavior as $N \rightarrow \infty$ without sending capacity to infinity. 
+\begin{itemize}
+\item
+	Normalize the power of the transmitted symbols and channel gains such that $\Expect[\vy^H \vy ] = 1$.
+	Enforcing that $\Expect[\vx^H \vx ] = \frac{1}{N}\mathbf{I}$ and $\Expect[\|h_{i,j}\|^2_2 ]= \frac{1}{M}$ for $\mH \in \complex^{M \times N} $ satisfies this.
+\end{itemize}
+\item
+	The popular result that MIMO capacity grows as $\min{N_T, N_R}$ does not include such a normalization. Which type of gain is now eliminated?
+\item
+	Allows for a matrix to have a converging eigenvalue distribution despite $N, M \rightarrow \infty$.
+\end{itemize}
+\end{frame}
+
+\subsection{IRS Channel}
 \begin{frame}
@@
 
-\end{frame}
-
-\begin{frame}
-\frametitle{Channel Correlation}
-	\begin{itemize}
-		\item 
-			FIR filters can represent correlation and their AED can be found. (Can just consider subset for FIR length?)
-		\item
-			The exponential correlation model is one such FIR filter with a single parameter. (Closed form expression for the S-Transform)
-		\item 
-			Measurement results for similar models show that $\rho$ can take on values up to $.6$ in $x$ GHz (a band relevant to 5G)
-	\end{itemize}
 \end{frame}
@@
 
+\begin{frame}[allowframebreaks]
+        \frametitle{References}
+        \bibliography{bibliography.bib}
+\end{frame}
+
 \end{document}

### 2020-07-10 simulation/MATLAB/projection_s_transform.m

@@ -1,21 +1,29 @@
 rows = 100;
 cols = 100;
 steps = 100;
-bottom = 1e-3;
-top = 20; 
+bottom = 1e-1;
+top = 4; 
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
-s_values = x_values + 1i*1e-3;
-betas = linspace(1e-4, 1, 5);
+s_values = x_values + 1i*1e-7;
+betas = linspace(0.05, 1, 10);
 numeric_capacity = zeros(size(betas));
 asymptotic_capacity = zeros(size(betas));
+figure(1)
 for i = 1:length(betas)
-    numeric_capacity(i) = average_numeric_capacity(rows, cols, betas(i), 200);
+    numeric_capacity(i) = average_numeric_capacity(rows, cols, betas(i), 10);
     stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, betas(i)));
     pdf = 1/pi .* imag(stieltjes_values);
-    pdf = abs(pdf);
+%     pdf = abs(pdf);
     asymptotic_capacity(i)  = aed_capacity(x_values, pdf, 1/cols, rows, step_size);
+    plot(x_values,pdf,'DisplayName', '\beta = ' + string(betas(i)));
+    title('AED');
+    hold on
 end
+legend();
+f = gca;
+exportgraphics(f,'results\AED.png')
+figure(2)
 plot(betas, numeric_capacity, '-s');
 hold on
 plot(betas, asymptotic_capacity, '-o');
@@ -40,27 +48,34 @@ function x = gamma_s_func(eval_point, beta)
         F = z.*S_func(z, beta) - eval_point - z*eval_point;
     end
 init_point = rand() + 1j*rand();
-x = fsolve(@root1,init_point);
+% x = fsolve(@root1,init_point);
+polynomial_vector = [eval_point, (2*eval_point-1), eval_point];
+check = roots(polynomial_vector);
+%   Need to pick out positive part of the polynomial
+x = check(1);
 end
 
 function val = S_func(z, beta)
-    wishart = 1./power((1+z),1);
+%     wishart = 1./power((1+z),1);
+    wishart = 1./(1+z); 
     projection = (1+z)/(z+beta);
-    val = projection*wishart*projection*wishart;
+%     val = projection*wishart*projection*wishart;
+%     val = wishart*projection;
+    val = wishart;
 end
 
[…]

### 2020-07-10 simulation/MATLAB/useful_functions/deformed_qc.m

@@ -1,0 +1,7 @@
+function [outputArg1] = deformed_qc(s, alpha)
+%DEFORMED_QC Summary of this function goes here
+%   Detailed explanation goes here
+outputArg1 = sqrt((1-alpha).^2./(4.*s.^2) - (1+alpha)./(2.*s) + 1/4) - 1/2 - (1-alpha)./(2.*s);
+% outputArg1 = sqrt((1-alpha).^2./(4.*s.^2) - (1+alpha)./(2.*s) + 1/4) + 1/2 + (1-alpha)./(2.*s);
+end 
+

### 2020-07-10 simulation/OFDM/ofdm_tools.py

@@ -9,4 +9,4 @@ def DFT_matrix(size):
     for i in range(size):
         for j in range(size):
             mat[i,j] = i*j
-    return np.exp(1j*2*np.pi*mat/size)/100
+    return np.exp(-1j*2*np.pi*mat/size)/100

### 2020-07-10 simulation/src/stieltjes_eqn.py

@@ -19,7 +19,6 @@ def steiltjes_from_gamma(s_vector, gamma_function, itr=200):
     return (1/s_vector)*(1+fixed_point(gamma_function, 1/s_vector, itr, 1e-2))
 
 
-
 # Move from inverse gamme to s-transform
 def s_transform_inverse_gamma(z, inverse_gamma):
     return ((1+z)/z)*inverse_gamma(z)

### 2020-07-10 simulation/testing/Correlation_tests/correlation_convergence.py

@@ -10,21 +10,13 @@ from src.lin_alg import *
 from src.network_simulation_tools import water_filling
 import seaborn
 from scipy.linalg import toeplitz
-
-
-size = 100
-bins = int(size/2)
 irs_cor = []
 rank = 100
-phi = rank/size
-correlation_block_sizes = [1, 5, 25, 100]
-correlated_capacity_averages = np.zeros(len(correlation_block_sizes))
-
 fig = plt.figure()
 ave_plt = fig.add_subplot(2, 1, 1)
 var_plt = fig.add_subplot(2, 1, 2)
-average = 100
-sizes = [5, 10, 20, 50, 100, 200]
+average = 10
+sizes = [10, 20, 50, 100]
 correlations = [0, .5, .8, .9, .98]
 for cor_rho in correlations:
     capacity_variance = []
@@ -34,17 +26,17 @@ for cor_rho in correlations:
         s_correlation_matrix = exponential_correlation(size, cor_rho)
         r_correlation_matrix = exponential_correlation(size, cor_rho)
         cor_cap = []
+        G = c_rand(size, size)
+        G2 = c_rand(size, size)
         for i in range(average):
-            G = c_rand(size, size)
-            G2 = c_rand(size, size)
             # H_cor = r_correlation_matrix@G@t_correlation_matrix
-            H_cor = r_correlation_matrix@G2@s_correlation_matrix@G@t_correlation_matrix
+            H_cor = r_correlation_matrix@G2@s_correlation_matrix@random_phase(size)@G@t_correlation_matrix
             HH_cor = H_cor@hermetian(H_cor)
             # e_val_cor = np.linalg.eigvalsh(HH_cor)
             # irs_cor.append(np.real(e_val_cor))
             # cor_trace.append(np.trace(HH_cor))
-            cor_cap.append(capacity(HH_cor, 1/size))
-            # cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
+            # cor_cap.append(capacity(HH_cor, 1/size))
+            cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
         capacity_average.append(np.average(cor_cap))
         capacity_variance.append(np.var(cor_cap))
 

### 2020-07-10 simulation/testing/theory_testing/deformed_qc_eqn.py

@@ -1,0 +1,31 @@
+from src.stieltjes_eqn import *
+import matplotlib.pyplot as plt
+from src.circle_laws import *
+from src.lin_alg import *
+cols = 500
+alpha_list = [.1, .5, .8, 1]
+bins = 50
+los_rank = 1
+average = 10
+plt.figure()
+for alpha in alpha_list:
+    x_values, step = np.linspace(1e-2, 10, 500, retstep=True)
+    distribution = deformed_qc_eigen(x_values, alpha)
+    y = 1e-9
+    s_vector = x_values + 1j * y
+    check = estimated_pdf(deformed_quarter_circle(s_vector, alpha))
+    # plt.plot(x_values, check, label=f"check {alpha}")
+
+    eigen_values = []
+    rows = int(cols * alpha)
+    for i in range(average):
+        H = c_rand(rows, cols, var=1/(2*cols))
+        eigen_values.append(np.linalg.eigvalsh(H@hermetian(H)))
+    eigen_value_vec, bins = np.histogram(np.asarray(eigen_values), bins=bins)
+    plt.plot(bins[:-1], eigen_value_vec / (rows * average), label=f"{alpha}")
+
+plt.title("deformed quarter circle")
+# plt.plot(values, st_distribution)
+plt.legend(loc="upper right")
+plt.show()
+

### 2020-07-10 simulation/testing/theory_testing/stieltjes_practice.py

@@ -1,37 +1,0 @@
-# A test for looking at impact of adding correlation through concatenated channels on the channel capacity
-# Takeaway:
-# If power is allocated equally the ave capacity doesn't seem to be affected however variance increase
-# with increasing correlation. However if water-filling
-# is used then the correlated channels seem to have increasing capacity as correlation increases
-
-import matplotlib.pyplot as plt
-from src.lin_alg import *
-from src.stieltjes_eqn import *
-import seaborn
-from scipy.linalg import toeplitz
-from src.circle_laws import *
-from scipy.optimize import *
-from scipy.optimize import fsolve
-
-
-rows = 100
-cols = 100
-x_values, step = np.linspace(.01, 10, 100, retstep=True)
-s_values = x_values + 1j*1e-6
-
-def R_Transform(s, past):
-    return 1/(1/(1+past) - s)
-
-stieltjes_values = steiltjes_from_r_transform(s_values, R_Transform)
-pdf = estimated_pdf(stieltjes_values)
-print(f"asymptotic capacity: {aed_capacity(x_values, pdf, 1/rows, rows, step=step)}")
-
-fig, ax = plt.subplots()
-plt.title("IRS Channels Components: i.i.d, $\mathbb{N}(0,1/N), H = H_1 \Phi H_2$")
-ax.plot(x_values, pdf, label='asymptotic')
-plt.legend(loc="upper right")
-ax.grid(True, which='both')
-seaborn.despine(ax=ax, offset=0)
-plt.show()
-
-

### 2020-07-14 documentation/Paper/abstract.tex
Checked secrecy capacity for correlated. As expected secrecy capacity increases with correlation for the IRS channel.

@@
-This work investigates the impact of passive reflective surfaces on the capacity of MIMO systems.
-In particular, the system is analyzed in the asymptotic size domain in order to employ the tools of free probability.
+This work investigates the impact of passive reflective surfaces on the capacity of MIMO communication systems.
+This using the tools of free probability, this system is analyzed in the asymptotic domain to derive an expression for the system capacity.
 A system is considered first in the case of fully uncorrelated channels.

### 2020-07-14 documentation/Paper/paper.tex
Checked secrecy capacity for correlated. As expected secrecy capacity increases with correlation for the IRS channel.

@@
 \section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards.
-Industry research for 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
-This expansion is motivated by large bands of currently unused frequency
+The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
+Industry research for the 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
+This shift is motivated by large bands of currently unoccupied frequency
 and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-A primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
-For frequency bands relevant for upcoming wireless standards, energy attenuation due to free space path loss  molecular absorption loss
-has been shown to exceed 100 dB. Countering this loss is a key steps towards meeting the increased data rates and reliability required for future networks.
+One primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
+(Justify previous statement by comparing attenuation from previous to new standards).
+Countering this loss is a key step towards meeting the increased data rates and reliability required for future networks.
 \par
-In previous generations of wireless communications, the problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
+In previous generations of wireless communications, attenuation due to path loss (since not at high enough frequencies for absorption) has been alleviated by reducing the effective transmission distance
 and supplementing the transmission power using relays \cite{dahlman20134g}.
@@
 (Would be nice to cite some actual costs here).
-Deploying this solution on a scale to counter
-the attenuation levels relevant to new standards would pose impractical costs (why).
-One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves.
+Deploying enough relays to counter
+the attenuation levels relevant to new standards would pose impractical costs.
+One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many, sub-wavelength sized elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients can be controlled to select the phase of reflected waves.
 In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
@@
 associated with a relay.
-The absence of an rf-chain in the RIS offers additional benefits over relays.
+(Compare performance to relays) (detail relay model) as show in \cite{bjornson2019demystifying}, the absence of an rf-chain in the RIS offers some benefits over relays.
 First, reflection does not introduce the noise associated with the rf-chain.
 Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay.
+Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
 The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
 For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?).
+Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
+It is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
+
 For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
@@
 Example of how these surfaces are implemented (Cite).
+
 Discuss how it has be looked at from MISO and MIMO perspective.
@@
 \par
-The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will requires systems 
+The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will require systems 
 sizes on the order of x. Now motivate using asymptotic analysis.
@@
 \chapter{System Model}\label{system_model}
+\section{Power Normalization}
+Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
+the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
+In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
+Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
+In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
+This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
+
 \section{The Correlated IRS Channel}

### 2020-07-14 simulation/src/network_simulation_tools.py

@@ -113,7 +113,7 @@ def water_filling(channel_matrix, power_constraint, sigma_square=1, vals=False):
     covariance_matrix = hermetian(V_H)@np.diag(variables.value)@V_H
     check = np.trace(covariance_matrix)
     if vals:
-        return variables.value
+        return np.diag(variables.value)
     else:
         return channel_matrix@covariance_matrix@hermetian(channel_matrix)
 

### 2020-07-14 simulation/testing/secrecy_capacity/secrecy_correlation_convergence.py

@@ -1,0 +1,68 @@
+# A test for looking at impact of adding correlation through concatenated channels on the channel capacity
+# Takeaway:
+# If power is allocated equally the ave capacity doesn't seem to be affected however variance increase
+# with increasing correlation. However if water-filling
+# is used then the correlated channels seem to have increasing capacity as correlation increases
+
+import matplotlib.pyplot as plt
+from src.fixed_point import *
+from src.lin_alg import *
+from src.network_simulation_tools import water_filling
+import seaborn
+
+irs_cor = []
+rank = 100
+fig = plt.figure()
+ave_plt = fig.add_subplot(2, 1, 1)
+var_plt = fig.add_subplot(2, 1, 2)
+average = 1000
+sizes = [10, 20, 50, 100, 200]
+# average = 10
+# sizes = [10, 20, 50, 100]
+correlations = [0, .5, .8, .9, .98]
+for cor_rho in correlations:
+    capacity_variance = []
+    capacity_average = []
+    for size in sizes:
+        t_correlation_matrix = exponential_correlation(size, cor_rho)
+        s_correlation_matrix = exponential_correlation(size, cor_rho)
+        r_correlation_matrix = exponential_correlation(size, cor_rho)
+        sec_cap = []
+        G = c_rand(size, size)
+        G2 = c_rand(size, size)
+        G3 = c_rand(size, size)
+        for i in range(average):
+            # H_cor = r_correlation_matrix@G@t_correlation_matrix
+            H_cor = r_correlation_matrix@G2@s_correlation_matrix@random_phase(size)@G@t_correlation_matrix
+            H_cor_eve = r_correlation_matrix@G3@s_correlation_matrix@random_phase(size)@G@t_correlation_matrix
+            HH_cor = H_cor@hermetian(H_cor)
+            HH_cor_eve = H_cor_eve@hermetian(H_cor_eve)
+            #   Optimize for the intended received
+            # powers = water_filling(H_cor, 1, vals=True)
+            # sec_cap.append(capacity_water_filled(H_cor@powers@hermetian(H_cor)) - capacity_water_filled(H_cor_eve@powers@hermetian(H_cor_eve)))
+            sec_cap.append(capacity(HH_cor, 1/size) - capacity(HH_cor_eve, 1/size))
+        capacity_average.append(np.average(sec_cap))
[…]

### 2020-07-16 documentation/Paper/correlation_review.tex
working on writing

@@
 \par
-Investigation leading to the current interest in MIMO systems showed the key property that (with constant transmit power), capacity (and mutual information) in an $N \times N$ MIMO system with independent channels scales linearly in $N$ \cite{foschini1998limits}.
-However the assumption of independent channels between each pair of transmit and receive antennas is not generally reasonable.
-This prompted much investigation into the impact of correlation to this scaling.
+Investigation leading to the current interest in MIMO systems showed the key property that (with constant transmit power), capacity (and mutual information) in an $N \times N$ MIMO system with independent, i.i.d gaussian channels scales linearly in $N$ \cite{foschini1998limits}.
+While the assumption of \emph{identically distributed} channels is intuitively reasonable when adding more antennas, the assumption of \emph{independent} channels between each pair of transmit and receive antennas is not generally reasonable.
+As this point could determine the ability to leverage capacity scaling of massive MIMO, it prompted much investigation.
 In the following, a number of these studies are reviewed.
@@
 1. Looking at correlation for finite size systems.
-In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?).
-(Is this a bigger problem for higher frequencies?).
-Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
+In test data collected with channel parameters (carrier frequency, bandwidth, power...) relevant to 3G and 4G, wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?).
+Need to describe the propagation environment a bit more.
+As increased antenna density is one of the primary motivations for moving to higher frequencies, future antenna spacing with respect to wavelength is likely to remain consistent and therefore similar antenna correlation can be expected (proof of this?).
+Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome for large numbers of antennas. Describe early results for MIMO.
+\par
+Explain reason for using power scaling here.
+
 \par
 One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. 
+(TODO need to explain the power normalization needed for this to work)
 This is motivated by the observation that the eigenvalues of a channel covariance matrix converge to a deterministic distribution for many random 

### 2020-07-16 documentation/Paper/paper.tex
working on writing

@@
 The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
-Industry research for the 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency
-and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-One primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
-(Justify previous statement by comparing attenuation from previous to new standards).
-Countering this loss is a key step towards meeting the increased data rates and reliability required for future networks.
+Industry research for the 6G standard has already investigated systems using frequency bands in the 120, 200 and 340 GHz range \cite{Koziol}.
+This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
+One primary hurdle towards efficient communication at these frequencies is the higher net attenuation of radio energy due both to scattering effects and molecular absorption.
+\cite{TODO} show through extensive data collection that relevant attenuation will be (x).
+As a relevant example, radar systems have long used this band of frequencies due to the resulting high resolution.
+For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO}.
+Countering this type of loss is a key step towards meeting the increased data rates and reliability required for future networks. 
 \par
-In previous generations of wireless communications, attenuation due to path loss (since not at high enough frequencies for absorption) has been alleviated by reducing the effective transmission distance
+In previous generations of wireless communications, attenuation due to path loss has been alleviated by reducing the effective transmission distance 
 and supplementing the transmission power using relays \cite{dahlman20134g}.
 Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
-(Would be nice to cite some actual costs here).
 Deploying enough relays to counter
 the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many, sub-wavelength sized elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients can be controlled to select the phase of reflected waves.
+One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable electromagnetic reflection and whose reflective coefficient can be controlled to select the phase of the reflected wave.
 In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
+While this work focuses on the theory surrounding such surfaces, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an RIS with 224 elements was built and tested.
 \par
@@
 associated with a relay.
-(Compare performance to relays) (detail relay model) as show in \cite{bjornson2019demystifying}, the absence of an rf-chain in the RIS offers some benefits over relays.
-First, reflection does not introduce the noise associated with the rf-chain.
+The absence of an rf-chain on the RIS offers some benefits over relays.
+First, reflection does not introduce the noise associated with an rf-chain.
 Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
@@
 Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
-It is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
-
-For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
-(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
-Example of how these surfaces are implemented (Cite).
-
-Discuss how it has be looked at from MISO and MIMO perspective.
+While such results are indeed motivating, it is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
 \par
-Need to show how many surfaces are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
+Need to show how many surfaces (are the corresponding size) are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
 \par
@@
 The following sections are now previewed.
-In section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
+Section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
 In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
@@
 \end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
+If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is also circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
 \begin{equation}\label{entropy}
@@
 
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect[\vn\vn^H] = \mI$, substituting equation \eqref{entropy} into equation
-\eqref{mut_ent} gives
+For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$, substituting equation \eqref{entropy} into equation
+\eqref{mut_ent} and simplifying gives
 \begin{equation}\label{mut_inf}
-\Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
+\Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right].
 \end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to the
-transmit covariance matrix and the phases at the IRS gives and expression for the capacity of the system
+Using the model for $\mHt$ from the previous section, and maximizing with respect to both the
+transmit covariance matrix and phases at the IRS gives and expression for the system capacity
 
 \begin{equation}\label{capacity}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
+\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right]
 \end{equation}
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
-
-If $\mHt$ is not known at the transmitter, the transmit power can be allocated
-equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
+or equivalently,
+\begin{equation}
+\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(\mathbf{I} + \lambda_i)\right].
 \end{equation}
 where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-Notice that in this form, the linear scaling of capacity with N is clear.
-Expanding the expectation gives the expression
+
+Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
+(Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).
+If $\mHt$ is not known at the transmitter, the transmit power can be allocated
+equally such that $\mQ_x = \mathbf{I}$ (Need to provided some evidence for this choice and extend to the case that we choose it in some other way using a distribution).
+This allows for simplifying equation \eqref{capacity} to
+\begin{equation}\label{no_csi_capacity}
+\mathbb{C} = \underset{\phase}{\mathop{max}} N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)].
+\end{equation}
+
+In this form, the linear scaling of capacity with N is clear.
+Expanding the expectation gives 
 \begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
+\mathbb{C} = \underset{\phase}{\mathop{max}} N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
 \end{equation}
-in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
+in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
 
@@
 \begin{equation}\label{mut_inf}
-\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda_i])
+\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda])
 \end{equation}
-Now discuss capacity and consider equal power but optimize over phase
-We now want to consider the case in
+
+\section{Explanation of Asymptotic Analysis}
+Equation \eqref{no_csi_capacity_aed} highlights the important point that knowledge of the eigenvalue distribution of the received signal co-variance matrix is sufficient information to find the capacity of a MIMO channel.
+As we are generally interested in random channels, finding this eigenvalue distribution is not a straightforward task.
 
@@
 Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = Expect[\Tr{(mHt \mHt^H})] = $.
+For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = \Expect[\Tr{(mHt \mHt^H})] = $.
 One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of

### 2020-07-17 documentation/Paper/paper.tex
added model details to report

@@
 Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
-
 \par
@@
 Using the model for $\mHt$ from the previous section, and maximizing with respect to both the
-transmit covariance matrix and phases at the IRS gives and expression for the system capacity
+transmit covariance matrix and phases at the IRS gives an expression for the system capacity
 
@@
 or equivalently,
-\begin{equation}
+\begin{equation}\label{capacity_tricky}
 \mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(\mathbf{I} + \lambda_i)\right].
 \end{equation}
+where $\lambda_i$ are the eigenvalues of $\mHt \mQ_x \mHt^H$.
+
+Recalling that in general we are interested in evaluating the capacity in order to determine a feasible transmission rate for a given communication channel, equation \eqref{capacity_tricky} poses a couple of hurdles to evaluation.
+First, we need to be able to evaluate an expectation over the joint pdf $p(\lambda_1 \cdots	 \lambda_N)$.
+Second, there is an optimization problem whose solution will generally depend on the distribution of the channel which would require a closed-form of the optimization problem solution.
+Two simplifications will be used in order to move forward. 
+First, we will enforce that the transmitter chooses a deterministic $\mQ_x$. 
+By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the optimization.
+Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge.
+
+If $\mHt$ is not known at the transmitter, a common choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$, giving the simplified form of equation \eqref{capacity}
+\begin{equation}\label{no_csi_capacity}
+\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R \Expect[\Log(1 + \frac{P_{\text{total}}}{\sigma_nN_t}\lambda_i)].
+\end{equation}
 where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
+In this form, the linear scaling of capacity with N is clear.
+If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation to get the expression
+\begin{equation}\label{no_csi_capacity_aed}
+\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
+\end{equation}
+in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
+In the following we will refer to this distribution as the Asymptotic Eigenvalue Distribution (AED).
+
+From this, we see that if the normalized channel has a converging eigenvalue distribution we can simply scale the capacity for a single channel by the system size $N_R$.
 
@@
 (Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).
-If $\mHt$ is not known at the transmitter, the transmit power can be allocated
-equally such that $\mQ_x = \mathbf{I}$ (Need to provided some evidence for this choice and extend to the case that we choose it in some other way using a distribution).
-This allows for simplifying equation \eqref{capacity} to
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)].
-\end{equation}
-
-In this form, the linear scaling of capacity with N is clear.
-Expanding the expectation gives 
-\begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
-\end{equation}
-in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
 
@@
 This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{Proof of Freeness}
+\section{Free Probability}\label{Free_Prob_Intro}
 \begin{itemize}
@@
 
+\section{The IRS Channel}
+Using the approach for finding asyptotic eigenvalue distributions detailed in \ref{Free_Prob_Intro} we will then use the capacity analysis detailed in \ref{sectiond:mut_info}.
+We will now develop a channel model on which to perform this analysis.
+Specifically we develop a channel model that is representative of a MIMO communication system with an RIS.
+\par
+In Section \cite{Background} previous studies were reviewed which showed the impact and relevance of correlation to massive MIMO systems, therefore we begin with an RIS channel given by 
+	\begin{equation}
+	\mH_{Total} = \mR_{R}^{\frac{1}{2}}(\underbrace{\mH_{2}\boldsymbol{\Phi}\mR_{S}^{\frac{1}{2}}\mH_{1}}_{\text{IRS}}.
+	\end{equation}
+
+The random matrices $\mH_{2}$ and $ \mH_{1}$ represent the channels from the transmitter to the IRS and from the IRS to the receiver respectively.
+Both are assumed to have i.i.d.
+elements with variance $\frac{1}{N}$.
+The matrices $\mR_{R}, \mR_{S}$ and $\mR_{T}$ represent a deterministic model for the correlation at the receiver, IRS and transmitter respectively.
+The resulting covariance matrix looks quite similar to the one used for the example in Section \ref{Free_Prob_Intro} but first we want to verify that 
+the resulting components will be free.
+
+(TODO provide references to prove that the components are free)
+
+(Will need to detail any constraints embedded in this system model)
+
 \chapter{Results}\label{Results}
 \section{Asymptotic Results}
-We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
+Begin this section using the model from the last
+The final part of this section should show that phases will always cancel out for free probability analysis.
+
+\subsection{Asymptotic Results Discussion}
+
+We first consider the case in which all matrices of the channel given by (ref) are free. As shown in (Appendix), in this case the 
 phase shifts at the RIS, $\phase$, cancels out of the AED expression. 

### 2020-07-17 simulation/testing/Correlation_tests/correlation_convergence.py

@@ -32,11 +32,12 @@ for cor_rho in correlations:
             # H_cor = r_correlation_matrix@G@t_correlation_matrix
             H_cor = r_correlation_matrix@G2@s_correlation_matrix@random_phase(size)@G@t_correlation_matrix
             HH_cor = H_cor@hermetian(H_cor)
-            # e_val_cor = np.linalg.eigvalsh(HH_cor)
-            # irs_cor.append(np.real(e_val_cor))
-            # cor_trace.append(np.trace(HH_cor))
-            # cor_cap.append(capacity(HH_cor, 1/size))
-            cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
+            cor_cap.append(capacity(HH_cor, 1/size))
+            #   Check alternative transmit correlation matrix (normalize such that the covariance trace is 1
+            # transmit_covariance = t_correlation_matrix/np.sum(np.diag(t_correlation_matrix))
+            # HH_cor = H_cor@transmit_covariance@hermetian(H_cor)
+            # cor_cap.append(capacity(HH_cor, 1))
+            # cor_cap.append(capacity_water_filled(water_filling(H_cor, 1)))
         capacity_average.append(np.average(cor_cap))
         capacity_variance.append(np.var(cor_cap))
 

### 2020-07-21 documentation/Paper/background_opening.tex
cleaning up report stucture

The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards. Industry research for the 6G standard has already investigated systems using frequency bands in the 120, 200 and 340 GHz range \cite{Koziol}. This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}. One primary hurdle towards efficient communication at these frequencies is the high attenuation of radio energy due both to scattering effects and molecular absorption. \cite{TODO} show through extensive data collection that relevant attenuation will be (x). 
As a relevant example, radar systems have long used this band of frequencies due to the resulting high resolution. For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO}. 
Countering this type of loss is a key step towards meeting the increased data rates and reliability required for future networks. 
\par
In previous generations of wireless communications, attenuation due to path loss has been alleviated by reducing the effective transmission distance 
and supplementing the transmission power using relays \cite{dahlman20134g}. Two, key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission (CITE?). Deploying enough relays to counter
the attenuation levels relevant to new standards would pose impractical costs. One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable electromagnetic reflection and whose reflective coefficient can be controlled to select the phase of the reflected wave. In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
While this work focuses on the theory of communications channels with such surfaces, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an RIS with 224 elements was built and tested. 
\par
The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
associated with a relay. In addition, the absence of an rf-chain on the RIS offers some benefits over relays. First, reflection from an RIS does not introduce the noise associated with an rf-chain. Second, without the band-pass filters of an rf-chain, RISs have a frequency response over the entire spectrum (explain). Third, the RIS is inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay. 
The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
Consider a single antenna, single receiver channel using an RIS with $N$ reflective elements. In this case, the phase control at the RIS can be used to add the multiple paths coherently at the receiver and intuitively increase the received signal power. While such results are motivating, there are two points that should be noted. First, in this example from \cite{basar2019wireless}, the received power would still scale with $N^2$ even if the reflective surfaces had no control over the reflection phase. Second, it is important to note that scaling of received SNR proportional to $N^2$ will only hold to the point limited by energy conservation \cite{bjornson2019demystifying}.
This leaves us with a question. What is the true potential benefits for the RIS in future systems?
Before investigating this question we review key aspects of MIMO systems need to fully develop a system model with which to explore the RIS.

### 2020-07-21 documentation/Paper/correlation_review.tex
cleaning up report stucture

--- 2020-06-24 documentation/Paper/correlation_review.tex
@@
 \par
-Investigation leading to the current interest in MIMO systems showed the key property that (with constant transmit power), capacity (and mutual information) in an $N \times N$ MIMO system with independent channels scales linearly in $N$ \cite{foschini1998limits}.
-However the assumption of independent channels between each pair of transmit and receive antennas is not generally reasonable.
-This prompted much investigation into the impact of correlation to this scaling.
+Investigation leading to the current interest in MIMO systems showed the key property that with constant transmit power, the capacity (and mutual information) in an $N \times N$ MIMO system with independent, i.i.d Gaussian channels scales linearly in $N$ \cite{foschini1998limits}.
+While the assumption of \emph{identically distributed} channels is reasonable when adding more antennas, the assumption of \emph{independent} channels between each pair of transmit and receive antennas is not.
+To understand the potential for leveraging this capacity scaling in deployed massive MIMO systems, the investigation of correlated MIMO channels has gained substantial interest.
 In the following, a number of these studies are reviewed.
-These chosen works use channel correlation models similar to the one we will develop for our system model.
-In particular, this models the correlation between antennas as decreasing  exponentially with the distance between antennas.
+Some of these chosen works use channel correlation models similar to the one we will develop for the RIS system model.
+In particular, this method models the correlation between antennas as decreasing  exponentially with the distance between antennas.
 Definitely need to show this is a valid model (Cite).
 \par
-1. Looking at correlation for finite size systems.
-In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?).
-(Is this a bigger problem for higher frequencies?).
-Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
+Might need to give more motivation for the use of asymptotic analysis.
 \par
 One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. 
-This is motivated by the observation that the eigenvalues of a channel covariance matrix converge to a deterministic distribution for many random 
-matrices as  $N \rightarrow \infty$. In the remainder, this will be referred to as the Asymptotic Eigenvalue Distribution (AED). 
-By making this assumption the impact of $N$ is only a scaling for some deterministic value. 
+This is motivated by the observation that if the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution as $N \rightarrow \infty$.
+In the remainder, this will be referred to as the Asymptotic Eigenvalue Distribution (AED).
+In order to perform comparisons across various system configurations, the power of the received signal is held constant by using constant transmit power and by normalizing the variance of the channel elements.
+(Example in which received power is always 1).
+Results found while using this normalization be extended by simply scaling the asymptotic result with $N$.
+(Still need to clarify the types of gains that are observed under this normalization).
 \par
-In \cite{loyka2001channel}, an upper bound on the capacity is found using Jensen's inequality.
-For this upper bound, increasing correlation between antennas can be interpreted as a decrease in the average receiver SNR.
-Similarly, \cite{skupch2005free} used multiplicative free convolution from free probability theory to analyze the exponential correlation model at either the transmitting or receiving antenna.
+Evaluating the capacity for the case in which the transmitter has channel state information (CSIT ) can be difficult to analyze as the solution to this optimization is not generally in closed form.
+In \cite{loyka2001channel}, this problem is avoided using an upper bound on the capacity of a general MIMO channel derived from Jensen's inequality.
+Using this upper bound and the exponential correlation model, an interpretation is given in which increasing correlation between antennas corresponds to a decrease in the average receiver SNR.
+Similarly, \cite{chuah2002capacity} investigates the exponentially correlated MIMO system approximations and compares optimal versus equal power distribution for cases with high and low SNR.
+One interesting finding from this work is that for the case with low
+SNR (what does this mean?) the water-filling capacity is actually increased with correlation while the equal power case is only impact by the SNR and not the distribution of the channel eigenvalues. 
+Finally, \cite{skupch2005free} used multiplicative free convolution from free probability theory to analyze the exponential correlation model at either the transmitting or receiving antenna.
+Similarly, they also observe a decrease in the capacity as correlation increase but only at high values of correlation in the system.
 Part of this result will be used in our analysis.
-
-\cite{taricco2008asymptotic} analyzed the correlated ricean fading channel using the Replica Method. 
+Finally,
+\cite{taricco2008asymptotic} analyzed the correlated Ricean fading channel using the Replica Method. 
 In general it is observed that the slope of the linear scaling of (mutual information) is reduced by the addition of correlation. 
-One important point to note is that the above works do not consider the case of CSIT. When the water-filling capacity of the 
-correlated MIMO channel is considered,  \cite{chuah2002capacity} notes an exception to this trend. For the case with low
-SNR (what does this mean?) the water-filling capacity is actually increased with correlation.
 \par
-Before moving on to use any of these asymptomatic results, it is fitting to consider whether or not the assumption of  $N \rightarrow \infty$ is reasonable.
+Before moving on to apply any of these asymptomatic results with our own system model, it is fitting to consider two questions. 
+First, whether or not the assumption of  $N \rightarrow \infty$ is reasonable?
 Intuitively, we would expect that under correlation this convergence may occur only at very high $N$.
 This result was confirmed in  \cite{martin2004asymptotic} using the exponential correlation model at both the transmitter and receiver.
+Second, what values for the parameters of the correlation model are reasonable for a real-world system?
+In test data collected with channel parameters (carrier frequency, bandwidth, power...) relevant to 3G and 4G, wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ (linear array?) system \cite{martin2000multiple} (TODO justify if this is a comparison for current systems).

### 2020-07-21 documentation/Paper/free_probability.tex
cleaning up report stucture

--- 2020-07-14 documentation/Paper/paper.tex
@@
-\documentclass[12pt,a4paper]{report}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{float}
-\input defs.tex
-\bibliographystyle{alpha}
-\graphicspath{ {./figures/} }
-%\hyphenpenalty=10000
-
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
-\author{Peter Hartig}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-%
-%\newpage
-\tableofcontents
-\chapter{Introduction}
-\section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
-Industry research for the 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency
-and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-One primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
-(Justify previous statement by comparing attenuation from previous to new standards).
-Countering this loss is a key step towards meeting the increased data rates and reliability required for future networks.
-\par
-In previous generations of wireless communications, attenuation due to path loss (since not at high enough frequencies for absorption) has been alleviated by reducing the effective transmission distance
-and supplementing the transmission power using relays \cite{dahlman20134g}.
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
-(Would be nice to cite some actual costs here).
-Deploying enough relays to counter
-the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many, sub-wavelength sized elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients can be controlled to select the phase of reflected waves.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-\par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-(Compare performance to relays) (detail relay model) as show in \cite{bjornson2019demystifying}, the absence of an rf-chain in the RIS offers some benefits over relays.
-First, reflection does not introduce the noise associated with the rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
-It is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
-
-For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
-(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
-Example of how these surfaces are implemented (Cite).
-
-Discuss how it has be looked at from MISO and MIMO perspective.
-\par
-Need to show how many surfaces are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
-\par
-The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will require systems 
-sizes on the order of x. Now motivate using asymptotic analysis.
-\par
-\input{correlation_review}
-\par
-Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
-
-\par
-Now discuss all of the previous work looking at the optimization of the phases. 
-General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
-
-\section{Notation}
-The following notation is used throughout the remainder. 
-$\Expect[x]$ is the expected value of a random variable $x$.
-Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
-The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
-$\mI_N$ denotes the identity matrix of dimension $N$.
-For square $\mX$ with dimension $N\times N$ the normalized trace is denoted by $\Tr(\mX)  = \underset{N \rightarrow \infty}{\text{lim}}
-\Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
-
-\section{Outline}
-The following sections are now previewed.
-In section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
-
-\chapter{System Model}\label{system_model}
-\section{Power Normalization}
-Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
-the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
-In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
-Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
-In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
-This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
-
-\section{The Correlated IRS Channel}
-In the most general case we consider the channel given by
-
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
-
-\section{Mutual Information}\label{sectiond:mut_info}
-A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel $\mHt$ is known at the receiver, the mutual information is given by
-\begin{equation}\label{mut_ent}
-I(\vy;\vx) = h(\vy) - h(\vy|\vx).
-\end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
-\begin{equation}\label{entropy}
-h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
-\end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
-
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect[\vn\vn^H] = \mI$, substituting equation \eqref{entropy} into equation
-\eqref{mut_ent} gives
-\begin{equation}\label{mut_inf}
-\Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to the
-transmit covariance matrix and the phases at the IRS gives and expression for the capacity of the system
-
-\begin{equation}\label{capacity}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
-
-If $\mHt$ is not known at the transmitter, the transmit power can be allocated
-equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
-\end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-Notice that in this form, the linear scaling of capacity with N is clear.
-Expanding the expectation gives the expression
-\begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
-\end{equation}
-in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
-
-Note that because $\Log$ is a concave function, we can upper bound the mutual information with
-\begin{equation}\label{mut_inf}
-\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda_i])
-\end{equation}
-Now discuss capacity and consider equal power but optimize over phase
-We now want to consider the case in
-
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{Proof of Freeness}
 \begin{itemize}
@@
 \end{itemize}
-
-\chapter{Results}\label{Results}
-\section{Asymptotic Results}
-We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
-phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
-By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
-the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
-\par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity.
-One notable observation is that this rate
-decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
-\begin{itemize}
-\item
-
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
-\end{itemize}
-\subsection{Asymptotic Assumptions Under Correlation}
-First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.
-
-\section{Non-Asymptotic Results}
-In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
-RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
-As seen in \ref{Background} current large systems have been shown for
-sizes (X). We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).
-
-\chapter{Conclusion}
-\chapter{Appendix}
-Proofs
-\begin{enumerate}
-\item
-	Prove freeness of the relevant points. 
-\item 
-	rotations have same non-zero eigenvalues
-\item 
-Proof of split using singular values.\\
-Consider the channel given by 
-\begin{equation}
-\mH_{Total} = \mH_{2} + \mH_{1}.
-\end{equation}
-with covariance matrix 
-\begin{equation}
-\mC =
-\mH_{2}\mathbf{H}_{2}^H + 
-\mH_{1}\mathbf{H}_{1}^H + 
-\mH_{2}\mathbf{H}_{1}^H +
-\mH_{1}\mathbf{H}_{2}^H .
-\end{equation}
-Because the components of this sum are not free (CITE) (and some are non-hermetian), additive free convolution cannot be used.
-We note that using the singular value decomposition $\mH_{Total} = \mU \boldsymbol{\Sigma} \mV^H$ we can
-see that the covariance matrix can be written 
-\begin{equation}
-\mC = \mU \boldsymbol{\Sigma}\boldsymbol{\Sigma}^H \mU^H
-= \mU \boldsymbol{\Lambda} \mU^H.
-\end{equation}
-With $\mC \succeq 0 \implies \lambda(\mC\mC)_i \geq 0$ we see that the definition of the singular values 
-$\sigma(\mC)_i = \sqrt{\lambda(\mC\mC)_i}$ leads to an expression relating the density function of the singular values of a channel matrix
-, $p_{\sigma}(x)$, and the eigenvalues of the channel covariance matrix, $ p_{\lambda}(x)$,  given by $p_{\sigma}(x^2) = p_{\lambda}(x)$.
-Defining a new, symmetric density function of the singular value distribution
- 	 \begin{equation}\label{symmetric}
- 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2},
- 	 \end{equation}
- 	 we now substitute $\tilde{p}_{\sigma}(x)$ into the definition of the Stieljes transform \eqref{stieltjes}
- 	 to give the relationship
- 	 	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}dx
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-With the R-transform definition unchanged for the symmetric distribution, and the additional useful property (CITE)
-\begin{equation}
-\tilde{R}_{\mH_{Total}}(w) = \tilde{R}_{\mH_{2}}(w) + \tilde{R}_{\mH_{1}}
-\end{equation}
-We find that we only need the Stieltjes transform for the covariance matrix of each component of the sum which can in general be found using the tools
-of multiplicative free convolution.
-
-\item 
-Proof of canceling phases.\\
-Consider the channel given by 
-\begin{equation}
-\mathbf{H}_{Total} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}.
-\end{equation}
-This product of matrices suggests the use of multiplicative free convolution to solve for the AED using the S-Transform. 
-Following the same procedure show in \cite{muller2002asymptotic}, we begin with the covariance matrix
-\begin{equation}
-\mC_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H
-\end{equation}
-and note that this has the same non-zero eigenvalues as the matrix (Proof)
-\begin{equation}
-\tilde{\mC}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H.
-\end{equation}
-The relationship between the S-Transform of $\mC_2$ and $\tilde{\mC}_2$ is given by 
-\begin{equation}\label{rotation_property}
-S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
-\end{equation}
-With \begin{equation}
-\mC_1 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
-\end{equation}
-we see that $S_{\tilde{\mC}_2}(z) = S_{\mC_1}(z) S_{\mathbf{H}_{1}\mathbf{H}_{1}^H}(z)$.  Repeating the above procedure on $\mC_1$ we find
-\begin{equation}
-\tilde{\mC}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
-\end{equation}
-with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$ we are left with 
-$S_{\mC_1}(z) =  S_{\mathbf{H}_{2}^H\mathbf{H}_{2}}(z)$. As a result we see that 
-the phase matrix cancels. The same canceling applies to any unitary matrix $\mathbf{A}$ such that   $\mathbf{A}\mathbf{A}^\dagger =\mathbf{A}^\dagger \mathbf{A} = \mI$.
-\end{enumerate}
-\bibliography{bibliography}
-\end{document}

### 2020-07-21 documentation/Paper/information_theory_background.tex
cleaning up report stucture

--- 2020-07-17 documentation/Paper/paper.tex
@@
-\documentclass[12pt,a4paper]{report}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{float}
-\input defs.tex
-\bibliographystyle{alpha}
-\graphicspath{ {./figures/} }
-%\hyphenpenalty=10000
-
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
-\author{Peter Hartig}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-%
-%\newpage
-\tableofcontents
-\chapter{Introduction}
-\section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
-Industry research for the 6G standard has already investigated systems using frequency bands in the 120, 200 and 340 GHz range \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-One primary hurdle towards efficient communication at these frequencies is the higher net attenuation of radio energy due both to scattering effects and molecular absorption.
-\cite{TODO} show through extensive data collection that relevant attenuation will be (x).
-As a relevant example, radar systems have long used this band of frequencies due to the resulting high resolution.
-For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO}.
-Countering this type of loss is a key step towards meeting the increased data rates and reliability required for future networks. 
-\par
-In previous generations of wireless communications, attenuation due to path loss has been alleviated by reducing the effective transmission distance 
-and supplementing the transmission power using relays \cite{dahlman20134g}.
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
-Deploying enough relays to counter
-the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable electromagnetic reflection and whose reflective coefficient can be controlled to select the phase of the reflected wave.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-While this work focuses on the theory surrounding such surfaces, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an RIS with 224 elements was built and tested.
-\par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-The absence of an rf-chain on the RIS offers some benefits over relays.
-First, reflection does not introduce the noise associated with an rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
-While such results are indeed motivating, it is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
-\par
-Need to show how many surfaces (are the corresponding size) are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
-\par
-The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will require systems 
-sizes on the order of x. Now motivate using asymptotic analysis.
-\par
-\input{correlation_review}
-\par
-Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
-\par
-Now discuss all of the previous work looking at the optimization of the phases. 
-General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
-
-\section{Notation}
-The following notation is used throughout the remainder. 
-$\Expect[x]$ is the expected value of a random variable $x$.
-Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
-The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
-$\mI_N$ denotes the identity matrix of dimension $N$.
-For square $\mX$ with dimension $N\times N$ the normalized trace is denoted by $\Tr(\mX)  = \underset{N \rightarrow \infty}{\text{lim}}
-\Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
-
-\section{Outline}
-The following sections are now previewed.
-Section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
-
-\chapter{System Model}\label{system_model}
-\section{Power Normalization}
-Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
-the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
-In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
-Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
-In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
-This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
-
-\section{The Correlated IRS Channel}
-In the most general case we consider the channel given by
-
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
-
-\section{Mutual Information}\label{sectiond:mut_info}
 A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
@@
 \end{equation}
-
-\section{Explanation of Asymptotic Analysis}
-Equation \eqref{no_csi_capacity_aed} highlights the important point that knowledge of the eigenvalue distribution of the received signal co-variance matrix is sufficient information to find the capacity of a MIMO channel.
-As we are generally interested in random channels, finding this eigenvalue distribution is not a straightforward task.
-
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = \Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{Free Probability}\label{Free_Prob_Intro}
-\begin{itemize}
-
-\item 
-	A brief introduction to free probability.
-	\par 
-	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
-	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
-	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
-	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
-	for this investigation but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
-	 \begin{itemize}
-	 \item 
-	 	The cornerstone of free probability is the concept freeness of two random variables in the context of a projection. In our case, freeness between two 
-	 	matrices is given by 
-	 	We now introduce useful tools from free probability  that can be applied to free random variables.
-
-	 	\item 
-	 	Stieltjes Transform: 
-	 	Generally, the projection of matrix is not given in closed form. The projection can be uniquely represented using
-		\begin{equation}\label{stieltjes}
-		G(s) = \int_{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
-		\end{equation}	
-		This representation will allow us to find the projection of the sum and project of matrices using addition and multiplication.
-		\item 
-		R-Transform:
-		Mention self-adjoint
-		Given the Stieltjes transform of a RV, the corresponding R-Transform given by 
-		X. 
-		The R-Transform of the sum of two, free RVs can be found using
-		Y
-		This process use known as free additive convolution.
-		\item
-		S-Transform:
-		Mention self-adjoint
-		Similarly, given the Stieltjes transform of a RV, the S-Transform given by 
-		X, X;
-		The S-Transform of the product of two, free RVs can be found using
-		Y
-		This process is known as free multiplicative convolution.
-	 	\item 
-	 	If there is a case in which finding the singular value distribution is easier than the eigenvalue distribution we can use equation  
-	 	 \eqref{sigular_eigen} to move between the two. Considering a symmetric function of the singular value distribution
-	 	 \begin{equation}\label{symmetric}
-	 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2}
-	 	 \end{equation}
-	 	 the definition of the Steiltjes transform gives the following steps, leading to a relationship between the Stieltjes transforms. 
-	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-	 	\item 
-	 		Another useful property that will be used is the relationship between the eigenvalues of a matrix and its rotation.
-	 		Specifically, consider the covariance matrix given by $\mH_2\mH_1 \mH1^H \mH2^H$.
-	 		By the rotational property of the trace, 
-	 		\begin{align*}
-	 		trace({\mH_2\mH_1 \mH1^H \mH2^H }) = \trace({\mH_1^H \mH_2^H \mH_2 \mH_1 }).
-	 		\end{align*}
-	 		... Not sure if I need to prove this here.  
-	 	\item 
-	  	build up to show that phases cancel via s-transform
-	 	Stieltjes of AED (IE for hermetian matrices)
-	 	Stieltjes of SVD (when they are components covariance matrix)
-	 	Getting Steiltes of SVD using R-Transform
-	 	Finding R-Transform of Individual components
-	 	Using Stieltjes of individuals to get R-transforms
-	 	Using aed stieltjes to get svd individual
-	 	Using S-transform to get stieltjes of individual aed.
-	\end{itemize}
-
-	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
-
-\item 
-	Show that at least for uncorrelated case, the matrices in the sum are free.
-
-\item 
-	Give proofs in an appendix?
-
-\end{itemize}
-
-\section{The IRS Channel}
-Using the approach for finding asyptotic eigenvalue distributions detailed in \ref{Free_Prob_Intro} we will then use the capacity analysis detailed in \ref{sectiond:mut_info}.
-We will now develop a channel model on which to perform this analysis.
-Specifically we develop a channel model that is representative of a MIMO communication system with an RIS.
-\par
-In Section \cite{Background} previous studies were reviewed which showed the impact and relevance of correlation to massive MIMO systems, therefore we begin with an RIS channel given by 
-	\begin{equation}
-	\mH_{Total} = \mR_{R}^{\frac{1}{2}}(\underbrace{\mH_{2}\boldsymbol{\Phi}\mR_{S}^{\frac{1}{2}}\mH_{1}}_{\text{IRS}}.
-	\end{equation}
-
-The random matrices $\mH_{2}$ and $ \mH_{1}$ represent the channels from the transmitter to the IRS and from the IRS to the receiver respectively.
-Both are assumed to have i.i.d.
-elements with variance $\frac{1}{N}$.
-The matrices $\mR_{R}, \mR_{S}$ and $\mR_{T}$ represent a deterministic model for the correlation at the receiver, IRS and transmitter respectively.
-The resulting covariance matrix looks quite similar to the one used for the example in Section \ref{Free_Prob_Intro} but first we want to verify that 
-the resulting components will be free.
-
-(TODO provide references to prove that the components are free)
-
-(Will need to detail any constraints embedded in this system model)
-
-\chapter{Results}\label{Results}
-\section{Asymptotic Results}
-Begin this section using the model from the last
-The final part of this section should show that phases will always cancel out for free probability analysis.
-
-\subsection{Asymptotic Results Discussion}
-
-We first consider the case in which all matrices of the channel given by (ref) are free. As shown in (Appendix), in this case the 
-phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
-By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
-the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
-\par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity.
-One notable observation is that this rate
-decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
-\begin{itemize}
-\item
-
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
-\end{itemize}
-\subsection{Asymptotic Assumptions Under Correlation}
-First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.
-
-\section{Non-Asymptotic Results}
-In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
-RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
-As seen in \ref{Background} current large systems have been shown for
-sizes (X). We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).
-
-\chapter{Conclusion}
-\chapter{Appendix}
-Proofs
-\begin{enumerate}
-\item
-	Prove freeness of the relevant points. 
-\item 
-	rotations have same non-zero eigenvalues
-\item 
-Proof of split using singular values.\\
-Consider the channel given by 
-\begin{equation}
-\mH_{Total} = \mH_{2} + \mH_{1}.
-\end{equation}
-with covariance matrix 
-\begin{equation}
-\mC =
-\mH_{2}\mathbf{H}_{2}^H + 
-\mH_{1}\mathbf{H}_{1}^H + 
-\mH_{2}\mathbf{H}_{1}^H +
-\mH_{1}\mathbf{H}_{2}^H .
-\end{equation}
-Because the components of this sum are not free (CITE) (and some are non-hermetian), additive free convolution cannot be used.
-We note that using the singular value decomposition $\mH_{Total} = \mU \boldsymbol{\Sigma} \mV^H$ we can
-see that the covariance matrix can be written 
-\begin{equation}
-\mC = \mU \boldsymbol{\Sigma}\boldsymbol{\Sigma}^H \mU^H
-= \mU \boldsymbol{\Lambda} \mU^H.
-\end{equation}
-With $\mC \succeq 0 \implies \lambda(\mC\mC)_i \geq 0$ we see that the definition of the singular values 
-$\sigma(\mC)_i = \sqrt{\lambda(\mC\mC)_i}$ leads to an expression relating the density function of the singular values of a channel matrix
-, $p_{\sigma}(x)$, and the eigenvalues of the channel covariance matrix, $ p_{\lambda}(x)$,  given by $p_{\sigma}(x^2) = p_{\lambda}(x)$.
-Defining a new, symmetric density function of the singular value distribution
- 	 \begin{equation}\label{symmetric}
- 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2},
- 	 \end{equation}
- 	 we now substitute $\tilde{p}_{\sigma}(x)$ into the definition of the Stieljes transform \eqref{stieltjes}
- 	 to give the relationship
- 	 	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}dx
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-With the R-transform definition unchanged for the symmetric distribution, and the additional useful property (CITE)
-\begin{equation}
-\tilde{R}_{\mH_{Total}}(w) = \tilde{R}_{\mH_{2}}(w) + \tilde{R}_{\mH_{1}}
-\end{equation}
-We find that we only need the Stieltjes transform for the covariance matrix of each component of the sum which can in general be found using the tools
-of multiplicative free convolution.
-
-\item 
-Proof of canceling phases.\\
-Consider the channel given by 
-\begin{equation}
-\mathbf{H}_{Total} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}.
-\end{equation}
-This product of matrices suggests the use of multiplicative free convolution to solve for the AED using the S-Transform. 
-Following the same procedure show in \cite{muller2002asymptotic}, we begin with the covariance matrix
-\begin{equation}
-\mC_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H
-\end{equation}
-and note that this has the same non-zero eigenvalues as the matrix (Proof)
-\begin{equation}
-\tilde{\mC}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H.
-\end{equation}
-The relationship between the S-Transform of $\mC_2$ and $\tilde{\mC}_2$ is given by 
-\begin{equation}\label{rotation_property}
-S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
-\end{equation}
-With \begin{equation}
-\mC_1 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
-\end{equation}
-we see that $S_{\tilde{\mC}_2}(z) = S_{\mC_1}(z) S_{\mathbf{H}_{1}\mathbf{H}_{1}^H}(z)$.  Repeating the above procedure on $\mC_1$ we find
-\begin{equation}
-\tilde{\mC}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
-\end{equation}
-with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$ we are left with 
-$S_{\mC_1}(z) =  S_{\mathbf{H}_{2}^H\mathbf{H}_{2}}(z)$. As a result we see that 
-the phase matrix cancels. The same canceling applies to any unitary matrix $\mathbf{A}$ such that   $\mathbf{A}\mathbf{A}^\dagger =\mathbf{A}^\dagger \mathbf{A} = \mI$.
-\end{enumerate}
-\bibliography{bibliography}
-\end{document}

### 2020-07-21 documentation/Paper/irs_optimization.tex
cleaning up report stucture

The previous section looked at a select few entries from the substantial literature investigating the capacity of MIMO systems.
This analysis can often be difficult because the definition of capacity implies an optimization over any controlled variables. 
While in select cases this optimization may yield a closed form solution that may be used in the analysis, this is not generally the case. 
In fact, the current literature on RIS system has skipped this analysis entirely because the RIS optimization is a non-convex problem. Rather 
numerous algorithms have been proposed and analyzed numerically. We now briefly review a few selected results from the current RIS literature. 
\par
Go through and review
\par
One key point to note is that in none of the above works and indeed so far as the author's knowledge have these works considered relevant case of a MIMO channel with channel correlation.

### 2020-07-21 documentation/Paper/notation.tex
cleaning up report stucture

--- 2020-07-14 documentation/Paper/paper.tex
@@
-\documentclass[12pt,a4paper]{report}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{float}
-\input defs.tex
-\bibliographystyle{alpha}
-\graphicspath{ {./figures/} }
-%\hyphenpenalty=10000
-
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
-\author{Peter Hartig}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-%
-%\newpage
-\tableofcontents
-\chapter{Introduction}
-\section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
-Industry research for the 6G standard has already investigated 120, 200 and 340 GHz \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency
-and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-One primary hurdle towards efficient communication at these frequencies is an increased attenuation of radio energy.
-(Justify previous statement by comparing attenuation from previous to new standards).
-Countering this loss is a key step towards meeting the increased data rates and reliability required for future networks.
-\par
-In previous generations of wireless communications, attenuation due to path loss (since not at high enough frequencies for absorption) has been alleviated by reducing the effective transmission distance
-and supplementing the transmission power using relays \cite{dahlman20134g}.
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
-(Would be nice to cite some actual costs here).
-Deploying enough relays to counter
-the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many, sub-wavelength sized elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients can be controlled to select the phase of reflected waves.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-\par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-(Compare performance to relays) (detail relay model) as show in \cite{bjornson2019demystifying}, the absence of an rf-chain in the RIS offers some benefits over relays.
-First, reflection does not introduce the noise associated with the rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
-It is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
-
-For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
-(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
-Example of how these surfaces are implemented (Cite).
-
-Discuss how it has be looked at from MISO and MIMO perspective.
-\par
-Need to show how many surfaces are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
-\par
-The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will require systems 
-sizes on the order of x. Now motivate using asymptotic analysis.
-\par
-\input{correlation_review}
-\par
-Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
-
-\par
-Now discuss all of the previous work looking at the optimization of the phases. 
-General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
-
-\section{Notation}
 The following notation is used throughout the remainder. 
@@
 \Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
-
-\section{Outline}
-The following sections are now previewed.
-In section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
-
-\chapter{System Model}\label{system_model}
-\section{Power Normalization}
-Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
-the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
-In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
-Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
-In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
-This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
-
-\section{The Correlated IRS Channel}
-In the most general case we consider the channel given by
-
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
-
-\section{Mutual Information}\label{sectiond:mut_info}
-A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel $\mHt$ is known at the receiver, the mutual information is given by
-\begin{equation}\label{mut_ent}
-I(\vy;\vx) = h(\vy) - h(\vy|\vx).
-\end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
-\begin{equation}\label{entropy}
-h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
-\end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
-
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect[\vn\vn^H] = \mI$, substituting equation \eqref{entropy} into equation
-\eqref{mut_ent} gives
-\begin{equation}\label{mut_inf}
-\Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to the
-transmit covariance matrix and the phases at the IRS gives and expression for the capacity of the system
-
-\begin{equation}\label{capacity}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect [\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)].
-\end{equation}
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of $h(\vx) $ and $h(\vn)$.
-
-If $\mHt$ is not known at the transmitter, the transmit power can be allocated
-equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
-\end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-Notice that in this form, the linear scaling of capacity with N is clear.
-Expanding the expectation gives the expression
-\begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
-\end{equation}
-in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
-
-Note that because $\Log$ is a concave function, we can upper bound the mutual information with
-\begin{equation}\label{mut_inf}
-\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda_i])
-\end{equation}
-Now discuss capacity and consider equal power but optimize over phase
-We now want to consider the case in
-
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{Proof of Freeness}
-\begin{itemize}
-
-\item 
-	A brief introduction to free probability.
-	\par 
-	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
-	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
-	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
-	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
-	for this investigation but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
-	 \begin{itemize}
-	 \item 
-	 	The cornerstone of free probability is the concept freeness of two random variables in the context of a projection. In our case, freeness between two 
-	 	matrices is given by 
-	 	We now introduce useful tools from free probability  that can be applied to free random variables.
-
-	 	\item 
-	 	Stieltjes Transform: 
-	 	Generally, the projection of matrix is not given in closed form. The projection can be uniquely represented using
-		\begin{equation}\label{stieltjes}
-		G(s) = \int_{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
-		\end{equation}	
-		This representation will allow us to find the projection of the sum and project of matrices using addition and multiplication.
-		\item 
-		R-Transform:
-		Mention self-adjoint
-		Given the Stieltjes transform of a RV, the corresponding R-Transform given by 
-		X. 
-		The R-Transform of the sum of two, free RVs can be found using
-		Y
-		This process use known as free additive convolution.
-		\item
-		S-Transform:
-		Mention self-adjoint
-		Similarly, given the Stieltjes transform of a RV, the S-Transform given by 
-		X, X;
-		The S-Transform of the product of two, free RVs can be found using
-		Y
-		This process is known as free multiplicative convolution.
-	 	\item 
-	 	If there is a case in which finding the singular value distribution is easier than the eigenvalue distribution we can use equation  
-	 	 \eqref{sigular_eigen} to move between the two. Considering a symmetric function of the singular value distribution
-	 	 \begin{equation}\label{symmetric}
-	 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2}
-	 	 \end{equation}
-	 	 the definition of the Steiltjes transform gives the following steps, leading to a relationship between the Stieltjes transforms. 
-	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-	 	\item 
-	 		Another useful property that will be used is the relationship between the eigenvalues of a matrix and its rotation.
-	 		Specifically, consider the covariance matrix given by $\mH_2\mH_1 \mH1^H \mH2^H$.
-	 		By the rotational property of the trace, 
-	 		\begin{align*}
-	 		trace({\mH_2\mH_1 \mH1^H \mH2^H }) = \trace({\mH_1^H \mH_2^H \mH_2 \mH_1 }).
-	 		\end{align*}
-	 		... Not sure if I need to prove this here.  
-	 	\item 
-	  	build up to show that phases cancel via s-transform
-	 	Stieltjes of AED (IE for hermetian matrices)
-	 	Stieltjes of SVD (when they are components covariance matrix)
-	 	Getting Steiltes of SVD using R-Transform
-	 	Finding R-Transform of Individual components
-	 	Using Stieltjes of individuals to get R-transforms
-	 	Using aed stieltjes to get svd individual
-	 	Using S-transform to get stieltjes of individual aed.
-	\end{itemize}
-
-	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
-
-\item 
-	Show that at least for uncorrelated case, the matrices in the sum are free.
-
-\item 
-	Give proofs in an appendix?
-
-\end{itemize}
-
-\chapter{Results}\label{Results}
-\section{Asymptotic Results}
-We first consider the case in which all components of the channel given by (ref) are free. As shown in (Appendix), in this case the 
-phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
-By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
-the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
-\par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity.
-One notable observation is that this rate
-decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
-\begin{itemize}
-\item
-
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
-\end{itemize}
-\subsection{Asymptotic Assumptions Under Correlation}
-First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.
-
-\section{Non-Asymptotic Results}
-In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
-RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
-As seen in \ref{Background} current large systems have been shown for
-sizes (X). We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).
-
-\chapter{Conclusion}
-\chapter{Appendix}
-Proofs
-\begin{enumerate}
-\item
-	Prove freeness of the relevant points. 
-\item 
-	rotations have same non-zero eigenvalues
-\item 
-Proof of split using singular values.\\
-Consider the channel given by 
-\begin{equation}
-\mH_{Total} = \mH_{2} + \mH_{1}.
-\end{equation}
-with covariance matrix 
-\begin{equation}
-\mC =
-\mH_{2}\mathbf{H}_{2}^H + 
-\mH_{1}\mathbf{H}_{1}^H + 
-\mH_{2}\mathbf{H}_{1}^H +
-\mH_{1}\mathbf{H}_{2}^H .
-\end{equation}
-Because the components of this sum are not free (CITE) (and some are non-hermetian), additive free convolution cannot be used.
-We note that using the singular value decomposition $\mH_{Total} = \mU \boldsymbol{\Sigma} \mV^H$ we can
-see that the covariance matrix can be written 
-\begin{equation}
-\mC = \mU \boldsymbol{\Sigma}\boldsymbol{\Sigma}^H \mU^H
-= \mU \boldsymbol{\Lambda} \mU^H.
-\end{equation}
-With $\mC \succeq 0 \implies \lambda(\mC\mC)_i \geq 0$ we see that the definition of the singular values 
-$\sigma(\mC)_i = \sqrt{\lambda(\mC\mC)_i}$ leads to an expression relating the density function of the singular values of a channel matrix
-, $p_{\sigma}(x)$, and the eigenvalues of the channel covariance matrix, $ p_{\lambda}(x)$,  given by $p_{\sigma}(x^2) = p_{\lambda}(x)$.
-Defining a new, symmetric density function of the singular value distribution
- 	 \begin{equation}\label{symmetric}
- 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2},
- 	 \end{equation}
- 	 we now substitute $\tilde{p}_{\sigma}(x)$ into the definition of the Stieljes transform \eqref{stieltjes}
- 	 to give the relationship
- 	 	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}dx
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-With the R-transform definition unchanged for the symmetric distribution, and the additional useful property (CITE)
-\begin{equation}
-\tilde{R}_{\mH_{Total}}(w) = \tilde{R}_{\mH_{2}}(w) + \tilde{R}_{\mH_{1}}
-\end{equation}
-We find that we only need the Stieltjes transform for the covariance matrix of each component of the sum which can in general be found using the tools
-of multiplicative free convolution.
-
-\item 
-Proof of canceling phases.\\
-Consider the channel given by 
-\begin{equation}
-\mathbf{H}_{Total} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}.
-\end{equation}
-This product of matrices suggests the use of multiplicative free convolution to solve for the AED using the S-Transform. 
-Following the same procedure show in \cite{muller2002asymptotic}, we begin with the covariance matrix
-\begin{equation}
-\mC_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H
-\end{equation}
-and note that this has the same non-zero eigenvalues as the matrix (Proof)
-\begin{equation}
-\tilde{\mC}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H.
-\end{equation}
-The relationship between the S-Transform of $\mC_2$ and $\tilde{\mC}_2$ is given by 
-\begin{equation}\label{rotation_property}
-S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
-\end{equation}
-With \begin{equation}
-\mC_1 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
-\end{equation}
-we see that $S_{\tilde{\mC}_2}(z) = S_{\mC_1}(z) S_{\mathbf{H}_{1}\mathbf{H}_{1}^H}(z)$.  Repeating the above procedure on $\mC_1$ we find
-\begin{equation}
-\tilde{\mC}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
-\end{equation}
-with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$ we are left with 
-$S_{\mC_1}(z) =  S_{\mathbf{H}_{2}^H\mathbf{H}_{2}}(z)$. As a result we see that 
-the phase matrix cancels. The same canceling applies to any unitary matrix $\mathbf{A}$ such that   $\mathbf{A}\mathbf{A}^\dagger =\mathbf{A}^\dagger \mathbf{A} = \mI$.
-\end{enumerate}
-\bibliography{bibliography}
-\end{document}

### 2020-07-21 documentation/Paper/paper.tex
cleaning up report stucture

@@
 \section{Background}\label{Background}
-The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
-Industry research for the 6G standard has already investigated systems using frequency bands in the 120, 200 and 340 GHz range \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
-One primary hurdle towards efficient communication at these frequencies is the higher net attenuation of radio energy due both to scattering effects and molecular absorption.
-\cite{TODO} show through extensive data collection that relevant attenuation will be (x).
-As a relevant example, radar systems have long used this band of frequencies due to the resulting high resolution.
-For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO}.
-Countering this type of loss is a key step towards meeting the increased data rates and reliability required for future networks. 
-\par
-In previous generations of wireless communications, attenuation due to path loss has been alleviated by reducing the effective transmission distance 
-and supplementing the transmission power using relays \cite{dahlman20134g}.
-Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission.
-Deploying enough relays to counter
-the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable electromagnetic reflection and whose reflective coefficient can be controlled to select the phase of the reflected wave.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-While this work focuses on the theory surrounding such surfaces, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an RIS with 224 elements was built and tested.
-\par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-The absence of an rf-chain on the RIS offers some benefits over relays.
-First, reflection does not introduce the noise associated with an rf-chain.
-Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain).
-Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (?
-Does this hold if we remove the ability to choose the angle?)(Also does this normalize the path power at all?).
-While such results are indeed motivating, it is important to note that scaling of received SNR proportional to $N^2$ will not hold asymptotically due to the conservation of transmitted energy \cite{bjornson2019demystifying}.
-\par
-Need to show how many surfaces (are the corresponding size) are needed in order to be on the same power gain order that is lost by the frequency increase (100dB).
-\par
-The previous section has shown that countering the increased attenuation of high frequencies using RIS and MIMO gains will require systems 
-sizes on the order of x. Now motivate using asymptotic analysis.
+\input{background_opening}
 \par
@@
 \par
-Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation
+\input{irs_optimization}
 \par
-Now discuss all of the previous work looking at the optimization of the phases. 
-General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
+In summary, this section has highlight two primary points.
+First, that channel correlation can degrade the capacity scaling of massive MIMO systems.
+Second, that an RIS may provide additional degrees of freedom with which to increase capacity.
+We now investigate these questions be formulating and analyzing and RIS-aided, MIMO system with channel correlation.
 
 \section{Notation}
-The following notation is used throughout the remainder. 
-$\Expect[x]$ is the expected value of a random variable $x$.
-Bold font, lower-case letters ($\vx$) and upper-case letters ($\mX$) denote vector and matrices respectively and all vectors are assumed to be column vectors unless noted otherwise. 
-The transpose and hermetian of $\mX$ are denoted by $\mX^T$ and $\mX^H$ respectively.
-$\mI_N$ denotes the identity matrix of dimension $N$.
-For square $\mX$ with dimension $N\times N$ the normalized trace is denoted by $\Tr(\mX)  = \underset{N \rightarrow \infty}{\text{lim}}
-\Expect[\frac{1}{N}\sum_{i=1}^N x_{i,i}]$
-
+\input{notation}
 \section{Outline}
 The following sections are now previewed.
-Section \ref{system_model} covers the system model used throughout as well as a review of the necessary theory to characterize the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results where appropriate are discussed.
+Section \ref{system_model} covers the system model used in the remainder as well as a review of the necessary theory to analyze the resulting channel.
+In section \ref{Results}, the theoretical results and corresponding numerical results are discussed.
 
@@
 \section{Power Normalization}
-Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
-the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
-In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
-Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
-In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
-This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
+\input{power_normalization}
+\section{The Correlated IRS Channel}
+In general, the received signal from a line of sight path between the transmitter and receiver would contain much more power than any reflected signal.
+We consider only the reflected portion of this path in the system model given by
 
-\section{The Correlated IRS Channel}
-In the most general case we consider the channel given by
-
+%	\begin{equation}
+%	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+%	\end{equation}
 	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}}\mathbf{R}_{T}^{\frac{1}{2}}.
 	\end{equation}
-
+The random channels $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ represent the paths from transmitter to RI and from RIS to receiver, respectively.
+Similarly, $ \mathbf{R}_{R},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively.
+This correlation will always be modeled by the so-called exponential correlation model. As seen in the sequence this model for correlation will enable a very clean analysis.
 \section{Mutual Information}\label{sectiond:mut_info}
-A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel $\mHt$ is known at the receiver, the mutual information is given by
-\begin{equation}\label{mut_ent}
-I(\vy;\vx) = h(\vy) - h(\vy|\vx).
-\end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is also circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
-\begin{equation}\label{entropy}
-h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
-\end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
-
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$, substituting equation \eqref{entropy} into equation
-\eqref{mut_ent} and simplifying gives
-\begin{equation}\label{mut_inf}
-\Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right].
-\end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to both the
-transmit covariance matrix and phases at the IRS gives an expression for the system capacity
-
-\begin{equation}\label{capacity}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right]
-\end{equation}
-or equivalently,
-\begin{equation}\label{capacity_tricky}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(\mathbf{I} + \lambda_i)\right].
-\end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mQ_x \mHt^H$.
-
-Recalling that in general we are interested in evaluating the capacity in order to determine a feasible transmission rate for a given communication channel, equation \eqref{capacity_tricky} poses a couple of hurdles to evaluation.
-First, we need to be able to evaluate an expectation over the joint pdf $p(\lambda_1 \cdots	 \lambda_N)$.
-Second, there is an optimization problem whose solution will generally depend on the distribution of the channel which would require a closed-form of the optimization problem solution.
-Two simplifications will be used in order to move forward. 
-First, we will enforce that the transmitter chooses a deterministic $\mQ_x$. 
-By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the optimization.
-Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge.
-
-If $\mHt$ is not known at the transmitter, a common choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$, giving the simplified form of equation \eqref{capacity}
-\begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R \Expect[\Log(1 + \frac{P_{\text{total}}}{\sigma_nN_t}\lambda_i)].
-\end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-In this form, the linear scaling of capacity with N is clear.
-If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation to get the expression
-\begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
-\end{equation}
-in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
-In the following we will refer to this distribution as the Asymptotic Eigenvalue Distribution (AED).
-
-From this, we see that if the normalized channel has a converging eigenvalue distribution we can simply scale the capacity for a single channel by the system size $N_R$.
-
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
-(Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).
-
-Note that because $\Log$ is a concave function, we can upper bound the mutual information with
-\begin{equation}\label{mut_inf}
-\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda])
-\end{equation}
-
+\input{information_theory_background}
 \section{Explanation of Asymptotic Analysis}
-Equation \eqref{no_csi_capacity_aed} highlights the important point that knowledge of the eigenvalue distribution of the received signal co-variance matrix is sufficient information to find the capacity of a MIMO channel.
-As we are generally interested in random channels, finding this eigenvalue distribution is not a straightforward task.
-
+Equation \eqref{no_csi_capacity_aed} highlights the important point that knowledge of the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
+For general, random channels, finding this eigenvalue distribution is not a straightforward task.
 \section{Motivation for AED as a projection of Matrices}
@@
 This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
+
 \section{Free Probability}\label{Free_Prob_Intro}
-\begin{itemize}
-
-\item 
-	A brief introduction to free probability.
-	\par 
-	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
-	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
-	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
-	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
-	for this investigation but for a thorough coverage of these topics, see \cite{tulino2004random} and \cite{mingo2017free}.
-	 \begin{itemize}
-	 \item 
-	 	The cornerstone of free probability is the concept freeness of two random variables in the context of a projection. In our case, freeness between two 
-	 	matrices is given by 
-	 	We now introduce useful tools from free probability  that can be applied to free random variables.
-
-	 	\item 
-	 	Stieltjes Transform: 
-	 	Generally, the projection of matrix is not given in closed form. The projection can be uniquely represented using
-		\begin{equation}\label{stieltjes}
-		G(s) = \int_{-\infty}^{\infty} \frac{dp(x)}{(x - s)} \; \text{for} \; \img(s) > 0
-		\end{equation}	
-		This representation will allow us to find the projection of the sum and project of matrices using addition and multiplication.
-		\item 
-		R-Transform:
-		Mention self-adjoint
-		Given the Stieltjes transform of a RV, the corresponding R-Transform given by 
-		X. 
-		The R-Transform of the sum of two, free RVs can be found using
-		Y
-		This process use known as free additive convolution.
-		\item
-		S-Transform:
-		Mention self-adjoint
-		Similarly, given the Stieltjes transform of a RV, the S-Transform given by 
-		X, X;
-		The S-Transform of the product of two, free RVs can be found using
-		Y
-		This process is known as free multiplicative convolution.
-	 	\item 
-	 	If there is a case in which finding the singular value distribution is easier than the eigenvalue distribution we can use equation  
-	 	 \eqref{sigular_eigen} to move between the two. Considering a symmetric function of the singular value distribution
-	 	 \begin{equation}\label{symmetric}
-	 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2}
-	 	 \end{equation}
-	 	 the definition of the Steiltjes transform gives the following steps, leading to a relationship between the Stieltjes transforms. 
-	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-	 	\item 
-	 		Another useful property that will be used is the relationship between the eigenvalues of a matrix and its rotation.
-	 		Specifically, consider the covariance matrix given by $\mH_2\mH_1 \mH1^H \mH2^H$.
-	 		By the rotational property of the trace, 
-	 		\begin{align*}
-	 		trace({\mH_2\mH_1 \mH1^H \mH2^H }) = \trace({\mH_1^H \mH_2^H \mH_2 \mH_1 }).
-	 		\end{align*}
-	 		... Not sure if I need to prove this here.  
-	 	\item 
-	  	build up to show that phases cancel via s-transform
-	 	Stieltjes of AED (IE for hermetian matrices)
-	 	Stieltjes of SVD (when they are components covariance matrix)
-	 	Getting Steiltes of SVD using R-Transform
-	 	Finding R-Transform of Individual components
-	 	Using Stieltjes of individuals to get R-transforms
-	 	Using aed stieltjes to get svd individual
-	 	Using S-transform to get stieltjes of individual aed.
-	\end{itemize}
-
-	Now describe with S-Transform and show that it can have a singular distribution in the limit at we go to rank 1.
-
-\item 
-	Show that at least for uncorrelated case, the matrices in the sum are free.
-
-\item 
-	Give proofs in an appendix?
-
-\end{itemize}
-
+\input{free_probability}
 \section{The IRS Channel}

### 2020-07-21 documentation/Paper/power_normalization.tex
cleaning up report stucture

Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.

### 2020-07-22 documentation/Paper/free_probability.tex
added some analysis results to the report and setup for next steps

@@
 	\par 
-	Free probability is the counterpart to classical probability for the case in which random variables (in this case random matrices) do 
-	not commute. In the same way that classical free probability provides methods to find the pdf for the products and sums of commuting random
-	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables. It is important to note
-	 that non-commutativity of matrices affects not only the pdf of matrix multiplications but also matrix sums. 
+	Free probability is the counterpart to classical probability used to analyze cases in which random variables (in this case random matrices) do 
+	not commute. In the same way that classical probability provides methods to find the pdf for the sums and products of commuting random
+	 variables, free probability provides tools to find the pdfs for products and sums of non-commuting random variables.
 	In the following work, a number of tools from free probability will be used. These tools are now introduced with appropriate context and intuition 
@@
 	 	matrices is given by
-	 	We now introduce useful tools from free probability  that can be applied to free random variables.
 

### 2020-07-22 documentation/Paper/information_theory_background.tex
added some analysis results to the report and setup for next steps

@@
-A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal.
-If the channel $\mHt$ is known at the receiver, the mutual information is given by
+Consider the general MIMO communication system given by 
+	\begin{equation}\label{com_system}
+	\vy = \mH_{Total}\vx + \vn
+	\end{equation}
+	in which $\mH_{Total}$ is a random channel (we can always represent linearly?).
+A key metric of interest for this system is the mutual information between the transmitted and received signal given by 
 \begin{equation}\label{mut_ent}
@@
 \end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
-
+ for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$.
+Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
+(explain what implication this has)
 For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$, substituting equation \eqref{entropy} into equation
@@
 \end{equation}
-Using the model for $\mHt$ from the previous section, and maximizing with respect to both the
-transmit covariance matrix and phases at the IRS gives an expression for the system capacity
+Maximizing the mutual informationwith respect to both the
+transmit covariance matrix and phases at the RIS gives an expression for the system capacity
 
@@
 \begin{equation}\label{capacity_tricky}
-\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(\mathbf{I} + \lambda_i)\right].
+\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(1 + \lambda_i)\right].
 \end{equation}
@@
 First, we need to be able to evaluate an expectation over the joint pdf $p(\lambda_1 \cdots	 \lambda_N)$.
-Second, there is an optimization problem whose solution will generally depend on the distribution of the channel which would require a closed-form of the optimization problem solution.
-Two simplifications will be used in order to move forward. 
+Second, there is an optimization problem whose solution may in general depend on the distribution of the channel.
+Including this solution in the capacity expression without a closed form may be difficult.
+To resolve these hurdles, two simplifications will be used.
 First, we will enforce that the transmitter chooses a deterministic $\mQ_x$. 
-By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the optimization.
+By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the expectation.
 Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge.
 
-If $\mHt$ is not known at the transmitter, a common choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$, giving the simplified form of equation \eqref{capacity}
+If $\mHt$ is not known at the transmitter, an intuitive choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$ which leads to the form of equation \eqref{capacity}
 \begin{equation}\label{no_csi_capacity}
-\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R \Expect[\Log(1 + \frac{P_{\text{total}}}{\sigma_nN_t}\lambda_i)].
+\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R \Expect \left[\Log(1 + \frac{P_{\text{total}}}{\sigma_nN_t}\lambda_i)\right].
 \end{equation}
 where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-In this form, the linear scaling of capacity with N is clear.
-If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation to get the expression
+In this form, we can clearly see the linear scaling of capacity with $N_R$ that has motivated MIMO research.
+If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation over the AED to get the expression
 \begin{equation}\label{no_csi_capacity_aed}
-\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
+\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R  \int_{0}^{\infty}\Log(1 + \frac{1}{\sigma_n}x)p_{\lambda\lambda^H}(x) dx
 \end{equation}
-in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
-In the following we will refer to this distribution as the Asymptotic Eigenvalue Distribution (AED).
-
-From this, we see that if the normalized channel has a converging eigenvalue distribution we can simply scale the capacity for a single channel by the system size $N_R$.
-
-Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
+in which $p_{\lambda\lambda^H}(x)$ is the channel AED and the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
+Note that because the choice for the distributions of $\vx$ and $\vn$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
 (Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).

### 2020-07-22 documentation/Paper/irs_channel.tex
added some analysis results to the report and setup for next steps

Using the approach for finding asymptotic eigenvalue distributions detailed in \ref{Free_Prob_Intro} we now analyze a channel model that is representative of a MIMO communication system with an RIS. 
\par
In general, the received signal from a line of sight path between a transmitter and receiver would contain  sufficient energy to disregard the contribution of any reflected signal.  Therefore, we focus on a channel model in which all signal power to the receiver is through the path of the RIS  given by

%	\begin{equation}
%	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
%	\end{equation}
	\begin{equation}\label{correlated_channel}
	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}\mathbf{H}_{2}\underbrace{\phase}_{\text{IRS}}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}\mathbf{R}_{T}^{\frac{1}{2}}.
	\end{equation}
The random channels $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ represent the random channel gains from transmitter to RIS and from RIS to receiver, respectively. Similarly, $ \mathbf{R}_{T},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively. 
This correlation will always be modeled by the so-called exponential correlation model. As seen in the later, this model for correlation will enable a very clean analysis.
\par

To begin using the tools of free probability, we first ensure that the components of the covariance matrix given by 
	\begin{equation}
	\mC = \mR_{R}^{\frac{1}{2}}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}^{\frac{1}{2}}
	\left(\mR_{R}^{\frac{1}{2}}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}^{\frac{1}{2}}\right)^H.
	\end{equation}
are free. 
Noting from Equation \eqref{no_csi_capacity_aed} that the capacity is only affected by the non-zero eigenvalues of the covariance matrix allows us to equivalently 
consider the AED of the covariance matrix given by 
	\begin{align}
	\mC &  =  \mH_{1}^H \mR_{S}^{\frac{H}{2}}\phase^H \mH_{2}^H \mR_{R}^{\frac{H}{2}}
	 \mR_{R}^{\frac{1}{2}}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}^{\frac{1}{2}}\mR_{T}^{\frac{H}{2}}
	 \\
	 &  =  
	 \mH_{1}^H \mR_{S}^{\frac{H}{2}}\phase^H \mH_{2}^H \mR_{R}^{\frac{H}{2}}
	 \mR_{R}^{\frac{1}{2}}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}.
	\end{align}
This structure appears to admit analysis through the use of the S-Transform and multiplicative free convolution. In order to continue decomposing this covariance matrix into subcomponents  with known S-Transforms, we exploit a property of the S-Transform given by (proven in Appendix \ref{rotation_property})
\begin{equation}
S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
\end{equation}
If all system dimensions are $N$, these S-Trasforms (and AEDs) identical so for this section we will assume that all system dimensions are $N$.
Iterating the rotational property gives the equivalent matrix
	\begin{equation}
	\mC =
\underbrace{\phase^H \phase}_{\mI}
	\mR_{S}
	 \mH_{2}^H \mH_{2}
	 \mH_{1}\mH_{1}^H 
	  \mR_{R}
	 \mR_{T}.
	\end{equation}
Without any further analysis this expression shows the interesting result that in the asymptotic domain, the phases at the RIS (or indeed any unitary matrix in the channel) cancels out. For the remaining terms, we simply need an expression for the S-Transform in order to evaluate the asymptotic system capacity.
\par 
While we do not consider a line of sight component to this channel, it is worth noting that using the tools from a similar analysis in \cite{}, it can be shown that 
a similar cancellation of phases will occur. 
(also mention the non-zero mean aspect here?)

### 2020-07-22 documentation/Paper/paper.tex
added some analysis results to the report and setup for next steps

@@
 \input{power_normalization}
-\section{The Correlated IRS Channel}
-In general, the received signal from a line of sight path between the transmitter and receiver would contain much more power than any reflected signal.
-We consider only the reflected portion of this path in the system model given by
-
-%	\begin{equation}
-%	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-%	\end{equation}
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}}\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
-The random channels $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ represent the paths from transmitter to RI and from RIS to receiver, respectively.
-Similarly, $ \mathbf{R}_{R},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively.
-This correlation will always be modeled by the so-called exponential correlation model. As seen in the sequence this model for correlation will enable a very clean analysis.
 \section{Mutual Information}\label{sectiond:mut_info}
@@
 \section{Explanation of Asymptotic Analysis}
-Equation \eqref{no_csi_capacity_aed} highlights the important point that knowledge of the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
+Equation \eqref{no_csi_capacity_aed} highlights the important point that the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
 For general, random channels, finding this eigenvalue distribution is not a straightforward task.
+We now introduce some of the key tools that will be used in order to perform this analysis.
+\section{Free Probability}\label{Free_Prob_Intro}
+`\input{free_probability}
 \section{Motivation for AED as a projection of Matrices}
@@
 This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-
-\section{Free Probability}\label{Free_Prob_Intro}
-\input{free_probability}
 \section{The IRS Channel}
-Using the approach for finding asyptotic eigenvalue distributions detailed in \ref{Free_Prob_Intro} we will then use the capacity analysis detailed in \ref{sectiond:mut_info}.
-We will now develop a channel model on which to perform this analysis.
-Specifically we develop a channel model that is representative of a MIMO communication system with an RIS.
-\par
-In Section \cite{Background} previous studies were reviewed which showed the impact and relevance of correlation to massive MIMO systems, therefore we begin with an RIS channel given by 
-	\begin{equation}
-	\mH_{Total} = \mR_{R}^{\frac{1}{2}}(\underbrace{\mH_{2}\boldsymbol{\Phi}\mR_{S}^{\frac{1}{2}}\mH_{1}}_{\text{IRS}}.
-	\end{equation}
-
-The random matrices $\mH_{2}$ and $ \mH_{1}$ represent the channels from the transmitter to the IRS and from the IRS to the receiver respectively.
-Both are assumed to have i.i.d.
-elements with variance $\frac{1}{N}$.
-The matrices $\mR_{R}, \mR_{S}$ and $\mR_{T}$ represent a deterministic model for the correlation at the receiver, IRS and transmitter respectively.
-The resulting covariance matrix looks quite similar to the one used for the example in Section \ref{Free_Prob_Intro} but first we want to verify that 
-the resulting components will be free.
-
-(TODO provide references to prove that the components are free)
-
-(Will need to detail any constraints embedded in this system model)
+\input{irs_channel}
 
@@
 \section{Asymptotic Results}
-Begin this section using the model from the last
-The final part of this section should show that phases will always cancel out for free probability analysis.
 
@@
 RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
+Second, that for matrices of random elements, the  convergence of the matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
+Aligning these results with previous results showing the benefits of exploiting RIS phases in Section \ref{Background} requires further defining the parameters of systems which can be considered asymptotic.
+In order to at least numerically investigate this question we now compare the performance of RIS MIMO systems with parameters reflecting the state of the art. 
  As seen in \ref{Background} current large systems have been shown for 
-sizes (X). We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).
-
+sizes S= 224. We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify). 
 \chapter{Conclusion}

### 2020-07-23 documentation/Paper/appendix.tex
Report content on evaluating AED

--- 2020-07-22 documentation/Paper/paper.tex
@@
-\documentclass[12pt,a4paper]{report}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
-\usepackage{float}
-\input defs.tex
-\bibliographystyle{alpha}
-\graphicspath{ {./figures/} }
-%\hyphenpenalty=10000
-
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
-\author{Peter Hartig}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-%
-%\newpage
-\tableofcontents
-\chapter{Introduction}
-\section{Background}\label{Background}
-\input{background_opening}
-\par
-\input{correlation_review}
-\par
-\input{irs_optimization}
-\par
-In summary, this section has highlight two primary points.
-First, that channel correlation can degrade the capacity scaling of massive MIMO systems.
-Second, that an RIS may provide additional degrees of freedom with which to increase capacity.
-We now investigate these questions be formulating and analyzing and RIS-aided, MIMO system with channel correlation.
-
-\section{Notation}
-\input{notation}
-\section{Outline}
-The following sections are now previewed.
-Section \ref{system_model} covers the system model used in the remainder as well as a review of the necessary theory to analyze the resulting channel.
-In section \ref{Results}, the theoretical results and corresponding numerical results are discussed.
-
-\chapter{System Model}\label{system_model}
-\section{Power Normalization}
-\input{power_normalization}
-\section{Mutual Information}\label{sectiond:mut_info}
-\input{information_theory_background}
-\section{Explanation of Asymptotic Analysis}
-Equation \eqref{no_csi_capacity_aed} highlights the important point that the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
-For general, random channels, finding this eigenvalue distribution is not a straightforward task.
-We now introduce some of the key tools that will be used in order to perform this analysis.
-\section{Free Probability}\label{Free_Prob_Intro}
-`\input{free_probability}
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = \Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{The IRS Channel}
-\input{irs_channel}
-
-\chapter{Results}\label{Results}
-\section{Asymptotic Results}
-
-\subsection{Asymptotic Results Discussion}
-
-We first consider the case in which all matrices of the channel given by (ref) are free. As shown in (Appendix), in this case the 
-phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
-By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
-the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free.
-\par
-While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity.
-One notable observation is that this rate
-decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
-\begin{itemize}
-\item
-
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
-\end{itemize}
-\subsection{Asymptotic Assumptions Under Correlation}
-First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.
-
-\section{Non-Asymptotic Results}
-In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
-RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of the matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
-Aligning these results with previous results showing the benefits of exploiting RIS phases in Section \ref{Background} requires further defining the parameters of systems which can be considered asymptotic.
-In order to at least numerically investigate this question we now compare the performance of RIS MIMO systems with parameters reflecting the state of the art. 
- As seen in \ref{Background} current large systems have been shown for 
-sizes S= 224. We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify). 
-\chapter{Conclusion}
-\chapter{Appendix}
 Proofs
 \begin{enumerate}
+\item 
+	Eigen/Singular Relationship
 \item
@@
 \end{enumerate}
-\bibliography{bibliography}
-\end{document}

### 2020-07-23 documentation/Paper/correlation_review.tex
Report content on evaluating AED

@@
 \par
-Investigation leading to the current interest in MIMO systems showed the key property that with constant transmit power, the capacity (and mutual information) in an $N \times N$ MIMO system with independent, i.i.d Gaussian channels scales linearly in $N$ \cite{foschini1998limits}.
+Early investigation which lead to the current interest in MIMO systems showed the key property that with constant transmit power, the capacity (and mutual information) in an $N \times N$ MIMO system with independent, i.i.d Gaussian channels scales linearly in $N$ \cite{foschini1998limits}.
 While the assumption of \emph{identically distributed} channels is reasonable when adding more antennas, the assumption of \emph{independent} channels between each pair of transmit and receive antennas is not.
@@
 \par
-One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. 
-This is motivated by the observation that if the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution as $N \rightarrow \infty$.
-In the remainder, this will be referred to as the Asymptotic Eigenvalue Distribution (AED).
-In order to perform comparisons across various system configurations, the power of the received signal is held constant by using constant transmit power and by normalizing the variance of the channel elements.
-(Example in which received power is always 1).
-Results found while using this normalization be extended by simply scaling the asymptotic result with $N$.
-(Still need to clarify the types of gains that are observed under this normalization).
+One method that has been shown to reduce the complexity of this calculation is to allow the system size $N \rightarrow \infty$ while normalizing the received power. 
+This is motivated by the observation that when the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution which will be referred to as the Asymptotic Eigenvalue Distribution (AED).
+Because the received power would generally not be normalized as the system size in increased, results found while using this normalization be extended by simply scaling the asymptotic result with $N$.
+(Still need to clarify the types of MIMO gains that are observed under this normalization and how to deal with non-zero mean components).
 \par
-Evaluating the capacity for the case in which the transmitter has channel state information (CSIT ) can be difficult to analyze as the solution to this optimization is not generally in closed form.
+Evaluating the capacity for the case in which the transmitter has channel state information (CSIT ) can be difficult as the solution to this optimization is not generally in a closed-form.
 In \cite{loyka2001channel}, this problem is avoided using an upper bound on the capacity of a general MIMO channel derived from Jensen's inequality.
-Using this upper bound and the exponential correlation model, an interpretation is given in which increasing correlation between antennas corresponds to a decrease in the average receiver SNR.
-Similarly, \cite{chuah2002capacity} investigates the exponentially correlated MIMO system approximations and compares optimal versus equal power distribution for cases with high and low SNR.
+Using this upper bound and the exponential correlation model, an interpretation is given in which increasing correlation between antennas corresponds to a decrease in the average SNR at the receiver.
+Similarly, \cite{chuah2002capacity} investigates the exponentially correlated MIMO system using close approximations to compares optimal versus equal power distribution at the transmitter for cases with high and low SNR.
 One interesting finding from this work is that for the case with low
-SNR (what does this mean?) the water-filling capacity is actually increased with correlation while the equal power case is only impact by the SNR and not the distribution of the channel eigenvalues. 
+SNR (what does this mean?) the water-filling capacity is actually increased with correlation while in all other cases, correlation has a negative impact on the capacity.
+The intuition because this finding is that the equal power case is only impacted by the SNR and not the distribution of the channel eigenvalues (unlike the water-filling case).
 Finally, \cite{skupch2005free} used multiplicative free convolution from free probability theory to analyze the exponential correlation model at either the transmitting or receiving antenna.

### 2020-07-23 documentation/Paper/free_probability.tex
Report content on evaluating AED

@@
+Equation \eqref{no_csi_capacity_aed} highlights the important point that the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
+For general, random channels, finding this eigenvalue distribution is not a straightforward task.
+We now introduce some of the key tools that will be used in order to perform this analysis.
 \begin{itemize}
-
 \item 

### 2020-07-23 documentation/Paper/information_theory_background.tex
Report content on evaluating AED

@@
 To resolve these hurdles, two simplifications will be used.
-First, we will enforce that the transmitter chooses a deterministic $\mQ_x$. 
+First, we will enforce that the transmitter chooses $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$. 
 By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the expectation.
 Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge. 
-
-If $\mHt$ is not known at the transmitter, an intuitive choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$ which leads to the form of equation \eqref{capacity}
+These assumptions lead to the form of equation \eqref{capacity}
 \begin{equation}\label{no_csi_capacity}
@@
 (Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).
-
+\par
+Discuss if $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$ is optimal at high SNR (are we in the low SNR regeime?)
+\par
 Note that because $\Log$ is a concave function, we can upper bound the mutual information with

### 2020-07-23 documentation/Paper/irs_channel.tex
Report content on evaluating AED

@@
 Similarly, $ \mathbf{R}_{T},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively.
-This correlation will always be modeled by the so-called exponential correlation model. As seen in the later, this model for correlation will enable a very clean analysis.
+The structure of $\mathbf{H}_{Total}$ suggests that we will want to use multiplicative free convolution in order to evalute the AED and therefore we would like to choose a   valid correlation model with a known S-Transform expression.
+The exponential correlation model is defined by a single parameter $\rho$ with elements given by $\rho^{i-j}$.
+As seen in the later, this model for correlation will enable a very clean analysis.
 \par
@@
 	\end{equation}
-Without any further analysis this expression shows the interesting result that in the asymptotic domain, the phases at the RIS (or indeed any unitary matrix in the channel) cancels out.
+Now give the S-Transforms explicitly (given the above definitions of the matrices) and show that freeness applies. 
+\begin{enumerate}
+\item 
+	For both $\mH_{2}^H \mH_{2}$ and $\mH_{1}\mH_{1}^H$ with all system dimensions equal, the S-Transform is derived in \cite{} and given by 
+	\begin{equation}
+	S(z) = \frac{1}{1+z}
+	\end{equation}
+\item 
+	As shown in \cite{}, as $N \rightarrow \infty$ the  Toeplitz matrix has a spectrum (and thus AED) of a circulant matrix. Circulant matrices have an AED 
+	given by the (discrete fourier transform ?) of the first row.
+Using the definition of the Stieltjes transform and the definition of the S-Transform in terms of the Stieljes transform the S-Transform of the exponential correlation model is shown to be
+	\begin{equation}
+	x
+	\end{equation}
+\item
+	The final step to evaluating this expression is ensuring that the individual components whose S-Transforms we have found are free and thus able to be use with multiplicative free convolution. 	
+\end{enumerate}
+\par
+Without any further analysis this expression shows the interesting result that under the assumptions of Free Probability and equal power transmission, the phases at the RIS (or indeed any unitary matrix in the channel) cancel out.
 For the remaining terms, we simply need an expression for the S-Transform in order to evaluate the asymptotic system capacity.
 \par 
-While we do not consider a line of sight component to this channel, it is worth noting that using the tools from a similar analysis in \cite{}, it can be shown that 
-a similar cancellation of phases will occur. 
-(also mention the non-zero mean aspect here?)
+Using the previous analysis we can observe an extension of this result that may be relevant to variations of this system model. 
+While we do not consider a line of sight component to this channel, using the tools from a similar analysis in \cite{}, it can be shown that 
+a similar cancellation of phases will occur for the RIS channel with an LOS (or non-zero mean) component of the signal.
+Note that if we are simply adding a non-zero mean component.
+We can find the corresponding AED by looking considering the channel without the non-zero mean and then adding the single eigenvalue from the non-zero mean component; in which case the phases will again cancel out.

### 2020-07-23 documentation/Paper/non_asymptotic.tex
Report content on evaluating AED

Up to this point we have shown an analysis which seems to contradict the observations from previous investigations review in Section \ref{Background} showing the benefits of exploiting RIS phases. The key to the coexistence of these results is determining when the asymptotic assumptions required by Free Probability hold. 
As mentioned in in Section \ref{Background}, \cite {} have shown that for matrices of random elements, the  convergence of the matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements. We would now like to at least numerically investigate this question to investigate the performance of RIS MIMO systems with parameters reflecting state of the art.

 As seen in \ref{Background} current large systems have been shown for 
sizes S= 224. We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).

### 2020-07-23 documentation/Paper/paper.tex
Report content on evaluating AED

@@
 \par
+\subsection{MIMO with correlation}\label{mimo_corr}
 \input{correlation_review}
 \par
+\subsection{RIS Optimization}\label{irs_opt}
 \input{irs_optimization}
@@
 
-\chapter{System Model}\label{system_model}
-\section{Power Normalization}
+\chapter{Analyzing the RIS System}\label{system_model}
+\section{Analysis Preliminaries}
+\subsection{Power Normalization}
 \input{power_normalization}
-\section{Mutual Information}\label{sectiond:mut_info}
+\subsection{Mutual Information}\label{sectiond:mut_info}
 \input{information_theory_background}
-\section{Explanation of Asymptotic Analysis}
-Equation \eqref{no_csi_capacity_aed} highlights the important point that the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
-For general, random channels, finding this eigenvalue distribution is not a straightforward task.
-We now introduce some of the key tools that will be used in order to perform this analysis.
-\section{Free Probability}\label{Free_Prob_Intro}
-`\input{free_probability}
-\section{Motivation for AED as a projection of Matrices}
-Equation \ref{no_csi_capacity_aed} shows that the capacity of a communication channel can be found using the AED of the channel covariance matrix.
-For matrices with converging eigenvalue distributions we can use $p_{\lambda\lambda^H} = \Expect[\Tr{(mHt \mHt^H})] = $.
-One property that will be useful in the following is the relationship between the eigenvalues of a channel covariance matrix and the singular values of
-the channel.
-For a general matrix $\mH \in \complex^{n \times m}$ the singular values are given by 
-\begin{equation}\label{sigular_eigen}
-\sigma(\mH)_i = \sqrt{\lambda({\mH\mH^H}_i)}.
-\end{equation}
-This also gives the relationship between the pdf for the singular and and eigenvalues $p_{\sigma}(x^2) = p_{\lambda}(x)$
-\section{The IRS Channel}
+\subsection{Free Probability}\label{Free_Prob_Intro}
+\input{free_probability}
+\section{RIS Analysis}
+\subsection{The Asyptotic IRS Channel}
 \input{irs_channel}
-
+\subsection{Non-Asymptotic Results}
+\input{non_asymptotic}
 \chapter{Results}\label{Results}
@@
 
-	Show that in the asymptotic case where free probability holds, the phase matrix no longer matters
-	We first show that under assumptions of the channel which allow for use of free probability, $\phase$
-	cancels out of \eqref{capacity}.
-
-\item
-	Show that numerical agrees with theoretical for uncorrelated 
-\item
-	Now consider the case where the assumptions of free probability no longer apply 
-	Show how numerical has larger deviation from theoretic for the case of the correlated channel and perhaps 
-	look at methods of phase optimization that can help counter this loss. Also need to show proof that for the
-	correlated case, free probability assumptions no longer hold. 
 \end{itemize}
@@
 
-\section{Non-Asymptotic Results}
-In the previous section, two notable observations were made. First, that under the assumptions of Free Probability (Ref), the choice of $\phase$ at the 
-RIS has no impact on the capacity of the system.
-Second, that for matrices of random elements, the  convergence of the matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
-Aligning these results with previous results showing the benefits of exploiting RIS phases in Section \ref{Background} requires further defining the parameters of systems which can be considered asymptotic.
-In order to at least numerically investigate this question we now compare the performance of RIS MIMO systems with parameters reflecting the state of the art. 
- As seen in \ref{Background} current large systems have been shown for 
-sizes S= 224. We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify). 
 \chapter{Conclusion}
 \chapter{Appendix}
-Proofs
-\begin{enumerate}
-\item
-	Prove freeness of the relevant points. 
-\item 
-	rotations have same non-zero eigenvalues
-\item 
-Proof of split using singular values.\\
-Consider the channel given by 
-\begin{equation}
-\mH_{Total} = \mH_{2} + \mH_{1}.
-\end{equation}
-with covariance matrix 
-\begin{equation}
-\mC =
-\mH_{2}\mathbf{H}_{2}^H + 
-\mH_{1}\mathbf{H}_{1}^H + 
-\mH_{2}\mathbf{H}_{1}^H +
-\mH_{1}\mathbf{H}_{2}^H .
-\end{equation}
-Because the components of this sum are not free (CITE) (and some are non-hermetian), additive free convolution cannot be used.
-We note that using the singular value decomposition $\mH_{Total} = \mU \boldsymbol{\Sigma} \mV^H$ we can
-see that the covariance matrix can be written 
-\begin{equation}
-\mC = \mU \boldsymbol{\Sigma}\boldsymbol{\Sigma}^H \mU^H
-= \mU \boldsymbol{\Lambda} \mU^H.
-\end{equation}
-With $\mC \succeq 0 \implies \lambda(\mC\mC)_i \geq 0$ we see that the definition of the singular values 
-$\sigma(\mC)_i = \sqrt{\lambda(\mC\mC)_i}$ leads to an expression relating the density function of the singular values of a channel matrix
-, $p_{\sigma}(x)$, and the eigenvalues of the channel covariance matrix, $ p_{\lambda}(x)$,  given by $p_{\sigma}(x^2) = p_{\lambda}(x)$.
-Defining a new, symmetric density function of the singular value distribution
- 	 \begin{equation}\label{symmetric}
- 	 \tilde{p}_{\sigma}(x) = \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{2},
- 	 \end{equation}
- 	 we now substitute $\tilde{p}_{\sigma}(x)$ into the definition of the Stieljes transform \eqref{stieltjes}
- 	 to give the relationship
- 	 	 	 \begin{align*}
-	 	 \tilde{G}_\sigma (s) & =  \frac{1}{2} \int_{-\infty}^{\infty} \frac{p_{\sigma}(x) + p_{\sigma}(-x)}{x-s}dx
-	 	 \\&  =  
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{x - s} + 
-	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{dp_{\sigma}(x)}{-x - s}
-	 	 \\&  =  
-	 	 	 	 \frac{1}{2} \int_{-\infty}^{\infty} \frac{-2s dp_{\sigma}(x)}{(-x - s)(x - s)}
-	  	 \\&  =  
-	 	 	 	\int_{-\infty}^{\infty} \frac{-s dp_{\sigma}(x)}{-x^2 + s^2} 	
- 	 	  	 \\&  =  
- 	 	 	-s  \int_{-\infty}^{\infty}  \frac{dp_{\sigma}(x)}{ x^2 - s^2} 	 
- 	 	 \\&  =  
-			-s  \int_{-\infty}^{\infty}  \frac{dp_{\lambda}(x)}{ x - s^2} 
-	 	 \\&  = 
-	 	 	-s G_{\lambda}(s^2)
-	 	 \end{align*}
-With the R-transform definition unchanged for the symmetric distribution, and the additional useful property (CITE)
-\begin{equation}
-\tilde{R}_{\mH_{Total}}(w) = \tilde{R}_{\mH_{2}}(w) + \tilde{R}_{\mH_{1}}
-\end{equation}
-We find that we only need the Stieltjes transform for the covariance matrix of each component of the sum which can in general be found using the tools
-of multiplicative free convolution.
-
-\item 
-Proof of canceling phases.\\
-Consider the channel given by 
-\begin{equation}
-\mathbf{H}_{Total} = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}.
-\end{equation}
-This product of matrices suggests the use of multiplicative free convolution to solve for the AED using the S-Transform. 
-Following the same procedure show in \cite{muller2002asymptotic}, we begin with the covariance matrix
-\begin{equation}
-\mC_2 = \mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H\boldsymbol{\Phi}^H\mathbf{H}_{2}^H
-\end{equation}
-and note that this has the same non-zero eigenvalues as the matrix (Proof)
-\begin{equation}
-\tilde{\mC}_2 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}\mathbf{H}_{1}^H.
-\end{equation}
-The relationship between the S-Transform of $\mC_2$ and $\tilde{\mC}_2$ is given by 
-\begin{equation}\label{rotation_property}
-S_{C_N}(z) = \frac{z+1}{z+\chi_2} S_{\tilde{C}_N}(\frac{z}{\chi_2}).
-\end{equation}
-With \begin{equation}
-\mC_1 = \boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}\boldsymbol{\Phi}
-\end{equation}
-we see that $S_{\tilde{\mC}_2}(z) = S_{\mC_1}(z) S_{\mathbf{H}_{1}\mathbf{H}_{1}^H}(z)$.  Repeating the above procedure on $\mC_1$ we find
-\begin{equation}
-\tilde{\mC}_{1} = \boldsymbol{\Phi}\boldsymbol{\Phi}^H\mathbf{H}_{2}^H\mathbf{H}_{2}
-\end{equation}
-with $\boldsymbol{\Phi}\boldsymbol{\Phi}^H = \mathbf{I}$ we are left with 
-$S_{\mC_1}(z) =  S_{\mathbf{H}_{2}^H\mathbf{H}_{2}}(z)$. As a result we see that 
-the phase matrix cancels. The same canceling applies to any unitary matrix $\mathbf{A}$ such that   $\mathbf{A}\mathbf{A}^\dagger =\mathbf{A}^\dagger \mathbf{A} = \mI$.
-\end{enumerate}
+\input{appendix}
 \bibliography{bibliography}

### 2020-07-23 documentation/Paper/irs_channel.tex
TODO finish details of system model and assuptions for freeness

@@
 \item 
-	For both $\mH_{2}^H \mH_{2}$ and $\mH_{1}\mH_{1}^H$ with all system dimensions equal, the S-Transform is derived in \cite{} and given by 
+	For both $\mH_{2}^H \mH_{2}$ and $\mH_{1}\mH_{1}^H$ with all system dimensions equal, the S-Transform as shown in \cite{muller2002random} is given by 
 	\begin{equation}
@@
 \item 
-	As shown in \cite{}, as $N \rightarrow \infty$ the  Toeplitz matrix has a spectrum (and thus AED) of a circulant matrix. Circulant matrices have an AED 
-	given by the (discrete fourier transform ?) of the first row.
-Using the definition of the Stieltjes transform and the definition of the S-Transform in terms of the Stieljes transform the S-Transform of the exponential correlation model is shown to be
+	As shown in \cite{gray2006toeplitz}, with $N \rightarrow \infty$ Toeplitz matrices have the spectrum (and thus AED) of an equivalent circulant matrix (whose spectrum and thus AED is given by the discrete Fourier transform of the first row).
+Using the definition of the Stieltjes transform and the definition of the S-Transform in terms of the Stieljes transform, the S-Transform of the exponential correlation model is shown to be \cite{skupch2005free}
 	\begin{equation}
-	x
+	S(z) = \frac{\alpha z^2 + z \sqrt{alpha^2 z^2 -(z^2-1)}}{z-1}
 	\end{equation}

### 2020-07-23 documentation/Paper/paper.tex
TODO finish details of system model and assuptions for freeness

@@
 \par
-In summary, this section has highlight two primary points.
+In summary, this section has highlighted two primary points.
 First, that channel correlation can degrade the capacity scaling of massive MIMO systems.
 Second, that an RIS may provide additional degrees of freedom with which to increase capacity.
-We now investigate these questions be formulating and analyzing and RIS-aided, MIMO system with channel correlation.
-
+We now investigate these questions by formulating and analyzing an RIS-aided, MIMO system with channel correlation.
 \section{Notation}

### 2020-07-27 documentation/Optimization Work/07_24_20.tex
Numerial evidence for phase optimization at high correlations

--- 2020-07-10 documentation/Meeting Notes/07_07_20.tex
@@
 
-\section{Asymptotic Vs. Numerical Results}
-First I confirm that the methods I use to obtain the AED given a closed-form S-Transform work for basic cases.
-The numerical results  used for comparison are generated by averaging 100 realizations of the system for size $N=100 \times 100$.
-As seen in the figures below, the solution becomes incorrect if there is too much correlation in the channel. 
+\section{Numerical Results}
+I use the system model given by 
+\begin{equation}
+\vy = \mH_{Total} \mW \vs  + \vn
+\end{equation}
+with
+
+	\begin{equation}
+	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+	\end{equation}
+
+Using a matched filter  and precoder which satisfies the transmission power constraint gives (normalize each column)
+\begin{equation}
+\mW = \mathbf{H}_{Total}^H
+\end{equation}
+
+Numerical results for $s_k \in \mathcal{NC}(0,1)$, $n_k \in \mathcal{NC}(0,1)$
 \begin{itemize}
 \item
-	Channel 
-	\begin{equation}
-	\mH  \sim \CNO
-	\end{equation}
-	For equal power allocation 
-	\\
-	Numeric Capacity = $  1.4286$ 
-	\\
-	Asymptotic Capacity = $ 1.4129 $
+zf mu: 202.0989509146119
+zf std: 57.23864582855936
+\item
+matched mu: 476.69650954785794
+matched std: 239.3529067968275
+\end{itemize}
 
-\item
-	Channel 
-	\begin{equation}
-	\mH_2 \phase \mH_1  
-	\end{equation}
-	with $\mH \sim \CNO$.
-	For equal power allocation 
-	\\
-	Numeric Capacity = $ 1.4227$ 
-	\\
-	Asymptotic Capacity = $ 1.4060 $
-
-	* This difference between numeric and asymptotic  can be reduced by increasing the AED sampling density.
-
-\item
-	The single Rayleigh channel with receiver correlation given by 
-	\begin{equation}
-	\mC\mH  
-	\end{equation}
-	with $\mC$ as the exponential correlation matrix  with elements $c_{i,j} = \rho^{i-j}$. 
-	For equal power allocation the capacities are plotted below.
-Because the s-transform for the correlation has a quadratic form, I show the solutions for both the $+/-$ forms of the quadratic expression.
-
-	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison}
-	  \caption{Equal power capacity with correlation form 1.}
-	\end{figure}
-	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison2}
-	  \caption{Equal power capacity with correlation form 2.}
-\end{figure}
-
-\item
-	The IRS channel without LOS and correlation only at receiver
-	\begin{equation}
-	\mC\mH_2 \phase\mH_1
-	\end{equation}
-	with $\mC$ as the exponential correlation matrix. 
-	For equal power allocation the capacities are plotted below. 
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/equation1_Rx_only}
-	  \caption{Equal power capacity for IRS channel without LOS and correlation only at receiver.}
-	\end{figure}
-\item
-	The IRS channel without LOS and correlation at receiver and IRS elements
-	\begin{equation}
-	\mC_2 \mH_2\phase\mC_1 \mH_1
-	\end{equation}
-	with $\mC_i$ as an exponential correlation matrix. 
-	For equal power allocation the capacities are plotted below. 
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/Rx_IRS_Correlation}
-	  \caption{Equal power capacity for IRS channel without LOS and correlation at receiver and IRS.}
-	\end{figure}
-
-\item
-	The IRS channel without LOS and correlation at receiver, IRS elements and transmitter
-	\begin{equation}
-	\mC_3 \mH_2\mC_2 \phase\mC_1 \mH_1
-	\end{equation}
-	with $\mC_i$ as the exponential correlation matrix. 
-	For equal power allocation the capacities are plotted below. 
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/Rx_IRS_Tx_Correlation}
-	  \caption{Equal power capacity for IRS channel without LOS and correlation at receiver, IRS and transmitter.}
-	\end{figure}
-
-\item
-	These are the steps I use to move from the closed-form S-Transform of the channel to the AED.
-	The S-Tranform of the exponential correlation model is derived in \cite{skupch2005free}
-	\begin{enumerate}
-	\item
-		Beginning with the S-Transform, $S(z)$, I find $\gamma(s)$ using the fixed-point equation
-		\begin{equation}
-			S(\gamma(z))\gamma(z) - z - z\gamma(z) = 0 
-		\end{equation}
-	\item
-		I find the Stieltjes transform using
-		\begin{equation}
-			G(s) = \frac{1}{s} (1+\gamma(\frac{1}{s}))
-		\end{equation}			
-	\item 
-		I recover the AED using 
-\begin{equation}
-p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
-\end{equation}
-	\end{enumerate}
-
-\item
-	I checked the equation provided for the gamma transform of just the the exponential correlation model and observed the following AEDs.
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/exponential_correlation_aed}
-	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
-	\end{figure}
-
-\end{itemize}
-\section{Capacity Convergence}
-As there is not a method for evaluating the rate of convergence of finite systems towards the asymptotic behavior I investigate this numerically in the following figures by looking at the variance of the capacity of numerical simulations for the IRS channel given by
-	\begin{equation}
-	\mC_3 \mH_2\mC_2 \phase\mC_1 \mH_1
-	\end{equation}
-Figures \ref{equal_power} and \ref{water_filling} show the average capacity and capacity variance over different, random, values of  $\phase$ with all channels held constant for equal power allocation and optimal power allocation respectively.
-Note that \emph{neither} uses any optimization of the phases at the IRS.
-These results reflect the intuition that smaller finite system sizes with significant correlation are not well represented by the asymptotic result that shows the phase matrix at the IRS canceling.
-It might be interesting to see if an optimization of the IRS phases reduces this variance in the high correlation case.
-
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_equal_power_capacity}
-	  \caption{Capacity average and variance for system size $100$ equal power allocation over $N=1000$ channel realizations}
-	  \label{equal_power}
-	\end{figure}
-
-		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_water_filling_capacity}
-	  \caption{Capacity average and variance for system size $100$ equal power allocation over $N=100$ channel realizations}
-	  	  \label{water_filling}
-	\end{figure}
-
-\section{Questions}
-\begin{enumerate}
-\item
-	Because we can show that the phases cancel for the asymptotic analysis, it seems that any further look at this is just extending capacity results without 
-	IRS to a more complicated channel model. Can I focus the remainder of the work on just looking at the when the asymptotic assumption is valid?
-\item
-	To start considering the "zero-tapping" that was suggested in our last meeting.I began by using the S-Transform of the "projector" matrix (essentially resulting in the deformed quarter circle law for $\mH \in \complex^{N \times K}$ with $N \leq K$).
-There are two points, I was hoping to discuss on this.
-	\begin{enumerate}
-	\item
-		First, one problem is that this is also changing the power of the received signal to $\frac{N}{K}$ 
-		 because the deformed quarter circle assumes elements with variance $\frac{1}{K}$ rather than $\frac{1}{N}$.
-\item
-		One other strange thing I noticed is that when I perform numerical simulations with the deformed quarter circle the distributions I have, as shown in Figure \ref{AED} 
-		appear to not integrate to 1 but nevertheless  the resulting capacity they agree with numerical results as shown in Figure \ref{projection}.
-				\begin{figure}[H]
-	\includegraphics[width= 12cm,height = 8cm]{figures/projection}
-	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
-	  \label{projection}
-	\end{figure}
-		\begin{figure}[H]
-	\includegraphics[width= 12cm,height = 8cm]{figures/AED}
-	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
-	  	  \label{AED}
-	\end{figure}
-
-	\end{enumerate}
-\item
-I was also preparing some slides for the mid-term presentation and a question on the background.
-\begin{itemize}
-\item
-	Will it be okay to just assume flat-fading using OFDM rather than including a model with multiple delays? Will OFDM still be possible at very high frequencies?
-\end{itemize}
-\end{enumerate}
 \bibliography{bibliography}

### 2020-07-27 documentation/Paper/irs_optimization.tex
Numerial evidence for phase optimization at high correlations

@@
-The previous section looked at a select few entries from the substantial literature investigating the capacity of MIMO systems.
-This analysis can often be difficult because the definition of capacity implies an optimization over any controlled variables. 
-While in select cases this optimization may yield a closed form solution that may be used in the analysis, this is not generally the case. 
-In fact, the current literature on RIS system has skipped this analysis entirely because the RIS optimization is a non-convex problem. Rather 
+During the review of select entries from the literature investigating the capacity of MIMO systems in the previous section, it was noted that this analysis can be difficult because the definition of capacity implies an optimization over any controlled variables.
+While in select cases this optimization may yield a closed form solution that may be used in the analysis, this is generally not the case.
+As a result, bounds on the capacity are often analyzed instead.
+Modeling the RIS as a diagonal matrix whose diagonal elements are given by 
+	\begin{equation}
+	\phi_i = \exp{j\theta_i}
+	\end{equation}
+	it is clear that any optimization over the possible values of $phi_i$ includes this non-convex constraint to say nothing of the objective function.
+
+In fact, the current literature on RIS systems has skipped this analysis entirely because the RIS optimization is a non-convex problem. Rather, 
 numerous algorithms have been proposed and analyzed numerically. We now briefly review a few selected results from the current RIS literature. 
 \par
-Go through and review
+\begin{itemize}
+\item
+
+	As is often the case with non-convex optimization, different reformulations of the problem have been used in order to find local optima of the problem.
+
+First show that in general the non-convex constraint of the phases requires relaxation or the problem
+\item
+\end{itemize}
 \par

### 2020-07-27 simulation/src/lin_alg.py

@@ -70,6 +70,15 @@ def normalize_matrix(matrix, rows=False):
             ret_mat[row, :] /= np.linalg.norm(matrix[row, :], 2)
         return ret_mat
 
+def precode_mf(channel, normalize=True):
+    ret = hermetian(channel)
+    for col in range(ret.shape[1]):
+        ret[:, col] = ret[:, col]/np.power(np.linalg.norm(ret[:, col]), 2)
+    return ret
+
+def precode_zf(channel):
+    return hermetian(channel) @ np.linalg.pinv(channel @ hermetian(channel))
+
 
 def normalize(x):
     normalized = x/np.linalg.norm(x)

### 2020-07-27 simulation/testing/Optimization/MSE_phase_optimization.py

@@ -1,0 +1,69 @@
+from src.lin_alg import *
+import matplotlib.pyplot as plt
+
+num_tx = 10
+num_rx = 10
+num_reflectors = 50
+average = 100
+cor_mf_mu = []
+cor_mf_var = []
+cor_zf_mu = []
+cor_zf_var = []
+
+#   Measure over average
+rhos = [.01, .5, .9, .99]
+# rhos = [.99]
+for cor_rho in rhos:
+    MSE_zf_mu = []
+    MSE_matched_mu = []
+    MSE_zf_var = []
+    MSE_matched_var = []
+    t_correlation_matrix = exponential_correlation(num_tx, cor_rho)
+    s_correlation_matrix = exponential_correlation(num_reflectors, cor_rho)
+    r_correlation_matrix = exponential_correlation(num_rx, cor_rho)
+    for i in range(average):
+        transmit_symbols = np.random.randn(num_rx)
+        G = c_rand(num_reflectors, num_tx)
+        G2 = c_rand(num_rx, num_reflectors)
+        G3 = c_rand(num_rx, num_tx)
+        MSE_matched = []
+        MSE_zf = []
+        noise = c_rand(num_rx, 1, var=1).flatten()
+        for j in range(average):
+            phase = random_phase(num_reflectors)
+            los_path = G3
+            irs_path = r_correlation_matrix @ G2 @ s_correlation_matrix @ phase @ G @ t_correlation_matrix
+            channel = (irs_path + los_path)
+            transmit_signal_matched = precode_mf(channel) @ transmit_symbols
+            received_matched = channel @ (transmit_signal_matched + noise)
+            MSE_matched.append(np.power(np.linalg.norm(received_matched - transmit_symbols), 2))
+            zf_precoding = precode_zf(channel) @ transmit_symbols
+            received_zf = channel @(zf_precoding + noise)
+            # received_zf = channel @ precode_zf(channel) @ transmit_symbols
+            MSE_zf.append(np.power(np.linalg.norm(received_zf - transmit_symbols), 2))
+        MSE_zf_mu.append(np.average(MSE_zf))
+        MSE_zf_var.append(np.var(MSE_zf))
+        MSE_matched_mu.append(np.average(MSE_matched))
[…]

### 2020-07-28 documentation/Optimization Work/07_24_20.tex
Reduced to single paraemter for IRS phases and added corresponding results to a progress report

@@
 
-Using a matched filter  and precoder which satisfies the transmission power constraint gives (normalize each column)
-\begin{equation}
-\mW = \mathbf{H}_{Total}^H
-\end{equation}
+The following numerical results evaluate the MSE the channel above with $s_k \in \mathcal{NC}(0,\frac{1}{N_R})$, $n_k \in \mathcal{NC}(0,\frac{1}{N_R})$.
+For each channel realization, I evaluate the MSE for the matched filter and zero-forcing filter using $\phase = e^{j\theta}\mI$ for 10 uniform spaced values of $\theta \in [0, 2\pi]$.
+Then I collect the mean and variance of the 10 MSEs for the channel realization.
+I repeat and average across 1000 channel realizations.
+I perform this same analysis across different values of $\rho$ in the exponential correlation model (i.e.
+$r_{i,j}=\rho^(i-j)$.
+The results in Figure \ref{MSE_correlation} indicate that the variance of the MSE over different values of the phase matrix parameter, $\theta$,  can be quite large.
+In particular when the correlation parameter, $\rho$ is close to 1.
+		\begin{figure}[H]
+	\includegraphics[width= 15cm,height = 10cm]{figures/over_10_thetas}
+	  \caption{Comparison MSE over different amounts of correlation.}
+	  	  \label{MSE_correlation}
+	\end{figure}
 
-Numerical results for $s_k \in \mathcal{NC}(0,1)$, $n_k \in \mathcal{NC}(0,1)$
+Based on the measurement data collected in \cite{martin2000multiple}, the correlation between antennas can reach $.5$ even in a, $4 \times 4 $ MIMO systems. 
+Since the IRS introduces the potential for additional correlation it seems reasonable to allow for $\rho = .5$.
+
+\section{Next Steps: Methods for Phase Optimization}
+\begin{enumerate}
+\item
+	Method from \cite{huang2018achievable}.
 \begin{itemize}
 \item
-zf mu: 202.0989509146119
-zf std: 57.23864582855936
+	No LOS of correlation
 \item
-matched mu: 476.69650954785794
-matched std: 239.3529067968275
+	Requires full CSIT at IRS
+\item 
+	Assumes a ZF precoder in order to simplify sum-capacity objective function.
+\item
+	Alternating algorithm approach with two main steps
+	\begin{enumerate}
+	\item 
+		Minimize transmission power w.r.t phases
+	\item
+		Maximize rate wrt power
+	\end{enumerate}
 \end{itemize}
 
+\item 
+	Method from \cite{huang2018achievable}.
+
+\end{enumerate}
 \bibliography{bibliography}
-
 \end{document}

### 2020-07-28 documentation/Optimization Work/07_24_20.tex
adding notes on progress report for reviewing methods of phase optimization

@@
 \item
-	No LOS of correlation
+	No LOS or correlation considered in system model.
 \item
-	Requires full CSIT at IRS
+	Algorithm requires full CSIT at IRS
 \item 
@@
 \item 
-	Method from \cite{huang2018achievable}.
+	Method from \cite{kammoun2020asymptotic}.
+	\begin{itemize}
+		\item 
+			Algorithm only requires channel statistics for the RIS update
+		\item
+			System model includes correlation.
+		\item
+			Does transmitter need perfect CSI or can they make do with just knowing the composite channel.
+		\item
+			Uses max-min SINR as objective
+		\item
+			Optimizes the beamformer, the powers and RIS (continuous phases allowed)
+		\item
+			Question: \cite[Section III ]{kammoun2020asymptotic} makes a claim that for a single rank channel the RIS only offers a "substantial" benefit if there
+			is a single user.  This makes sense for a Decision-Feedback equalization sense but it is still not clear to me why the benefit
+	\end{itemize}
 

### 2020-07-30 documentation/Optimization Work/07_24_20.tex
Changes for progress report on simulations/optimization

--- 2020-07-27 documentation/Optimization Work/07_24_20.tex
@@
 with 
+	\begin{equation}
+	\mH_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+	\end{equation}
+Elements of $\mH_{2},\mH_{1}$ and $\mG$ are normalized so as to have a normalized trace of 1 and $\mathbf{R}_{k}$ represents the exponential correlation model with elements $r_{i,j}=\rho^{i-j}$.
+\par
+The following numerical results evaluate the MSE for the channel above with $s_k \in \mathcal{NC}(0,\frac{1}{N_T})$, $n_k \in \mathcal{NC}(0,\frac{1}{N_R})$
+$N_T = 10$ (transmitters), $N_R = 10$ (receivers) and $N_S = 50$ (IRS surfaces).
+For each channel realization, I evaluate the MSE for the matched filter and zero-forcing filter using $\phase = e^{j\theta}\mI$ for 10 uniform spaced values of $\theta \in [0, 2\pi]$.
+I then find the mean and variance of the MSE over the 10 values of $\theta$.
+I repeat and average this across 1000 channel realizations.
+This numerical test is performed for different values of $\rho$ in the exponential correlation model.
+The results in Figure \ref{MSE_correlation} indicate that the variance of the MSE over different values of the phase matrix parameter, $\theta$,  is significant and increases as the correlation parameter, $\rho$ approaches 1.
+		\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/over_10_thetas}
+	  \caption{Comparison MSE over different amounts of correlation.}
+	  	  \label{MSE_correlation}
+	\end{figure}
 
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
+Based on the measurement data collected in \cite{martin2000multiple}, the correlation between antennas can reach $.5$ even in a $4 \times 4 $ MIMO system without an IRS. 
+The IRS should introduce the potential for additional correlation in the system so seems reasonable to allow for $\rho = .5$ (or maybe even higher).
+I have not been able to find any correlation measurement data from larger systems or with an IRS.
 
-Using a matched filter  and precoder which satisfies the transmission power constraint gives (normalize each column)
-\begin{equation}
-\mW = \mathbf{H}_{Total}^H
-\end{equation}
-
-Numerical results for $s_k \in \mathcal{NC}(0,1)$, $n_k \in \mathcal{NC}(0,1)$
+\section{Next Steps: Methods for Phase Optimization}
+The above results motivate the investigation of phase optimization to reduce the MSE. A couple of previous works looking at such optimization are now reviewed. 
+\begin{enumerate}
+\item
+	Method from \cite{huang2018achievable}.
 \begin{itemize}
 \item
-zf mu: 202.0989509146119
-zf std: 57.23864582855936
+	No LOS or correlation component in the channel model.
 \item
-matched mu: 476.69650954785794
-matched std: 239.3529067968275
+	Algorithm requires full CSIT at IRS.
+\item 
+	Assumes a ZF precoder in order to simplify the sum-capacity objective function.
+\item
+	Uses an alternating algorithm approach with two main steps
+	\begin{enumerate}
+	\item 
+		Minimize transmission power w.r.t $\theta$.
+	\item
+		Maximize rate with respect transmission power of each transmit antenna.
+	\end{enumerate}
 \end{itemize}
 
+\item 
+	Method from \cite{kammoun2020asymptotic}.
+	\begin{itemize}
+		\item 
+			Algorithm only requires channel statistics for the IRS update.
+		\item
+			Unlike \cite{huang2018achievable}, optimization occurs over a beamforming matrix and then optimizes powers and phases. 
+		\item
+			Channel uses a correlation model sort of like the exponential model except for that I don't think the matrix is normalized for power.
+		\item
+			Uses max-min $\text{SINR}_K$ as objective.
+		\item
+			For case with single rank channel (fully correlated channel) they use constant phase at all elements. Different phases are used for full-rank case.
+		\item 
+			Figure 9. of this work is particular interesting as it provides evidence that their algorithm can significantly increase the minimum user rate when compared
+			to another method of selecting the phases at the IRS. Note that here they allow the number of reflecting surfaces to be quite high but keep $N_R = N_T = 8$.
+	\end{itemize}
+\end{enumerate}
+Assuming a matched filter at the transmitter the MSE problem for the choice of the IRS phases becomes
+		\begin{subequations}
+	\label{optim}
+	\begin{align}
+	    \underset{\phase}{\text{minimize }}
+	    & \; \| \mH_{Total}(\mH_{Total}^H\vs + \vn) - \vs \| \label{potential_game} \\
+	    \text{subject to  } \; &
+	    | \theta_i | = 1 & \forall i \in {1 \cdots	 N_S}
+		\label{pos_power_const}
+	\end{align}
+	\end{subequations}
+	Is the utility function at least convex?
+	Do I need to include the power constraint here? Probably
 \bibliography{bibliography}
-
 \end{document}

### 2020-07-30 simulation/testing/Optimization/MSE_phase_optimization.py

@@ -52,16 +52,16 @@ for cor_rho in rhos:
     cor_zf_mu.append(np.average(MSE_zf_mu))
     cor_zf_var.append(np.average(MSE_matched_mu))
 
-ave_plt.plot(rhos, cor_mf_mu, '-o', label='MF')
-ave_plt.plot(rhos, cor_zf_mu, '-o', label='ZF')
-var_plt.plot(rhos, cor_mf_var, '-o', label='MF')
-var_plt.plot(rhos, cor_zf_var, '-o', label='ZF')
+ave_plt.plot(rhos, cor_mf_mu, '-o', label='Matched Filter')
+ave_plt.plot(rhos, cor_zf_mu, '-o', label='Zero-Forcing')
+var_plt.plot(rhos, cor_mf_var, '-o', label='Matched Filter')
+var_plt.plot(rhos, cor_zf_var, '-o', label='Zero-Forcing')
 ave_plt.set_ylabel("Average MSE  (dB)")
 var_plt.set_ylabel("Average MSE Variance (dB)")
 ave_plt.set_yscale('log')
-ave_plt.set_xlabel("rho")
+ave_plt.set_xlabel("Correlation Parameter (rho)")
 var_plt.set_yscale('log')
-var_plt.set_xlabel("rho")
+var_plt.set_xlabel("Correlation Parameter (rho)")
 ave_plt.legend(loc="upper left")
 var_plt.legend(loc="upper left")
 ave_plt.set_title(f'MSE comparison for N = {average}')

### 2020-08-07 documentation/Optimization Work/08_03_20.tex
Adding SINR/achievable with corrected model variance

--- 2020-07-28 documentation/Optimization Work/07_24_20.tex
@@
 \section{Numerical Results}
-I use the system model given by 
-\begin{equation}
-\vy = \mH_{Total} \mW \vs  + \vn
-\end{equation}
-with
-
-	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
-	\end{equation}
-
-The following numerical results evaluate the MSE the channel above with $s_k \in \mathcal{NC}(0,\frac{1}{N_R})$, $n_k \in \mathcal{NC}(0,\frac{1}{N_R})$.
-For each channel realization, I evaluate the MSE for the matched filter and zero-forcing filter using $\phase = e^{j\theta}\mI$ for 10 uniform spaced values of $\theta \in [0, 2\pi]$.
-Then I collect the mean and variance of the 10 MSEs for the channel realization.
-I repeat and average across 1000 channel realizations.
-I perform this same analysis across different values of $\rho$ in the exponential correlation model (i.e.
-$r_{i,j}=\rho^(i-j)$.
-The results in Figure \ref{MSE_correlation} indicate that the variance of the MSE over different values of the phase matrix parameter, $\theta$,  can be quite large.
-In particular when the correlation parameter, $\rho$ is close to 1.
+Below are the simulation results using the simulation system parameters as shown in the previous results but now plotted over $\theta \in [0, 2\pi]$ for $\phase = e^{j\theta}\mI$.
+This result suggests that the optimal $\phase$ with respect to the MSE depends  primarily on the fast-fading characteristics of the channel.
 		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/over_10_thetas}
+	\includegraphics[width= 16cm,height = 13cm]{figures/plotted_over_theta}
 	  \caption{Comparison MSE over different amounts of correlation.}
@@
 
-Based on the measurement data collected in \cite{martin2000multiple}, the correlation between antennas can reach $.5$ even in a, $4 \times 4 $ MIMO systems. 
-Since the IRS introduces the potential for additional correlation it seems reasonable to allow for $\rho = .5$.
-
-\section{Next Steps: Methods for Phase Optimization}
-\begin{enumerate}
-\item
-	Method from \cite{huang2018achievable}.
-\begin{itemize}
-\item
-	No LOS of correlation
-\item
-	Requires full CSIT at IRS
-\item 
-	Assumes a ZF precoder in order to simplify sum-capacity objective function.
-\item
-	Alternating algorithm approach with two main steps
-	\begin{enumerate}
-	\item 
-		Minimize transmission power w.r.t phases
-	\item
-		Maximize rate wrt power
-	\end{enumerate}
-\end{itemize}
-
-\item 
-	Method from \cite{huang2018achievable}.
-
-\end{enumerate}
-\bibliography{bibliography}
 \end{document}

### 2020-08-07 documentation/Optimization Work/08_07_20.tex
Adding SINR/achievable with corrected model variance

--- 2020-07-28 documentation/Optimization Work/07_24_20.tex
@@
 with 
-
 	\begin{equation}
-	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
+	\mH_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
 	\end{equation}
-
-The following numerical results evaluate the MSE the channel above with $s_k \in \mathcal{NC}(0,\frac{1}{N_R})$, $n_k \in \mathcal{NC}(0,\frac{1}{N_R})$.
-For each channel realization, I evaluate the MSE for the matched filter and zero-forcing filter using $\phase = e^{j\theta}\mI$ for 10 uniform spaced values of $\theta \in [0, 2\pi]$.
-Then I collect the mean and variance of the 10 MSEs for the channel realization.
-I repeat and average across 1000 channel realizations.
-I perform this same analysis across different values of $\rho$ in the exponential correlation model (i.e.
-$r_{i,j}=\rho^(i-j)$.
-The results in Figure \ref{MSE_correlation} indicate that the variance of the MSE over different values of the phase matrix parameter, $\theta$,  can be quite large.
-In particular when the correlation parameter, $\rho$ is close to 1.
+Below are the simulation results using the simulation system parameters as shown in the previous results but now plotted over $\theta \in [0, 2\pi]$ for $\phase = e^{j\theta}\mI$.
+This result suggests that the optimal $\phase$ with respect to the MSE depends  primarily on the fast-fading characteristics of the channel.
 		\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/over_10_thetas}
+	\includegraphics[width= 16cm,height = 13cm]{figures/plotted_over_theta}
 	  \caption{Comparison MSE over different amounts of correlation.}

### 2020-08-07 documentation/Paper/abstract.tex
Adding SINR/achievable with corrected model variance

--- 2020-06-24 documentation/Paper/abstract.tex
@@
-This work investigates the impact of passive reflective surfaces on the capacity of MIMO systems.
-In particular, the system is analyzed in the asymptotic size domain in order to employ the tools of free probability.
-A system is considered first in the case of fully uncorrelated channels.
-In this case it is shown that the passive control of such a reflective surface has no impact on the capacity of the system.
-The system is then generalized to allow for correlation within the channel.
-In this case is it shown that the selection of the phases at the IRS still has an impact on the system.
-Finally we compare the capacity of the correlated channel with optimized phases at the IRS to the uncorrelated channel to show that....?
-
-This work provides an upper bound on the capacity of the RIS assisted MIMO system.
-This upper bound is a useful tool with which to analyze the performance of the various sub-optimal optimization approaches used.
-Question: would the capacity of the finite , phase-optimized system with correlation ever exceed that of the asymptotic case?
+Reflective surfaces with controlled reflected phase (IRS) have been the subject of interest in recent years in order to control the wireless propagation environment.
+This work investigates the impact of an IRS on the capacity of MIMO communication systems.
+Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity.
+A key component of this analysis is considering the correlation common to massive MIMO systems, an aspect of the MIMO system not yet considered in previous work on IRS assisted systems.
+A key finding is that the choice for the phase of reflected waves has no impact for the asymptotic analysis.
+Motivated the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior.
+We also investigate the use of phase optimization which require minimal channel state information.

### 2020-08-18 documentation/Optimization Work/08_17_20.tex
Made changes to see that pi is optimal phase for LOS/No CSI case.

--- 2020-08-07 documentation/Optimization Work/08_07_20.tex
@@
 Below are the simulation results using the simulation system parameters as shown in the previous results but now plotted over $\theta \in [0, 2\pi]$ for $\phase = e^{j\theta}\mI$.
-This result suggests that the optimal $\phase$ with respect to the MSE depends  primarily on the fast-fading characteristics of the channel.
+At the transmitter, I am precoding with a matched filter that does \emph{not} include the phases of the IRS.
+This result suggests that the optimal $\phase$ with respect to the achievable rate and SINR depends  primarily on the fast-fading characteristics of the channel but that there is an optimal choice when using MSE as the utility.
 		\begin{figure}[H]
-	\includegraphics[width= 16cm,height = 13cm]{figures/plotted_over_theta}
-	  \caption{Comparison MSE over different amounts of correlation.}
+	\includegraphics[width= 16cm,height = 13cm]{figures/MSE}
+	  \caption{Comparison MSE over different choice of IRS phase.}
 	  	  \label{MSE_correlation}
@@
 
+			\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/SINR}
+	  \caption{Comparison of SINR over different choice of IRS phase.}
+	  	  \label{SINR}
+	\end{figure}
+
+			\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/Achievable_Rate}
+	  \caption{Comparison achievable rate over different choice of IRS phase.}
+	  	  \label{Achievable_rate}
+	\end{figure}
+
 \end{document}

### 2020-08-18 documentation/Paper/information_theory_background.tex
Made changes to see that pi is optimal phase for LOS/No CSI case.

@@
 	\end{equation}
-	in which $\mH_{Total}$ is a random channel (we can always represent linearly?).
+	in which $\mH_{Total}$ is a random channel.
+Using methods such as OFDM to transform any multi-path channel into a flat-fading channel, this representation is sufficient to represent any linear channel.
 A key metric of interest for this system is the mutual information between the transmitted and received signal given by 

### 2020-08-18 documentation/Paper/irs_channel.tex
Made changes to see that pi is optimal phase for LOS/No CSI case.

--- 2020-07-22 documentation/Paper/irs_channel.tex
@@
-Using the approach for finding asymptotic eigenvalue distributions detailed in \ref{Free_Prob_Intro} we now analyze a channel model that is representative of a MIMO communication system with an RIS. 
+Using the approach from Section \ref{Free_Prob_Intro} to find the asymptotic eigenvalue distribution, we now analyze a channel model that is representative of a MIMO communication system with an RIS. 
 \par
-In general, the received signal from a line of sight path between a transmitter and receiver would contain  sufficient energy to disregard the contribution of any reflected signal.
+In general, the received signal from any line of sight path between a transmitter and receiver would contain  sufficient energy to disregard the contribution of any reflected signal (although this component may be very low rank).
 Therefore, we focus on a channel model in which all signal power to the receiver is through the path of the RIS  given by
@@
 Similarly, $ \mathbf{R}_{T},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively.
-This correlation will always be modeled by the so-called exponential correlation model. As seen in the later, this model for correlation will enable a very clean analysis.
+When compared with the example seen in Section \ref{Free_Prob_Intro}, the structure of $\mathbf{H}_{Total}$ suggests that we can to use multiplicative free convolution in order to evalute the AED.
+With this in mind, we would like to choose a correlation model  that not only represents the system but one with a known S-Transform expression.
+One such model is exponential correlation. Defined by a single parameter $\rho$, the elements of the corresponding correlation matrix are given by $\rho^{i-j}$.
 \par
-
-To begin using the tools of free probability, we first ensure that the components of the covariance matrix given by 
+To begin using the tools of free probability, we must first separate out the free sub-components of the covariance matrix given by 
 	\begin{equation}
@@
 	\end{equation}
-are free. 
-Noting from Equation \eqref{no_csi_capacity_aed} that the capacity is only affected by the non-zero eigenvalues of the covariance matrix allows us to equivalently 
+Given the S-transform of these sub-components, the total S-transform is simply the resulting product.
+Noting from Equation \eqref{no_csi_capacity_aed} that the capacity is only affected by the non-zero eigenvalues of the covariance matrix, we equivalently
 consider the AED of the covariance matrix given by 
@@
 	 &  =  
-	 \mH_{1}^H \mR_{S}^{\frac{H}{2}}\phase^H \mH_{2}^H \mR_{R}^{\frac{H}{2}}
-	 \mR_{R}^{\frac{1}{2}}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}.
+	 \mH_{1}^H \mR_{S}^{\frac{H}{2}}\phase^H \mH_{2}^H \mR_{R}\mH_{2}\phase \mR_{S}^{\frac{1}{2}}\mH_{1}\mR_{T}.
 	\end{align}
-This structure appears to admit analysis through the use of the S-Transform and multiplicative free convolution.
-In order to continue decomposing this covariance matrix into subcomponents  with known S-Transforms, we exploit a property of the S-Transform given by (proven in Appendix \ref{rotation_property})
+In order to continue decomposing this covariance matrix into sub-components, we exploit a property of the S-Transform given by (proven in Appendix \ref{rotation_property})
 \begin{equation}
@@
 If all system dimensions are $N$, these S-Trasforms (and AEDs) identical so for this section we will assume that all system dimensions are $N$.
-Iterating the rotational property gives the equivalent matrix
+Repeated iteration of the rotational property above gives the equivalent matrix
 	\begin{equation}
@@
 	\end{equation}
-Without any further analysis this expression shows the interesting result that in the asymptotic domain, the phases at the RIS (or indeed any unitary matrix in the channel) cancels out.
-For the remaining terms, we simply need an expression for the S-Transform in order to evaluate the asymptotic system capacity.
+Note that using the Eigen Value Decomposition of $\mR_{R}$ and $\mR_{T}$, we find the product $\mR_{R} \mR_{T}$ is also a Toeplitz matix (which can be represented by a circulant matrix in the asymptotic system size).
+In \cite[Section 4.3.2]{muller2013applications} it is shown that the hermetian matrix $\mH\mH_^H$ with $h_{i,j} \in \NC\{0, \frac{1}{N}\}$
+is free 
+Having shown the relative freeness of the matrix sub-components, we must now find the S-transform for each in order to find the product.
+\begin{enumerate}
+\item 
+	For both $\mH_{2}^H \mH_{2}$ and $\mH_{1}\mH_{1}^H$ with all system dimensions equal, the S-Transform as shown in \cite{muller2002random} is given by 
+	\begin{equation}
+	S(z) = \frac{1}{1+z}
+	\end{equation}
+\item 
+	As shown in \cite{gray2006toeplitz}, with $N \rightarrow \infty$ Toeplitz matrices have the spectrum (and thus AED) of an equivalent circulant matrix (whose spectrum and thus AED is given by the discrete Fourier transform of the first row).
+Using the definition of the Stieltjes transform and the definition of the S-Transform in terms of the Stieljes transform, the S-Transform of the exponential correlation model is shown to be \cite{skupch2005free}
+	\begin{equation}
+	S(z) = \frac{\alpha z^2 + z \sqrt{alpha^2 z^2 -(z^2-1)}}{z-1}
+	\end{equation}
+\end{enumerate}
 \par
-While we do not consider a line of sight component to this channel, it is worth noting that using the tools from a similar analysis in \cite{}, it can be shown that 
-a similar cancellation of phases will occur. 
-(also mention the non-zero mean aspect here?)
+Without further analysis, this expression shows the interesting result that under the assumptions of Free Probability and equal power transmission, the phases at the RIS (or indeed any unitary matrix in the channel) cancel out of the expression for the AED.
+For the remaining terms, we simply need an expression for the S-Transform in order to evaluate the asymptotic system capacity.s
+\par 
+Using the previous analysis we can observe an extension of this result that may be relevant to variations of this system model. 
+While we do not consider a line of sight component to this channel, using the tools from a similar analysis in \cite{muller2012channel}, it can be shown that 
+a similar cancellation of phases will occur for the RIS channel with an LOS (or non-zero mean) component of the signal.
+Note that if we are simply adding a non-zero mean component.
+We can find the corresponding AED by looking considering the channel without the non-zero mean and then adding the single eigenvalue from the non-zero mean component; in which case the phases will again cancel out.

### 2020-08-18 documentation/Presentations/Final Presentation/final_presentation.tex
Made changes to see that pi is optimal phase for LOS/No CSI case.

--- 2020-07-10 documentation/Presentations/Mid-Presentation/mid_term_presentation.tex
@@
 \title{\large \bfseries Asymptotic Capacity of Systems with Intelligent Reflective Surfaces}
-\author{Peter Hartig \\ \and Supervisor: Saba Asaad}
+\author{Peter Hartig \\ \and Supervisor: Saba Asaad and Ali Bereyhi}
 
@@
 \subsection{MIMO Channel}
+\begin{frame}
+\frametitle{The Impact of Increasing Frequency}
+5G already has a 40 - 100 GHz band. Current/future work will extend beyond this. 
+\begin{itemize}
+\item
+	The $\frac{\lambda}{2}$ dipole antenna is smaller.
+\item
+	 RF components at this frequency are expensive. 
+\item 
+	\cdots but lots of available spectrum!
+\item
+	Higher free-space path loss. (Need to avoid molecular absorption). Comparison of free-space path loss at 1 km.  320 dB at 2.1 GHz (GSM) vs. 443 dB at 100 GHz.
+\item 
+	We focus on how to best counter this high path loss. 
+\end{itemize}
+\end{frame}
 

### 2020-08-18 simulation/testing/Correlation_tests/correlation_convergence.py

@@ -44,7 +44,6 @@ for cor_rho in correlations:
     var_plt.plot(sizes, capacity_variance, '-o', label=f'Correlation (rho): {cor_rho}')
     ave_plt.plot(sizes, capacity_average, '-o', label=f'Correlation (rho): {cor_rho}')
 
-
 # ave_plt.title("Capacity Average")
 # plt.title("Capacity Variance")
 var_plt.set_ylabel("Capacity Variance")

### 2020-08-18 simulation/testing/Optimization/MSE_vs_theta.py

@@ -5,7 +5,7 @@ num_tx = 50
 num_rx = 50
 num_reflectors = 50
 average = 1000
-average = 10
+# average = 10
 cor_mf_mu = []
 cor_mf_var = []
 cor_zf_mu = []
@@ -16,15 +16,17 @@ phases = np.linspace(0, 2 * np.pi, 10)
 fig = plt.figure()
 ave_plt = fig.add_subplot(2, 1, 1)
 var_plt = fig.add_subplot(2, 1, 2)
-transmit_cor = .7
-irs_cor = .9
-receiver_cor = .1
+transmit_cor = .5
+irs_cor = .5
+receiver_cor = .5
 MSE_zf_mu = []
 MSE_matched_mu = []
+MSE_matched_mu_rand = []
 SINR_matched_mu = []
 AR_matched_mu = []
 MSE_zf_var = []
 MSE_matched_var = []
+MSE_matched_var_rand = []
 SINR_matched_var = []
 AR_matched_var = []
 for theta in phases:
@@ -32,7 +34,9 @@ for theta in phases:
     s_correlation_matrix = exponential_correlation(num_reflectors, irs_cor)
     r_correlation_matrix = exponential_correlation(num_rx, receiver_cor)
     phase = np.diag(np.exp(1j * theta * np.ones(num_reflectors)))
+    phase_rand = random_phase(num_reflectors)
     MSE_matched = []
+    MSE_matched_rand = []
     MSE_zf = []
     SINR_matched = []
     AR_Matched = []
@@ -43,12 +47,17 @@ for theta in phases:
         # noise = c_rand(num_rx, 1).flatten()
         G = c_rand(num_reflectors, num_tx, var=1)
         G2 = c_rand(num_rx, num_reflectors, var=1)
-        los_path = c_rand(num_rx, num_tx, var=1)
+        los_path = c_rand(num_rx, num_tx, var=1)*0
         irs_path = G2 @ s_correlation_matrix @ phase @ G
+        irs_path_rand = G2 @ s_correlation_matrix @ phase_rand @ G
         channel = r_correlation_matrix @(irs_path + los_path)@ t_correlation_matrix
-        transmit_signal_matched = precode_mf(channel) @ transmit_symbols
+        channel_rand = r_correlation_matrix @(irs_path_rand + los_path)@ t_correlation_matrix
[…]

### 2020-08-19 documentation/Optimization Work/08_19_20.tex
Added files for optimizing phases using CSI estimation

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Numerical Results}
The system model is given by 
\begin{equation}
\vy = \mH_{Total} \mW \vs  + \vn
\end{equation}
with 
	\begin{equation*}
	\mH_{Total} = \underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}},
	\end{equation*}
%		\mH_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
	$\mW$ as a linear precoder and $\vs$ as the vector of transmit symbols.
	\par
	Using the method for CSI estimation from \cite{nadeem2019intelligent} to estimate the CSI to be used in the choice of $\mW$, the following procedure is used. 
	\begin{enumerate}
	\item
		With all IRS surfaces turned off LOS channel is estimated as $\hat{\mG}$ by 
		transmitting some pilot vector $\mathbf{p}$. This is done using the MMSE estimate for the columns of $\hat{\mG}$ using $\vy$.
		In this case the received signal is 
			\begin{equation*}
			\vy = \mG\mathbf{p}  + \vn.
			\end{equation*}

	\item 	
		In the next time interval, we turn on a single IRS element and transmit the pilot sequence again such that the received signal is 
			\begin{equation*}
			\vy = (\mG + \mathbf{h}_{2,1}\mathbf{h}^H_{1,1}  )\mathbf{p}  + \vn
			\end{equation*}
			where $\mathbf{h}_{2,1}$ is the 1st column of $ \mathbf{H}_{2}$ and $\mathbf{h}_{1,1}^H$  is the 1st row of $\mathbf{H}_{1}$.
			By subtracting off $\hat{\mG}\mathbf{p}$ from the received signal we have 
						\begin{equation*}
			\vy = (\mathbf{h}_{2,1}\mathbf{h}^H_{1,1})\mathbf{p}  + \vn
			\end{equation*}
			from which we can estimate the rank 1 matrix $\mathbf{h}_{2,1}\mathbf{h}^H_{1,1}$.
	\item 
		Iterating this procedure allows us to estimate both paths of the channel. 
		** Question: This is different from what was discussed in our meeting because in the method Ali mentioned. In that case, the IRS surfaces were turned on during the
		first round of estimation. We want to use the matched filter so it seems like we should have all IRS turned on in the first round like Ali mentioned. 
	\end{enumerate}
The precoded transmission $\vx = \mathbf{F}\mathbf{s}$ is a function of this estimate if we assume a matched filter (or other types of precoding).

\bibliography{bibliography}

\end{document}

### 2020-08-19 simulation/src/lin_alg.py

@@ -35,6 +35,12 @@ def c_rand(rows, cols, var=None, mean=0, normalize=True):
         else:
             return (np.random.randn(rows, cols) + 1j*np.random.randn(rows, cols))/np.sqrt(2*rows) + mean
 
+def BPSK(num):
+    ret = np.random.randint(0, 2, num)
+    ret += -1*(ret == 0)
+    return ret
+
+
 
 def block_matrix(block, size, rows=True):
     """

### 2020-08-19 simulation/src/optimization.py

@@ -1,0 +1,9 @@
+"""
+Optimization tools for selecting the phases as IRS elements
+"""
+import numpy as np
+
+def optimize_phases(num_elements, info):
+    opt_phases = np.ones(num_elements)
+
+    return opt_phases

### 2020-08-19 simulation/testing/Optimization/MSE_phase_optimization.py

@@ -1,9 +1,9 @@
 from src.lin_alg import *
 import matplotlib.pyplot as plt
 
-num_tx = 10
+num_tx = 16
 num_rx = 10
-num_reflectors = 50
+num_reflectors = 128
 average = 1000
 cor_mf_mu = []
 cor_mf_var = []

### 2020-08-19 simulation/testing/Optimization/MSE_vs_theta.py

@@ -41,13 +41,13 @@ for theta in phases:
     SINR_matched = []
     AR_Matched = []
     for j in range(average):
-        # transmit_symbols = np.random.randn(num_rx)
-        transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
+        # transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
+        transmit_symbols = BPSK(num_rx)
         noise = c_rand(num_rx, 1, var=1).flatten()
         # noise = c_rand(num_rx, 1).flatten()
         G = c_rand(num_reflectors, num_tx, var=1)
         G2 = c_rand(num_rx, num_reflectors, var=1)
-        los_path = c_rand(num_rx, num_tx, var=1)*0
+        los_path = c_rand(num_rx, num_tx, var=1)
         irs_path = G2 @ s_correlation_matrix @ phase @ G
         irs_path_rand = G2 @ s_correlation_matrix @ phase_rand @ G
         channel = r_correlation_matrix @(irs_path + los_path)@ t_correlation_matrix

### 2020-08-26 documentation/Optimization Work/08_24_20.tex
added heuristic algorithm for phase choices in sum. TODO is add reiterations with different ordering or the phase improvements

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{CSI Estimation Setup}
The system model is given by 
\begin{equation}
\vy = \mH_{Total} \mW \vs  + \vn
\end{equation}
with 
	\begin{equation*}
	\mH_{Total} = \underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}},
	\end{equation*}
%		\mH_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
	$\mW$ as a linear precoder and $\vs$ as the vector of transmit symbols.
	\par
	CSI is estimated for use in the choice of $\mW$ in following procedure. 
	In  , 
	\begin{enumerate}
	\item
		With all IRS surfaces turned off, the LOS channel is estimated as $\hat{\mG}$ by 
		transmitting a matrix of pilots $\mathbf{P}$ (with orthogonal rows)  and using the least squares solution with the matrix $\mathbf{Y}$ \cite{nadeem2019intelligent}.
		The resulting solution is given by
		\begin{equation}
			\hat{\mG} = [(\mathbf{P} \mathbf{P}^H)^{-1} \mathbf{P} \mathbf{Y}^H]^H
		\end{equation}

	\item 	
		In the next time interval, we turn on a single IRS element and transmit the pilot sequence again such that the received signal is 
			\begin{equation*}
			\mathbf{Y} = (\mG + \mathbf{h}_{2,1}\mathbf{h}^T_{1,1}  )\mathbf{P}  + \vn
			\end{equation*}
			where $\mathbf{h}_{2,1}$ is the 1st \emph{column} of $ \mathbf{H}_{2}$ and $\mathbf{h}_{1,1}^T$  is the 1st \emph{row} of $\mathbf{H}_{1}$.
			By subtracting off the known contribution from the LOS path, $\hat{\mG}\mathbf{P}$, from the received signal we have 
						\begin{equation*}
						\mathbf{Y} = (\mathbf{h}_{2,1}\mathbf{h}^T_{1,1})\mathbf{P}  + \vn
						\end{equation*}
			from which we can estimate the rank 1 matrix $\mathbf{h}_{2,1}\mathbf{h}^T_{1,1}$.
			Note that the matrix $\mathbf{P} $ need only have 2 columns in order to have sufficient data to estimate the rank 1 matrix.
			\par
			At this point, the estimate for the channel contains
				\begin{equation*}
				\hat{\mH_{Total}} = \hat{\mG} + \mathbf{h}_{2,1}\mathbf{h}^T_{1,1}.
				\end{equation*}
				Note that we can change the resulting sum of the matrices using the phase of components of the IRS matrix using  $\phi_i \hat{\mathbf{h}}_{2,1}\hat{\mathbf{h}^T_{1,1}} $.
	\item 
		Iterating this procedure allows us to estimate both paths of the channel. 
		such that for a system with $N_S$ IRS elements, the channel estimate after $N_S$ iterations of the above procedure yields the channel 
				\begin{equation*}
				\hat{\mH}_{Total} = \hat{\mG} + \Sigma_{i=1}^{N_S} \phi_i \hat{\mathbf{h}}_{2,i}\hat{\mathbf{h}^T_{1,i}} 
				\end{equation*}
				where $\phi_i$ is the phases selected by the IRS controller for element $i$. 
		From this representation, it is clear that if in the worst case, the matrices $\hat{\mathbf{h}}_{2,i}\hat{\mathbf{h}^T_{1,i}}$ were completely correlated
		then we would want these components to be added together coherently, implying that all phases $\phi_i$ should be the same.
	\end{enumerate}

\section{Optimization}
Using a matched filter at the transmitter with CSI found using the above procedure, the MSE problem for selecting the phases for each element is now given by
	\begin{subequations}
	\label{optim}
	\begin{align}
	    \underset{\phase}{\text{minimize }}
	    & \; E\left[\|  \vs - (\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\hat{\mG} + \Sigma_{i=1}^{N_S} \phi_i \hat{\mathbf{h}}_{2,i}\hat{\mathbf{h}^T_{1,i}})^H\vs) \|^2 \right]\\
	    \text{subject to  } \; &
	    | \phi_i | = 1 & \forall i \in {1 \cdots	 N_S}
	\end{align}
	\end{subequations}	
	Assuming the CSI is known at the transmitter such that the expectation applies only to the transmit symbols, this problem becomes
	\begin{subequations}
	\label{optim}
	\begin{align}
	    \underset{\phase}{\text{minimize }}
	    & \; trace\left((\mI - \mH\hat{\mH}_{Total}^H) \mathbf{Q}_{\vs\vs} 
	    (\mI - \mH\hat{\mH}_{Total}) ^H\right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1 & \forall i \in {1 \cdots	 N_S}
	\end{align}
	\end{subequations}

	To Discuss
	\begin{enumerate}
	\item
		Not convex in this formulation. The above utility can be seen as a sum such as 
		\begin{equation}
		 x_1x_2 c_1 + x_2x_3c_2 + x_1x_3c_3 + \cdots + x_1x_2x_3c_i \cdots
		\end{equation}
		in which the coefficients $c_i \in \mathcal{C}$ so that the result is non-convex (Should be a sum of concave and convex terms though).
		Clearly the constraint $ | \phi_i | = 1$ is also non-convex
	\item
		Need to clarify the order of estimation because using the above method, we wouldn't be able to have a precoder based on the total channel until all $N_S$ steps were 					complete.	Should the order of estimation be reversed (i.e. start with all elements turned on and then turn them off 1 by 1?).
	\item
		How to connect intuition of the resulting metric from MSE with a matched with a performance metric like SER?
		Particularly when the channels are correlated, this interference will dominate performance.
%	\item
%		Can we factor out $\vs$ and just consider how close we can get the combined filters to be the identity?
	\end{enumerate}
\bibliography{bibliography}
\end{document}

### 2020-08-26 documentation/Paper/abstract.tex
added heuristic algorithm for phase choices in sum. TODO is add reiterations with different ordering or the phase improvements

@@
-Reflective surfaces with controlled reflected phase (IRS) have been the subject of interest in recent years in order to control the wireless propagation environment.
+Surfaces with reflective phase control (IRS) have been the subject of interest in recent years as means to better control the wireless propagation environment.
 This work investigates the impact of an IRS on the capacity of MIMO communication systems.
 Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity.
-A key component of this analysis is considering the correlation common to massive MIMO systems, an aspect of the MIMO system not yet considered in previous work on IRS assisted systems.
-A key finding is that the choice for the phase of reflected waves has no impact for the asymptotic analysis.
+A key finding is that in the asymptotic domain, the choice for the phase of reflected waves has no impact for the asymptotic analysis.
 Motivated the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior.
-We also investigate the use of phase optimization which require minimal channel state information.
+We then investigate new methods for phase optimization for systems without full channel state information at the IRS.

### 2020-08-26 simulation/src/CSI_estimator.py

@@ -5,4 +5,5 @@ from src.lin_alg import *
 
 def ML_MIMO(received_pilots, pilots):
     #   Assume that the pilot matrix is orthogonal with respect to the vectors at each Tx antenna.
+    #   Basically we don't want any of our equations to be linearly dependent.
     return hermetian(np.linalg.pinv(pilots@hermetian(pilots))@pilots@hermetian(received_pilots))

### 2020-08-27 documentation/Optimization Work/08_27_20.tex
added new progress report with phase cancellation and also saw result that discrete optimization was performing at high N but not continuous

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents
\section{System Overview}
The communication system model is given by 
\begin{equation}
\vy = \mH_{Total} \mW \vs  + \vn
\end{equation}
with 
	\begin{equation*}
	\mH_{Total} = \underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}}.
	\end{equation*}
In the following, we will assume that a matched filter is used at transmitter. Further we will evaluate the performance of the system using the Expected Mean Square Error given by 
\begin{equation}
E\left[\|  \vs - (\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H\vs) \|^2 \right]
\end{equation}
in which we do not consider the noise at the receiver. 
Note that we are using the full information of the current channel in order to estimate the optimal phases at the IRS. 
In order to minimize the overhead of pilot acquisition, we estimate the CSE in the following manner. In the following we assume that the elements of the IRS can not only adjust the phases of the impinging waves but may also be "turned off" such that they are not included in the channel.
\begin{enumerate}
\item
	In order to first transmit matched filter we leave all IRS elements \emph{on}. Assuming that we perfectly estimate the entire channel $\mH_{Total}$, we will not have sufficient information to estimate the individual components of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
\item
	In the next iteration of pilots, we turn off all elements in order to estimate the LOS matrix $\mG $.
\item
	By turning on a specific IRS element (can we use patterns using the correlation?) and sending additional pilots, we can estimate 
	the path through a specific IRS element given by $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$.
	Using this information, we can now use the expression 
	\begin{equation}
E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}) \|^2 \right]
\end{equation}
	\item
		Repeating the procedure above, we estimate additional elements $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and perform the optimization over increasing number of phases.
	\item 
		We are investigating how this optimization enhances performance w.r.t the MSE as well a how the correlation in the channel may change this. 
		Note that once correlation is considered, we may want to decide the phase of elements in close proximity whose channels we have not yet estimated.
\end{enumerate}
\section{Optimization Problem}
First we consider equal power allocation at the transmitter such that after finding the first $K < N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem
is given by
	\begin{align}
	    \underset{\phase}{\text{minimize }}
	    & \; trace\left((\mI - \mathbf{A}\mathbf{A}^H)
	   (\mI - \mathbf{A}\mathbf{A}^H)^H \right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
	\end{align}
	with $\mathbf{A} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
	We classify the terms of the resulting polynomial into two types:
	cross types (e.g $(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$) and non-cross types (e.g $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$). For non-cross types, the phase terms will always cancel out ($\phi_i \phi_i^H = 1$). For the cross-terms, note that
	for each term in the polynomial, the hermetian of the term is also in the polynomial sum (e.g both $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$ and $ (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}\mG^H$). Because $= trace(\mH^H) = trace(\mH)^*$, the complex component of these terms will always cancel. The result is that the phases, $\phi_i \forall i \in {1 \cdots	 K}$ will cancel out of the MSE expression. This result would hold for any linear-beamformer is used at the transmitter.
\bibliography{bibliography}
\end{document}

### 2020-08-27 simulation/testing/Optimization/opt_test.py

@@ -1,12 +1,12 @@
 import numpy as np
 from src.sum_optimization import *
 from src.lin_alg import *
-num_elements = 10
+num_elements = 100
 num_distrete_vals = 4
 num_repetitions = 2
 rep_val = []
 complex_coefficient = np.random.uniform(-10, 10, 100) + 1j * np.random.uniform(-10, 10, 100)
-phase_adjacency = np.random.randint(0, 100, (num_elements, 4))
+phase_adjacency = np.random.randint(0, 100, (num_elements, 50))
 coded = np.zeros((num_elements, complex_coefficient.size))
 for i in range(phase_adjacency.shape[0]):
     for j in range(phase_adjacency.shape[1]):
@@ -17,24 +17,24 @@ o_complex_coefficient = complex_coefficient
 for rep_ind in range(num_repetitions):
     for elem_ind in range(num_elements):
         #   For each phase option choose the one resulting in the largest current sum
-        # best_phase_ind = 0
-        # min_power = np.inf
+        best_phase_ind = 0
+        min_power = np.inf
         poly_code = coded[elem_ind, :]
         off_elements = complex_coefficient*np.logical_not(poly_code)
-        # for phase_ind in range(num_distrete_vals):
-        #     phase_code = np.exp(1j*phase_ind*(np.pi/2))*poly_code
-        #     cur = np.abs(phase_code@complex_coefficient + np.sum(off_elements))
-        #     if cur < min_power:
-        #         min_power = cur
-        #         best_phase_ind = phase_ind
-        # adjustment = np.exp(1j*best_phase_ind*(np.pi/2))*poly_code*complex_coefficient
+        for phase_ind in range(num_distrete_vals):
+            phase_code = np.exp(1j*phase_ind*(np.pi/2))*poly_code
+            cur = np.abs(phase_code@complex_coefficient + np.sum(off_elements))
+            if cur < min_power:
+                min_power = cur
+                best_phase_ind = phase_ind
+        adjustment = np.exp(1j*best_phase_ind*(np.pi/2))*poly_code*complex_coefficient
         total_off = np.sum(off_elements)
-        best_phase = np.angle(total_off) + np.pi - np.angle(poly_code*complex_coefficient)
-        adjustment = np.exp(1j*best_phase)*poly_code*complex_coefficient
+        # best_phase = np.angle(total_off) + np.pi - np.angle(poly_code*complex_coefficient)
[…]

### 2020-08-28 documentation/Optimization Work/08_27_20.tex
Resolved bug causing conflict with theory and numerical (need perfect CSI at Tx)

@@
 	\end{equation*}
-In the following, we will assume that a matched filter is used at transmitter. Further we will evaluate the performance of the system using the Expected Mean Square Error given by 
+	The IRS component of the channel can be decomposed in the paths through the individual elements as
+	\begin{equation}
+	\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}= 
+	\Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
+	\end{equation}
+	for an IRS with $N_S$ reflective elements. 
+In the following, we will assume that a matched filter is used at the transmitter. Further we will evaluate the performance of the system using the Mean Square Error given by 
 \begin{equation}
@@
 in which we do not consider the noise at the receiver. 
-Note that we are using the full information of the current channel in order to estimate the optimal phases at the IRS. 
+In the following we assume that the elements of the IRS can not only adjust the phases of the impinging waves but may also be "turned off" such that they are not included in the channel.
 In order to minimize the overhead of pilot acquisition, we estimate the CSE in the following manner. 
-In the following we assume that the elements of the IRS can not only adjust the phases of the impinging waves but may also be "turned off" such that they are not included in the channel.
 \begin{enumerate}
 \item
-	In order to first transmit matched filter we leave all IRS elements \emph{on}.
-Assuming that we perfectly estimate the entire channel $\mH_{Total}$, we will not have sufficient information to estimate the individual components of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
+We leave all IRS elements \emph{on} to first find the full matched filter at the Tx.
+Assuming that we estimate the entire channel $\mH_{Total}$, we will not have sufficient information to estimate the individual components of the IRS component of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
 \item
@@
 \item
-	By turning on a specific IRS element (can we use patterns using the correlation?) and sending additional pilots, we can estimate 
+	Next, by turning on a specific IRS element (can we use patterns using the correlation?) and sending additional pilots, we can estimate 
 	the path through a specific IRS element given by $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$.
@@
 	\begin{equation}
-E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}) \|^2 \right]
+E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\vs  \|^2 \right]
 \end{equation}
@@
 \section{Optimization Problem}
+First, we rewrite the above utility function as
+	\begin{equation}
+E\left[ trace\left((\vs - \mathbf{H}\mathbf{F}\vs)(\vs - \mathbf{H}\mathbf{F}\vs)^H \right)\right]
+\end{equation}
+Assuming $\mathbf{H}$ and $\mathbf{F}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$ to give
+	\begin{equation}
+trace\left((\vs - \mathbf{H}\mathbf{F}\vs)(\vs - \mathbf{H}\mathbf{F}\vs)^H \right).
+\end{equation}
 First we consider equal power allocation at the transmitter such that after finding the first $K < N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem
@@
 	We classify the terms of the resulting polynomial into two types:
-	cross types (e.g $(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$) and non-cross types (e.g $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$).
-For non-cross types, the phase terms will always cancel out ($\phi_i \phi_i^H = 1$).
+	cross types (e.g $(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$) and non-cross terms (e.g $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$).
+For non-cross terms, the phases will always cancel out ($\phi_i \phi_i^H = 1$).
 For the cross-terms, note that
@@
 Because $= trace(\mH^H) = trace(\mH)^*$, the complex component of these terms will always cancel.
-The result is that the phases, $\phi_i \forall i \in {1 \cdots	 K}$ will cancel out of the MSE expression.
+The result is that the phases, $\phi_i  \;\forall i \in {1 \cdots	 K}$ will cancel out of the MSE expression.
 This result would hold for any linear-beamformer is used at the transmitter.
+\section{Numerical Inconsistency}
+Why is it that in the numerical results, the phase does have an impact on the MSE.
+
 \bibliography{bibliography}

### 2020-08-31 documentation/Optimization Work/08_27_20.tex
corrected phase error in update and corrected continuous version of optimization in code

@@
 	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
-	\end{align}
+	\end{align}\label{mse_problem}
 	with $\mathbf{A} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
@@
 	for each term in the polynomial, the hermetian of the term is also in the polynomial sum (e.g both $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$ and $ (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}\mG^H$).
-Because $= trace(\mH^H) = trace(\mH)^*$, the complex component of these terms will always cancel.
-The result is that the phases, $\phi_i  \;\forall i \in {1 \cdots	 K}$ will cancel out of the MSE expression.
-This result would hold for any linear-beamformer is used at the transmitter.
-\section{Numerical Inconsistency}
-Why is it that in the numerical results, the phase does have an impact on the MSE.
-
+Because $trace(\mH^H) = trace(\mH)^*$, the complex component of these terms will always cancel.
+In the case of minimizing the MSE, this means that we want to minimize the resulting real component for all of these cross terms.
+In order to investigate further, we decompose the problem further.
+Neglecting the non-cross terms and using the linearity $trace$ with respect to scaling, we can equivalently consider the problem
+	\begin{align}
+	    \underset{\phase}{\text{maximize }}
+	    & \; Img \left( \Sigma_{i=1}^{K} \Sigma_{j=i}^{K} c_{i,j} \phi_{i} \phi_{j}  \right)
+	     \\
+	    \text{subject to  } \; &
+	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}.
+	\end{align}\label{mse_problem_clear}
+	Note that because the polynomial in Problem \eqref{mse_problem} is the trace of a Hermetian product, only the imaginary component along the diagonal of each cross term is canceled by its corresponding Hermetian component in the sum.
+Replacing $ Img$ with a trigonometric function it is clear that this utility function is non-convex over the domain.
+\section{Phase Optimization}
+One potential method of performing this optimization is based on the ideas of coordinate descent in which a single variable is updated while all others are held constant.
+From problem \eqref{mse_problem_clear} it is clear that if all $\phi_{j} \; j \neq i $ are held constant, $\phi_{i}$ can be chosen as a phase which fulfills the objective function.
+The numerical results below indicate the impact of this optimization technique on the MSE of the IRS system described above.
 \bibliography{bibliography}

### 2020-08-31 simulation/testing/Optimization/MSE_vs_theta.py

@@ -4,7 +4,7 @@ import matplotlib.pyplot as plt
 num_tx = 16
 num_rx = 10
 num_reflectors = 50
-average = 100
+average = 1000
 # average = 10
 cor_mf_mu = []
 cor_mf_var = []
@@ -16,9 +16,9 @@ phases = np.linspace(0, 2 * np.pi, 10)
 fig = plt.figure()
 ave_plt = fig.add_subplot(2, 1, 1)
 var_plt = fig.add_subplot(2, 1, 2)
-transmit_cor = 0
-irs_cor = 0
-receiver_cor = 0
+transmit_cor = .5
+irs_cor = .5
+receiver_cor = .5
 MSE_zf_mu = []
 MSE_matched_mu = []
 MSE_matched_mu_rand = []
@@ -56,7 +56,8 @@ for theta in phases:
         # Using perfect
         transmit_signal_matched = precode_mf(channel) @ transmit_symbols
         received_matched = channel @ transmit_signal_matched + noise
-        received_matched_rand = channel_rand @ transmit_signal_matched + noise
+        transmit_signal_matched_rand = precode_mf(channel_rand) @ transmit_symbols
+        received_matched_rand = channel_rand @ transmit_signal_matched_rand + noise
         MSE_matched.append(10*np.log(np.power(np.linalg.norm(received_matched - transmit_symbols), 2)))
         MSE_matched_rand.append(10*np.log(np.power(np.linalg.norm(received_matched_rand - transmit_symbols), 2)))
         # zf_precoding = precode_zf(channel) @ transmit_symbols

### 2020-09-02 documentation/Optimization Work/08_27_20.tex
optimization for MMSE seems to be working some

@@
 \end{equation}
-First we consider equal power allocation at the transmitter such that after finding the first $K < N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem
+First we consider equal power allocation at the transmitter such that after finding the first $K \leq N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem
 is given by
@@
 In order to investigate further, we decompose the problem further.
-Neglecting the non-cross terms and using the linearity $trace$ with respect to scaling, we can equivalently consider the problem
+Neglecting the non-cross terms and using the linearity of the $trace$ operator with respect to scaling, we can equivalently consider the problem
 	\begin{align}
@@
 One potential method of performing this optimization is based on the ideas of coordinate descent in which a single variable is updated while all others are held constant.
-From problem \eqref{mse_problem_clear} it is clear that if all $\phi_{j} \; j \neq i $ are held constant, $\phi_{i}$ can be chosen as a phase which fulfills the objective function.
+From problem \eqref{mse_problem_clear} it is clear that if all $\phi_{j} \; j \neq i $ are held constant,$\phi_{j} \; j \neq i $can be chosen as a phase which fulfills the objective function. 
+For a given step of this algorithm in which $\phi_{i}$ the update is determined by a sum $\phi_{i}a + b$ and $\phi_{i}$ is chosen so at to maximize the 
+complex component of the sum (which we will be canceled out by its complex conjugate.
 The numerical results below indicate the impact of this optimization technique on the MSE of the IRS system described above. 

### 2020-09-02 simulation/testing/Optimization/opt_channel.py

@@ -4,15 +4,17 @@ from src.optimization import *
 
 
 average = 100
-num_tx = 4
-num_rx = 4
-num_reflectors = 4
-
+num_tx = 10
+num_rx = 10
+num_reflectors = 16
+MSE_rand = []
+MSE_optimized = []
 for j in range(average):
     #   Each IRS path is a column of F2 @ a row of F1
     F1 = c_rand(num_reflectors, num_tx, var=1)
     F2 = c_rand(num_rx, num_reflectors, var=1)
-    G = c_rand(num_rx, num_tx, var=1)
+    transmit_matched_opt = precode_mf(F2@F1)
+    transmit_matched_rand = precode_mf(F2@F1)
     coefficients = []
     for elem_ind in range(num_reflectors):
         ind = elem_ind * num_reflectors
@@ -22,9 +24,8 @@ for j in range(average):
             f1i = F1[elem_ind, :]
             f2j = F2[:, elem2_ind]
             f1j = F1[elem2_ind, :]
-            coefficient = f1i.T@np.conjugate(f1j)*hermetian(f2j)@f2i
+            coefficient = f1i.T@transmit_matched_opt@hermetian(transmit_matched_opt)@np.conjugate(f1j)*hermetian(f2j)@f2i
             coefficients.append(coefficient)
-
     phase_adjacency = np.zeros((num_reflectors, len(coefficients)))
     ind = 0
     for elem_ind in range(num_reflectors-1):
@@ -34,7 +35,17 @@ for j in range(average):
             phase_adjacency[elem3_ind, ind + elem2_ind] = 1
             pass
         ind += num_reflectors - elem_ind - 1
-    optimal_phases = optimize_phases(np.asarray(coefficients), phase_adjacency)
-    pass
+    # optimal_phases = optimize_phases(np.asarray(coefficients), phase_adjacency, 1)
+    optimal_phases = optimize_phases_max(-np.asarray(coefficients), phase_adjacency, 1)
+    #   Now perform comparison of optimized phases with random/uniform phases
+    rand = random_phase(num_reflectors)
+    transmit_symbols = c_rand(num_tx, 1, var=1).flatten()
+    received_rand = F2@rand@F1@transmit_matched_rand@transmit_symbols
+    received_opt = F2@np.diag(optimal_phases)@F1@transmit_matched_opt@transmit_symbols
+    # transmit_symbols = BPSK(num_rx)
[…]

### 2020-09-02 simulation/testing/Optimization/opt_test.py

@@ -18,7 +18,7 @@ for rep_ind in range(num_repetitions):
     for elem_ind in range(num_elements):
         #   For each phase option choose the one resulting in the largest current sum
         best_phase_ind = 0
-        min_power = np.inf
+        # min_power = np.inf
         poly_code = coded[elem_ind, :]
         off_elements = complex_coefficient*np.logical_not(poly_code)
         total_off = np.sum(off_elements)
@@ -33,8 +33,8 @@ for rep_ind in range(num_repetitions):
         best_phase = np.angle(total_off) + np.pi - np.angle(poly_code@complex_coefficient)
         adjustment = np.exp(1j*best_phase)*poly_code*complex_coefficient
         complex_coefficient = off_elements + adjustment
-        min_power = np.abs(np.sum(complex_coefficient))
-    rep_val.append(min_power)
+        # min_power = np.abs(np.sum(complex_coefficient))
+    rep_val.append(np.abs(np.sum(complex_coefficient)))
 gain = original - rep_val
 print(gain)
 #   Adjust the complex coefficients using the winning phase

### 2020-09-03 documentation/Optimization Work/08_27_20.tex
testing with complex conjugate for basic test

@@
 	\begin{equation}
-E\left[ trace\left((\vs - \mathbf{H}\mathbf{F}\vs)(\vs - \mathbf{H}\mathbf{F}\vs)^H \right)\right]
+E\left[ \trace\left((\vs - \mathbf{H}\mathbf{W}\vs)(\vs - \mathbf{H}\mathbf{W}\vs)^H \right)\right]
 \end{equation}
-Assuming $\mathbf{H}$ and $\mathbf{F}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$ to give
+Assuming $\mathbf{H}$ and $\mathbf{W}$ (The Tx linear precoder) to be constant over a coherence period, the expectation applies only to $\vs \vs^H$ 
+and we consider equal power allocation at the transmitter such that the utility function becomes
 	\begin{equation}
-trace\left((\vs - \mathbf{H}\mathbf{F}\vs)(\vs - \mathbf{H}\mathbf{F}\vs)^H \right).
+\trace\left((\mI - \mathbf{H}\mathbf{W})(\mI - \mathbf{H}\mathbf{W})^H \right).
 \end{equation}
-First we consider equal power allocation at the transmitter such that after finding the first $K \leq N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem
-is given by
+After finding the first $K \leq N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases for the case of a Matched Filter at the Tx is given by
 	\begin{align}
 	    \underset{\phase}{\text{minimize }}
-	    & \; trace\left((\mI - \mathbf{A}\mathbf{A}^H)
-	   (\mI - \mathbf{A}\mathbf{A}^H)^H \right)
+	    & \; \trace\left((\mI - \mathbf{H}\mathbf{H}^H)
+	   (\mI - \mathbf{H}\mathbf{H}^H)^H \right)
 	     \\
@@
 	\end{align}\label{mse_problem}
-	with $\mathbf{A} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
+	or further simplified into
+		\begin{align}
+	    \underset{\phase}{\text{minimize }}
+	    & \; - 2 \trace \left( \mathbf{H}\mathbf{H}^H\right) + \trace\left(\mathbf{H}\mathbf{H}^H \mathbf{H}\mathbf{H}^H\right)
+	     \\
+	    \text{subject to  } \; &
+	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
+	\end{align}\label{mse_problem_s}
+	with $\mathbf{H} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
 	We classify the terms of the resulting polynomial into two types:
@@
 The numerical results below indicate the impact of this optimization technique on the MSE of the IRS system described above.
+
+\begin{enumerate}
+\item
+	Always compare to the matched filter with random phases
+\end{enumerate}
 \bibliography{bibliography}

### 2020-09-03 simulation/src/optimization.py

@@ -34,7 +34,11 @@ def optimize_phases(coefficients, adjacency, num_repetitions=1):
             #         min_power = cur
             #         best_phase_ind = phase_ind
             # adjustment = np.exp(1j*best_phase_ind*(np.pi/2))*poly_code*complex_coefficient
-            best_phase = np.angle(total_off) + np.pi - np.angle(poly_code @ coefficients)
+            #   Adjust phase to minimize the real component of the resulting sum
+            # best_phase = np.angle(total_off) + np.pi - np.angle(poly_code @ coefficients)
+            best_phase = np.conjugate(np.angle(total_off)) - np.angle(poly_code @ coefficients)
+            # best_phase = np.angle(total_off) - np.angle(poly_code @ coefficients)
+            # best_phase = - np.angle(poly_code @ coefficients)
             final_phases[elem_ind] *= np.exp(1j * best_phase)
             adjustment = np.exp(1j * best_phase) * poly_code * coefficients
             coefficients = off_elements + adjustment

### 2020-09-03 simulation/testing/Optimization/opt_channel.py

@@ -3,7 +3,7 @@ from src.lin_alg import *
 from src.optimization import *
 import matplotlib.pyplot as plt 
 
-average = 10
+average = 200
 num_tx = 10
 num_rx = 10
 IRS_sizes = [10, 16, 64]
@@ -17,9 +17,10 @@ for size in IRS_sizes:
         #   Each IRS path is a column of F2 @ a row of F1
         F1 = c_rand(num_reflectors, num_tx, var=1)
         F2 = c_rand(num_rx, num_reflectors, var=1)
-        transmit_matched_opt = precode_mf(F2@F1)
-        transmit_matched_rand = precode_mf(F2@F1)
+        rand = random_phase(num_reflectors)
+        transmit_matched_rand = precode_mf(F2@rand@F1)
         coefficients = []
+        #   Choose the optimal IRS phases using the known channels with 0 phase at the IRS elements.
         for elem_ind in range(num_reflectors):
             ind = elem_ind * num_reflectors
             for elem2_ind in range(num_reflectors - elem_ind - 1):
@@ -28,7 +29,8 @@ for size in IRS_sizes:
                 f1i = F1[elem_ind, :]
                 f2j = F2[:, elem2_ind]
                 f1j = F1[elem2_ind, :]
-                coefficient = f1i.T@transmit_matched_opt@hermetian(transmit_matched_opt)@np.conjugate(f1j)*hermetian(f2j)@f2i
+                # coefficient = f1i.T@transmit_matched_rand@hermetian(transmit_matched_rand)@np.conjugate(f1j)*hermetian(f2j)@f2i
+                coefficient = f1i.T@np.conjugate(f1j)*hermetian(f2j)@f2i
                 coefficients.append(coefficient)
         phase_adjacency = np.zeros((num_reflectors, len(coefficients)))
         ind = 0
@@ -40,13 +42,16 @@ for size in IRS_sizes:
                 pass
             ind += num_reflectors - elem_ind - 1
         # optimal_phases = optimize_phases(np.asarray(coefficients), phase_adjacency, 1)
-        optimal_phases = optimize_phases_max(-np.asarray(coefficients), phase_adjacency, 1)
+        #TODO toggle negative
+        optimal_phases = optimize_phases(np.asarray(coefficients), phase_adjacency, 2)
+        check1 = np.trace(F2@rand@F1@hermetian(F2@rand@F1))
[…]

### 2020-09-03 simulation/testing/Optimization/opt_test.py

@@ -6,16 +6,20 @@ num_distrete_vals = 4
 num_repetitions = 2
 rep_val = []
 complex_coefficient = np.random.uniform(-10, 10, 100) + 1j * np.random.uniform(-10, 10, 100)
-phase_adjacency = np.random.randint(0, 100, (num_elements, 50))
+phase_adjacency = np.random.randint(0, 100, (num_elements, 10))
 coded = np.zeros((num_elements, complex_coefficient.size))
 for i in range(phase_adjacency.shape[0]):
     for j in range(phase_adjacency.shape[1]):
         coded[i, phase_adjacency[i, j]] = 1
+coded = np.concatenate((coded, coded), axis=1)
+complex_coefficient = np.concatenate((complex_coefficient, np.conjugate(complex_coefficient)))
 cur_sum = 0
 min_power = np.inf
 original = np.abs(np.sum(complex_coefficient))
+check = []
 for rep_ind in range(num_repetitions):
     for elem_ind in range(num_elements):
+        check.append(np.abs(np.sum(complex_coefficient)))
         #   For each phase option choose the one resulting in the largest current sum
         best_phase_ind = 0
         # min_power = np.inf
@@ -29,11 +33,12 @@ for rep_ind in range(num_repetitions):
         #         min_power = cur
         #         best_phase_ind = phase_ind
         # adjustment = np.exp(1j*best_phase_ind*(np.pi/2))*poly_code*complex_coefficient
-        check = np.angle(poly_code@complex_coefficient)
+        # check = np.angle(poly_code@complex_coefficient)
         best_phase = np.angle(total_off) + np.pi - np.angle(poly_code@complex_coefficient)
         adjustment = np.exp(1j*best_phase)*poly_code*complex_coefficient
         complex_coefficient = off_elements + adjustment
         # min_power = np.abs(np.sum(complex_coefficient))
+    # rep_val.append(np.abs(np.sum(complex_coefficient)))
     rep_val.append(np.abs(np.sum(complex_coefficient)))
 gain = original - rep_val
 print(gain)

### 2020-09-04 documentation/Optimization Work/08_27_20.tex
optimization test working and figured out why steps always improve

@@
 	with $\mathbf{H} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
-	We classify the terms of the resulting polynomial into two types:
-	cross types (e.g $(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$) and non-cross terms (e.g $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$).
-For non-cross terms, the phases will always cancel out ($\phi_i \phi_i^H = 1$).
-For the cross-terms, note that
-	for each term in the polynomial, the hermetian of the term is also in the polynomial sum (e.g both $\mG (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H$ and $ (\phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}\mG^H$).
-Because $trace(\mH^H) = trace(\mH)^*$, the complex component of these terms will always cancel.
-In the case of minimizing the MSE, this means that we want to minimize the resulting real component for all of these cross terms.
-In order to investigate further, we decompose the problem further.
-Neglecting the non-cross terms and using the linearity of the $trace$ operator with respect to scaling, we can equivalently consider the problem
+	Expanding the above utility function and neglecting the terms without any phase coefficients, the above utility function becomes
+	\begin{equation}
+	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H.
+	\end{equation}
+	From this expression it can be seen that the complex component of each term will be canceled by its complex conjugate.
+	One method to potentially perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
+	the resulting components of the sum have entirely negative and real. In this case, each step of the algorithm solves the problem
 			\begin{align}
-	    \underset{\phase}{\text{maximize }}
-	    & \; Img \left( \Sigma_{i=1}^{K} \Sigma_{j=i}^{K} c_{i,j} \phi_{i} \phi_{j}  \right)
+	    \underset{\phi_k}{\text{minimize }}
+	    & \; \phi_k c + \phi_k^* c^* + a 
 	     \\
 	    \text{subject to  } \; &
-	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}.
-	\end{align}\label{mse_problem_clear}
-	Note that because the polynomial in Problem \eqref{mse_problem} is the trace of a Hermetian product, only the imaginary component along the diagonal of each cross term is canceled by its corresponding Hermetian component in the sum.
-Replacing $ Img$ with a trigonometric function it is clear that this utility function is non-convex over the domain.
-\section{Phase Optimization}
-One potential method of performing this optimization is based on the ideas of coordinate descent in which a single variable is updated while all others are held constant.
-From problem \eqref{mse_problem_clear} it is clear that if all $\phi_{j} \; j \neq i $ are held constant,$\phi_{j} \; j \neq i $can be chosen as a phase which fulfills the objective function. 
-For a given step of this algorithm in which $\phi_{i}$ the update is determined by a sum $\phi_{i}a + b$ and $\phi_{i}$ is chosen so at to maximize the 
-complex component of the sum (which we will be canceled out by its complex conjugate.
-The numerical results below indicate the impact of this optimization technique on the MSE of the IRS system described above.
-
-\begin{enumerate}
-\item
-	Always compare to the matched filter with random phases
-\end{enumerate}
+	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
+	\end{align}\label{mse_problem_s}
+	in which $a \in \reals$ and $c \in \complex$ so the result is simply to choose $\phi_k$ to make $\phi_k c$ negative and real.
+	The primary difficulty to performing such and optimization is finding the all values of $ c_i $ for each $\phi_{k}$ as each phase will be associated with 
+	more than one $ c_i $ (particularly in the case of the matched filter precoder).
+\section{Optimization Implementation}
+To begin solving this problem I considered randomly selecting many complex numbers and corresponding phases in order to model the utility function 
+	\begin{equation}
+	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H.
+	\end{equation}
+	The algorithm then iterates through each $\phi_{k}$ shifts the phase such that the result is a negative real (only if this update improves the utility function).
 \bibliography{bibliography}

### 2020-09-04 simulation/src/optimization.py

@@ -6,7 +6,9 @@ import numpy as np
 def optimize_phases(coefficients, adjacency, num_repetitions=1):
     """
     Given a sum of complex numbers rotated by a number of phases with overlap, this is a sub-optimal algorithm for
-    finding a set of phases to minimize this sum.
+    finding a set of phases to minimize this sum. Note that the complex numbers are implied to be included in a
+    sum of the numbers (with phase offsets) and the complex conjugates of the nubmers. With that in mind this
+    always minimizes the real component of the sum in each step since the complex components would cancel anyways.
     :param coefficients:
     :param adjacency:
     :param num_repetitions:
@@ -14,45 +16,28 @@ def optimize_phases(coefficients, adjacency, num_repetitions=1):
     """
     #   Do not need to encode adjacency matrix since it is structured.
     num_elements = adjacency.shape[0]
-    final_phases = 1j*np.ones((adjacency.shape[0]))
+    final_phases = -1j*1j*np.ones((adjacency.shape[0]))
     #   First setup binary matrix for easier optimization
-    original = np.abs(np.sum(coefficients))
+    original = np.sum(coefficients) + np.sum(np.conjugate(coefficients))
     rep_val = []
+    progress = []
     for rep_ind in range(num_repetitions):
-        min_power = 0
         for elem_ind in range(num_elements):
             #   For each phase option choose the one resulting in the largest current sum
-            best_phase_ind = 0
-            min_power = np.inf
+            progress.append(np.sum(coefficients) + np.sum(np.conjugate(coefficients)))
             poly_code = adjacency[elem_ind, :]
             off_elements = coefficients * np.logical_not(poly_code)
-            total_off = np.sum(off_elements)
-            # for phase_ind in range(num_distrete_vals):
-            #     phase_code = np.exp(1j*phase_ind*(np.pi/2))*poly_code
-            #     cur = np.abs(phase_code@complex_coefficient + total_off)
-            #     if cur < min_power:
-            #         min_power = cur
-            #         best_phase_ind = phase_ind
-            # adjustment = np.exp(1j*best_phase_ind*(np.pi/2))*poly_code*complex_coefficient
-            #   Adjust phase to minimize the real component of the resulting sum
-            # best_phase = np.angle(total_off) + np.pi - np.angle(poly_code @ coefficients)
-            best_phase = np.conjugate(np.angle(total_off)) - np.angle(poly_code @ coefficients)
[…]

### 2020-09-08 documentation/Optimization Work/08_27_20.tex
Working on optimization for arbitrary beamformer

@@
 	for an IRS with $N_S$ reflective elements. 
-In the following, we will assume that a matched filter is used at the transmitter. Further we will evaluate the performance of the system using the Mean Square Error given by 
+In the following, we will assume that a matched filter is used at the transmitter and we will evaluate the performance of the system using the Mean Square Error given by 
 \begin{equation}
-E\left[\|  \vs - (\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})^H\vs) \|^2 \right]
+E\left[\|  \vs - (\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\mathbf{W}\vs) \|^2 \right]
 \end{equation}
-in which we do not consider the noise at the receiver. 
-In the following we assume that the elements of the IRS can not only adjust the phases of the impinging waves but may also be "turned off" such that they are not included in the channel.
+in which $\mathbf{W}$ is the linear precoder at the transmitter. We do not consider the noise at the receiver in this utility function. 
+In the following we assume that the elements of the IRS can adjust the phases of the impinging waves and may also be "turned off" such that they are not included in the channel.
 In order to minimize the overhead of pilot acquisition, we estimate the CSE in the following manner. 
@@
 \item
-We leave all IRS elements \emph{on} to first find the full matched filter at the Tx.
-Assuming that we estimate the entire channel $\mH_{Total}$, we will not have sufficient information to estimate the individual components of the IRS component of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
+We leave all IRS elements \emph{on} to first find the full channel   $\mH_{Total}$(and then transmit this to the Tx).
+At this point, we will not have sufficient information to estimate the individual components of the IRS component of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
 \item
@@
 	\begin{equation}
-E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})(\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\vs  \|^2 \right]
+E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\mathbf{W}\vs  \|^2 \right]
 \end{equation}
+in which $\mH_{Total} =  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$
 	\item
@@
 		We are investigating how this optimization enhances performance w.r.t the MSE as well a how the correlation in the channel may change this. 
-		Note that once correlation is considered, we may want to decide the phase of elements in close proximity whose channels we have not yet estimated.
+		Note that once correlation is considered, the order in which the channels through each element $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ may impact the optimization performance .
 \end{enumerate}
@@
 \end{equation}
-Assuming $\mathbf{H}$ and $\mathbf{W}$ (The Tx linear precoder) to be constant over a coherence period, the expectation applies only to $\vs \vs^H$ 
-and we consider equal power allocation at the transmitter such that the utility function becomes
+Assuming $\mathbf{H}$ and $\mathbf{W}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$.
+If we consider equal power allocation at the transmitter, the utility function becomes
 	\begin{equation}
@@
 \end{equation}
-After finding the first $K \leq N_S$ IRS channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases for the case of a Matched Filter at the Tx is given by
+After finding $K \leq N_S$ IRS element channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases for the case of a Matched Filter at the Tx is given by
 	\begin{align}
 	    \underset{\phase}{\text{minimize }}
-	    & \; \trace\left((\mI - \mathbf{H}\mathbf{H}^H)
-	   (\mI - \mathbf{H}\mathbf{H}^H)^H \right)
+	    & \; \trace\left((\mI - \mathbf{H}\mathbf{W})
+	   (\mI - \mathbf{H}\mathbf{W})^H \right)
 	     \\
@@
 	    \underset{\phase}{\text{minimize }}
-	    & \; - 2 \trace \left( \mathbf{H}\mathbf{H}^H\right) + \trace\left(\mathbf{H}\mathbf{H}^H \mathbf{H}\mathbf{H}^H\right)
+	    & \; - \trace \left( \mathbf{H}\mathbf{W}\right) - \trace \left( \mathbf{H}\mathbf{W}\right)^* +
+	    \trace\left(\mathbf{H}\mathbf{W} \mathbf{W}^H \mathbf{H}^H\right)
 	     \\
@@
 	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
-	\end{align}\label{mse_problem_s}
+	\end{align}\label{mse_problem_full}
 	with $\mathbf{H} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
-	Expanding the above utility function and neglecting the terms without any phase coefficients, the above utility function becomes
+	Expanding the above utility function and neglecting the terms without any phase coefficients, the cost function becomes
 	\begin{equation}
@@
 	One method to potentially perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
-	the resulting components of the sum have entirely negative and real. In this case, each step of the algorithm solves the problem
+	the corresponding component of the sum is real and negative. In this case, each step of the algorithm solves the problem
 			\begin{align}
@@
 	\end{align}\label{mse_problem_s}
-	in which $a \in \reals$ and $c \in \complex$ so the result is simply to choose $\phi_k$ to make $\phi_k c$ negative and real.
-	The primary difficulty to performing such and optimization is finding the all values of $ c_i $ for each $\phi_{k}$ as each phase will be associated with 
-	more than one $ c_i $ (particularly in the case of the matched filter precoder).
+	in which $a \in \reals$ and $c \in \complex$ so the result is simply to choose $\phi_k$ such that $\phi_k c$ is negative and real.
+
 \section{Optimization Implementation}
-To begin solving this problem I considered randomly selecting many complex numbers and corresponding phases in order to model the utility function 
+To begin solving this problem I considered randomly selecting many complex numbers and corresponding phases in order to model the cost function 
 	\begin{equation}
@@
 	\end{equation}
-	The algorithm then iterates through each $\phi_{k}$ shifts the phase such that the result is a negative real (only if this update improves the utility function).
+	The algorithm then iterates through each $\phi_{k}$ and shifts the phase such $\phi_{k} c_i$ a real and negative.
+	Below is an example in which 100 numbers and their complex conjugates from $\mathcal{NC}(0,1)$ are assigned phase coefficients which are then used in the above optimization 
+	problem. I have the algorithm for this test working but it is more difficult to apply this to the IRS case (numerical results) as I need to find the coefficient $c_i$ for each phase term.
+	\par
+	Next I considered the case in which an arbitrary linear precoding filter is used at the transmitter and the MSE is minimized.
+	In this case, each step of the algorithm is the same as Problem \eqref{mse_problem_s}.
+
+	 with the matched filter is finding the all values of $ c_i $ for each $\phi_{k}$ as each phase will be associated with more than one $ c_i $.
 \bibliography{bibliography}

### 2020-09-08 simulation/testing/Optimization/opt_channel.py

@@ -1,12 +1,11 @@
 import numpy as np
 from src.lin_alg import *
 from src.optimization import *
-import matplotlib.pyplot as plt 
-
-average = 200
-num_tx = 10
+import matplotlib.pyplot as plt
+average = 50
+num_tx = 16
 num_rx = 10
-IRS_sizes = [10, 16, 64]
+IRS_sizes = [16, 164, 128]
 MSE_opt_size = []
 MSE_rand_size = []
 for size in IRS_sizes:
@@ -17,48 +16,63 @@ for size in IRS_sizes:
         #   Each IRS path is a column of F2 @ a row of F1
         F1 = c_rand(num_reflectors, num_tx, var=1)
         F2 = c_rand(num_rx, num_reflectors, var=1)
+        H_t = F2@F1
         rand = random_phase(num_reflectors)
-        transmit_matched_rand = precode_mf(F2@F1)
+        beamformer = precode_mf(F2@F1)
         coefficients = []
-        #   Choose the optimal IRS phases using the known channels with 0 phase at the IRS elements.
+        # Fc = -1j*1j*np.zeros((num_rx, num_tx))
         for elem_ind in range(num_reflectors):
             ind = elem_ind * num_reflectors
+            f2i = F2[:, elem_ind]
+            f1i = F1[elem_ind, :]
+            i_mat = np.outer(f2i, f1i)
+            # Fc += i_mat
             for elem2_ind in range(num_reflectors - elem_ind - 1):
                 elem2_ind += elem_ind + 1
-                f2i = F2[:, elem_ind]
-                f1i = F1[elem_ind, :]
                 f2j = F2[:, elem2_ind]
                 f1j = F1[elem2_ind, :]
-                coefficient = f1i.T@transmit_matched_rand@hermetian(transmit_matched_rand)@np.conjugate(f1j)*hermetian(f2j)@f2i
+                j_mat = np.outer(f2j, f1j)
+                coefficient = f1i.T@beamformer@hermetian(beamformer)@np.conjugate(f1j)*hermetian(f2j)@f2i
                 # coefficient = f1i.T@np.conjugate(f1j)*hermetian(f2j)@f2i
+                # coefficient1 = np.trace(i_mat@transmit_matched_rand@hermetian(transmit_matched_rand)@hermetian(j_mat))
                 coefficients.append(coefficient)
-        phase_adjacency = np.zeros((num_reflectors, len(coefficients)))
[…]

### 2020-09-09 documentation/Optimization Work/08_27_20.tex
adding results to update

@@
 	for an IRS with $N_S$ reflective elements. 
-In the following, we will assume that a matched filter is used at the transmitter and we will evaluate the performance of the system using the Mean Square Error given by 
+In the following we will evaluate the performance of this system using the Mean Square Error given by 
 \begin{equation}
@@
 \end{equation}
-in which $\mathbf{W}$ is the linear precoder at the transmitter. We do not consider the noise at the receiver in this utility function. 
-In the following we assume that the elements of the IRS can adjust the phases of the impinging waves and may also be "turned off" such that they are not included in the channel.
-In order to minimize the overhead of pilot acquisition, we estimate the CSE in the following manner. 
+in which $\mathbf{W}$ is the linear precoder at the transmitter. Note that receiver noise is not considered in this utility function. 
+In the following ,we assume that the elements of the IRS can adjust the phases of the impinging waves and may also be "turned off" such that they reflect any wave.
+CSI is estimated using the following procedure.
 \begin{enumerate}
 \item
-We leave all IRS elements \emph{on} to first find the full channel   $\mH_{Total}$(and then transmit this to the Tx).
+We leave all IRS elements \emph{on} and set the phase to a known value (e.g.
+$\phi_i = 1 $) to first find the full channel   $\mH_{Total}$(and then transmit this to the Tx).
 At this point, we will not have sufficient information to estimate the individual components of the IRS component of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
@@
 \item
-	Next, by turning on a specific IRS element (can we use patterns using the correlation?) and sending additional pilots, we can estimate 
+	Next, by turning on a specific IRS element and sending additional pilots, we can estimate 
 	the path through a specific IRS element given by $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$.
@@
 \end{equation}
-Assuming $\mathbf{H}$ and $\mathbf{W}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$.
-If we consider equal power allocation at the transmitter, the utility function becomes
+Assuming $\mathbf{H}$ and $\mathbf{W}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$. If we consider equal power allocation at the transmitter, this becomes
 	\begin{equation}
@@
 \end{equation}
-After finding $K \leq N_S$ IRS element channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases for the case of a Matched Filter at the Tx is given by
+After finding $K \leq N_S$ IRS element channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases is given by
 	\begin{align}
@@
 	\end{align}\label{mse_problem}
-	or further simplified into
+	which is further simplified into
 		\begin{align}
@@
 	From this expression it can be seen that the complex component of each term will be canceled by its complex conjugate.
-	One method to potentially perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
+	One method to perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
 	the corresponding component of the sum is real and negative. In this case, each step of the algorithm solves the problem
@@
 \section{Optimization Implementation}
-To begin solving this problem I considered randomly selecting many complex numbers and corresponding phases in order to model the cost function 
-	\begin{equation}
-	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H.
-	\end{equation}
-	The algorithm then iterates through each $\phi_{k}$ and shifts the phase such $\phi_{k} c_i$ a real and negative.
-	Below is an example in which 100 numbers and their complex conjugates from $\mathcal{NC}(0,1)$ are assigned phase coefficients which are then used in the above optimization 
-	problem. I have the algorithm for this test working but it is more difficult to apply this to the IRS case (numerical results) as I need to find the coefficient $c_i$ for each phase term.
-	\par
-	Next I considered the case in which an arbitrary linear precoding filter is used at the transmitter and the MSE is minimized.
+	First, I considered the case in which an arbitrary linear precoding filter is used at the transmitter (in this case the matched filter to the channel when all IRS elements have phase $\phi_i = 1 $) and the MSE is minimized.
 	In this case, each step of the algorithm is the same as Problem \eqref{mse_problem_s}.
+	The results for a system using a matched filter (for the channel before IRS updates) at the transmitter are shown in \ref{MSE_opt} and indicate that increasing the number of elements at the IRS 
+	can allow for improved optimization gains. 
+	\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/MSE_optimized}
+	  \caption{MSE with/without optimization versus number of IRS elements without receiver noise.}
+	  	  \label{MSE_opt}
+	\end{figure}
 
-	 with the matched filter is finding the all values of $ c_i $ for each $\phi_{k}$ as each phase will be associated with more than one $ c_i $.
+	\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/MSE_optimized_noise}
+	  \caption{MSE with/without optimization versus number of IRS elements with receiver noise.}
+	  	  \label{MSE_opt_noise}
+	\end{figure}
+Seems strange because for infinite SNR case, Matched filter should have optimal MSE right?
+Next the same results are generated for the case in which the zero-forcing filter to the channel with all phase elements set to 0 is used. 
+\section{Next Steps}
+The first next step will be to include the impact of the phases on the choice of the beamformer matrix. To simplify the optimization the beamformer was fixed in the above results. 
 \bibliography{bibliography}

### 2020-09-09 simulation/testing/Optimization/opt_channel.py

@@ -2,10 +2,10 @@ import numpy as np
 from src.lin_alg import *
 from src.optimization import *
 import matplotlib.pyplot as plt
-average = 50
+average = 100
 num_tx = 16
 num_rx = 10
-IRS_sizes = [16, 64]
+IRS_sizes = [16, 64, 128]
 MSE_opt_size = []
 MSE_rand_size = []
 for size in IRS_sizes:
@@ -56,16 +56,16 @@ for size in IRS_sizes:
         optimal_phases, improvements = optimize_phases(np.asarray(coefficients), phase_adjacency, F2, F1, num_repetitions=2, improvements=True)
         H_opt = F2@optimal_phases@F1
         check_1 = np.trace(-H_t@beamformer) + np.conjugate(np.trace(-H_t@beamformer)) + np.trace(H_t@beamformer@hermetian(H_t@beamformer))
-        check_2 = improvements[-1]
-        check_3 = np.trace(-H_opt@beamformer) + np.conjugate(np.trace(-H_opt@beamformer)) + np.trace(H_opt@beamformer@hermetian(H_opt@beamformer))
+        check_2 = np.trace(-H_opt@beamformer) + np.conjugate(np.trace(-H_opt@beamformer)) + np.trace(H_opt@beamformer@hermetian(H_opt@beamformer))
+
         #   Now perform comparison of optimized phases with random/uniform phases
         # transmit_symbols = BPSK(num_rx)
         noise = c_rand(num_rx, 1, var=1).flatten()
         transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
-        received_rand = F2@F1@beamformer@transmit_symbols
+        received_rand = F2@F1@beamformer@transmit_symbols + noise
         #   Line below could be done by going through each element and shifting the phase. Line below is faster.
         transmit_matched_opt = precode_mf(F2@optimal_phases@F1)
-        received_opt = F2@optimal_phases@F1@beamformer@transmit_symbols
+        received_opt = F2@optimal_phases@F1@beamformer@transmit_symbols + noise
         MSE_rand.append(10 * np.log(np.power(np.linalg.norm(received_rand - transmit_symbols), 2)))
         MSE_optimized.append(10 * np.log(np.power(np.linalg.norm(received_opt - transmit_symbols), 2)))
[…]

### 2020-09-10 documentation/Optimization Work/09_10_20.tex
Added some details for progress report

--- 2020-09-09 documentation/Optimization Work/08_27_20.tex
@@
 \end{equation}
+or equivalently
+	\begin{equation}
+ \trace\left((\mI - \mathbf{H}\mathbf{W})E\left[\vs\vs^H\right](\mI - \mathbf{H}\mathbf{W})^H \right)
+\end{equation}
 Assuming $\mathbf{H}$ and $\mathbf{W}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$. If we consider equal power allocation at the transmitter, this becomes
@@
 \end{equation}
-After finding $K \leq N_S$ IRS element channels $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, the optimization problem with respect to the IRS phases is given by
+The MSE optimization with respect to the IRS phases is given by
 	\begin{align}
@@
 	which is further simplified into
+	\begin{subequations}
+	\label{mse_problem_s}
 	\begin{align}
-	    \underset{\phase}{\text{minimize }}
+	   	    \underset{\phi_k}{\text{minimize }}
 	    & \; - \trace \left( \mathbf{H}\mathbf{W}\right) - \trace \left( \mathbf{H}\mathbf{W}\right)^* +
@@
 	    \text{subject to  } \; &
-	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
-	\end{align}\label{mse_problem_full}
+	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
+	\end{align}
+	\end{subequations}
+
 	with $\mathbf{H} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
@@
 	\begin{equation}
-	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H.
+	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H*
 	\end{equation}
@@
 	One method to perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
-	the corresponding component of the sum is real and negative. In this case, each step of the algorithm solves the problem
+	the corresponding component of the sum is real and negative. Each step of the algorithm solves the problem
+	\begin{subequations}
+	\label{algo_step}
 	\begin{align}
@@
 	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
-	\end{align}\label{mse_problem_s}
+	\end{align}
+	\end{subequations}
 	in which $a \in \reals$ and $c \in \complex$ so the result is simply to choose $\phi_k$ such that $\phi_k c$ is negative and real.
+	Note that if the transmitter has found $K \leq N_S$ IRS element channels, $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, then the optimization can only be performed over
+	the elements whose channel , $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, is known (i.e. Step 3 of the pilot symbols has been performed for $\phi_i$).
 
 \section{Optimization Implementation}
-	First, I considered the case in which an arbitrary linear precoding filter is used at the transmitter (in this case the matched filter to the channel when all IRS elements have phase $\phi_i = 1 $) and the MSE is minimized.
-	In this case, each step of the algorithm is the same as Problem \eqref{mse_problem_s}.
-	The results for a system using a matched filter (for the channel before IRS updates) at the transmitter are shown in \ref{MSE_opt} and indicate that increasing the number of elements at the IRS 
-	can allow for improved optimization gains. 
+	First, I considered the case in which an arbitrary linear precoding filter is used at the transmitter (in this case the matched filter to the channel when all IRS elements have phase $\phi_i = 1 $).
+	Each step of the algorithm is the same as Problem \eqref{algo_step}.
+	All channel elements, transmit symbols and noise are generated using $\mathcal{CN}(0,1)$.
+	The results are shown in Figures \ref{MSE_opt} and \ref{MSE_opt_noise} for the cases with and without noise at the receiver.
+These results indicate that increasing the number of elements at the IRS can allow for improved optimization gains.
 	\begin{figure}[H]
@@
 	\end{figure}
-
+		The result in Figure \ref{MSE_opt} seems  strange because for the infinite SNR case (i.e. no receiver noise), the matched filter should have optimal MSE right?
 	\begin{figure}[H]
@@
 	\end{figure}
-Seems strange because for infinite SNR case, Matched filter should have optimal MSE right?
-Next the same results are generated for the case in which the zero-forcing filter to the channel with all phase elements set to 0 is used. 
+
+%For certain choices of $\mathbf{W}$ this phase optimization could be performed during the phase optimization step of an alternating optimization algorithm.
 \section{Next Steps}
-The first next step will be to include the impact of the phases on the choice of the beamformer matrix. To simplify the optimization the beamformer was fixed in the above results. 
+\begin{itemize}
+\item
+The next step is to extend the MSE/phase optimization to also include the impact of the phases on the beamformer. For example in the case of the matched filter, this would take the form of 
+		\begin{align}
+	    \underset{\phase}{\text{minimize }}
+	    & \; - \trace \left( \mathbf{H}\mathbf{H}\right) - \trace \left( \mathbf{H}\mathbf{H}\right)^* +
+	    \trace\left(\mathbf{H}\mathbf{H} \mathbf{H}^H \mathbf{H}^H\right)
+	     \\
+	    \text{subject to  } \; &
+	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}.
+	\end{align}\label{mse_problem_full}
+To simplify the initial optimization setup in the above results, the beamformer was fixed before the MSE optimization w.r.t.
+the phases.
+For a linear beamformer, this should result in a similar optimization setup as \ref{mse_problem_s}, however, finding the coefficients to the phases will be more difficult and there will be many more terms in the sum.
+\item
+	Next, the impact of CSI knowledge on the optimization gain should be studied.
+In this case, only some of the IRS paths should be used in the optimization and the result should be compared to the full CSI case.
+\item
+	Finally, the impact of correlation of the impact of the phase optimization should be studied.
+And perhaps look at the optimal order in which to estimate CSI for the correlated system if not all channels will be found.
+\end{itemize}
 \bibliography{bibliography}

### 2020-09-10 simulation/src/optimization.py

@@ -4,7 +4,7 @@ Optimization tools for selecting the phases as IRS elements
 import numpy as np
 from src.lin_alg import *
 
-def optimize_phases(coefficients, adjacency, F2, F1, num_repetitions=1, improvements=False):
+def optimize_phases(coefficients, adjacency, F2, F1,  csi = 1, num_repetitions=1, improvements=False):
     """
     Given a sum of complex numbers rotated by a number of phases with overlap, this is a sub-optimal algorithm for
     finding a set of phases to minimize this sum. Note that the complex numbers are implied to be included in a
@@ -15,15 +15,17 @@ def optimize_phases(coefficients, adjacency, F2, F1, num_repetitions=1, improvem
     :param num_repetitions:
     :return:
     """
+    #   Check how many or the IRS paths are known.
+
     #   Do not need to encode adjacency matrix since it is structured.
-    num_elements = adjacency.shape[0]
+    effective_num_elements = int(adjacency.shape[0]*csi)
     final_phases = -1j*1j*np.ones((adjacency.shape[0]))
     #   First setup binary matrix for easier optimization
     H_opt = F2 @ F1
     beamformer = precode_mf(H_opt)
     progress = [np.trace(-H_opt@beamformer) + np.conjugate(np.trace(-H_opt@beamformer)) + np.trace(H_opt@beamformer@hermetian(H_opt@beamformer))]
     for rep_ind in range(num_repetitions):
-        for elem_ind in range(num_elements):
+        for elem_ind in range(effective_num_elements):
             #   For each phase option choose the one resulting in the largest current sum
             poly_code = adjacency[elem_ind, :]
             off_elements = coefficients * np.logical_not(poly_code)

### 2020-09-10 simulation/testing/Optimization/opt_channel.py

@@ -19,6 +19,7 @@ for size in IRS_sizes:
         H_t = F2@F1
         rand = random_phase(num_reflectors)
         beamformer = precode_mf(F2@F1)
+        # beamformer = precode_zf(F2@F1)
         coefficients = []
         non_cross = -1j*1j*0
         for elem_ind in range(num_reflectors):
@@ -53,28 +54,35 @@ for size in IRS_sizes:
                 elem3_ind = elem2_ind + elem_ind + 1
                 phase_adjacency[elem3_ind, ind + elem2_ind] = 1
             ind += num_reflectors - elem_ind - 1
-        optimal_phases, improvements = optimize_phases(np.asarray(coefficients), phase_adjacency, F2, F1, num_repetitions=2, improvements=True)
+        optimal_phases, improvements = optimize_phases(np.asarray(coefficients),
+                                                       phase_adjacency, F2, F1, csi=.5, num_repetitions=2,
+                                                       improvements=True)
         H_opt = F2@optimal_phases@F1
         check_1 = np.trace(-H_t@beamformer) + np.conjugate(np.trace(-H_t@beamformer)) + np.trace(H_t@beamformer@hermetian(H_t@beamformer))
         check_2 = np.trace(-H_opt@beamformer) + np.conjugate(np.trace(-H_opt@beamformer)) + np.trace(H_opt@beamformer@hermetian(H_opt@beamformer))
-
         #   Now perform comparison of optimized phases with random/uniform phases
         # transmit_symbols = BPSK(num_rx)
-        noise = c_rand(num_rx, 1, var=1).flatten()
-        transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
-        received_rand = F2@F1@beamformer@transmit_symbols + noise
+        transmit_averge = 100
+        trasmit_ave_list_rand = []
+        trasmit_ave_list_opt = []
+        for i in range(transmit_averge):
+            noise = c_rand(num_rx, 1, var=1).flatten()
+            transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
+            received_rand = F2@F1@beamformer@transmit_symbols
+            trasmit_ave_list_rand.append(np.power(np.linalg.norm(received_rand - transmit_symbols), 2))
         #   Line below could be done by going through each element and shifting the phase. Line below is faster.
[…]

### 2020-09-19 documentation/Paper/abstract.tex
updates to report

@@
-Surfaces with reflective phase control (IRS) have been the subject of interest in recent years as means to better control the wireless propagation environment.
-This work investigates the impact of an IRS on the capacity of MIMO communication systems.
+Surfaces which can control the phase of reflected radio waves, known as intelligent reflective surfaces (IRS), have been the subject of interest in recent years as a means to better control the wireless propagation environment.
+This work investigates the impact of an IRS on the MIMO communication systems.
 Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity.
-A key finding is that in the asymptotic domain, the choice for the phase of reflected waves has no impact for the asymptotic analysis.
-Motivated the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior.
-We then investigate new methods for phase optimization for systems without full channel state information at the IRS.
+A key finding is that the phase control offered by the IRS has no impact in the asymptotic domain.
+Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior.
+We then propose an optimization method for the non-convex problem of minimizing the Mean Square Error with respect to the IRS phases.

### 2020-09-19 documentation/Paper/free_probability.tex
updates to report

--- 2020-07-22 documentation/Paper/free_probability.tex
@@
+Equation \eqref{no_csi_capacity_aed} illustrates that the eigenvalue distribution of the received signal covariance matrix is sufficient information to find the capacity of a MIMO channel.
+For general, random channels, finding this eigenvalue distribution is not a straightforward task.
+We now introduce some of the key tools that will be used in order to perform this analysis.
 \begin{itemize}
-
 \item 

### 2020-09-19 documentation/Paper/information_theory_background.tex
updates to report

@@
 \end{equation}
-If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is also circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
+If both $\vx$ and $\vn$ are circularly symmetric, Gaussian random vectors, $\vy$ is also circularly symmetric and Gaussian with entropy given by \cite{telatar1999capacity} 
 \begin{equation}\label{entropy}
@@
 \end{equation}
- for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$.
+ in which $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$.
 Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.
 (explain what implication this has)
-For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$, substituting equation \eqref{entropy} into equation
+For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$ (allow for this to be different), substituting equation \eqref{entropy} into equation
 \eqref{mut_ent} and simplifying gives
@@
 \end{equation}
-Maximizing the mutual informationwith respect to both the
+Maximizing the mutual information with respect to both the
 transmit covariance matrix and phases at the RIS gives an expression for the system capacity
@@
 \end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mQ_x \mHt^H$.
+in which $\lambda_i$ are the eigenvalues of $\mHt \mQ_x \mHt^H$.
 
@@
 First, we need to be able to evaluate an expectation over the joint pdf $p(\lambda_1 \cdots	 \lambda_N)$.
-Second, there is an optimization problem whose solution may in general depend on the distribution of the channel.
+Second, there is an optimization problem whose solution will generally depend on the distribution of the channel.
 Including this solution in the capacity expression without a closed form may be difficult.
 To resolve these hurdles, two simplifications will be used.
-First, we will enforce that the transmitter chooses $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$. 
+First, we will enforce that the transmitter chooses $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$ (Justify this by saying we assume sufficient SNR?). 
 By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the expectation.
 Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge. 
-These assumptions lead to the form of equation \eqref{capacity}
+With thesis assumptions, equation \eqref{capacity} becomes
 \begin{equation}\label{no_csi_capacity}
@@
 \end{equation}
-where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
-In this form, we can clearly see the linear scaling of capacity with $N_R$ that has motivated MIMO research.
+This form is useful in observing the linear scaling of capacity with $N_R$ that has motivated MIMO research.
 If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation over the AED to get the expression
@@
 \end{equation}
-in which $p_{\lambda\lambda^H}(x)$ is the channel AED and the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix.
-Note that because the choice for the distributions of $\vx$ and $\vn$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
-(Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).
+in which $p_{\lambda\lambda^H}(x)$ is the channel AED and the bounds of the integral reflect the range of eigenvalues for a positive semidefinite matrix.
+Note that because the choice for the distributions of $\vx$ and $\vn$ maximize $h(\vy)$, equation \eqref{no_csi_capacity_aed} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$.
 \par
-Discuss if $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$ is optimal at high SNR (are we in the low SNR regeime?)
-\par
-Note that because $\Log$ is a concave function, we can upper bound the mutual information with
+Also note that because $\Log$ is a concave function, we can upper bound the mutual information with
 \begin{equation}\label{mut_inf}

### 2020-09-19 documentation/Paper/irs_channel.tex
updates to report

@@
-Using the approach from Section \ref{Free_Prob_Intro} to find the asymptotic eigenvalue distribution, we now analyze a channel model that is representative of a MIMO communication system with an RIS. 
+Using the approach from Section \ref{Free_Prob_Intro} to find the asymptotic eigenvalue distribution, we now analyze a channel model that is representative of a MIMO communication system with an IRS. 
 \par
 In general, the received signal from any line of sight path between a transmitter and receiver would contain  sufficient energy to disregard the contribution of any reflected signal (although this component may be very low rank).
-Therefore, we focus on a channel model in which all signal power to the receiver is through the path of the RIS  given by
+Therefore, we focus on a channel model in which all signal power to the receiver is through the path of the IRS  given by
 
@@
 	\end{equation}
-The random channels $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ represent the random channel gains from transmitter to RIS and from RIS to receiver, respectively.
-Similarly, $ \mathbf{R}_{T},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, RIS and receiver, respectively.
+The random channels $\mathbf{H}_{1}$ and $\mathbf{H}_{2}$ represent the random channel gains from transmitter to IRS and from IRS to receiver, respectively.
+Similarly, $ \mathbf{R}_{T},  \mathbf{R}_{S} $ and  $\mathbf{R}_{S}$ represent the correlation introduced at the transmitter, IRS and receiver, respectively.
 When compared with the example seen in Section \ref{Free_Prob_Intro}, the structure of $\mathbf{H}_{Total}$ suggests that we can to use multiplicative free convolution in order to evalute the AED.
 With this in mind, we would like to choose a correlation model  that not only represents the system but one with a known S-Transform expression.
-One such model is exponential correlation. Defined by a single parameter $\rho$, the elements of the corresponding correlation matrix are given by $\rho^{i-j}$.
+One commonly used model exponential correlation. Defined by a single parameter $\rho$, the elements of the corresponding correlation matrix are given by $\rho^{i-j}$.
 \par
-To begin using the tools of free probability, we must first separate out the free sub-components of the covariance matrix given by 
+To begin using the tools of free probability, we first separate out the free sub-components of the total covariance matrix given by 
 	\begin{equation}
@@
 \end{equation}
-If all system dimensions are $N$, these S-Trasforms (and AEDs) identical so for this section we will assume that all system dimensions are $N$.
-Repeated iteration of the rotational property above gives the equivalent matrix
+If all system dimensions are $N$, these S-Trasforms equal. For simplified analysis and notation, we will assume that all system dimensions are $N$.
+Repeated iteration of the property above gives the equivalent matrix
 	\begin{equation}
@@
 Note that using the Eigen Value Decomposition of $\mR_{R}$ and $\mR_{T}$, we find the product $\mR_{R} \mR_{T}$ is also a Toeplitz matix (which can be represented by a circulant matrix in the asymptotic system size).
-In \cite[Section 4.3.2]{muller2013applications} it is shown that the hermetian matrix $\mH\mH_^H$ with $h_{i,j} \in \NC\{0, \frac{1}{N}\}$
-is free 
+In \cite[Section 4.3.2]{muller2013applications} it is shown that the hermetian matrix $\mH\mH^H$ with $h_{i,j} \in \CNO$
+is free (relative to what?)...
 Having shown the relative freeness of the matrix sub-components, we must now find the S-transform for each in order to find the product.
@@
 \par
-Without further analysis, this expression shows the interesting result that under the assumptions of Free Probability and equal power transmission, the phases at the RIS (or indeed any unitary matrix in the channel) cancel out of the expression for the AED.
+Without further analysis, this expression shows the interesting result that under the assumptions of Free Probability and equal power transmission, the phases at the IRS (or any unitary matrix in the channel) cancel out of the expression for the AED.
 For the remaining terms, we simply need an expression for the S-Transform in order to evaluate the asymptotic system capacity.s
@@
 While we do not consider a line of sight component to this channel, using the tools from a similar analysis in \cite{muller2012channel}, it can be shown that 
-a similar cancellation of phases will occur for the RIS channel with an LOS (or non-zero mean) component of the signal.
+a similar cancellation of phases will occur for the IRS channel with an LOS (or non-zero mean) component of the signal.
 Note that if we are simply adding a non-zero mean component.
-We can find the corresponding AED by looking considering the channel without the non-zero mean and then adding the single eigenvalue from the non-zero mean component; in which case the phases will again cancel out.
+We can find the corresponding AED by considering the channel without the non-zero mean and then adding the single eigenvalue from the non-zero mean component; in which case the phases will again cancel out.

### 2020-09-19 documentation/Paper/non_asymptotic.tex
updates to report

@@
-Up to this point we have shown an analysis which seems to contradict the observations from previous investigations review in Section \ref{Background} showing the benefits of exploiting RIS phases.
+Up to this point we have shown an analysis which seems to contradict the observations from previous investigations reviewed in Section \ref{Background} showing the benefits of exploiting phase control at the IRS.
 The key to the coexistence of these results is determining when the asymptotic assumptions required by Free Probability hold.
-As mentioned in in Section \ref{Background}, \cite {} have shown that for matrices of random elements, the  convergence of the matrix behavior towards that predicted by Free Probability is slowed by adding correlations between elements.
+As mentioned in in Section \ref{Background}, \cite {} have shown that for matrices of random elements, the  convergence of the matrix characteristics towards that predicted by Free Probability is slowed by adding correlations between elements.
 We would now like to at least numerically investigate this question to investigate the performance of RIS MIMO systems with parameters reflecting state of the art.
-
+(This is where being able to analyze the capacity using the AED from above would provide nice results)
  As seen in \ref{Background} current large systems have been shown for 

### 2020-09-19 documentation/Paper/paper.tex
Adding SINR/achievable with corrected model variance
updates to report

@@
 
-\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
+\title{Free Probability Analysis for Reflective Surface aided Massive MIMO systems}
 \author{Peter Hartig}
@@
 
-\chapter{Analyzing the RIS System}\label{system_model}
+\chapter{Analyzing the IRS System}\label{system_model}
 \section{Analysis Preliminaries}
@@
 \input{free_probability}
-\section{RIS Analysis}
+\section{Applying Free Probability to the IRS Chanel}
 \subsection{The Asyptotic IRS Channel}
@@
 \section{Asymptotic Results}
-
 \subsection{Asymptotic Results Discussion}

### 2020-09-19 documentation/Paper/power_normalization.tex
updates to report

@@
-Consider the SISO channel with an RIS given by $y = \vh_{rs}^H \phase \vh_{st}$. In this case, $ \phase$ should be chosen optimally such that 
-the phases all add constructively at the receiver (\ie $h_{rs,i} \phi_k h_{st,j}$ add constructively at the receiver). 
-In the case of an i.i.d Rayleigh channel with elements $\CNO$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
-Intuitively if the channel gains are normalized such that the received power is normalized for any system size $N$ the IRS offers no gain. 
-In this case we can achieve this normalization by enforcing that elements of the a channel matrix $\mH \in \CN^{N \times K}$ have variance $\frac{1}{N}$.
-This observation highlights an point that must be considered when performing asymptotic analysis; the received power must be normalized.
+Consider the SISO channel with an IRS given by $y = \vh_{rs}^H \phase \vh_{st}$. To maximize capacity in this case, $ \phase$ should be chosen such that 
+the each incoming signal at the receiver adds constructively (\eg $\angle h_{rs,i} \phi_k h_{st,j} = 0$). 
+In the case of an i.i.d Rayleigh channel with elements $\mathcal{CN} (0,1)$ and $\sigma^2_{\text{noise}} =1$, the SNR at the receiver is given by $\gamma = N$.
+This observation highlights the point that when performing asymptotic analysis (\ie when $N \rightarrow \infty$) the received power (or metric of interest) should be normalized with respect to the system size.
+Extending this to the case of a MIMO channel, for an $N_R \times N_T$ channel $\mH$, we can achieve this normalization by enforcing that elements of the a channel matrix  have variance $\frac{1}{N_R}$ such that even for concatenated MIMO channels (\eg $\mH_2 \mH_1$) the power of the received signal is still
+ $\Expect \left[ \trace \left(\mH_2 \mH_1\vx (\mH_2 \mH_1\vx)^H \right)\right] = 1$ when the total transmit power is normalized to 1. 
+ (Discuss why this is important to consider the conservation of energy?

### 2020-09-19 simulation/testing/Optimization/opt_num_CSI.py

@@ -1,0 +1,102 @@
+import numpy as np
+from src.lin_alg import *
+from src.optimization import *
+import matplotlib.pyplot as plt
+
+average = 100
+num_tx = 100
+num_rx = 100
+MSE_opt_size = []
+MSE_rand_size = []
+csi_percentages = [.01, .25, .75, 1]
+cor_rho = 0
+num_reflectors = 100
+for percent in csi_percentages:
+    MSE_rand = []
+    MSE_optimized = []
+    for j in range(average):
+        #   Each IRS path is a column of F2 @ a row of F1
+        s_correlation_matrix = exponential_correlation(num_reflectors, cor_rho)
+        r_correlation_matrix = exponential_correlation(num_rx, cor_rho)
+        # F1 = s_correlation_matrix@c_rand(num_reflectors, num_tx, var=1)
+        # F2 = r_correlation_matrix@c_rand(num_rx, num_reflectors, var=1)
+        F1 = s_correlation_matrix @ c_rand(num_reflectors, num_tx)
+        F2 = r_correlation_matrix @ c_rand(num_rx, num_reflectors)
+        H_t = F2 @ F1
+        rand = random_phase(num_reflectors)
+        # beamformer = precode_mf(H_t)
+        beamformer = precode_zf(F2@F1)
+        coefficients = []
+        non_cross = -1j * 1j * 0
+        for elem_ind in range(num_reflectors):
+            ind = elem_ind * num_reflectors
+            f2i = F2[:, elem_ind]
+            f1i = F1[elem_ind, :]
+            i_mat = np.outer(f2i, f1i)
+            non_cross += np.trace(i_mat @ beamformer @ hermetian(i_mat @ beamformer))
+            # Fc += i_mat
+            for elem2_ind in range(num_reflectors - elem_ind - 1):
+                elem2_ind += elem_ind + 1
+                f2j = F2[:, elem2_ind]
+                f1j = F1[elem2_ind, :]
+                j_mat = np.outer(f2j, f1j)
+                coefficient = f1i.T @ beamformer @ hermetian(beamformer) @ np.conjugate(f1j) * hermetian(f2j) @ f2i
+                # coefficient = f1i.T@np.conjugate(f1j)*hermetian(f2j)@f2i
+                # coefficient1 = np.trace(i_mat@transmit_matched_rand@hermetian(transmit_matched_rand)@hermetian(j_mat))
+                coefficients.append(coefficient)
+        # Check that the sum is equal to the original channel
[…]

### 2020-09-24 documentation/Optimization Work/09_23_20.tex
Improved normalization for MSE and updated progress report for taking expection over channels at Tx.

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents
\section{Points from Last Meeting}
\begin{enumerate}
\item{High MSE Issue:} \par
	As discussed with was a normalization error. Now I normalize the transmit vector $\vs$ to have total power 1. I normalize the noise vector $\vn$ in the same way.
	Finally, for the linear beamformer $\mathbf{W}$, in the case of the matched filter, I normalize each column using $\frac{\mathbf{w}_i}{\|\mathbf{w}_i\|^2}$ so that
	$\Expect\left[\mH\mW\right] = \mI$.

	\begin{figure}[H]
	\includegraphics[width= 16cm,height = 13cm]{figures/Corrected_MSE}
	  \caption{MSE with/without optimization versus number of IRS elements with receiver noise.}
	  	  \label{MSE_opt_noise}
	\end{figure}

\item{Expectation Over Elements of Channel:} \par
For the MSE Objective function with deterministic $\mW $ is 
	\begin{equation}
	\Expect\left[\trace\left((\mathbf{s} - \mH \mW \mathbf{s} )  (\mathbf{s} - \mH \mW \mathbf{s} ) ^H  \right)\right]
	\end{equation}
		\begin{equation}
	\trace\left(\Expect\left[ (\mI - \mH \mW  )  (\mI - \mH \mW  ) ^H \right] \Expect\left[ \mathbf{s}\mathbf{s}^H\right]\right)
	\end{equation}
	Assuming equal power allocation here since matrix size are relatively large this becomes
	\begin{equation}
	\trace\left(\Expect\left[ (\mI - \mH \mW  )  (\mI - \mH \mW  ) ^H \right]\right).
	\end{equation}
	Because the beamformer, $\mathbf{W}$, is deterministic at this point, the optimization problem becomes
		\begin{subequations}
	\label{mse_problem_total}
	\begin{align}
	   	    \underset{\boldsymbol{\phi}}{\text{minimize }}
	    & \; - \trace \left( \Expect\left[\mathbf{H} \right] \mathbf{W} \right) - \trace \left( \Expect\left[\mathbf{H} \right]  \mathbf{W}\right)^* +
	    \trace\left(\Expect\left[\mathbf{H}^H \mathbf{H} \right] \mathbf{W} \mathbf{W}^H \right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
	\end{align}
	\end{subequations}	
	Assuming channel components corresponding to the IRS path to be zero-mean, this becomes 
		\begin{subequations}
	\label{mse_problem_total}
	\begin{align}
	   	    \underset{\boldsymbol{\phi}}{\text{minimize }}
	    & \; \trace\left(\Expect\left[\mathbf{H}^H \mathbf{H} \right] \mathbf{W} \mathbf{W}^H \right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
	\end{align}
	\end{subequations}	
	In general we assume the channels through the elements of the IRS are correlated and in this case we use the exponential model.
	Now we decompose the channel $\mH$ into the individual paths through the IRS  $\Sigma_{i=1}^{N_S} \phi_i \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
	\mR_{T}^{\frac{1}{2}}$. Expanding the utility function in \ref{mse_problem_total} shows that all $\phi_i$ cancel out from any non-cross term. The resulting utility function only contains cross terms of the form 
	\begin{equation}
		\trace\left(\Expect\left[ \phi_i \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
	\mR_{T}^{\frac{1}{2}} (\phi_i \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,j}\mathbf{h}^T_{1,j}
	\mR_{T}^{\frac{1}{2}})\right] \right)
	\end{equation}
	or equivalently
		\begin{equation}
		\phi_j ^* \phi_i  \trace\left(\mR_{R} \Expect\left[ \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
	\mR_{T} (\mathbf{h}_{2,j}\mathbf{h}^T_{1,j})^H\right] \right).
	\end{equation}
	Assuming the all elements of $\mathbf{h}_{2,i}$ and $\mathbf{h}^T_{1,i}$ to be independent, these cross terms will also have expected value of zero and 
	thus the final expression does not include any $\phi_i$.
\item{Deterministic Elements of Channel:} \par
	In the above problem, if the transmitter has access to the channel state information for the channels through the IRS, the optimization problem from the last report can be used with results seen in figure \ref{MSE_opt_noise}
\end{enumerate}
\end{document}

### 2020-09-24 simulation/src/lin_alg.py

@@ -81,6 +81,7 @@ def precode_mf(channel, normalize=True):
     ret = hermetian(channel)
     for col in range(ret.shape[1]):
         ret[:, col] = ret[:, col]/np.power(np.linalg.norm(ret[:, col]), 2)
+        # ret[:, col] = ret[:, col]/np.linalg.norm(ret[:, col])
     return ret
 
 def precode_zf(channel):

### 2020-09-24 simulation/testing/Optimization/opt_channel.py

@@ -2,7 +2,7 @@ import numpy as np
 from src.lin_alg import *
 from src.optimization import *
 import matplotlib.pyplot as plt
-average = 100
+average = 10
 num_tx = 16
 num_rx = 10
 IRS_sizes = [16, 64, 128]
@@ -19,6 +19,7 @@ for size in IRS_sizes:
         H_t = F2@F1
         rand = random_phase(num_reflectors)
         beamformer = precode_mf(F2@F1)
+        check = H_t@beamformer
         # beamformer = precode_zf(F2@F1)
         coefficients = []
         non_cross = -1j*1j*0
@@ -62,31 +63,32 @@ for size in IRS_sizes:
         check_2 = np.trace(-H_opt@beamformer) + np.conjugate(np.trace(-H_opt@beamformer)) + np.trace(H_opt@beamformer@hermetian(H_opt@beamformer))
         #   Now perform comparison of optimized phases with random/uniform phases
         # transmit_symbols = BPSK(num_rx)
-        transmit_averge = 100
+        transmit_averge = 10
         trasmit_ave_list_rand = []
         trasmit_ave_list_opt = []
         for i in range(transmit_averge):
-            noise = c_rand(num_rx, 1, var=1).flatten()
-            transmit_symbols = c_rand(num_rx, 1, var=1).flatten()
-            received_rand = F2@F1@beamformer@transmit_symbols
+            noise = c_rand(num_rx, 1).flatten()
+            transmit_symbols = c_rand(num_rx, 1).flatten()
+            received_rand = H_t@beamformer@transmit_symbols + noise
+            check1 = np.power(np.linalg.norm(received_rand - transmit_symbols), 2)
             trasmit_ave_list_rand.append(np.power(np.linalg.norm(received_rand - transmit_symbols), 2))
         #   Line below could be done by going through each element and shifting the phase. Line below is faster.
         #     transmit_matched_opt = precode_mf(F2@optimal_phases@F1)
-            received_opt = F2@optimal_phases@F1@beamformer@transmit_symbols
-            trasmit_ave_list_opt.append(np.power(np.linalg.norm(received_opt - transmit_symbols), 2))
+            received_opt = F2@optimal_phases@F1@beamformer@transmit_symbols + noise
[…]

### 2020-12-17 documentation/Optimization Work/10_17_2020.tex
Back at the paper. Just added a couple of revisions

--- 2020-09-24 documentation/Optimization Work/09_23_20.tex
@@
 \begin{enumerate}
-\item{High MSE Issue:} \par
-	As discussed with was a normalization error. Now I normalize the transmit vector $\vs$ to have total power 1. I normalize the noise vector $\vn$ in the same way.
-	Finally, for the linear beamformer $\mathbf{W}$, in the case of the matched filter, I normalize each column using $\frac{\mathbf{w}_i}{\|\mathbf{w}_i\|^2}$ so that
-	$\Expect\left[\mH\mW\right] = \mI$.
-
-	\begin{figure}[H]
-	\includegraphics[width= 16cm,height = 13cm]{figures/Corrected_MSE}
-	  \caption{MSE with/without optimization versus number of IRS elements with receiver noise.}
-	  	  \label{MSE_opt_noise}
-	\end{figure}
-
-\item{Expectation Over Elements of Channel:} \par
-For the MSE Objective function with deterministic $\mW $ is 
+In the first iteration of CSI collection, 1 pilot is 
+\item{Issue with Expectation Over Elements of Channel:} \par
+	The MSE Objective function with deterministic $\mW $ is 
 	\begin{equation}
@@
 		\begin{equation}
-	\trace\left(\Expect\left[ (\mI - \mH \mW  )  (\mI - \mH \mW  ) ^H \right] \Expect\left[ \mathbf{s}\mathbf{s}^H\right]\right)
+	\trace\left(\Expect\left[ (\mI - \mH \mW  )^H  (\mI - \mH \mW  ) \right] \Expect\left[ \mathbf{s}\mathbf{s}^H\right]\right)
 	\end{equation}
-	Assuming equal power allocation here since matrix size are relatively large this becomes
+	Assuming equal power allocation this becomes
 	\begin{equation}
-	\trace\left(\Expect\left[ (\mI - \mH \mW  )  (\mI - \mH \mW  ) ^H \right]\right).
+	\trace\left(\Expect\left[ (\mI - \mH \mW  )^H  (\mI - \mH \mW  )  \right]\right).
 	\end{equation}
-	Because the beamformer, $\mathbf{W}$, is deterministic at this point, the optimization problem becomes
+	Because the beamformer, $\mathbf{W}$, is deterministic, the optimization problem becomes
 		\begin{subequations}
@@
 	\end{subequations}	
-	Assuming channel components corresponding to the IRS path to be zero-mean, this becomes 
+	Assuming channel components corresponding to the IRS path to be zero-mean (reasonable), this becomes 
 		\begin{subequations}
@@
 	\mR_{T}^{\frac{1}{2}}$.
-Expanding the utility function in \ref{mse_problem_total} shows that all $\phi_i$ cancel out from any non-cross term.
+Expanding the utility function in \ref{mse_problem_total} shows that all $\phi_i$  from the channel $\mH$ cancel out from the non-cross terms.
 The resulting utility function only contains cross terms of the form
@@
 		\trace\left(\Expect\left[ \phi_i \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
-	\mR_{T}^{\frac{1}{2}} (\phi_i \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,j}\mathbf{h}^T_{1,j}
-	\mR_{T}^{\frac{1}{2}})\right] \right)
+	\mR_{T}^{\frac{1}{2}} (\phi_j \mR_{R}^{\frac{1}{2}}\mathbf{h}_{2,j}\mathbf{h}^T_{1,j}
+	\mR_{T}^{\frac{1}{2}})^H \right] \right)
 	\end{equation}
-	or equivalently
+	in which $i \neq j$, or equivalently
 		\begin{equation}
@@
 	\end{equation}
-	Assuming the all elements of $\mathbf{h}_{2,i}$ and $\mathbf{h}^T_{1,i}$ to be independent, these cross terms will also have expected value of zero and 
-	thus the final expression does not include any $\phi_i$.
+	Assuming all elements of the random vectors $\mathbf{h}_{2,i}$ and $\mathbf{h}^T_{1,i}$ to be independent (also for $i \neq j$) these cross terms will also have expected value of zero.
+As a result the final expression does not include any $\phi_i$ meaning that without having knowledge of the actual channels, the MSE cannot be optimized with respect to the phases.
 \item{Deterministic Elements of Channel:} \par

### 2020-12-17 documentation/Paper/abstract.tex
Back at the paper. Just added a couple of revisions

@@
 Surfaces which can control the phase of reflected radio waves, known as intelligent reflective surfaces (IRS), have been the subject of interest in recent years as a means to better control the wireless propagation environment.
-This work investigates the impact of an IRS on the MIMO communication systems.
-Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity.
-A key finding is that the phase control offered by the IRS has no impact in the asymptotic domain.
-Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior.
+This work investigates the impact of an IRS on MIMO (Multiple Input Multiple Output)  communication systems.
+Using the tools of free probability, a MIMO communication system is analyzed in the asymptotic domain to derive an expression for the system capacity.
+A key finding is that the phase control offered by the IRS has no impact on system capacity in the asymptotic domain.
+Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards the derived asymptotic behavior.
 We then propose an optimization method for the non-convex problem of minimizing the Mean Square Error with respect to the IRS phases.

### 2020-12-17 documentation/Paper/background_opening.tex
Back at the paper. Just added a couple of revisions

@@
-The spectrum of electromagnetic frequencies used by wireless communication has been expanded to progressively higher frequencies with new standards.
+In new standards for wireless communication, the spectrum of electromagnetic frequencies used has been expanded to progressively higher frequencies.
 Industry research for the 6G standard has already investigated systems using frequency bands in the 120, 200 and 340 GHz range \cite{Koziol}.
-This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}.
+This shift is motivated by large bands of currently unoccupied frequency and the possibility for decreased antenna size (to enable increased antenna density) \cite{akyildiz2018combating}.
 One primary hurdle towards efficient communication at these frequencies is the high attenuation of radio energy due both to scattering effects and molecular absorption.
-\cite{TODO} show through extensive data collection that relevant attenuation will be (x).
+\cite{TODO} show through extensive empirical results that relevant attenuation can reach (x).
 As a relevant example, radar systems have long used this band of frequencies due to the resulting high resolution.
-For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO}.
-Countering this type of loss is a key step towards meeting the increased data rates and reliability required for future networks. 
+For example, one such system operating in the 2 - 4 GHz band requires 25 kW of power for a detection range of 400 km \cite{TODO} (NEED NEW EXAMPLE).
+Countering this power loss is a key step towards meeting the increased data rates and reliability required for future wireless communication networks. 
 \par
-In previous generations of wireless communications, attenuation due to path loss has been alleviated by reducing the effective transmission distance 
-and supplementing the transmission power using relays \cite{dahlman20134g}.
-Two, key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission (CITE?).
+In previous generations of wireless communications, attenuation due to path loss has been alleviated by incorporating relays into the system to reduce the point-to-point transmission distance and supplement the transmission power \cite{dahlman20134g}.
+Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission (CITE?).
+As of today, most relays are only capable of operating in a half-duplex mode and thus a third cost to the system (by reducing capacity) is often incurred by using relays.
 Deploying enough relays to counter
 the attenuation levels relevant to new standards would pose impractical costs.
-One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable electromagnetic reflection and whose reflective coefficient can be controlled to select the phase of the reflected wave.
-In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
-While this work focuses on the theory of communications channels with such surfaces, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an RIS with 224 elements was built and tested.
+One new solution that has attracted significant interest is the use of surfaces made of many, sub-wavelength sized elements with favorable reflection for the relevant electromagnetic spectrum and whose reflective coefficient can be controlled to select the phase of the reflected wave.
+In this work, such a surface will generally be referred to as an intelligent reflective surface (IRS).
+While this work focuses on the theory of IRS communication channels, it is worth noting that these surfaces have been successfully implemented in works such as \cite{tan2018enabling} in which an IRS with 224 elements was built and tested.
 \par
-The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with a relay.
-In addition, the absence of an rf-chain on the RIS offers some benefits over relays.
-First, reflection from an RIS does not introduce the noise associated with an rf-chain.
-Second, without the band-pass filters of an rf-chain, RISs have a frequency response over the entire spectrum (explain).
-Third, the RIS is inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
-The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
-Consider a single antenna, single receiver channel using an RIS with $N$ reflective elements.
-In this case, the phase control at the RIS can be used to add the multiple paths coherently at the receiver and intuitively increase the received signal power.
+The motivating use for the IRS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
+associated with an active relay.
+In addition, the absence of an rf-chain on the IRS offers some benefits over relays.
+First, reflection from an IRS does not introduce the noise associated with an rf-chain.
+Second, without the band-pass filters of an rf-chain, IRSs have a frequency response over the entire spectrum (explain).
+Third, the IRS is inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
+The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of IRSs and a few notable points are included here.
+Consider a single antenna, single receiver channel using an IRS with $N$ reflective elements.
+In this case, the phase control at the IRS can be used to add the multiple paths coherently at the receiver and intuitively increase the received signal power.
 While such results are motivating, there are two points that should be noted.
@@
 Second, it is important to note that scaling of received SNR proportional to $N^2$ will only hold to the point limited by energy conservation \cite{bjornson2019demystifying}.
-This leaves us with a question. What is the true potential benefits for the RIS in future systems?
-Before investigating this question we review key aspects of MIMO systems need to fully develop a system model with which to explore the RIS.
+This leaves us with a question. What is the true potential benefits for the IRS in future systems?
+Before investigating this question we review key aspects of MIMO systems need to fully develop a system model with which to explore the IRS.

### 2020-12-18 documentation/Paper/abstract.tex
Revisions on reread through

@@
 A key finding is that the phase control offered by the IRS has no impact on system capacity in the asymptotic domain.
-Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards the derived asymptotic behavior.
+Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards the derived asymptotic behavior`.
 We then propose an optimization method for the non-convex problem of minimizing the Mean Square Error with respect to the IRS phases.

### 2020-12-18 documentation/Paper/correlation_review.tex
Revisions on reread through

@@
 In the following, a number of these studies are reviewed.
-Some of these chosen works use channel correlation models similar to the one we will develop for the RIS system model.
+Some of these chosen works use channel correlation models similar to the one we will develop for the IRS system model.
 In particular, this method models the correlation between antennas as decreasing  exponentially with the distance between antennas.
-Definitely need to show this is a valid model (Cite).
+TODO Definitely need to show this is a valid model (Cite).
 \par
@@
 \par
-One method that has been shown to reduce the complexity of this calculation is to allow the system size $N \rightarrow \infty$ while normalizing the received power. 
-This is motivated by the observation that when the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution which will be referred to as the Asymptotic Eigenvalue Distribution (AED).
-Because the received power would generally not be normalized as the system size in increased, results found while using this normalization be extended by simply scaling the asymptotic result with $N$.
+One method that has been shown to reduce the complexity of theoretic evaluation of MIMO systems is to allow the system size $N \rightarrow \infty$ while normalizing the received power. 
+This is motivated by the observation that when the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution often referred to as the Asymptotic Eigenvalue Distribution (AED).
+Because the received power would generally not be normalized as the system size in increased, results found while using this normalization be extended by simply scaling the asymptotic result with $N$ (Can't we assume constant transmit power such that we do actually have this normalization?).
 (Still need to clarify the types of MIMO gains that are observed under this normalization and how to deal with non-zero mean components).
@@
 Second, what values for the parameters of the correlation model are reasonable for a real-world system?
-In test data collected with channel parameters (carrier frequency, bandwidth, power...) relevant to 3G and 4G, wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ (linear array?) system \cite{martin2000multiple} (TODO justify if this is a comparison for current systems).
+In test data collected with channel parameters (carrier frequency, bandwidth, power...) relevant to 3G and 4G, wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ (linear array?) system \cite{martin2000multiple} (TODO justify if this is a compaIRSon for current systems).

### 2021-02-14 documentation/Optimization Work/09_25_20.tex
pre pull commit 02/14/21

--- 2020-12-17 documentation/Optimization Work/10_17_2020.tex
@@
 \begin{enumerate}
-In the first iteration of CSI collection, 1 pilot is 
+\item{High MSE Issue:} \par
+	As we discussed, this issue was a normalization error. Now I am normalizing both the transmit vector $\vs$  and noise vector $\vn$ to have power 1. 
+	For the linear beamformer $\mathbf{W}$, in the case of the matched filter, I normalize each column using $\frac{\mathbf{w}_i}{\|\mathbf{w}_i\|^2}$ so that
+	$\Expect\left[\mH\mW\right] = \mI$ when using the matched filter with zero mean, independent channel elements.
+	Figure  \ref{MSE_opt_noise} shows the results for the MSE using randomly chosen phases versus  optimized phases (NOTE in this optimization the transmitter is assumed to 		know the channels corresponding to each IRS element).
+	\begin{figure}[H]
+	\includegraphics[width= 16cm,height = 13cm]{figures/Corrected_MSE}
+	  \caption{MSE with/without optimization versus number of IRS elements with receiver noise.}
+	  	  \label{MSE_opt_noise}
+	\end{figure}
+
 \item{Issue with Expectation Over Elements of Channel:} \par

### 2021-02-14 documentation/Paper/irs_channel.tex
pre pull commit 02/14/21

@@
 	\end{align}
+
+This equivalence is shown by applying the Weinstein-Aronszajn identity to the characteristic polynomials of the to matrices.
 In order to continue decomposing this covariance matrix into sub-components, we exploit a property of the S-Transform given by (proven in Appendix \ref{rotation_property})
@@
 \end{equation}
-If all system dimensions are $N$, these S-Trasforms equal. For simplified analysis and notation, we will assume that all system dimensions are $N$.
+If all system dimensions are $N$, these S-Trasforms are equal. For simplified analysis and notation, we will assume that all system dimensions are $N$.
 Repeated iteration of the property above gives the equivalent matrix

### 2021-02-14 documentation/Paper/background_opening.tex

@@
 \par
-The motivating use for the IRS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
-associated with an active relay.
-In addition, the absence of an rf-chain on the IRS offers some benefits over relays.
-First, reflection from an IRS does not introduce the noise associated with an rf-chain.
-Second, without the band-pass filters of an rf-chain, IRSs have a frequency response over the entire spectrum (explain).
-Third, the IRS is inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on an active relay.
+The motivating use for the IRS is to alleviate the power attenuation problem without incurring the costs of a relay rf-chain and transmission power.
+The absence of an rf-chain on the IRS also means the reflection from an IRS does not introduce the noise associated with an rf-chain (detail where this comes from) and that the IRS has a frequency response of the entire spectrum unlike band pass filter of rf-chains (explain).
 The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of IRSs and a few notable points are included here.
 Consider a single antenna, single receiver channel using an IRS with $N$ reflective elements.
-In this case, the phase control at the IRS can be used to add the multiple paths coherently at the receiver and intuitively increase the received signal power.
+In this case, the phase control at the IRS can be used to ensure the multiple paths add coherently at the receiver.
 While such results are motivating, there are two points that should be noted.
-First, in this example from \cite{basar2019wireless}, the received power would still scale with $N^2$ even if the reflective surfaces had no control over the reflection phase.
-Second, it is important to note that scaling of received SNR proportional to $N^2$ will only hold to the point limited by energy conservation \cite{bjornson2019demystifying}.
-This leaves us with a question. What is the true potential benefits for the IRS in future systems?
-Before investigating this question we review key aspects of MIMO systems need to fully develop a system model with which to explore the IRS.
+First, in this example from \cite{basar2019wireless}, the received power would still scale with $N$ even if the reflective surfaces had no control over the reflection phase.
+Second, it is important to note that the $N^2$ scaling of received SNR is constrained by the conservation of transmitted energy, a point rigorously treated in \cite{bjornson2019demystifying}.
+So what are the potential benefits for the IRS in future systems?
+Before investigating this question we review key aspects of MIMO systems needed to fully develop a system model with which to explore the IRS.

### 2021-02-15 documentation/Meeting Notes/02_14_21.tex
using Skupch's method for capacity analysis for single sided correlation. TODO extend to double sided correlation

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Status}
\begin{itemize}
\item
From 5/29/20 Update (yikes almost a year ago...)
\item
Known S-Transform for the exponential correlation model comes from \cite{skupch2005free}
\item
Ideally, the outline should go like:
\begin{enumerate}
\item 
 Show phases cancel and the capacity for exponential correlation (for high correlations). Note that the S-Transform of the correlation matrix is only valid for infinite size.
\item 
Look at ways of increasing capacity (and rate) for the small cases. Starting with $N_T = N_R = 1$.s
\item 
Conclude by looking at system sizes for which the asymptotic values represent the system well.
\item
For proving freeness, skupch also says that for H being gaussian, it is a wishart matrix and it is free w.r.t the correlation... Need to show that it is okay as long as the correlation matrices are not next to oneanother?
\end{enumerate}
\end{itemize}
\bibliography{bibliography}

\end{document}

### 2021-02-15 simulation/MATLAB/numerical_correlation.m

@@ -21,11 +21,13 @@ clear all;
 function ave = average_numeric_capacity(rows, cols, rho, average)
     vals = zeros(average,1);
     for i = 1:length(vals)
+        channel1 = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
         channel2 = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
         correlation = exponential_correlation(rows, rho);
 %         total_channel = channel1;
 %         total_channel = channel2*channel1;
-        total_channel = correlation*channel2*correlation*channel1*correlation;
+        total_channel = correlation*channel2*channel1*correlation;
+%         total_channel = correlation*channel2*correlation*channel1*correlation;
         total_cov = (total_channel*total_channel');
         vals(i) = MIMO_capacity(total_cov, 1/cols);
     end

### 2021-02-15 simulation/MATLAB/useful_functions/skupch.m

@@ -1,0 +1,1 @@
+function output = skupch(input, beta)

### 2021-02-22 documentation/Meeting Notes/02_21_21.tex
Updated a report and generated some figures

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/21_02_21} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Overview Status}\label{overview}
The hope is to show the following points in the paper to present both some analytic and numerical/optimization results.
\begin{enumerate}
\item \label{analysis_point}
Show the capacity for the IRS channel model with the exponential correlation model. See Section \ref{analysis_section} for status.
\item
Show that phases of the IRS will cancel out of the MIMO capacity term as $N \rightarrow \infty	$. (Done as shown in previous reports).
\item 
Look at ways of increasing capacity (or improving another metric like MSE) for the small system cases.
See Section \ref{opt_section} for status.
\item 
Conclude by looking at the convergence of the correlated IRS MIMO system towards the asymptotic results. This is easy to look at numerically but lacks much meaning without the comparison to the analytic results that I am having difficulty generating (from point 1 above).
\end{enumerate}
\section{Analysis Status}\label{analysis_section}
Using the S-Transform for the exponential correlation model shown in the work from \cite{skupch2005free} should allow me to find the analytic capacity, however, I have an issue when I attempt to recreate these results for the single rayleigh channel with receiver correlation (i.e. solving a quartic polynomial for the stieltjes transform $G(s)$). The problem I have is how to choose 1 of the 4 roots of the polynomial (for each point in the pdf, I get 4 roots to potentially evaluate). See resulting pdfs/capacity in figures \ref{AED} and \ref{capacity}). Similarly, if I want to extend this to the case with correlation at BOTH the transmitter and receiver, the polynomial is a 5th order polynomial and I'm not sure how to pick the correct root for the AED.

	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
	  \caption{AEDs generated using results from \cite{skupch2005free} }
	  \label{AED}
	\end{figure}
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/analytic_capacity}
	  \caption{Capacities corresponding to AEDs from Figure \ref{AED}}
	  \label{capacity}
	\end{figure}

\section{Optimization Status}\label{opt_section}
For the case $N_T = N_R = 1$ and full CSIT, if the values in $\boldsymbol{\Phi}$ (the IRS phases) are chosen such that all paths from the transmitter to receiver add coherently, this would result in an $N_S^2$ increase in SNR at the receiver (where $N_S$ is the number of IRS elements). However, for the case $N_T = N_R = N_S$ $\rightarrow \infty$ this coherent addition can only be done for a single Tx, Rx pair so the impact of this path will diminish as the power is distributed across an increasing number of antennas. This intuition agrees agrees with the analytic result that $\Phi$ cancels out of the AED.
\par
For small $N_T$, $N_R$ there is still be some benefit. 
\subsection{Full CSIT MSE optimization}
For the case with full CSIT, I wrote a method for the optimization of the phases assuming that the transmitter uses a matched filter (this can be replaced by another precoder). I'm not sure if it would be better to choose a metric other than mean-square-error to numerically evaluate this optimization but some initial results are in figure \ref{opt_results}.

		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/opt_results}
	  \caption{Capacities corresponding to AEDs from Figure \ref{AED}}
	  \label{opt_results}
	\end{figure}

This can be extended to the case in which pilots are collected for $< N_S$  IRS elements (rather than the full CSIT considered above.
\bibliography{bibliography}

\end{document}

### 2021-02-22 simulation/MATLAB/known_csi_capacity.m

@@ -1,0 +1,4 @@
+ function capacity = known_csi_capacity(x, pdf, snr, number_receivers, step)
+    probabilities = pdf.*step;
+    capacity = number_receivers.*sum(log2(1 + snr.*x).*probabilities);
+ end

### 2021-02-22 simulation/MATLAB/projection_s_transform.m

@@ -6,7 +6,7 @@ top = 7;
 x_values = linspace(bottom, top , steps);
 step_size = (top-bottom)/steps;
 s_values = x_values + 1i*1e-7; % The img component needs to go to zero to evaluate pdf.
-betas = linspace(0.5, 1, 10);
+betas = linspace(0.3, 1, 10);
 numeric_capacity = zeros(size(betas));
 asymptotic_capacity = zeros(size(betas));
 figure(1)
@@ -15,7 +15,8 @@ for i = 1:length(betas)
 %     stieltjes_values = (1./s_values).*(1+gamma_s(1./s_values, betas(i)));
     stieltjes_values = evaluate_points_skupch(s_values,betas(i));
     pdf = 1/pi .* imag(stieltjes_values);
-%     pdf = abs(pdf);
+    % Really cheating here because I'm not returning the proper roots.
+    pdf = abs(pdf);
     asymptotic_capacity(i)  = aed_capacity(x_values, pdf, 1/cols, rows, step_size);
     plot(x_values,pdf,'DisplayName', '\beta = ' + string(betas(i)));
     title('AED');
@@ -68,6 +69,9 @@ poly_vector = [eval_point^4  eval_point^3*(-2 -2*beta*eval_point) ...
 zeros = roots(poly_vector);
 % eval(check)
 % output = check(1);
+% Return zero with largest img component.
+% [max_imag, max_imag_ind] = max(imag(zeros));
+% output = zeros(max_imag_ind);
 output = zeros(3);
 if (imag(output) < 0)
 %     error('Strictly positive PDF??');
@@ -123,16 +127,12 @@ end
 function ave = average_numeric_capacity(rows, cols, beta, average)
     vals = zeros(average,1);
     for i = 1:length(vals)
-        awgn = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
+        instance_rayleigh = rayleigh_channel(rows, cols, 1/sqrt(2*rows));
 %         awgn = rayleigh_channel(rows, cols, 1/sqrt(2*(rows*beta)));
-        projector1 = projection_matrix(rows, beta);
-        projector2 = projection_matrix(rows, beta);
-%         correlation = exponential_correlation(rows, rho);
-%         total_channel = channel1;
-%         total_channel = channel2*channel1;
-%         total_channel = correlation*channel2*correlation*channel1*correlation;
-        total_channel = awgn*projector1;
-%         total_channel = projector2*awgn*projector1*awgn;
[…]

### 2021-02-22 simulation/testing/Optimization/opt_channel.py

@@ -3,9 +3,10 @@ from src.lin_alg import *
 from src.optimization import *
 import matplotlib.pyplot as plt
 average = 10
-num_tx = 16
-num_rx = 10
-IRS_sizes = [16, 64, 128]
+num_tx = 8
+num_rx = 8
+num_optimization_rounds = 6
+IRS_sizes = [4, 8, 16, 64, 128]
 MSE_opt_size = []
 MSE_rand_size = []
 for size in IRS_sizes:
@@ -19,34 +20,36 @@ for size in IRS_sizes:
         H_t = F2@F1
         rand = random_phase(num_reflectors)
         beamformer = precode_mf(F2@F1)
-        check = H_t@beamformer
-        # beamformer = precode_zf(F2@F1)
         coefficients = []
         non_cross = -1j*1j*0
         for elem_ind in range(num_reflectors):
-            ind = elem_ind * num_reflectors
+            #   Pick out rows and columns of the channels corresponding to specific IRS elements
             f2i = F2[:, elem_ind]
             f1i = F1[elem_ind, :]
             i_mat = np.outer(f2i, f1i)
             non_cross += np.trace(i_mat@beamformer@hermetian(i_mat@beamformer))
-            # Fc += i_mat
+            # Add all of the cross components of the polynomial to the list of coefficients.
             for elem2_ind in range(num_reflectors - elem_ind - 1):
                 elem2_ind += elem_ind + 1
                 f2j = F2[:, elem2_ind]
                 f1j = F1[elem2_ind, :]
                 j_mat = np.outer(f2j, f1j)
+                # Below wraps around to get complex scalar by the trace rotation property.
                 coefficient = f1i.T@beamformer@hermetian(beamformer)@np.conjugate(f1j)*hermetian(f2j)@f2i
                 # coefficient = f1i.T@np.conjugate(f1j)*hermetian(f2j)@f2i
                 # coefficient1 = np.trace(i_mat@transmit_matched_rand@hermetian(transmit_matched_rand)@hermetian(j_mat))
                 coefficients.append(coefficient)
-        # Check that the sum is equal to the original channel
+        # Create matrix to pick out the MSE terms influenced by specific IRS elements.
         phase_adjacency = np.concatenate((np.zeros((num_reflectors, len(coefficients))), np.eye(num_reflectors)), axis=1)
         # correct = np.allclose(Fc, F2@F1)
+        #TODO check if this should actually be 0j
[…]

### 2021-02-22 simulation/testing/Optimization/opt_num_CSI.py

@@ -61,7 +61,7 @@ for percent in csi_percentages:
                 elem3_ind = elem2_ind + elem_ind + 1
                 phase_adjacency[elem3_ind, ind + elem2_ind] = 1
             ind += num_reflectors - elem_ind - 1
-        optimal_phases, improvements = optimize_phases(np.asarray(coefficients),
+        optimal_phases, improvements = optimize_phases_preselected(np.asarray(coefficients),
                                                        phase_adjacency, F2, F1, csi=percent, num_repetitions=2,
                                                        improvements=True)
         H_opt = F2 @ optimal_phases @ F1

### 2021-02-22 simulation/testing/Optimization/small_system_evaluation.py

@@ -1,0 +1,103 @@
+# A test for looking at impact of adding correlation through concatenated channels on the channel capacity
+# Takeaway:
+# If power is allocated equally the ave capacity doesn't seem to be affected however variance increase
+# with increasing correlation. However if water-filling
+# is used then the correlated channels seem to have increasing capacity as correlation increases
+
+import matplotlib.pyplot as plt
+from src.lin_alg import *
+from src.optimization import *
+from src.network_simulation_tools import water_filling
+import seaborn
+
+fig = plt.figure()
+ave_plt = fig.add_subplot(2, 1, 1)
+var_plt = fig.add_subplot(2, 1, 2)
+num_channel_average = 100
+num_symbol_average = 100
+num_opt_rounds = 3
+IRS_sizes = [16, 32]
+num_rx = 8
+num_tx = 8
+
+correlations = np.linspace(0, 1-10e-6, 5)   # Don't want to go all the way to 1, just close.
+# Evaluate over different correlations.
+for num_reflectors in IRS_sizes:
+    MSE_variance = []
+    MSE_average = []
+    MSE_variance_opt = []
+    MSE_average_opt = []
+    capacity_average = []
+    capacity_variance = []
+    capacity_average_opt = []
+    capacity_variance_opt = []
+    # For each correlation level, evaluate over different system sizes.
+    for cor_rho in correlations:
+        t_correlation_matrix = exponential_correlation(num_tx, cor_rho)
+        r_correlation_matrix = exponential_correlation(num_rx, cor_rho)
+        cor_cap = []
+        cor_cap_opt = []
+        for i in range(num_channel_average): #  Average over channel realizations
+            F1 = c_rand(num_reflectors, num_tx, var=1)
+            F2 = c_rand(num_rx, num_reflectors, var=1)
+            H_t = F2@F1
+            H_cor = r_correlation_matrix@F2@random_phase(num_reflectors)@F1@t_correlation_matrix
+            # Initial beamformer -> can be updated in subsequent iterations.
+            beamformer = precode_mf(H_cor)
+            # Loop for iterating between optimizing MF then phases.
+            # Because phase optimization is just an improvement, reiterate (need to make sure this is in a different
+            # order each time!!?)
+            coefficients = get_polynomial_coefficients(F1@t_correlation_matrix, r_correlation_matrix@F2, beamformer, num_reflectors)
[…]

### 2021-03-16 documentation/Meeting Notes/03_15_21.tex
adding figures and tex for report

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/21_02_21} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Optimization Results}\label{overview}
These results consider a Multiple input Single output channel for which the transmitter sends the same signal across all antennas and using a Maximum Ratio Transmission scheme once the the phases at the IRS have been chosen (two step optimization). The LOS path is $ $
\begin{enumerate}
\item 
First just a figure verifying the basic MRT system is setup correctly for the case without any IRS.
When power is normalized at the transmitter for increasing numbers of antennas, the difference is quite small. 
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
	  \caption{MISO system without IRS using Maximum Ratio Transmission}
	  \label{basic_MRT}
	\end{figure}

\item 
Comparison of MISO system with and without IRS. Note that the IRS effectively adds power to the received signal at the received because of the new path.
Also note that the phases in this case are chosen randomly.
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
	  \caption{ Comparison of MISO channel with/without an IRS path (non-optimized phases)}
	  \label{AED}
	\end{figure}

\item 
Comparison of MISO system with IRS for cases with and without the two-stage IRS phase optimization. For the IRS optimized case, we consider full CSI at the transmitter.
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
	  \caption{MISO, IRS system with/without phase optimization}
	  \label{AED}
	\end{figure}

\item 
Comparison of MISO system with IRS and two-stage IRS phase optimization across different number of pilot symbols sent by the transmitter. 
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
	  \caption{MISO, IRS system with phase optimization for different # of pilot symbols.}
	  \label{AED}
	\end{figure}

\end{enumerate}

\bibliography{bibliography}

\end{document}

### 2021-03-16 simulation/testing/Optimization/csi_comparison_mrt_with_irs.py

@@ -11,19 +11,19 @@ import time
 
 num_simulations = 100
 per_channel_symbols_num = 1000
-SNRs_dB = np.linspace(-10, 15, 5)
+SNRs_dB = np.linspace(-10, 15, 10)
 SNRs = np.power(10, SNRs_dB/10)
 num_rx = 1
 surface_num_list = [16, 64, 256]
-num_surfaces  = 16
+num_surfaces = 16
 num_tx = 8
 alphabet = np.round(np.linspace(-1, +1, 2))
 IRS_partial = 1
-CSI_amount_list = [.25, .5, .75, 1]
+CSI_amount_list = [0, .5, .75, 1]
 MRT_partial = 1
 fig = plt.figure(1)
 for IRS_partial in CSI_amount_list:
-    BER_list = []
+    # BER_list = []
     BER_list_opt = []
     for snr in SNRs:
         errors = 0
@@ -33,33 +33,33 @@ for IRS_partial in CSI_amount_list:
             H1 = c_rand(num_surfaces, num_tx)
             h2 = np.random.standard_normal((num_surfaces, 1))/np.sqrt(2) + 1j*np.random.standard_normal((num_surfaces, 1))/np.sqrt(2)
             phase = optimize_MISO_phase(np.sum(los_channel), h2.T*np.sum(H1, 1), relative_pilots=IRS_partial)
-            beamformer_mrt = maximum_ratio_transmission((h2.T@np.eye(num_surfaces)@H1+los_channel).T)
+            # beamformer_mrt = maximum_ratio_transmission((h2.T@np.eye(num_surfaces)@H1+los_channel).T)
             beamformer_opt = maximum_ratio_transmission((h2.T@np.diag(phase)@H1+los_channel).T, partial=MRT_partial)
             for i in range(per_channel_symbols_num):
                     symbol = alphabet[np.random.randint(0, 2)]
                     noise = np.random.standard_normal()*np.sqrt((1 / (snr)))
-                    received_signal = (h2.T@np.eye(num_surfaces)@H1+los_channel)@\
-                                      (symbol*beamformer_mrt/np.sqrt(num_tx)) + noise
+                    # received_signal = (h2.T@np.eye(num_surfaces)@H1+los_channel)@\
+                    #                   (symbol*beamformer_mrt/np.sqrt(num_tx)) + noise
                     received_signal_opt = (h2.T@np.diag(phase)@H1+los_channel)@\
[…]

### 2021-04-09 documentation/Meeting Notes/03_15_21.tex
post 4_09_21 meeting with Ali

@@
 \section{Optimization Results}\label{overview}
-These results consider a Multiple input Single output channel for which the transmitter sends the same signal across all antennas and using a Maximum Ratio Transmission scheme once the the phases at the IRS have been chosen (two step optimization).
-The LOS path is $ $
+These results consider a Multiple input Single output channel for which the transmitter sends the same signal across all antennas using a Maximum Ratio Transmission scheme once the the phases at the IRS have been chosen (two step optimization).
+NOTE: The first four figures are just validating that the simulation was setup correctly. The fifth figure shows comparisons of the system with/without optimization.
 \begin{enumerate}
@@
 	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
-	  \caption{MISO system without IRS using Maximum Ratio Transmission}
+	\includegraphics[width= 15cm,height = 10cm]{figures/16_03_21/MRT_vs_Tx_normalized}
+	  \caption{MISO system without IRS using Maximum Ratio Transmission. Power normalized across number of Transmitters}
+	  \label{basic_MRT}
+	\end{figure}
+		\begin{figure}[H]
+	\includegraphics[width= 15cm,height = 10cm]{figures/16_03_21/MRT_vs_Tx_not_normalized}
+	  \caption{MISO system without IRS using Maximum Ratio Transmission. Power increasing similar to Maximum Ratio Combining for SIMO.}
 	  \label{basic_MRT}
@@
 	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
+	\includegraphics[width= 15cm,height = 10cm]{figures/16_03_21/MRT_With_Without_IRSPath}
 	  \caption{ Comparison of MISO channel with/without an IRS path (non-optimized phases)}
@@
 	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
+	\includegraphics[width= 15cm,height = 10cm]{figures/16_03_21/Optimized_vs_unoptimized}
 	  \caption{MISO, IRS system with/without phase optimization}
@@
 	\begin{figure}[H]
-	\includegraphics[width= 15cm,height = 10cm]{figures/21_02_21/aed}
-	  \caption{MISO, IRS system with phase optimization for different # of pilot symbols.}
+	\includegraphics[width= 15cm,height = 10cm]{figures/16_03_21/Optimized_vs_unoptimized}
+	  \caption{MISO, IRS system with phase optimization for different numbers of pilot symbols.}
 	  \label{AED}
@@
 
-\bibliography{bibliography}
-
 \end{document}

### 2021-04-09 simulation/testing/Optimization/partial_csi_mrt_no_irs.py

@@ -4,13 +4,14 @@ from src.optimization import optimize_MISO_phase
 from src.lin_alg import c_rand
 import matplotlib.pyplot as plt
 import numpy as np
+import time
 
 
 # BER evaluation framework
 
 num_simulations = 100
 per_channel_symbols_num = 1000
-SNRs_dB = np.linspace(-10, 15, 5)
+SNRs_dB = np.linspace(-10, 15, 8)
 SNRs = np.power(10, SNRs_dB/10)
 num_rx = 1
 num_surfaces = 1
@@ -53,6 +54,6 @@ title = "IRS + LOS "
 plt.title(title)
 plt.grid(True)
 plt.legend(loc='lower left')
-plt.show()
-time_path = f"Output/MRT_With_Without_IRSPath.png"
+# plt.show()
+time_path = f"Output/MRT_With_Without_IRSPath+" + f"{time.time()}" + ".png"
 fig.savefig(time_path, format="png")
